### 2022-07-06 Test parsing of Proff table.Rmd
Parsing experiments - Proff.no table. Creating Links folder. Uploading source html file.

---
title: "Test parsing of Proff.no table (Kreftforeningen)"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

NOT a model to be followed...

```{r}
library(tidyverse)
library(rvest)
```

```{r}
# Reading saved html
table_kreft <- read_html("Links/Kreftforeningen.html") |> 
  html_node("#inner-frame") |> 
  html_table()

# Removing white space
table_kreft2$REGNSKAPSPERIODE <- table_kreft$REGNSKAPSPERIODE |> str_squish()

# Removing empty column
table_kreft3 <- table_kreft2 |> select(REGNSKAPSPERIODE:`2002`)

# Removing duplicate rows
table_kreft4 <- table_kreft3 |> filter(!grepl("Lukk", REGNSKAPSPERIODE))

# tidying data:

table_kreft5 <- table_kreft4 |>
  pivot_longer(`2021`:`2002`, names_to = "year")

table_kreft6 <- table_kreft5

# Turning "-" into NA
table_kreft6 <- table_kreft6 |> 
  mutate(value = na_if(value, "-"))

# Turning years into integers
table_kreft6$year <-as.integer(table_kreft6$year)

# Removing unnecessary dates
table_kreft7 <- table_kreft6 |> 
  filter(
    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
      REGNSKAPSPERIODE != "Valutakode")

# Removing whitespace
table_kreft8 <- table_kreft7 |> 
  mutate(
    value = str_replace_all(value, "\\s", "")
  )

# Converting to numbers
table_kreft8 <- table_kreft8 |> 
  mutate(
    value = as.numeric(value)
  )

# Viewing table
table_kreft8 |>
  arrange(desc(year))

# Quick visualisation
table_kreft8 |> 
  filter(year == 2021) |> 
  ggplot(aes(REGNSKAPSPERIODE, value)) +
  geom_col() +
  coord_flip()
```

Note that we might have to convert implicitly negative numbers.

```{r}

```

### 2022-07-07 Test parsing of Proff table.Rmd
Fixed missing intermediate df. Style fixes.
Testing with different organisation. Unfortunately errors. Creating function.

@@
 
+table_kreft2 <- table_kreft
+
 # Removing white space
-table_kreft2$REGNSKAPSPERIODE <- table_kreft$REGNSKAPSPERIODE |> str_squish()
+table_kreft2$REGNSKAPSPERIODE <- table_kreft$REGNSKAPSPERIODE |>
+  str_squish()
 
 # Removing empty column
-table_kreft3 <- table_kreft2 |> select(REGNSKAPSPERIODE:`2002`)
+table_kreft3 <- table_kreft2 |>
+  select(REGNSKAPSPERIODE:`2002`)
 
 # Removing duplicate rows
-table_kreft4 <- table_kreft3 |> filter(!grepl("Lukk", REGNSKAPSPERIODE))
+table_kreft4 <- table_kreft3 |>
+  filter(!grepl("Lukk", REGNSKAPSPERIODE))
 
@@
 
-# Viewing table
-table_kreft8 |>
-  arrange(desc(year))
-
 # Quick visualisation
@@
   coord_flip()
+
+# Removing values that are years
+table_kreft9 <- table_kreft8 |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+
+# Viewing table
+table_kreft9 |>
+  arrange(desc(year))
 ```
@@
 
-```{r}
+```{r opera}
+# Reading saved html
+opera <- read_html("Links/opera.html") |> 
+  html_node("#inner-frame") |> 
+  html_table()
+
+opera2 <- opera
+
+# Removing white space
+opera2$REGNSKAPSPERIODE <- opera2$REGNSKAPSPERIODE |>
+  str_squish()
+
+# Removing empty column
+opera3 <- opera2 |>
+  select(REGNSKAPSPERIODE:`2002`)
+
+# Removing duplicate rows
+opera4 <- opera3 |>
+  filter(!grepl("Lukk", REGNSKAPSPERIODE))
+
+# tidying data:
+
+opera5 <- opera4 |>
+  pivot_longer(`2021`:`2002`, names_to = "year")
+
+opera6 <- opera5
+
+# Turning "-" into NA
+opera6 <- opera6 |> 
+  mutate(value = na_if(value, "-"))
+
+# Turning years into integers
+opera6$year <-as.integer(opera6$year)
+
+# Removing unnecessary dates
+opera7 <- opera6 |> 
+  filter(
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode")
+
+# Removing whitespace
+opera8 <- opera7 |> 
+  mutate(
+    value = str_replace_all(value, "\\s", "")
+  )
+
+# Converting to numbers
+opera8 <- opera8 |> 
+  mutate(
+    value = as.numeric(value)
+  )
+
+# Removing values that are years
+opera9 <- opera8 |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+
+# Viewing table
+opera9 |>
+  arrange(desc(year))
+
+opera9 |> 
+  mutate(
+    error = year - value
+  ) |> 
+  filter(error == 1)
+
+# Result: duplicated entries with wrong year
+```
+
+```{r single operation}
+opera_test <- read_html("Links/opera.html") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select(
+    REGNSKAPSPERIODE:`2002`) |> 
+  mutate(
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+  filter(
+    !grepl("Lukk", REGNSKAPSPERIODE)) |> 
+  pivot_longer(
+    `2021`:`2002`, names_to = "year") |> 
+  mutate(
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  filter(
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate(
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate(
+    value = as.numeric(value)
+  ) |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+
+read_proff <- function(file){
+  read_html(file) |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select(
+    REGNSKAPSPERIODE:`2002`) |> 
+  mutate(
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+  filter(
+    !grepl("Lukk", REGNSKAPSPERIODE)) |> 
+  pivot_longer(
+    `2021`:`2002`, names_to = "year") |> 
+  mutate(
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  filter(
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate(
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate(
+    value = as.numeric(value)
+  ) |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+}
+
+read_proff("Links/opera.html")
 

### 2022-07-07 Test parsing of Proff table.Rmd
Hackish removal of duplicate rows. Noticed that these go from 1998...

--- 2022-07-06 Test parsing of Proff table.Rmd
@@
 
+table_kreft
+
+table_kreft2 <- table_kreft
+
 # Removing white space
-table_kreft2$REGNSKAPSPERIODE <- table_kreft$REGNSKAPSPERIODE |> str_squish()
+table_kreft2$REGNSKAPSPERIODE <- table_kreft$REGNSKAPSPERIODE |>
+  str_squish()
 
 # Removing empty column
-table_kreft3 <- table_kreft2 |> select(REGNSKAPSPERIODE:`2002`)
+table_kreft3 <- table_kreft2 |>
+  select(REGNSKAPSPERIODE:`2002`)
 
 # Removing duplicate rows
-table_kreft4 <- table_kreft3 |> filter(!grepl("Lukk", REGNSKAPSPERIODE))
+table_kreft4 <- table_kreft3 |>
+  filter(!grepl("Lukk", REGNSKAPSPERIODE))
 
@@
 
-```{r}
+```{r opera}
+# Reading saved html
+opera <- read_html("Links/opera.html") |> 
+  html_element("#inner-frame") |> 
+  html_table(trim = TRUE)
+
+# Removing white space
+opera$REGNSKAPSPERIODE <- opera$REGNSKAPSPERIODE |>
+  str_squish()
+
+# Removing empty column
+opera <- opera |>
+  select(REGNSKAPSPERIODE:`1998`)
+
+# Removing duplicate rows (hack)
+opera <- opera |>
+  filter(
+  !row_number() > 176
+)
+
+# Removing duplicate rows
+opera <- opera |>
+  filter(!grepl("Lukk", REGNSKAPSPERIODE))
+
+# tidying data:
+opera <- opera |>
+  pivot_longer(`2021`:`1998`, names_to = "year")
+
+# Turning "-" into NA
+opera <- opera |> 
+  mutate(value = na_if(value, "-"))
+
+# Turning years into integers
+opera$year <-as.integer(opera$year)
+
+# Removing unnecessary dates
+opera <- opera |> 
+  filter(
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode")
+
+# Removing whitespace
+opera <- opera |> 
+  mutate(
+    value = str_replace_all(value, "\\s", "")
+  )
+
+# Converting to numbers
+opera <- opera |> 
+  mutate(
+    value = as.numeric(value)
+  )
+
+# Removing values that are years
+opera <- opera |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+
+# Viewing table
+opera |>
+  arrange(desc(year))
 
 ```
+
+```{r single operation}
+opera_test <- read_html("Links/opera.html") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select(
+    REGNSKAPSPERIODE:`1998`) |> 
+  mutate(
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+  filter(
+  !row_number() > 176
+  ) |> 
+  filter(
+    !grepl("Lukk", REGNSKAPSPERIODE)) |> 
+  pivot_longer(
+    `2021`:`1998`, names_to = "year") |> 
+  mutate(
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  filter(
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate(
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate(
+    value = as.numeric(value)
+  ) |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+
+identical(opera, opera_test)
+
+read_proff <- function(file){
+  read_html(file) |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select(
+    REGNSKAPSPERIODE:`1998`) |> 
+  mutate(
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+    filter(
+      !row_number() > 176
+    ) |>
+  filter(
+    !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer(
+    `2021`:`1998`, names_to = "year") |> 
+  mutate(
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  filter(
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate(
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate(
+    value = as.numeric(value)
+  ) |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+}
+
+opera_func <- read_proff("Links/opera.html")
+
+identical(opera, opera_func)
+```

### 2022-07-07 Test parsing of Proff table.Rmd
Commented

@@
 
+# Testing identity
 identical(opera, opera_test)
 
+# Function
 read_proff <- function(file){
@@
 
+#Testing identity
 identical(opera, opera_func)

### 2022-07-07 Test parsing of Proff table.Rmd
Naming chunks.

@@
 
-```{r}
+```{r libraries}
 library(tidyverse)
@@
 
-```{r}
+```{r kreftforeningen}
 # Reading saved html
@@
 identical(opera, opera_test)
+```
 
+```{r function}
 # Function

### 2022-07-07 Test parsing of Proff table.Rmd
Slightly more robust function added. Added headings. Added html for Rode Kors.

@@
 date: '2022-07-06'
-output: html_document
+output: 
+  html_document: 
+    df_print: paged
 ---
@@
 ```
+
+### Reading in stages
 
@@
 
-Note that we might have to convert implicitly negative numbers.
+>*Note that we might have to convert implicitly negative numbers.*
+
+### Testing with different file
 
@@
 
+### As a single operation
+
 ```{r single operation}
@@
 ```
+
+### Function 1
 
@@
 ```
+
+### Slightly more robust function
+
+- `file`: name and location of html file
+- `start`: first available year as a string ("")
+- `end`: last available year as a string ("")
+
+```{r function_modified}
+read_proff2 <- function(file, start, end){
+  read_html(file, encoding = "UTF-8") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select(
+    REGNSKAPSPERIODE:start) |> 
+  mutate(
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+    filter(
+      !row_number() > 176
+    ) |>
+  filter(
+    !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer(
+    end:start, names_to = "year") |> 
+  mutate(
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  filter(
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate(
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate(
+    value = as.numeric(value)
+  ) |> 
+  mutate(
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  )
+}
+
+read_proff2("Links/Kors.html", "2001", "2020") |> 
+  arrange(desc(year))
+```

### 2022-07-07 Test parsing of Proff table.Rmd
Visualisations.

@@
 
-read_proff2("Links/Kors.html", "2001", "2020") |> 
+Kors <- read_proff2("Links/Kors.html", "2001", "2020") |> 
   arrange(desc(year))
 ```
+
+### Visualisations
+```{r}
+Kors |> 
+  filter(REGNSKAPSPERIODE == "Årsresultat") |> 
+  ggplot(aes(year, value)) +
+  geom_col()
+
+Kors |> 
+  filter(REGNSKAPSPERIODE == "SUM EGENKAPITAL OG GJELD") |> 
+  ggplot(aes(year, value)) +
+  geom_col()
+```

### 2022-07-07 Test parsing of Proff table.Rmd
Update title.

@@
 ---
-title: "Test parsing of Proff.no table (Kreftforeningen)"
+title: "Test parsing of Proff.no table"
 author: "Torbjørn Skinnemoen Ottersen"

### 2022-07-08 Test parsing of Proff table.Rmd
Link to Brønnøysund åpnde data.
Adding currency column.

@@
 table_kreft6$year <-as.integer(table_kreft6$year)
+
+# currency
+table_kreft6b <- table_kreft6 |> mutate(
+  valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA))
+
+table_kreft6b |> 
+  fill(valutakode, .direction = "updown")
 
@@
   html_table() |> 
-  select(
+  select( # removing empty column
     REGNSKAPSPERIODE:start) |> 
-  mutate(
+  mutate( # removing whitespace
     REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
   ) |> 
-    filter(
+  filter( # removing duplicate table
       !row_number() > 176
     ) |>
-  filter(
+  filter( # removing duplicate rows
     !grepl("Lukk", REGNSKAPSPERIODE)
     ) |> 
-  pivot_longer(
+  pivot_longer( # tidying data
     end:start, names_to = "year") |> 
-  mutate(
+  mutate( # changing to real NAs and turning years into numbers
     value = na_if(value, "-"),
     year = as.integer(year)) |> 
-  filter(
+  mutate( # adding currency column
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |> 
+  fill( # filling the currency column
+    valutakode, .direction = "updown"
+    ) |> 
+  filter( # removing dates and redundant currency
     REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
       REGNSKAPSPERIODE != "Valutakode") |> 
-  mutate(
+  mutate( # removing whitespace in numbers
     value = str_replace_all(value, "\\s", "")
   ) |> 
-  mutate(
+  mutate( # turning into numbers
     value = as.numeric(value)
   ) |> 
-  mutate(
+  mutate( # removing years as values
     value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
@@
   arrange(desc(year))
+
+Kors
 ```
 
-### Visualisations
+### Basic visualisations
 ```{r}

### 2022-07-08 Test parsing of Proff table.Rmd
Adding plotly visualisation.

@@
 library(rvest)
+library(plotly)
 ```
@@
   geom_col()
+
+plot <- Kors |> 
+  filter(REGNSKAPSPERIODE == "Årsresultat") |> 
+  ggplot(aes(year, value)) +
+  geom_col()
+
+ggplotly(plot)
 ```

### 2022-07-09 PDF madness.Rmd
PDF experiments.

---
title: "PDF madness"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-08'
output: 
  html_document: 
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(readtext)
library(pdftools)
library(readr)
```

## `readtext`, `pdf_text`

```{r}
# kreft <- readtext("Data/Kreftforeningen2021.pdf")
# kreft

krefttable <- pdf_text("Data/Krefttable2021.pdf")
summary(krefttable)
kreft_split <- str_split(krefttable, "\\n")
kreft_split <- trimws(kreft_split)
summary(kreft_split)
```

## Testing API

```{r}
library(httr)

getresponse <- GET("https://data.brreg.no/regnskapsregisteret/regnskap/951812528?%C3%A5r=2019&regnskapstype=SELSKAP")

getresponse

```

Doesn't seem to work.

## Testing Tabula

[Tabula](https://tabula.technology)

### One page

```{r}
tabula_kreft <- read_csv("Data/tabula-Krefttable2021.csv")
tabula_kreft
tabula_kreft_long <- tabula_kreft |> 
  pivot_longer(cols = `2021`:`2020`, names_to = "year")
tabula_kreft_long <- tabula_kreft_long |> 
  mutate(
    value = as.numeric(str_replace_all(value, "\\s", "")),
    year = as.integer(year)
  )

tabula_kreft_long <- tabula_kreft_long |> 
  arrange(desc(year))
tabula_kreft_long
```

### All pages
```{r}
tabula_kreft_full <- read_csv("Data/tabula-Krefttable2021-full.csv")
tabula_kreft_full_long <- tabula_kreft_full |> 
  pivot_longer(`2021`:`2020`, names_to = "year")
tabula_kreft_full_long
tabula_kreft_full_long <- tabula_kreft_full_long |> 
  mutate(
    value = as.numeric(str_replace_all(value, "\\s", "")),
    year = as.integer(year)
  )
tabula_kreft_full
tabula_kreft_full_long <- tabula_kreft_full_long |> 
  arrange(desc(year))
tabula_kreft_full_long |> 
  ggplot(aes(`ANSKAFFELSE AV MIDLER`, value, fill = as.factor(year))) +
  geom_col(position = "dodge2") +
  coord_flip()

tabula_kreft_full_long

tabula_kreft_full_long |> 
  filter(`ANSKAFFELSE AV MIDLER` %in% c("Medlemsinntekter", "Sum tilskudd",
                                        "Sum innsamlede midler og gaver",
                                        "Sum operasjonelle aktiviteter",
                                        "Finansinntekter")) |> 
  ggplot(aes(`ANSKAFFELSE AV MIDLER`, value, fill = as.factor(year))) +
  geom_col(position = "dodge2") +
  coord_flip()
```

### 2022-07-09 Test parsing of Proff table.Rmd
Added link to Tabula. More visualisations.
More visualisations.

@@
 
+plot2 <- Kors |> 
+  filter(REGNSKAPSPERIODE == "SUM EGENKAPITAL OG GJELD") |> 
+  ggplot(aes(year, value)) +
+  geom_col()
+
 ggplotly(plot)
+ggplotly(plot2)
+
+Kors |> 
+  filter(REGNSKAPSPERIODE == "Driftsresultat") |> 
+  ggplot(aes(year, value)) +
+  geom_col()
+
+Kors |> 
+  filter(REGNSKAPSPERIODE == "Lønn") |> 
+  ggplot(aes(year, value)) +
+  geom_col()
+
+Kors |> 
+  filter(REGNSKAPSPERIODE == "Årsresultat") |> 
+  mutate(
+    profit = ifelse(value >= 0, TRUE, FALSE)
+  ) |> 
+  ggplot(aes(year, value * 1000, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "", big.mark = ".",
+                                                   decimal.mark = ",")) +
+  theme(legend.position = "none", axis.title.y = element_blank())
+
+Kors |> 
+  filter(REGNSKAPSPERIODE == "Ordinært resultat") |> 
+  mutate(
+    profit = ifelse(value >= 0, TRUE, FALSE)
+  ) |> 
+  ggplot(aes(year, value * 1000, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "", big.mark = ".",
+                                                   decimal.mark = ",")) +
+  theme(legend.position = "none", axis.title.y = element_blank())
+
+?element
+?theme
 ```

### 2022-07-10 Test parsing of Proff table.Rmd
More plotly.

@@
 
-?element
-?theme
+plot3 <- Kors |> 
+  filter(REGNSKAPSPERIODE == "Ordinært resultat") |> 
+  mutate(
+    profit = ifelse(value >= 0, TRUE, FALSE)
+  ) |> 
+  ggplot(aes(year, value * 1000, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "", big.mark = ".",
+                                                   decimal.mark = ",")) +
+  theme(legend.position = "none", axis.title.y = element_blank())
+
+ggplotly(plot3)
 ```

### 2022-07-10 PDF madness.Rmd
Brønnøysund pdf.
Experimenting with a Brønnøysund pdf.

@@
 ```
+
+## Brønnøysund pdf
+```{r}
+riksmaal <- pdf_text("Data/20220001922542-1.pdf")
+riksmaal
+riksmaal_split <- strsplit(riksmaal, "\\n")
+riksmaal_split
+riksmaal_tibble <- tibble(riksmaal = riksmaal_split)
+riksmaal2 <- riksmaal_tibble |> unnest_auto(col = 1)
+
+riksmaal2$numbers <- riksmaal2$riksmaal |> str_remove_all("[^0-9]")
+riksmaal2
+```

### 2022-07-11 Test parsing of Proff table.Rmd
Excel

--- 2022-07-08 Test parsing of Proff table.Rmd
@@
 ---
-title: "Test parsing of Proff.no table"
+title: "Test parsing of Proff.no table (Kreftforeningen)"
 author: "Torbjørn Skinnemoen Ottersen"
 date: '2022-07-06'
-output: 
-  html_document: 
-    df_print: paged
+output: html_document
 ---
@@
 
-```{r libraries}
+```{r}
 library(tidyverse)
 library(rvest)
-library(plotly)
 ```
 
-### Reading in stages
-
-```{r kreftforeningen}
+```{r}
 # Reading saved html
@@
   html_table()
-
-table_kreft
 
@@
 # Removing duplicate rows
-table_kreft4 <- table_kreft3 |>
-  filter(!grepl("Lukk", REGNSKAPSPERIODE))
+table_kreft4 <- table_kreft3 |> filter(!grepl("Lukk", REGNSKAPSPERIODE))
 
@@
 table_kreft6$year <-as.integer(table_kreft6$year)
-
-# currency
-table_kreft6b <- table_kreft6 |> mutate(
-  valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA))
-
-table_kreft6b |> 
-  fill(valutakode, .direction = "updown")
 
@@
 
+# Viewing table
+table_kreft8 |>
+  arrange(desc(year))
+
 # Quick visualisation
@@
   coord_flip()
-
-# Removing values that are years
-table_kreft9 <- table_kreft8 |> 
-  mutate(
-    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-  )
-
-# Viewing table
-table_kreft9 |>
-  arrange(desc(year))
 ```
 
->*Note that we might have to convert implicitly negative numbers.*
+Note that we might have to convert implicitly negative numbers.
 
-### Testing with different file
-
-```{r opera}
-# Reading saved html
-opera <- read_html("Links/opera.html") |> 
-  html_element("#inner-frame") |> 
-  html_table(trim = TRUE)
-
-# Removing white space
-opera$REGNSKAPSPERIODE <- opera$REGNSKAPSPERIODE |>
-  str_squish()
-
-# Removing empty column
-opera <- opera |>
-  select(REGNSKAPSPERIODE:`1998`)
-
-# Removing duplicate rows (hack)
-opera <- opera |>
-  filter(
-  !row_number() > 176
-)
-
-# Removing duplicate rows
-opera <- opera |>
-  filter(!grepl("Lukk", REGNSKAPSPERIODE))
-
-# tidying data:
-opera <- opera |>
-  pivot_longer(`2021`:`1998`, names_to = "year")
-
-# Turning "-" into NA
-opera <- opera |> 
-  mutate(value = na_if(value, "-"))
-
-# Turning years into integers
-opera$year <-as.integer(opera$year)
-
-# Removing unnecessary dates
-opera <- opera |> 
-  filter(
-    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-      REGNSKAPSPERIODE != "Valutakode")
-
-# Removing whitespace
-opera <- opera |> 
-  mutate(
-    value = str_replace_all(value, "\\s", "")
-  )
-
-# Converting to numbers
-opera <- opera |> 
-  mutate(
-    value = as.numeric(value)
-  )
-
-# Removing values that are years
-opera <- opera |> 
-  mutate(
-    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-  )
-
-# Viewing table
-opera |>
-  arrange(desc(year))
+```{r}
 
 ```
-
-### As a single operation
-
-```{r single operation}
-opera_test <- read_html("Links/opera.html") |> 
-  html_node("#inner-frame") |> 
-  html_table() |> 
-  select(
-    REGNSKAPSPERIODE:`1998`) |> 
-  mutate(
-    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-  ) |> 
-  filter(
-  !row_number() > 176
-  ) |> 
-  filter(
-    !grepl("Lukk", REGNSKAPSPERIODE)) |> 
-  pivot_longer(
-    `2021`:`1998`, names_to = "year") |> 
-  mutate(
-    value = na_if(value, "-"),
-    year = as.integer(year)) |> 
-  filter(
-    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-      REGNSKAPSPERIODE != "Valutakode") |> 
-  mutate(
-    value = str_replace_all(value, "\\s", "")
-  ) |> 
-  mutate(
-    value = as.numeric(value)
-  ) |> 
-  mutate(
-    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-  )
-
-# Testing identity
-identical(opera, opera_test)
-```
-
-### Function 1
-
-```{r function}
-# Function
-read_proff <- function(file){
-  read_html(file) |> 
-  html_node("#inner-frame") |> 
-  html_table() |> 
-  select(
-    REGNSKAPSPERIODE:`1998`) |> 
-  mutate(
-    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-  ) |> 
-    filter(
-      !row_number() > 176
-    ) |>
-  filter(
-    !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |> 
-  pivot_longer(
-    `2021`:`1998`, names_to = "year") |> 
-  mutate(
-    value = na_if(value, "-"),
-    year = as.integer(year)) |> 
-  filter(
-    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-      REGNSKAPSPERIODE != "Valutakode") |> 
-  mutate(
-    value = str_replace_all(value, "\\s", "")
-  ) |> 
-  mutate(
-    value = as.numeric(value)
-  ) |> 
-  mutate(
-    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-  )
-}
-
-opera_func <- read_proff("Links/opera.html")
-
-#Testing identity
-identical(opera, opera_func)
-```
-
-### Slightly more robust function
-
-- `file`: name and location of html file
-- `start`: first available year as a string ("")
-- `end`: last available year as a string ("")
-
-```{r function_modified}
-read_proff2 <- function(file, start, end){
-  read_html(file, encoding = "UTF-8") |> 
-  html_node("#inner-frame") |> 
-  html_table() |> 
-  select( # removing empty column
-    REGNSKAPSPERIODE:start) |> 
-  mutate( # removing whitespace
-    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-  ) |> 
-  filter( # removing duplicate table
-      !row_number() > 176
-    ) |>
-  filter( # removing duplicate rows
-    !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |> 
-  pivot_longer( # tidying data
-    end:start, names_to = "year") |> 
-  mutate( # changing to real NAs and turning years into numbers
-    value = na_if(value, "-"),
-    year = as.integer(year)) |> 
-  mutate( # adding currency column
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |> 
-  fill( # filling the currency column
-    valutakode, .direction = "updown"
-    ) |> 
-  filter( # removing dates and redundant currency
-    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-      REGNSKAPSPERIODE != "Valutakode") |> 
-  mutate( # removing whitespace in numbers
-    value = str_replace_all(value, "\\s", "")
-  ) |> 
-  mutate( # turning into numbers
-    value = as.numeric(value)
-  ) |> 
-  mutate( # removing years as values
-    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-  )
-}
-
-Kors <- read_proff2("Links/Kors.html", "2001", "2020") |> 
-  arrange(desc(year))
-
-Kors
-```
-
-### Basic visualisations
-```{r}
-Kors |> 
-  filter(REGNSKAPSPERIODE == "Årsresultat") |> 
-  ggplot(aes(year, value)) +
-  geom_col()
-
-Kors |> 
-  filter(REGNSKAPSPERIODE == "SUM EGENKAPITAL OG GJELD") |> 
-  ggplot(aes(year, value)) +
-  geom_col()
-
-plot <- Kors |> 
-  filter(REGNSKAPSPERIODE == "Årsresultat") |> 
-  ggplot(aes(year, value)) +
-  geom_col()
-
-ggplotly(plot)
-```

### 2022-07-11 excel_hack.R

@@ -1,0 +1,23 @@
+rm(list = ls())
+
+library(tidyverse)
+library(readxl)
+library(janitor)
+
+
+# Leser inn excel-fil og angir kolonnenavn
+jukseark <- read_excel("./OPXfinans.xlsx", col_names = c("entry", "value"))
+ 
+
+juksedata <- jukseark %>%
+  na.omit() %>% # Dropper NA (det som var overskrifter i excel)
+  pivot_wider(names_from = entry, values_from = value) %>% # Gjør om rader til kolonner (variabler)
+  clean_names() %>% # Lager mer r-vennlige kolonnenavn
+  mutate(
+    KPI = sum_total_revenue+total_assets/total_liabilities) # Legger til en kolonne med KPI,
+# regnet ut som summen av inntekter delt på liabilities. Detter er bare et eksempel.
+# Jeg vet fortsatt ingenting om å lage KPI.
+
+
+
+

### 2022-07-11 XML annoyance.Rmd

---
title: "XML annoyance"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(tidyverse)
library(xml2)
library(XML)
```

```{r}
x <- read_xml("Data/20191113105110-masse.xml")
xml_name(x)
xml_name(xml_parent(x))
x |> 
  xml_find_all("//*") |> 
  xml_path()
x1 <- xml_find_first(x, "//del")
x1
xml_name(xml_parent(x1))
xml_name(xml_children(x1))
xml_text(xml_find_first(x, "//hode"))

orgnr <- xml_text(xml_find_all(x, "//orgnr"))
xml_text(xml_find_first(x, "//info/feltkode"))
xml_text(xml_find_all(x, "//info/feltkode"))
xml_path(xml_find_all(x, "//info/feltkode"))

mass <- xmlParse("Data/20191113105110-masse.xml")

# xmlToDataFrame(mass, nodes = getNodeSet(mass, "//del"))

# Error in `[<-.data.frame`(`*tmp*`, i, names(nodes[[i]]), value = c(hode = "945011459S2018STORE14PLNRES2018010120181231MRNN201910059820190903BRLNJNNN", : 
# duplicate subscripts for columns

xmlToDataFrame(mass, nodes = getNodeSet(mass, "//del/hode"))
# works

# xmlToDataFrame(mass, nodes = getNodeSet(mass, "//del/info"))

# Error in `[<-.data.frame`(`*tmp*`, i, names(nodes[[i]]), value = c(feltkode = "7709", : 
# duplicate subscripts for columns

```

### 2022-07-11 Excel fun.Rmd
Excel long/short format and tidying.

---
title: "Excel fun"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-11'
output: 
  html_document: 
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(readxl)
library(tidyverse)
```

## Reading long file

```{r}
test <- read_excel("Data/OPXfinans test.xlsx")
test
```

## Tidying

```{r}
test2 <- test |> 
  mutate(
    category = ifelse(is.na(value), entry, NA)
  ) |> 
  fill(category) |> 
  fill(year) |> 
  filter(!is.na(value))
test2
```

## Wide file

```{r}
read_excel("Data/Book1.xlsx") |> 
  filter(!is.na(Year))
```

### 2022-07-11 Excel fun.Rmd
Filtering and summarising.

@@
   filter(!is.na(value))
-test2
+test2 |> 
+  filter()
+
+test2_no_totals <- test2 |> 
+  filter(
+    !str_detect(entry, "^Sum") & !str_detect(entry, "^Total")
+  )
+
+test2_no_totals
+test2_no_totals |>
+  group_by(category) |> 
+  summarise(
+    sum = sum(value)
+  )
 ```

### 2022-07-11 Excel fun.Rmd
Image explanation of long vs. wide data.
Headings and explanations.

--- 2022-07-11 Excel fun.Rmd
@@
 
+## Long vs. wide data
+![](Images/wideLong1-1.png)
+
+Fra [Statology](https://www.statology.org/long-vs-wide-data/).
+
 ## Reading long file
@@
   mutate(
-    category = ifelse(is.na(value), entry, NA)
+    category = ifelse(is.na(value), entry, NA) # transferring headings to category
   ) |> 
-  fill(category) |> 
-  fill(year) |> 
-  filter(!is.na(value))
+  fill(category) |> # filling category column
+  fill(year) |> # filling year column
+  filter(!is.na(value)) # removing headings
 test2
+```
+
+### Removing rows containing 'sum' and 'total'
+```{r}
+test2_no_totals <- test2 |> 
+  filter(
+    !str_detect(entry, "^Sum") & !str_detect(entry, "^Total")
+  )
+
+test2_no_totals
+test2_no_totals |>
+  group_by(category) |> 
+  summarise(
+    sum = sum(value)
+  )
 ```
@@
 ```{r}
-read_excel("Data/Book1.xlsx") |> 
+wide_data <- read_excel("Data/Book1.xlsx")
+wide_data
+wide_data |> 
   filter(!is.na(Year))

### 2022-07-12 Excel fun.Rmd
pivot_wider - entries as columns.
Deleting unnecessary file read. Adding logical test.

@@
 ```
-
-## Long vs. wide data
-![](Images/wideLong1-1.png)
-
-Fra [Statology](https://www.statology.org/long-vs-wide-data/).
 
@@
 
-## Wide file
+## `pivot_wider()`
+```{r}
+test4 <- test2 |>
+  unite(entry, c(category, entry), sep = " ") |> 
+  select(-year) |> 
+  pivot_wider(names_from = entry, values_from = value)
 
-```{r}
-wide_data <- read_excel("Data/Book1.xlsx")
-wide_data
-wide_data |> 
-  filter(!is.na(Year))
+test4
+
+test4 |> 
+  summarise(
+    sum = sum(`Revenue Private contributions`:`Revenue Other`)
+  ) |> 
+  summarise(
+    test = sum == test4$`Revenue Sum total revenue`)
 ```

### 2022-07-12 excel_hack.R
Alternativt xlsx

@@ -5,19 +5,32 @@ library(readxl)
 library(janitor)
 
 
-# Leser inn excel-fil og angir kolonnenavn
-jukseark <- read_excel("./OPXfinansAGA.xlsx", col_names = c("entry", "value"))
- 
 
-juksedata <- jukseark %>%
-  na.omit() %>% # Dropper NA (det som var overskrifter i excel)
-  pivot_wider(names_from = entry, values_from = value) %>% # Gjør om rader til kolonner (variabler)
-  clean_names() %>% # Lager mer r-vennlige kolonnenavn
+
+
+jukseark <- read_excel(
+  "C:/Users/Gusia/Desktop/HACKATHON/ISSSV1337/OPXfinans2.xlsx")
+
+jukseark_1 <- read_excel(
+  "C:/Users/Gusia/Desktop/HACKATHON/ISSSV1337/OPXfinansAGA.xlsx",
+                         col_names = c("entry", "value"))
+  
+
+juksedata <- jukseark %>% 
+  pivot_wider(names_from = Name, values_from = Value) %>% 
+  clean_names() %>% 
+    mutate(
+    KPI_ROI = administrative+fundraising/private_contributions/membership_fees)
+
+
+juksedata_1 <- jukseark_1 %>% 
+  pivot_wider(names_from = entry, values_from = value) %>% 
+  na.omit() %>% 
+  clean_names() %>% 
   mutate(
-    KPI = sum_total_revenue+total_assets/total_liabilities) # Legger til en kolonne med KPI,
-# regnet ut som summen av inntekter delt på liabilities. Dette er bare et eksempel.
-# Jeg vet fortsatt ingenting om å lage KPI.
+    KPI_ROI = administrative+fundraising/private_contributions/membership_fees)
 
+  
 
-
-
+  
+ 

### 2022-07-13 excel_hack.R
Delete excel_hack.R

@@ -1,23 +1,0 @@
-rm(list = ls())
-
-library(tidyverse)
-library(readxl)
-library(janitor)
-
-
-# Leser inn excel-fil og angir kolonnenavn
-jukseark <- read_excel("./OPXfinansAGA.xlsx", col_names = c("entry", "value"))
- 
-
-juksedata <- jukseark %>%
-  na.omit() %>% # Dropper NA (det som var overskrifter i excel)
-  pivot_wider(names_from = entry, values_from = value) %>% # Gjør om rader til kolonner (variabler)
-  clean_names() %>% # Lager mer r-vennlige kolonnenavn
-  mutate(
-    KPI = sum_total_revenue+total_assets/total_liabilities) # Legger til en kolonne med KPI,
-# regnet ut som summen av inntekter delt på liabilities. Dette er bare et eksempel.
-# Jeg vet fortsatt ingenting om å lage KPI.
-
-
-
-

### 2022-07-13 excel_hack.Rmd
Alternativt xlsx

---
title: "Excel hack med alternativt fil"
author: "Aga"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(janitor)

```

Jeg har et forslag til et alternativt excel-ark som jeg synes er enklere å bruke. Det er lastet opp i Github (håper jeg) og heter OPXfinansAGA

```{r alternativt excel-fil}
# Leser inn excel-fil og angir kolonnenavn
jukseark <- read_excel("./OPXfinansAGA.xlsx", col_names = c("entry", "value"))

# Lager et nytt objekt
juksedata <- jukseark %>%

  # Dropper NA (det som var overskrifter i excel)
    na.omit %>%

  # Gjør om rader til kolonner (variabler)
  pivot_wider(names_from = entry, values_from = value) %>%

    # Lager mer r-vennlige kolonnenavn
  clean_names() %>%

  # Ved pivot ble alle kolonnene satt til character pga. den første kolonnen med 
  # organisasjonsnavn. Det gjør det umulig å regne, så alle kolonnene unntatt den første
  # må konverteres til numeric
  mutate_at(vars(-name), as.numeric) %>%

  # Legger til en kolonne med KPI
  mutate(KPI_ROI = administrative+fundraising/private_contributions+membership_fees)
```

Med den fila vi har brukt til nå, får mange variabler flere rader med NA-verdier og flere av de verdiene vi skal regne sammen ligger på ulike rader og blir istedenfor regnet sammen med en NA-verdi, som gir NA i KPI-variabelen.

```{r pressure, echo=FALSE}
# Leser inn excel-fil
jukseark_1 <- read_excel("./OPXfinans2.xlsx")

# Lager et nytt objekt
juksedata_1 <- jukseark_1 %>%

  # # Gjør om rader til kolonner (variabler)
  pivot_wider(names_from = Name, values_from = Value) %>%

   # Lager mer r-vennlige kolonnenavn
  clean_names() %>%

  # Legger til en kolonne med KPI men alt er NA
  mutate(KPI_ROI = administrative+fundraising/private_contributions+membership_fees)

```

Jeg har prøvd å løse dette med å droppe NA-verdier, men det har ikke fungert.

```{r}

# na.omit fjerner alt innhold fra tabellen
juksedata_1 <- jukseark_1 %>% 
  pivot_wider(names_from = Name, values_from = Value) %>% 
  na.omit()

# na.rm legger til en variabel som heter na.rm XD
# KPI er fortsatt NA
juksedata_1 <- jukseark_1 %>% 
  pivot_wider(names_from = Name, values_from = Value) %>%
  clean_names() %>% 
  mutate(KPI_ROI = administrative+fundraising/private_contributions+membership_fees, 
         na.rm = TRUE)

```

Solveig foreslo å legge til as.numeric i i mutate foran hver variabel vi bruker til å regne ut KPI, men mente også at det blir mye styr og at det alternative arket generelt var enklere å bruke. Så vi bør kanskje vurdere om vi har gode grunner til å bruke det vi har brukt så langt.

### 2022-07-13 Excel fun.Rmd
Added code for final(?) version.
Excel experiments.

@@
 library(tidyverse)
+library(janitor)
 ```
@@
 ```
+
+## Final version?
+```{r}
+excel <- read_excel("Data/OPXfinans2.xlsx")
+excel <- clean_names(excel)
+
+```
+
+## Final final?
+```{r}
+excel <- read_excel("Data/OPXfinans_final.xlsx")
+excel <- clean_names(excel)
+excel_united <- excel |> 
+  unite(entry, c(category, sub_category, name), sep = " ")
+excel_united_wide <- excel_united |> 
+  pivot_wider(names_from = entry, values_from = value)
+excel_wide <- excel |> 
+  pivot_wider(names_from = name, values_from = value) |> 
+  arrange(desc(year))
+excel_wide <- clean_names(excel_wide)
+
+excel_united_wide <- clean_names(excel_united_wide)
+
+excel_united_wide
+united <- excel |> 
+  unite(entry, c(category, name), sep = "_")
+united <- clean_names(united)
+united <- united |> 
+  select(-sub_category)
+united_wide <- united |> 
+  pivot_wider(names_from = entry, values_from = value) |> 
+  clean_names()
+united_wide
+united_wide |> 
+  mutate(
+    KPI = expenses_fundraising / revenue_private_contributions * 100
+  ) |> 
+  select(year, firm, KPI)
+
+united_wide |> 
+  mutate(
+    KPI = (revenue_private_contributions - expenses_fundraising) / revenue_private_contributions * 100
+  ) |> 
+  select(year, firm, KPI)
+```

### 2022-07-13 Excel fun.Rmd
More Excel experiments.

--- 2022-07-12 Excel fun.Rmd
@@
 ```
+
+## Final version?
+```{r}
+excel <- read_excel("Data/OPXfinans2.xlsx")
+excel <- clean_names(excel)
+
+```
+
+## Final final?
+```{r}
+excel <- read_excel("Data/OPXfinans_final.xlsx")
+excel <- clean_names(excel)
+
+excel
+united <- excel |> 
+  unite(entry, c(category, name), sep = "_")
+united <- clean_names(united)
+united <- united |> 
+  select(-sub_category)
+united_wide <- united |> 
+  pivot_wider(names_from = entry, values_from = value) |> 
+  clean_names()
+united_wide
+
+united_wide |> 
+  mutate(
+    KPI = expenses_fundraising / revenue_private_contributions * 100
+  ) |> 
+  select(year, firm, KPI) |> 
+  ggplot(aes(as.factor(year), KPI)) +
+  geom_col()
+
+test <- excel
+test$result <- excel |> 
+  group_by(category) |> 
+  summarise(
+    value = sum(value)
+  ) |> 
+  pivot_wider(names_from = category, values_from = value) |> 
+  summarise(
+    result = Revenue - Expenses
+  ) |>
+  toString() |> 
+  as.numeric()
+test
+
+excel |> 
+  mutate(
+    value = ifelse(category == "Expenses" | category == "Liability", -value, value)
+  )
+excel |> 
+  filter(year == 2020) |> 
+  mutate(
+    total = sum(value)
+  )
+
+excel
+```

### 2022-07-13 USA Series 990.Rmd
USA charities notes. Smh. Example files.

---
title: "Series 990"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-13'
output: html_document
---

IRS Series 990 er ekstremt komplisert å analysere, viser det seg. Finnes en R-pakke - med varsel om at steg én tar flere dager... Ikke oppdatert på noen år. Charity Navigator har open-sourced litt av sine data, men ikke nok. Kan se mer på det, men da må vi evnt. bestemme oss for at vi ønsker å se på amerikanske charities også.

Laster opp eksempel på xml-fil (det er veldig mange tusen av dem bare i siste data dump) og index.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(XML)
library(xml2)
```

```{r}
xmlToDataFrame("Data/202100079349100000_public.xml")
xml <- read_xml("Data/202100079349100000_public.xml")
xml_name(xml)
xml_children(xml)
xml_text(xml)
```

```{r}
index <- read_csv("Data/index_2022.csv")
index
index |> 
  filter(grepl("^HUMAN RIGHTS", TAXPAYER_NAME))
```

### 2022-07-13 Test parsing of Proff table.Rmd
Improved function.

--- 2022-07-10 Test parsing of Proff table.Rmd
@@
 
+## Improved function
+
+```{r}
+read_proff3 <- function(file, start, end){
+  read_html(file, encoding = "UTF-8") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select( # removing empty column
+    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+  mutate( # removing whitespace
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+  filter( # removing duplicate table
+      !row_number() > 176
+    ) |>
+  filter( # removing duplicate rows
+    !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer( # tidying data
+    all_of(end):all_of(start), names_to = "year") |> # using all_of()
+  mutate( # changing to real NAs and turning years into numbers
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  mutate( # adding currency column
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |> 
+  fill( # filling the currency column
+    valutakode, .direction = "updown"
+    ) |> 
+  filter( # removing dates and redundant currency
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate( # removing whitespace in numbers
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate( # turning into numbers
+    value = as.numeric(value)
+  ) |> 
+  mutate( # removing years as values
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  ) |> 
+    rename( # more 'tidy' name
+      entry = REGNSKAPSPERIODE
+    ) |> 
+    arrange(desc(year)) # arranging by year
+}
+?rename
+read_proff3("Links/Kors.html", "2001", "2020")
+```
+
 ### Basic visualisations

### 2022-07-13 Test parsing of Proff table.Rmd
Testing function.

--- 2022-07-10 Test parsing of Proff table.Rmd
@@
 
+## Improved function
+
+```{r}
+read_proff3 <- function(file, start, end){
+  read_html(file, encoding = "UTF-8") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select( # removing empty column
+    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+  mutate( # removing whitespace
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+  filter( # removing duplicate table
+      !row_number() > 176
+    ) |>
+  filter( # removing duplicate rows
+    !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer( # tidying data
+    all_of(end):all_of(start), names_to = "year") |> # using all_of()
+  mutate( # changing to real NAs and turning years into numbers
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  mutate( # adding currency column
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |> 
+  fill( # filling the currency column
+    valutakode, .direction = "updown"
+    ) |> 
+  filter( # removing dates and redundant currency
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate( # removing whitespace in numbers
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate( # turning into numbers
+    value = as.numeric(value)
+  ) |> 
+  mutate( # removing years as values
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  ) |> 
+    rename( # more 'tidy' name
+      entry = REGNSKAPSPERIODE
+    ) |> 
+    arrange(desc(year)) # arranging by year
+}
+
+read_proff3("Links/Kors.html", "2001", "2020") # Testing
+read_proff3("Links/Riksmaal.html", "2011", "2021") # Testing
+```
+
 ### Basic visualisations

### 2022-07-14 Test parsing of Proff table.Rmd
Improved function.

--- 2022-07-13 Test parsing of Proff table.Rmd
@@
 library(plotly)
+library(janitor)
 ```
@@
 
-```{r}
+```{r, eval=FALSE}
 read_proff3 <- function(file, start, end){
@@
   ) |> 
-    rename( # more 'tidy' name
-      entry = REGNSKAPSPERIODE
+    distinct() |> # removing duplicate rows
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
       ) |> 
+    clean_names() |> 
     arrange(desc(year)) # arranging by year
 }
-?rename
-read_proff3("Links/Kors.html", "2001", "2020")
+
+read_proff3("Links/Kors.html", "2001", "2020") # Testing
+riksmaal <- read_proff3("Links/Riksmaal.html", "2011", "2021") # Testing
+clean_names(riksmaal)
+riksmaal |> 
+  pivot_wider(names_from = entry, values_from = value) |> 
+  clean_names()
+riksmaal
+```
+
+## Even more improved function
+```{r}
+read_proff4 <- function(file, start, end){
+  read_html(file, encoding = "UTF-8") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select( # removing empty column
+    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+  mutate( # removing whitespace
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+  filter( # removing duplicate rows
+    !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer( # tidying data
+    all_of(end):all_of(start), names_to = "year") |> # using all_of()
+  mutate( # changing to real NAs and turning years into numbers
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  mutate( # adding currency column
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |> 
+  fill( # filling the currency column
+    valutakode, .direction = "updown"
+    ) |> 
+  filter( # removing dates and redundant currency
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate( # removing whitespace in numbers
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate( # turning into numbers
+    value = as.numeric(value)
+  ) |> 
+  mutate( # removing years as values
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  ) |> 
+    distinct() |> # removing duplicate rows
+    pivot_wider( # pivoting
+      names_from = REGNSKAPSPERIODE, values_from = value
+      ) |> 
+    clean_names() |> # tidying names
+    arrange(desc(year)) # arranging by year
+}
+
+riksmaal2 <- read_proff4("Links/Riksmaal.html", "2011", "2021") # Testing
+
+?require
+```
+
+## Function with libraries
+```{r}
+read_proff5 <- function(file, start, end){
+  require(rvest)
+  require(tidyverse)
+  require(janitor)
+  read_html(file, encoding = "UTF-8") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select( # removing empty column
+    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+  mutate( # removing whitespace
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+  filter( # removing duplicate rows
+    !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer( # tidying data
+    all_of(end):all_of(start), names_to = "year") |> # using all_of()
+  mutate( # changing to real NAs and turning years into numbers
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  mutate( # adding currency column
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |> 
+  fill( # filling the currency column
+    valutakode, .direction = "updown"
+    ) |> 
+  filter( # removing dates and redundant currency
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate( # removing whitespace in numbers
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate( # turning into numbers
+    value = as.numeric(value)
+  ) |> 
+  mutate( # removing years as values
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  ) |> 
+    distinct() |> # removing duplicate rows
+    pivot_wider( # pivoting
+      names_from = REGNSKAPSPERIODE, values_from = value
+      ) |> 
+    clean_names() |> # tidying names
+    arrange(desc(year)) # arranging by year
+}
+
+read_proff5("Links/Riksmaal.html", "2011", "2021")
+
 ```

### 2022-07-14 Test parsing of Proff table.Rmd
New and improved function. Hurrah.

@@
 ## Function with libraries
+
+- `file`: name and location of html file
+- `start`: first available year as a string ("")
+- `end`: last available year as a string ("")
 ```{r}
@@
 
-read_proff5("Links/Riksmaal.html", "2011", "2021")
+riksmaal <- read_proff5("Links/Riksmaal.html", "2011", "2021")
+names(riksmaal)
+summary(riksmaal)
+
+rodekors <- read_proff5("Links/Kors.html", "2001", "2020")
+
+rodekors
+```
+
+## New function - ffs
+```{r}
+read_proff6 <- function(file, start, end){
+  require(rvest)
+  require(tidyverse)
+  require(janitor)
+  read_html(file, encoding = "UTF-8") |> 
+  html_node("#inner-frame") |> 
+  html_table() |> 
+  select( # removing empty column
+    REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+  mutate( # removing whitespace
+    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+  ) |> 
+    filter( # removing duplicate table - no idea why it's necessary
+      !row_number() > 176
+    ) |>
+  filter( # removing duplicate rows
+    !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer( # tidying data
+    all_of(end):all_of(start), names_to = "year") |> # using all_of()
+  mutate( # changing to real NAs and turning years into numbers
+    value = na_if(value, "-"),
+    year = as.integer(year)) |> 
+  mutate( # adding currency column
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |> 
+  fill( # filling the currency column
+    valutakode, .direction = "updown"
+    ) |> 
+  filter( # removing dates and redundant currency
+    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+      REGNSKAPSPERIODE != "Valutakode") |> 
+  mutate( # removing whitespace in numbers
+    value = str_replace_all(value, "\\s", "")
+  ) |> 
+  mutate( # turning into numbers
+    value = as.numeric(value)
+  ) |> 
+  mutate( # removing years as values
+    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+  ) |> 
+  mutate( # values no longer in 1000
+      value = value * 1000
+    ) |> 
+  distinct() |> # removing duplicate rows
+  pivot_wider( # pivoting
+      names_from = REGNSKAPSPERIODE, values_from = value
+      ) |> 
+  clean_names() |> # tidying names
+  select(-lederlonn_i_hele_1000,
+           -resultatregnskap_i_hele_1000,
+           -balanseregnskap_i_hele_1000) |> # removing headings
+  rename( # making it clear what lonn refers to
+      lederlonn = lonn
+    ) |> 
+  arrange(desc(year)) # arranging by year
+}
+
+read_proff6("Links/Kors.html", "2001", "2020")
+read_proff6("Links/opera.html", "1998", "2021")
+read_proff6("Links/Riksmaal.html", "2011", "2021")
 
@@
 ```
+
+## Visualisations - 5th function
+```{r}
+riksmaal |> 
+  ggplot(aes(year, arsresultat)) +
+  geom_path()
+
+riksmaal |> 
+  ggplot(aes(year, sum_egenkapital_og_gjeld)) +
+  geom_col()
+
+```

### 2022-07-14 Test parsing of Proff table.Rmd
Visualisations.

@@
 ```
+
+`read_proff6` is the current leading contender...
 
@@
 ## New function - ffs
-```{r}
+```{r read_proff6}
 read_proff6 <- function(file, start, end){
@@
     ) |> 
-  distinct() |> # removing duplicate rows
+  distinct() |> # removing any remaining duplicate rows
   pivot_wider( # pivoting
@@
 read_proff6("Links/Kors.html", "2001", "2020")
-read_proff6("Links/opera.html", "1998", "2021")
+opera <- read_proff6("Links/opera.html", "1998", "2021")
 read_proff6("Links/Riksmaal.html", "2011", "2021")
@@
 ```
+
+## Function 6 visualisations
+```{r}
+ggplot(opera, aes(year, lederlonn)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "", big.mark = ".",
+                                                   decimal.mark = ","))
+opera |> 
+  group_by(year) |> 
+  transmute(
+    sum = sum(lederlonn, leder_annen_godtgjorelse, na.rm = TRUE)
+  ) |> 
+  ggplot(aes(year, sum)) +
+  geom_col()
+opera
+opera_lonn <- opera |> 
+  group_by(year) |> 
+  transmute(
+    sum = sum(lederlonn, leder_annen_godtgjorelse, na.rm = TRUE) / arsresultat,
+    profit = ifelse(arsresultat >= 0, TRUE, FALSE)
+  ) |> 
+  ggplot(aes(year, sum, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_percent(scale = 1),
+                     name = "Godtgjørelse") +
+  xlab("År") +
+  theme(legend.position = "none")
+
+opera_lonn
+ggplotly(opera_lonn)
+```

### 2022-07-15 Test parsing of Proff table.Rmd
Adding Shiny links. Minor corrections.

@@
   transmute(
-    sum = sum(lederlonn, leder_annen_godtgjorelse, na.rm = TRUE) / arsresultat,
+    percentage = sum(lederlonn, leder_annen_godtgjorelse, na.rm = TRUE) / arsresultat,
     profit = ifelse(arsresultat >= 0, TRUE, FALSE)
   ) |> 
-  ggplot(aes(year, sum, fill = profit)) +
+  ggplot(aes(year, percentage, fill = profit)) +
   geom_col() +

### 2022-07-16 Excel fun.Rmd
Excel function. Visualisation.

@@
 excel <- clean_names(excel)
+excel
+
+excel2
 
@@
 excel
+read_excel("Data/OPXfinans_final.xlsx") |> 
+  mutate(
+    Value = Value * 1000
+  )
+
 ```
+
+## Function
+```{r}
+widen_excel <- function(file){
+  require(readxl)
+  read_excel(file) |> 
+    mutate(
+      Value = Value * 1000
+    ) |> 
+  unite(
+    entry, c(Category, Name), sep = "_"
+  ) |> 
+  select(-`Sub-category`) |> 
+  pivot_wider(names_from = entry, values_from = Value) |> 
+  clean_names()
+}
+
+widen_excel("Data/OPXfinans_final.xlsx")
+```

### 2022-07-16 Excel fun.Rmd
Minor testing.

--- 2022-07-13 Excel fun.Rmd
@@
 excel <- clean_names(excel)
+excel |> 
+  filter(year == 2020)
+
+excel2
 
@@
 excel
+read_excel("Data/OPXfinans_final.xlsx") |> 
+  mutate(
+    Value = Value * 1000
+  )
+
 ```
+
+## Function
+```{r}
+widen_excel <- function(file){
+  require(readxl)
+  require(tidyverse)
+  read_excel(file) |> 
+    mutate(
+      Value = Value * 1000
+    ) |> 
+  unite(
+    entry, c(Category, Name), sep = "_"
+  ) |> 
+  select(-`Sub-category`) |> 
+  pivot_wider(names_from = entry, values_from = Value) |> 
+  clean_names()
+}
+
+View(wide)
+
+wide <- widen_excel("Data/OPXfinans_final.xlsx")
+
+wide$expenses_fundraising / (wide$revenue_private_contributions + wide$revenue_public_grants + wide$revenue_membership_fees)
+
+wide$expenses_administrative / wide$expenses_program_services * 100
+
+```

### 2022-07-16 Shiny/Test/Flexdashboard test.Rmd
Uploading Shiny and flexdashboard tests.

---
title: "Untitled"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl, flexdashboard)

read_proff6 <- function(file, start, end){
    read_html(file, encoding = "UTF-8") |> 
      html_node("#inner-frame") |> 
      html_table() |> 
      select( # removing empty column
        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
      mutate( # removing whitespace
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |> 
      filter( # removing duplicate table - no idea why it's necessary
        !row_number() > 176
      ) |>
      filter( # removing duplicate rows
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |> 
      pivot_longer( # tidying data
        all_of(end):all_of(start), names_to = "year") |> # using all_of()
      mutate( # changing to real NAs and turning years into numbers
        value = na_if(value, "-"),
        year = as.integer(year)) |> 
      mutate( # adding currency column
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |> 
      fill( # filling the currency column
        valutakode, .direction = "updown"
      ) |> 
      filter( # removing dates and redundant currency
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode") |> 
      mutate( # removing whitespace in numbers
        value = str_replace_all(value, "\\s", "")
      ) |> 
      mutate( # turning into numbers
        value = as.numeric(value)
      ) |> 
      mutate( # removing years as values
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |> 
      mutate( # values no longer in 1000
        value = value * 1000
      ) |> 
      distinct() |> # removing any remaining duplicate rows
      pivot_wider( # pivoting
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
      clean_names() |> # tidying names
      select(-lederlonn_i_hele_1000,
             -resultatregnskap_i_hele_1000,
             -balanseregnskap_i_hele_1000) |> # removing headings
      rename( # making it clear what lonn refers to
        lederlonn = lonn
      ) |> 
      arrange(desc(year)) # arranging by year
  }

  # Function to read Excel file
  widen_excel <- function(file){
    read_excel(file) |> 
      mutate(
        Value = Value * 1000
      ) |> 
      unite(
        entry, c(Category, Name), sep = "_"
      ) |> 
      select(-`Sub-category`) |> 
      pivot_wider(names_from = entry, values_from = Value) |> 
      clean_names()
  }
```

Column {.sidebar}
-----------------------------------------------------------------------

### File upload

```{r}
# HTML
fileInput("html", label = "Upload html file", accept = ".html")

numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Button
actionButton("click", "Analyse")

observeEvent(input$click, { # wait for button click

    # creating data frame from html file
    htmldata <- reactive({
      read_proff6(input$html$datapath,
                  as.character(input$start), as.character(input$end))
      })

    # Creating data frame from Excel file
    exceldata <- reactive({
      widen_excel(input$excel$datapath)
    })

    # creating plots

    # ggplot as reactive function
    lonnplot <- reactive({
      ggplot(data = htmldata(), aes(year, lederlonn)) +
        geom_col() +
        scale_y_continuous(
          labels = scales::label_dollar(prefix = ""))
    })

    # Creating plotly plot
    output$plotlyplot <- renderPlotly(
      ggplotly(lonnplot())
    )
    output$htmlplot <- renderPlot(
      lonnplot()
    )
  })
```

Column 
-----------------------------------------------------------------------

### Chart B

```{r}
plotlyOutput("plotlyplot")
```

### Chart C
```{r}
plotOutput("htmlplot")
```

### 2022-07-16 Shiny/Test/Test data.Rmd
Uploading Shiny and flexdashboard tests.

---
title: "Test data"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-16'
output: 
  html_document: 
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Testing ideas for the Shiny app.

# Background

## Libraries
```{r}
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny,tidyverse, scales, rvest, janitor, plotly, readxl, MetBrewer)
```

## Functions

### HTML
```{r}
read_proff6 <- function(file, start, end){
    read_html(file, encoding = "UTF-8") |> 
      html_node("#inner-frame") |> 
      html_table() |> 
      select( # removing empty column
        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
      mutate( # removing whitespace
        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
      ) |> 
      filter( # removing duplicate table - no idea why it's necessary
        !row_number() > 176
      ) |>
      filter( # removing duplicate rows
        !grepl("Lukk", REGNSKAPSPERIODE)
      ) |> 
      pivot_longer( # tidying data
        all_of(end):all_of(start), names_to = "year") |> # using all_of()
      mutate( # changing to real NAs and turning years into numbers
        value = na_if(value, "-"),
        year = as.integer(year)) |> 
      mutate( # adding currency column
        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
      ) |> 
      fill( # filling the currency column
        valutakode, .direction = "updown"
      ) |> 
      filter( # removing dates and redundant currency
        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
          REGNSKAPSPERIODE != "Valutakode") |> 
      mutate( # removing whitespace in numbers
        value = str_replace_all(value, "\\s", "")
      ) |> 
      mutate( # turning into numbers
        value = as.numeric(value)
      ) |> 
      mutate( # removing years as values
        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
      ) |> 
      mutate( # values no longer in 1000
        value = value * 1000
      ) |> 
      distinct() |> # removing any remaining duplicate rows
      pivot_wider( # pivoting
        names_from = REGNSKAPSPERIODE, values_from = value
      ) |> 
      clean_names() |> # tidying names
      select(-lederlonn_i_hele_1000,
             -resultatregnskap_i_hele_1000,
             -balanseregnskap_i_hele_1000) |> # removing headings
      rename( # making it clear what lonn refers to
        lederlonn = lonn
      ) |> 
      arrange(desc(year)) # arranging by year
}
```

### Excel wide
```{r}
widen_excel <- function(file){
  read_excel(file) |> 
    mutate(
      Value = Value * 1000
    ) |> 
  unite(
    entry, c(Category, Name), sep = "_"
  ) |> 
  select(-`Sub-category`) |> 
  pivot_wider(names_from = entry, values_from = Value) |> 
  clean_names()
}
```

## Loading data
```{r}
kreft_html <- read_proff6("Data/Kreftforeningen.html", "2002", "2021")
kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
```

## Viewing data

### `View()`
```{r view, eval=FALSE}
View(kreft_html)
View(kreft_excel)
```

### kabled
```{r}
kreft_html
kreft_excel
```

# Testing

## Visualisation ideas

### CEO compensation
```{r}
compensation_p <- kreft_html |> 
  group_by(year) |> 
  mutate(
    leder_total = sum(lederlonn, leder_annen_godtgjorelse, na.rm = TRUE)
  ) |> 
  ggplot(aes(year, leder_total, fill = leder_total)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
ggplotly(compensation_p)

compensation_p2 <- kreft_html |> 
  select(
    year, lederlonn, leder_annen_godtgjorelse
  ) |> 
  rename(
    Year = year,
    Salary = lederlonn,
    `Other compensation` = leder_annen_godtgjorelse
  ) |> 
  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
  ggplot(aes(Year, value, fill = Compensation)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
  scale_fill_met_d("Troy") +
  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
ggplotly(compensation_p2)
```

### Result
```{r}
arsresultat_p <- kreft_html |> 
  mutate(
    profit = ifelse(arsresultat >= 0, TRUE, FALSE)
  ) |> 
  ggplot(aes(year, arsresultat, fill = profit)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = ""))
ggplotly(arsresultat_p)

ordinaert_p <- kreft_html |> 
  mutate(
    profit = ifelse(ordinaert_resultat >= 0, TRUE, FALSE)
  ) |> 
  ggplot(aes(year, ordinaert_resultat, fill = profit)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
  theme(legend.position = "none")
ggplotly(ordinaert_p)
```

## KPI ideas

### Fundraising ROI
```{r}
kreft_excel |> 
  mutate(
    KPI_ROI = expenses_fundraising / (revenue_private_contributions + revenue_membership_fees)
  ) |> 
  select(year, KPI_ROI)
```

### Administrative expenses
```{r}
kreft_excel |> 
  mutate(
    KPI_admin =
      str_c(
        format(
        (expenses_administrative / expenses_program_services *100), digits = 3
        ), "%"
        )
  ) |> 
  select(year, KPI_admin)
```

### Overhead
```{r}

```

### 2022-07-16 Shiny/Test/app.R

@@ -1,0 +1,167 @@
+# Using pacman to load libraries because so many are needed
+
+if (!require("pacman")) install.packages("pacman"); library(pacman)
+p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl)
+
+# Creating ui
+ui <- fluidPage(
+
+    # Application title
+    titlePanel("Charity analysis"),
+    
+    # Explanation of inputs
+    htmlOutput("explanation"),
+    
+    # Inputs
+      # need more strict validation for files?
+    # html
+    fileInput("html", label = "Upload html file", accept = ".html"),
+    numericInput("start", label = "Start", value = 2002),
+    numericInput("end", label = "End", value = 2021),
+    # Excel
+    fileInput("excel", "Upload Excel file", accept = ".xlsx"),
+    # Button
+    actionButton("click", "Analyse"),
+    
+    # Outputs (Planning on removing the tables once testing is done)
+      # need to work on how to display KPIs
+    
+    tableOutput("htmltable"),
+    tableOutput("exceltable"),
+    plotOutput("htmlplot"),
+    plotlyOutput("plotlyplot")
+)
+
+# Define server logic
+server <- function(input, output) {
+  
+  # function to read html file
+  
+  read_proff6 <- function(file, start, end){
+    read_html(file, encoding = "UTF-8") |> 
+      html_node("#inner-frame") |> 
+      html_table() |> 
+      select( # removing empty column
+        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+      mutate( # removing whitespace
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |> 
+      filter( # removing duplicate table - no idea why it's necessary
+        !row_number() > 176
+      ) |>
+      filter( # removing duplicate rows
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |> 
+      pivot_longer( # tidying data
+        all_of(end):all_of(start), names_to = "year") |> # using all_of()
+      mutate( # changing to real NAs and turning years into numbers
+        value = na_if(value, "-"),
+        year = as.integer(year)) |> 
+      mutate( # adding currency column
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |> 
+      fill( # filling the currency column
+        valutakode, .direction = "updown"
+      ) |> 
+      filter( # removing dates and redundant currency
[…]

### 2022-07-16 Shiny/Test/Flexdashboard test.Rmd
Commenting.

@@
 
+# Function to read html
 read_proff6 <- function(file, start, end){

### 2022-07-17 Shiny/Test/Flexdashboard test.Rmd
Title and comments.

@@
 ---
-title: "Untitled"
+title: "Charity analysis"
 output: 
@@
 }
-```
 
-Column {.sidebar}
------------------------------------------------------------------------
-
-### File upload
-
-```{r}
-# HTML
-fileInput("html", label = "Upload html file", accept = ".html")
-
-numericInput("start", label = "Start", value = 2002)
-numericInput("end", label = "End", value = 2021)
-
-# Excel
-fileInput("excel", "Upload Excel file", accept = ".xlsx")
-
-# Button
-actionButton("click", "Analyse")
-
+# Shiny actions
 observeEvent(input$click, { # wait for button click
@@
 
+Column {.sidebar}
+-----------------------------------------------------------------------
+
+### File upload
+
+```{r}
+# HTML
+fileInput("html", label = "Upload html file", accept = ".html")
+
+numericInput("start", label = "Start", value = 2002)
+numericInput("end", label = "End", value = 2021)
+
+# Excel
+fileInput("excel", "Upload Excel file", accept = ".xlsx")
+
+# Button
+actionButton("click", "Analyse")
+
+```
+
 Column 
@@
 
-### Chart B
+### Plotly
 
@@
 
-### Chart C
+### ggplot
 ```{r}

### 2022-07-17 Shiny/Test/Test data.Rmd
Playing with colour.

@@
 
-Testing ideas for the Shiny app.
+Testing ideas for the Shiny app/dashboard.
 
@@
 
+## Data manipulation
+```{r}
+kreft_excel |> 
+  filter(year == max(year))
+
+kreft_html |> 
+  filter(year == max(kreft_excel$year))
+
+kreft_excel |> 
+  left_join(kreft_html, by = "year")
+```
+
 ## Visualisation ideas
@@
 
+arsresultat_p_met <- kreft_html |> 
+  mutate(
+    profit = ifelse(arsresultat >= 0, TRUE, FALSE)
+  ) |> 
+  ggplot(aes(year, arsresultat, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
+  scale_fill_met_d("Tsimshian")
+ggplotly(arsresultat_p_met)
+
 ordinaert_p <- kreft_html |> 
@@
 ggplotly(ordinaert_p)
+
+ordinaert_p_met <- kreft_html |> 
+  mutate(
+    profit = ifelse(ordinaert_resultat >= 0, TRUE, FALSE)
+  ) |> 
+  ggplot(aes(year, ordinaert_resultat, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
+  theme(legend.position = "none") +
+  scale_fill_met_d("Tsimshian")
+ggplotly(ordinaert_p_met)
 ```

### 2022-07-17 Shiny/Test/Test data.Rmd
Currency tests.

@@
   geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix =
+                                    str_c(kreft_html$valutakode[1], " "))) +
   scale_fill_met_d("Troy") +
@@
 ggplotly(compensation_p2)
+
+# if else structure to check for uniform currency code?
+length(unique(kreft_html$valutakode)) == 1
+
+if (length(unique(kreft_html$valutakode)) == 1) {
+  kreft_html |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix =
+                                    str_c(kreft_html$valutakode[1], " "))) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
+} else {
+    kreft_html |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix = "? ")) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
+}
+currency_test <- kreft_html |> 
+  mutate(
+    valutakode = ifelse(year == 2019, "USD", valutakode)
+  )
+
+if (length(unique(currency_test$valutakode)) == 1) {
+  currency_test |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix =
+                                    str_c(kreft_html$valutakode[1], " "))) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
+} else {
+    currency_test |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix = "? ")) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+       subtitle = "Note: not all values are denominated in the same currency")
+}
 ```

### 2022-07-17 Shiny/Test/Test data.Rmd
Commenting.

--- 2022-07-17 Shiny/Test/Test data.Rmd
@@
 
-### CEO compensation
+### CEO compensation and currency testing
 ```{r}
@@
 ggplotly(compensation_p2)
+
+# if else structure to check for uniform currency code?
+length(unique(kreft_html$valutakode)) == 1
+
+# Testing with uniform currency
+if (length(unique(kreft_html$valutakode)) == 1) {
+  kreft_html |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix =
+                                    str_c(kreft_html$valutakode[1], " "))) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
+} else {
+    kreft_html |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix = "? ")) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
+}
+
+# Creating data frame with multiple currencies
+currency_test <- kreft_html |> 
+  mutate(
+    valutakode = ifelse(year == 2019, "USD", valutakode)
+  )
+
+# Testing new data frame
+if (length(unique(currency_test$valutakode)) == 1) {
+  currency_test |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix =
+                                    str_c(kreft_html$valutakode[1], " "))) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
+} else {
+    currency_test |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix = "? ")) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+       subtitle = "Note: not all values are denominated in the same currency")
+}
 ```

### 2022-07-17 Shiny/Test/Test data.Rmd
Fixes.

@@
 ## Libraries
-```{r}
+```{r libraries}
 if (!require("pacman")) install.packages("pacman"); library(pacman)
@@
   scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+       subtitle = "Note: not all values are denominated in the same currency")
 }

### 2022-07-17 Shiny/Test/Flexdashboard test.Rmd
Plot and explanation. ifelse() rewrite.

@@
 if (!require("pacman")) install.packages("pacman"); library(pacman)
-p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl, flexdashboard)
+p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl, flexdashboard, MetBrewer)
 
@@
     )
+
+    # Creating ggplot
     output$htmlplot <- renderPlot(
@@
     )
+
+    # Creating fancier plot
+    fancyplot <- reactive(
+      htmldata() |> 
+        select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+                    label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+                    label_dollar(prefix = "? "))
+      ) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+       subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
+                         "",
+                         "Note: not all values are denominated in the same currency"))
+    )
+
+    output$fancyplot <- renderPlotly(
+      ggplotly(fancyplot())
+    )
   })
+
 ```
@@
 ### File upload
+
+Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.
+
+Then upload an Excel file in the set format.
+
+Click 'Analyse' to analyse the data.
 
@@
 ```
+
+### Transparency
+```{r}
+plotlyOutput("fancyplot")
+```

### 2022-07-17 Shiny/Test/Test data.Rmd
Plot and explanation. ifelse() rewrite.

@@
 }
+
+# Could probably be done with two internal ifelse() functions
+
+# ifelse() version
+
+# Testing
+currency_test |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = ifelse(length(unique(currency_test$valutakode)) == 1,
+                    label_dollar(prefix = str_c(kreft_html$valutakode[1], " ")),
+                    label_dollar(prefix = "? "))
+      ) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+       subtitle = ifelse(length(unique(currency_test$valutakode)) == 1,
+                         element_blank(),
+                         "Note: not all values are denominated in the same currency"))
+
+# Original
+kreft_html |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = ifelse(length(unique(kreft_html$valutakode)) == 1,
+                    label_dollar(prefix = str_c(kreft_html$valutakode[1], " ")),
+                    label_dollar(prefix = "? "))
+      ) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+       subtitle = ifelse(length(unique(kreft_html$valutakode)) == 1,
+                         "",
+                         "Note: not all values are denominated in the same currency"))
 ```

### 2022-07-17 Shiny/Test/Flexdashboard test.Rmd
Adding Value Box.

@@
       widen_excel(input$excel$datapath)
+    })
+
+    # Creating value box
+    KPI_admin <- reactive({exceldata() |> 
+           mutate(
+    KPI_admin =
+      str_c(
+        format(
+        (expenses_administrative / expenses_program_services *100), digits = 3
+        ), "%"
+        )
+  ) |> 
+  filter(year == max(year)) |> 
+  select(KPI_admin) |> 
+  str_flatten()
+    })
+
+    output$KPI_box <- renderValueBox({
+      valueBox(
+        value = KPI_admin(),
+        caption = "admin",
+        color = "info",
+        icon = "fa-random"
+      )
     })
@@
 ```
+
+### KPI
+```{r}
+valueBoxOutput("KPI_box")
+```

### 2022-07-17 Shiny/Test/Test data.Rmd
Adding Value Box.

@@
   select(year, KPI_admin)
+
+kreft_excel |> 
+  mutate(
+    KPI_admin =
+      str_c(
+        format(
+        (expenses_administrative / expenses_program_services *100), digits = 3
+        ), "%"
+        )
+  ) |> 
+  filter(year == max(year)) |> 
+  select(KPI_admin) |> 
+  str_flatten()
+
+kreft_excel |> mutate(
+            KPI_admin =
+              expenses_administrative / expenses_program_services *100) |> 
+           filter(year == max(year)) |> 
+          select(KPI_admin)
 ```

### 2022-07-18 Shiny/Test/Flexdashboard test.Rmd
Moving server functions to the end of document (recommended style).

--- 2022-07-17 Shiny/Test/Flexdashboard test.Rmd
@@
 if (!require("pacman")) install.packages("pacman"); library(pacman)
-p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl, flexdashboard, MetBrewer)
+p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
+       flexdashboard, MetBrewer)
+```
 
-# Function to read html
+Column {.sidebar}
+-----------------------------------------------------------------------
+
+### File upload
+
+Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.
+
+Then upload an Excel file in the set format.
+
+Click 'Analyse' to analyse the data.
+
+```{r sidebar}
+# HTML
+fileInput("html", label = "Upload html file", accept = ".html")
+
+numericInput("start", label = "Start", value = 2002)
+numericInput("end", label = "End", value = 2021)
+
+# Excel
+fileInput("excel", "Upload Excel file", accept = ".xlsx")
+
+# Button
+actionButton("click", "Analyse")
+
+```
+
+Column 
+-----------------------------------------------------------------------
+
+### Plotly
+
+```{r}
+plotlyOutput("plotlyplot")
+```
+
+### ggplot
+```{r}
+plotOutput("htmlplot")
+```
+
+### Transparency
+```{r}
+plotlyOutput("fancyplot")
+```
+
+### KPI
+```{r}
+valueBoxOutput("KPI_box")
+```
+
+```{r shiny, include=FALSE}
+# Shiny actions
+observeEvent(input$click, { # wait for button click
+
+    # creating data frame from html file
+    htmldata <- reactive({
+      read_proff6(input$html$datapath,
+                  as.character(input$start), as.character(input$end))
+      })
+
+    # Creating data frame from Excel file
+    exceldata <- reactive({
+      widen_excel(input$excel$datapath)
+    })
+#---- Value Box ----------------------------------------------------------------    
+    # Creating value box
+    KPI_admin <- reactive({exceldata() |> 
+           mutate(
+    KPI_admin =
+      str_c(
+        format(
+        (expenses_administrative / expenses_program_services *100), digits = 3
+        ), "%"
+        )
+  ) |> 
+  filter(year == max(year)) |> 
+  select(KPI_admin) |> 
+  str_flatten()
+    })
+
+    output$KPI_box <- renderValueBox({
+      valueBox(
+        value = KPI_admin(),
+        caption = "admin",
+        color = "info",
+        icon = "fa-random"
+      )
+    })
+#---- Plots --------------------------------------------------------------------    
+    # creating plots
+
+    # ggplot as reactive function
+    lonnplot <- reactive({
+      ggplot(data = htmldata(), aes(year, lederlonn)) +
+        geom_col() +
+        scale_y_continuous(
+          labels = scales::label_dollar(prefix = ""))
+    })
+
+    # Creating plotly plot
+    output$plotlyplot <- renderPlotly(
+      ggplotly(lonnplot())
+    )
+
+    # Creating ggplot
+    output$htmlplot <- renderPlot(
+      lonnplot()
+    )
+
+    # Creating fancier plot
+    fancyplot <- reactive(
+      htmldata() |> 
+        select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |> 
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |> 
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+                    label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+                    label_dollar(prefix = "? "))
+      ) +
+  scale_fill_met_d("Troy") +
+  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+       subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
+                         "",
+                         "Note: not all values are denominated in the same currency"))
+    )
+
+    output$fancyplot <- renderPlotly(
+      ggplotly(fancyplot())
+    )
+  })
+```
+
+```{r functions, include=FALSE}
+# Function to read html file
 read_proff6 <- function(file, start, end){
@@
 
-# Shiny actions
-observeEvent(input$click, { # wait for button click
-
-    # creating data frame from html file
-    htmldata <- reactive({
-      read_proff6(input$html$datapath,
-                  as.character(input$start), as.character(input$end))
-      })
-
-    # Creating data frame from Excel file
-    exceldata <- reactive({
-      widen_excel(input$excel$datapath)
-    })
-
-    # creating plots
-
-    # ggplot as reactive function
-    lonnplot <- reactive({
-      ggplot(data = htmldata(), aes(year, lederlonn)) +
-        geom_col() +
-        scale_y_continuous(
-          labels = scales::label_dollar(prefix = ""))
-    })
-
-    # Creating plotly plot
-    output$plotlyplot <- renderPlotly(
-      ggplotly(lonnplot())
-    )
-
-    # Creating ggplot
-    output$htmlplot <- renderPlot(
-      lonnplot()
-    )
-
-    # Creating fancier plot
-    fancyplot <- reactive(
-      htmldata() |> 
-        select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |> 
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |> 
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
-                    label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
-                    label_dollar(prefix = "? "))
-      ) +
-  scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
-       subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
-                         "",
-                         "Note: not all values are denominated in the same currency"))
-    )
-
-    output$fancyplot <- renderPlotly(
-      ggplotly(fancyplot())
-    )
-  })
-
 ```
-
-Column {.sidebar}
------------------------------------------------------------------------
-
-### File upload
-
-Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.
-
-Then upload an Excel file in the set format.
-
-Click 'Analyse' to analyse the data.
-
-```{r}
-# HTML
-fileInput("html", label = "Upload html file", accept = ".html")
-
-numericInput("start", label = "Start", value = 2002)
-numericInput("end", label = "End", value = 2021)
-
-# Excel
-fileInput("excel", "Upload Excel file", accept = ".xlsx")
-
-# Button
-actionButton("click", "Analyse")
-
-```
-
-Column 
------------------------------------------------------------------------
-
-### Plotly
-
-```{r}
-plotlyOutput("plotlyplot")
-```
-
-### ggplot
-```{r}
-plotOutput("htmlplot")
-```
-
-### Transparency
-```{r}
-plotlyOutput("fancyplot")
-```

### 2022-07-18 Shiny/Test/Flexdashboard rows.Rmd
Uploading row-based version - which I think is better.

--- 2022-07-18 Shiny/Test/Flexdashboard test.Rmd
@@
   flexdashboard::flex_dashboard:
-    orientation: columns
-    vertical_layout: fill
+    orientation: rows
+    vertical_layout: scroll
 runtime: shiny
@@
 
-Column 
+Row 
+-----------------------------------------------------------------------
+
+### KPI
+```{r}
+valueBoxOutput("KPI_box")
+```
+
+### KPI 2
+```{r}
+valueBoxOutput("KPI_box2")
+```
+
+### KPI 3
+```{r}
+valueBoxOutput("KPI_box3")
+```
+
+### KPI 4
+```{r}
+valueBoxOutput("KPI_box4")
+```
+
+Row
 -----------------------------------------------------------------------
@@
 ### Plotly
-
 ```{r}
@@
 
+Row 
+-----------------------------------------------------------------------
+
 ### Transparency
@@
 plotlyOutput("fancyplot")
-```
-
-### KPI
-```{r}
-valueBoxOutput("KPI_box")
 ```
@@
     })
+
+  output$KPI_box2 <- renderValueBox({
+      valueBox(
+        value = KPI_admin(),
+        caption = "admin",
+        color = "info",
+        icon = "fa-random"
+      )
+    })
+    output$KPI_box3 <- renderValueBox({
+      valueBox(
+        value = KPI_admin(),
+        caption = "admin",
+        color = "info",
+        icon = "fa-random"
+      )
+    })
+
+    output$KPI_box4 <- renderValueBox({
+      valueBox(
+        value = KPI_admin(),
+        caption = "admin",
+        color = "info",
+        icon = "fa-random"
+      )
+    })
+
 #---- Plots --------------------------------------------------------------------    
@@
 
+  # Creating fancy plotlyplot
+    fancyplotly <- reactive({
+      ggplotly(fancyplot(), layout(
+                                legend = list(
+                                orientation = 'h', x = 0.3, y = -0.1, 
+                                title = list(text = 'My legend title'))))
+    })
+
     output$fancyplot <- renderPlotly(
-      ggplotly(fancyplot())
+      fancyplot()
     )

### 2022-07-18 Shiny/Test/Fontawesome.Rmd
Fontawesome details.

---
title: "Fontawesome"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(fontawesome, tidyverse)
```

[Fontawesome v5](https://fontawesome.com/v5/search?m=free)

Flexdashboard har ikke alle Fontawesome-ikonene. Pakken fontawesome har opp til ca. v. 5.15.3. Det enkleste for å finne ikoner er kanskje å bruke linken over og så sjekke navnet mot `fontawesome` ved å bruke koden nedenfor(`df$icon_names_full |> str_view_all("søketerm"), match = TRUE`)

```{r pressure, echo=FALSE}
# Creating df
df <- fontawesome::fa_metadata()
# Version check
df$version
# Example search
df$icon_names_full |> str_view_all("battery", match = TRUE)
```

### 2022-07-18 Shiny/Test/Flexdashboard rows.Rmd
Icon tests, tabs.

@@
 p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-       flexdashboard, MetBrewer)
+       flexdashboard, MetBrewer, DT, fontawesome)
+htmltools::tagList(fontawesome::fa_html_dependency())
 ```
 
-Column {.sidebar}
------------------------------------------------------------------------
+Sidebar {.sidebar}
+=======================================================================
 
@@
 
+Dashboard
+=======================================================================
+
 Row 
@@
 
+### KPI 5
+```{r}
+valueBoxOutput("KPI_box5")
+```
+
 Row
@@
 
+Details
+=======================================================================
+
+Row 
+-----------------------------------------------------------------------
+
+### Information 1
+
+The KPIs are based on the following calculations
+
+- **admin** is based on dividing
+```{r}
+
+```
+
+Row 
+-----------------------------------------------------------------------
+
+### Information 2
+```{r}
+
+```
+
 ```{r shiny, include=FALSE}
@@
     })
-#---- Value Box ----------------------------------------------------------------    
+
+#---- Value Boxes ----------------------------------------------------------------
+
     # Creating value box
@@
         caption = "admin",
-        color = "info",
+        color = "primary",
         icon = "fa-random"
@@
 
+    # Duplicate value boxes for testing
   output$KPI_box2 <- renderValueBox({
@@
         value = KPI_admin(),
-        caption = "admin",
-        color = "info",
+        caption = "danger",
+        color = "danger",
         icon = "fa-random"
@@
         value = KPI_admin(),
-        caption = "admin",
-        color = "info",
+        caption = "success",
+        color = "success",
         icon = "fa-random"
@@
         value = KPI_admin(),
-        caption = "admin",
+        caption = "info",
         color = "info",
         icon = "fa-random"
-      )
-    })
+      )})
+
+      output$KPI_box5 <- renderValueBox({
+      valueBox(
+        value = KPI_admin(),
+        caption = "admin",
+        color = "warning",
+        icon = "fal fa-abacus"
+)})
 
@@
 })
+
 ```
@@
 }
-
 ```

### 2022-07-18 Shiny/Test/Flexdashboard rows.Rmd
Fixing wrong code chunk. Plot and ValueBox fixes.

@@
 ```{r}
-plotlyOutput("fancyplot")
+plotlyOutput("fancyplotly")
 ```
@@
         value = KPI_admin(),
-        caption = "admin",
+        caption = "primary",
         color = "primary",
@@
         value = KPI_admin(),
-        caption = "admin",
+        caption = "warning",
         color = "warning",
-        icon = "fal fa-abacus"
+        icon = "fas fa-battery-empty"
 )})
@@
     fancyplotly <- reactive({
-      ggplotly(fancyplot(), layout(
-                                legend = list(
-                                orientation = 'h', x = 0.3, y = -0.1, 
-                                title = list(text = 'My legend title'))))
+      ggplotly(fancyplot())
     })
-
-    output$fancyplot <- renderPlotly(
-      fancyplot()
+    output$fancyplotly <- renderPlotly(
+      fancyplotly()
     )

### 2022-07-18 Shiny/Test/Fontawesome.Rmd
More instructions.
Fixing wrong code chunk. Plot and ValueBox fixes.

@@
 Flexdashboard har ikke alle Fontawesome-ikonene.
-Pakken fontawesome har opp til ca.
+Pakken `fontawesome` har opp til ca.
 v.
@@
 
-```{r pressure, echo=FALSE}
+Lag evnt. en kopi av Flexdashboard rows.Rmd og bytt ut icon-navnet i en ValueBox for å sjekke om det fungerer:
+
+```{r eval=FALSE}
+output$KPI_box5 <- renderValueBox({
+      valueBox(
+        value = KPI_admin(),
+        caption = "admin",
+        color = "warning",
+        icon = "fal fa-abacus"  # <-- Icon
+})
+```
+
+```{r}
 # Creating df

### 2022-07-18 Shiny/Test/Flexdashboard rows.Rmd
Commenting.

@@
 ```{r setup, include=FALSE}
+# Packages using pacman
 if (!require("pacman")) install.packages("pacman"); library(pacman)
 p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-       flexdashboard, MetBrewer, DT, fontawesome)
+       flexdashboard, MetBrewer, DT, fontawesome, htmltools)
+# Using package fontawesome for fontawesome icons
 htmltools::tagList(fontawesome::fa_html_dependency())
@@
 
+# Year inputs
 numericInput("start", label = "Start", value = 2002)
@@
 
+#---- Data frames --------------------------------------------------------------      
     # creating data frame from html file
@@
 
-#---- Value Boxes ----------------------------------------------------------------
+#---- Value Boxes --------------------------------------------------------------
 
-    # Creating value box
+    # Creating KPI_admin value
     KPI_admin <- reactive({exceldata() |> 
@@
 
+    # Creating the first value box
     output$KPI_box <- renderValueBox({
@@
     })
+  # Creating fancy plotly output  
     output$fancyplotly <- renderPlotly(
@@
     read_excel(file) |> 
-      mutate(
+      mutate( # Changing values
         Value = Value * 1000
       ) |> 
-      unite(
+      unite( # uniting Category and Name to deal with multiple 'Other'
         entry, c(Category, Name), sep = "_"
       ) |> 
-      select(-`Sub-category`) |> 
+      select(-`Sub-category`) |> # deleting Sub-category
+    # pivoting
       pivot_wider(names_from = entry, values_from = Value) |> 
-      clean_names()
+      clean_names() # cleaning names
 }

### 2022-07-18 Shiny/Test/Flexdashboard rows.Rmd
Removing DT package as it's not used. Commenting.

@@
 p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-       flexdashboard, MetBrewer, DT, fontawesome, htmltools)
+       flexdashboard, MetBrewer, fontawesome, htmltools)
 # Using package fontawesome for fontawesome icons
@@
 #---- Plots --------------------------------------------------------------------    
-    # creating plots
+    # creating plots - CEO compensation
 
@@
 
-    # Creating fancier plot
+    # fancier plot function
     fancyplot <- reactive(
@@
          ) |> 
-        rename(
+        rename( # renaming for the purposes of the plot
           Year = year,
@@
         ) |> 
+        # pivoting to join the two values to use fill
         pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
@@
         scale_y_continuous(
+          # Checking for uniform currency code
         labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+                        # Using the currency code if uniform
                     label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+                        # alternative prefix if multiple currencies in the data
                     label_dollar(prefix = "? "))
          ) +
-        scale_fill_met_d("Troy") +
+        scale_fill_met_d("Troy") + # pretty colours
         labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+             # Checking again for currency; including warning if not uniform
         subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
@@
     })
+
   # Creating fancy plotly output  

### 2022-07-19 Shiny/Test/Flexdashboard rows.Rmd
Icon tests.

@@
         color = "danger",
-        icon = "fa-random"
+        icon = "fas fa-dog"
       )
@@
         color = "success",
-        icon = "fa-random"
+        icon = "fas fa-coffee"
       )
@@
         color = "info",
-        icon = "fa-random"
+        icon = "fas fa-skull-and-crossbones"
       )})

### 2022-07-19 Shiny/Test/Fontawesome.Rmd
Icon tests.

@@
 # Example search
-df$icon_names_full |> str_view_all("battery", match = TRUE)
+df$icon_names_full |> str_view_all("skull", match = TRUE)
 ```

### 2022-07-19 Shiny/Test/Fontawesome.Rmd
Icon idea.
Icon ideas.

@@
 # Example search
-df$icon_names_full |> str_view_all("skull", match = TRUE)
+df$icon_names_full |> str_view_all("hands-helping", match = TRUE)
 ```
+
+### Ideas
+
+#### Checked
+
+- fas fa-seedling
+- fas fa-donate
+- fas fa-hand-holding-usd
+- fas fa-hands-helping
+
+#### Not yet checked
+
+#### Not available

### 2022-07-19 excel_hack.Rmd
Det nye Excel-arket mangler - beskjed til Aga.
Svar til Torbjørn
Oppdatert excel hack og xlsx-fil

@@
 
+>Du har ikke lastet opp det nye Excel-arket :) (Plasser det gjerne i Data-mappen ;)) --Torbjørn
+
+>Nå er arket lastet opp i Data-mappen
+--Aga
+
 Jeg har et forslag til et alternativt excel-ark som jeg synes er enklere å bruke. Det er lastet opp i Github (håper jeg) og heter OPXfinansAGA
@@
 # Leser inn excel-fil og angir kolonnenavn
-jukseark <- read_excel("./OPXfinansAGA.xlsx", col_names = c("entry", "value"))
+jukseark <- read_excel("./Data/OPXfinansAGA.xlsx", col_names = c("entry", "value"))
 
@@
   # må konverteres til numeric
-  mutate_at(vars(-name), as.numeric) %>%
+  mutate_at(vars(-org_name), as.numeric) %>%
 
   # Legger til en kolonne med KPI
-  mutate(KPI_ROI = administrative+fundraising/private_contributions+membership_fees)
+  mutate(KPI_ROI = fundraising/(private_contributions+membership_fees))
 ```
@@
   # Legger til en kolonne med KPI men alt er NA
-  mutate(KPI_ROI = administrative+fundraising/private_contributions+membership_fees)
+  mutate(KPI_ROI = fundraising/(private_contributions+membership_fees))
 
@@
   clean_names() %>% 
-  mutate(KPI_ROI = administrative+fundraising/private_contributions+membership_fees, 
+  mutate(KPI_ROI = undraising/(private_contributions+membership_fees), 
          na.rm = TRUE)

### 2022-07-19 Shiny/Test/Flexdashboard rows.Rmd
Icon fix.

--- 2022-07-18 Shiny/Test/Flexdashboard rows.Rmd
@@
 ```{r setup, include=FALSE}
-# Packages using pacman
+# Loading packages using pacman to ensure they're all installed
 if (!require("pacman")) install.packages("pacman"); library(pacman)
@@
        flexdashboard, MetBrewer, fontawesome, htmltools)
-# Using package fontawesome for fontawesome icons
+# Using fontawesome for more up-to-date fontawesome icons
 htmltools::tagList(fontawesome::fa_html_dependency())
@@
       KPI_admin =
-        str_c(
-         format(
+        str_c( # Using str_c() to concatenate value and % sign
+         format( # shortening(lengthening) the resulting value
+           # calculating the value
           (expenses_administrative / expenses_program_services *100), digits = 3
@@
       ) |> 
-      filter(year == max(year)) |> 
-       select(KPI_admin) |> 
-        str_flatten()
+      filter(year == max(year)) |> # selecting latest available year
+       select(KPI_admin) |> # dropping all other variables
+        str_flatten() # turning into a single string
          })
@@
         color = "danger",
-        icon = "fa-random"
+        icon = "fas fa-dog"
       )
+
     })
@@
         color = "info",
-        icon = "fa-random"
+        icon = "fas fa-skull-crossbones"
       )})

### 2022-07-19 Shiny/Test/Flexdashboard rows.Rmd
Fundraising ROI.

@@
 
+    # Creating Fundraising ROI value
+    KPI_fundraising <- reactive({
+      exceldata() |> 
+          mutate(
+    KPI_admin =
+      str_c(
+        format(
+        ((revenue_private_contributions +                  revenue_membership_fees) / expenses_fundraising * 100), digits = 3
+        ), "%"
+        )
+  ) |> 
+  filter(year == max(year)) |> 
+  select(KPI_admin) |> 
+  str_flatten()
+    })
+
     # Creating the first value box
@@
       valueBox(
-        value = KPI_admin(),
+        value = KPI_fundraising(),
         caption = "success",

### 2022-07-19 Shiny/Test/Test data.Rmd
Fundraising ROI.

@@
   select(year, KPI_ROI)
+
+kreft_excel |> 
+          mutate(
+            KPI_fundraising =
+              str_c(format((revenue_private_contributions +                  revenue_membership_fees) / expenses_fundraising) * 100, digits = 3
+          ), "%"
+         ) |> 
+  filter(year == max(year)) |> # selecting latest available year
+             select(KPI_fundraising) |> 
+        str_flatten() # turning into a single string
+
+kreft_excel |> 
+  mutate(
+    KPI_admin =
+      str_c(
+        format(
+        ((revenue_private_contributions +                  revenue_membership_fees) / expenses_fundraising * 100), digits = 3
+        ), "%"
+        )
+  ) |> 
+  filter(year == max(year)) |> 
+  select(KPI_admin) |> 
+  str_flatten()
+
+(kreft_excel$revenue_private_contributions + kreft_excel$revenue_membership_fees) / kreft_excel$expenses_fundraising
+
 ```

### 2022-07-19 Shiny/Test/Flexdashboard rows.Rmd
Improved value boxes. Annual result plot.

--- 2022-07-19 Shiny/Test/Flexdashboard rows.Rmd
@@
 
-### ggplot
+### Annual result
 ```{r}
-plotOutput("htmlplot")
+plotlyOutput("arsresultat_plot")
 ```
@@
     # Creating KPI_admin value
-    KPI_admin <- reactive({exceldata() |> 
+    KPI_admin <- reactive({
+      exceldata() |> 
            mutate(
@@
 
-    # Creating the first value box
+    # Creating Fundraising ROI value
+    KPI_fundraising_num <- reactive({
+      exceldata() |> 
+          mutate(
+            # Calculating the value as a percentage
+            KPI_fundraising_num =
+              (revenue_private_contributions +
+                 revenue_membership_fees - expenses_fundraising) /
+              expenses_fundraising * 100) |> 
+        # most recent year
+          filter(year == max(year)) |> 
+        # just want the single number
+          select(KPI_fundraising_num) |> 
+          pull()
+    })
+
+    # Creating the value box for administrative expenses
     output$KPI_box <- renderValueBox({
@@
         value = KPI_admin(),
-        caption = "primary",
+        caption = "Admin spend",
         color = "primary",
@@
 
-    # Duplicate value boxes for testing
+    # Creating value box for fundraising ROI
+      # Unsure about colour levels
     output$KPI_box2 <- renderValueBox({
       valueBox(
-        value = KPI_admin(),
-        caption = "danger",
-        color = "danger",
-        icon = "fas fa-dog"
+        # shortening(lengthening) the value with format()
+        # Using str_c() to concatenate value and % sign;
+        value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
+        caption = "Fundraising ROI",
+        # Colour depends on value
+        color = if (KPI_fundraising_num() < 100) {"danger"
+          } else if (KPI_fundraising_num() < 200) {"warning"
+            } else {"success"},
+        icon = "fas fa-donate"
       )
+    })
 
-    })
+    # Duplicate value boxes for testing
     output$KPI_box3 <- renderValueBox({
       valueBox(
-        value = KPI_admin(),
+        value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
         caption = "success",
@@
          ) |> 
-        rename( # renaming for the purposes of the plot
+        # renaming for the purposes of the plot
+        rename( 
           Year = year,
@@
          ) +
-        scale_fill_met_d("Troy") + # pretty colours
+        # pretty colours
+        scale_fill_met_d("Troy") + 
         labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
@@
 
-  # Creating fancy plotlyplot
-    fancyplotly <- reactive({
+  # Creating fancy plotly output  
+    output$fancyplotly <- renderPlotly(
       ggplotly(fancyplot())
+    )
+
+# Creating plots - annual result
+    arsresultat_plot <- reactive({
+      htmldata() |> 
+      mutate(
+        profit = ifelse(arsresultat >= 0, TRUE, FALSE)
+      ) |> 
+      ggplot(aes(year, arsresultat, fill = profit)) +
+      geom_col() +
+      scale_y_continuous(
+        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+                        # Using the currency code if uniform
+                    label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+                        # alternative prefix if multiple currencies in the data
+                    label_dollar(prefix = "? "))
+        ) +
+      scale_fill_met_d("Tsimshian") +
+      theme(legend.position = "none") +
+      labs(x = "Year", y = "", title = "Annual result",
+             # Checking again for currency; including warning if not uniform
+        subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
+                         "",
+                         "Note: not all values are denominated in the same currency"))
     })
 
-  # Creating fancy plotly output  
-    output$fancyplotly <- renderPlotly(
-      fancyplotly()
+    output$arsresultat_plot <- renderPlotly(
+      ggplotly(arsresultat_plot())
     )
+
 })
-
 ```

### 2022-07-19 Shiny/Test/Test data.Rmd
Improved value boxes. Annual result plot.

@@
   scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
-  scale_fill_met_d("Tsimshian")
+  scale_fill_met_d("Tsimshian") +
+  theme(legend.position = "none")
 ggplotly(arsresultat_p_met)
@@
   theme(legend.position = "none") +
-  scale_fill_met_d("Tsimshian")
+  scale_fill_met_d("Tsimshian") +
+  theme(legend.position = "none") +
+  labs(x = "", y = "")
 ggplotly(ordinaert_p_met)
@@
 
+# Numeric, for further transformation
+
+KPI_fundraising_num <- kreft_excel |> 
+          mutate(
+            KPI_fundraising_num =
+              (revenue_private_contributions +
+                 revenue_membership_fees - expenses_fundraising) /
+              expenses_fundraising * 100) |> 
+  filter(year == max(year)) |> 
+  select(KPI_fundraising_num) |> 
+  pull()
+
+# To turn into percentage
+str_c(format(KPI_fundraising_num, digits = 3), "%")
+
+KPI_fundraising_num
+
+if (KPI_fundraising_num < 100) {"danger"} else if (KPI_fundraising_num < 200) {"warning"} else {"success"}
+
 kreft_excel |> 
@@
         format(
-        ((revenue_private_contributions +                  revenue_membership_fees) / expenses_fundraising * 100), digits = 3
+        ((revenue_private_contributions + revenue_membership_fees - expenses_fundraising) / expenses_fundraising * 100), digits = 3
         ), "%"

### 2022-07-19 Shiny/Test/Flexdashboard rows.Rmd
Changing functions to keep numbers in thousands; changing CEO plot to keep salary.

@@
          ) |> 
+        mutate(
+          lederlonn = lederlonn * 1000,
+          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
+        ) |> 
         # renaming for the purposes of the plot
@@
       theme(legend.position = "none") +
-      labs(x = "Year", y = "", title = "Annual result",
+      labs(x = "Year", y = "", title = "Annual result (in thousands)",
              # Checking again for currency; including warning if not uniform
@@
       ) |> 
-      mutate( # values no longer in 1000
-        value = value * 1000
-      ) |> 
       distinct() |> # removing any remaining duplicate rows
@@
     read_excel(file) |> 
-      mutate( # Changing values
-        Value = Value * 1000
-      ) |> 
       unite( # uniting Category and Name to deal with multiple 'Other'

### 2022-07-19 Shiny/Test/Flexdashboard rows.Rmd
Colours are driving me insane.

@@
 
-The KPIs are based on the following calculations
+The KPIs are based on the following calculations:
 
-- **admin** is based on dividing
+- **admin** is based on dividing - expenses_administrative / expenses_program_services *100
+- **Fundraising ROI** - (revenue_private_contributions +
+                 revenue_membership_fees - expenses_fundraising) /
+              expenses_fundraising * 100
 ```{r}

### 2022-07-19 Shiny/Test/Test data.Rmd
Colours are driving me insane.

@@
 
+### Egenkapital og gjeld
+```{r}
+kreft_html2 <- kreft_html |> 
+  mutate(
+    profit = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+  )
+kreft_html2 |> 
+  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = label_dollar()) +
+  scale_fill_met_d("Tsimshian",
+                   direction = if (mean(kreft_html2$profit) == 1){-1} else {1})
+
+kreft_html2 |> 
+  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = label_dollar()) +
+  scale_fill_met_d("Tsimshian",
+                   direction = ifelse(mean(kreft_html2$profit) == 1,-1, 1))
+
+# Need to add profit variables
+
+```
+
 ## KPI ideas

### 2022-07-19 Shiny/Test/Test data.Rmd
Slightly nicer html function.

@@
 }
+
+read_proff7 <- function(file, start, end){
+    read_html(file, encoding = "UTF-8") |> 
+      html_node("#inner-frame") |> 
+      html_table() |> 
+      select( # removing empty column
+        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+      mutate( # removing whitespace
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |> 
+      filter( # removing duplicate table - no idea why it's necessary
+        !row_number() > 176
+      ) |>
+      filter( # removing duplicate rows
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |> 
+      pivot_longer( # tidying data
+        all_of(end):all_of(start), names_to = "year"
+        ) |> # using all_of()
+      mutate( # changing to real NAs and turning years into numbers
+        value = na_if(value, "-"),
+        year = as.integer(year)) |> 
+      mutate( # adding currency column
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |> 
+      fill( # filling the currency column
+        valutakode, .direction = "updown"
+      ) |> 
+      filter( # removing dates and redundant currency
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode") |> 
+      mutate( # removing whitespace in numbers
+        value = str_replace_all(value, "\\s", ""),
+        value = as.numeric(value),
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value),
+        value = value * 1000
+      ) |> 
+      distinct() |> # removing any remaining duplicate rows
+      pivot_wider( # pivoting
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |> 
+      clean_names() |> # tidying names
+      select(-lederlonn_i_hele_1000,
+             -resultatregnskap_i_hele_1000,
+             -balanseregnskap_i_hele_1000) |> # removing headings
+      rename( # making it clear what lonn refers to
+        lederlonn = lonn
+      ) |> 
+      arrange(desc(year)) # arranging by year
+}
+
 ```
@@
 kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
+
+kreft_html3 <- read_proff7("Data/Kreftforeningen.html", "2002", "2021")
+
+identical(kreft_html, kreft_html3)
 ```

### 2022-07-20 Shiny/Test/Flexdashboard rows.Rmd
read_proff8 function. met.brewer test

@@
         caption = "Admin spend",
-        color = "primary",
+        color = met.brewer("Wissing", n = 1),
         icon = "fa-random"

### 2022-07-20 Shiny/Test/Test data.Rmd
read_proff8 function. met.brewer test

@@
 
+read_proff8 <- function(file, start, end){
+    read_html(file, encoding = "UTF-8") |> 
+      html_node("#inner-frame") |> 
+      html_table() |> 
+      select( # removing empty column
+        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+      mutate( # removing whitespace
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |> 
+      filter( # removing duplicate table - no idea why it's necessary
+        !row_number() > 176
+      ) |>
+      filter( # removing duplicate rows
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |> 
+      pivot_longer( # tidying data
+        all_of(end):all_of(start), names_to = "year"
+        ) |> # using all_of()
+      mutate( # changing to real NAs and turning years into numbers
+        value = na_if(value, "-"),
+        year = as.integer(year), 
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |> 
+      fill( # filling the currency column
+        valutakode, .direction = "updown"
+      ) |> 
+      filter( # removing dates and redundant currency
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode") |> 
+      mutate( # removing whitespace in numbers
+        value = str_replace_all(value, "\\s", ""),
+        value = as.numeric(value),
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value),
+        value = value * 1000
+      ) |> 
+      distinct() |> # removing any remaining duplicate rows
+      pivot_wider( # pivoting
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |> 
+      clean_names() |> # tidying names
+      select(-lederlonn_i_hele_1000,
+             -resultatregnskap_i_hele_1000,
+             -balanseregnskap_i_hele_1000) |> # removing headings
+      rename( # making it clear what lonn refers to
+        lederlonn = lonn
+      ) |> 
+      arrange(desc(year)) # arranging by year
+}
+
 ```
@@
 identical(kreft_html, kreft_html3)
+
+kreft_html5 <- read_proff8("Data/Kreftforeningen.html", "2002", "2021")
+
+identical(kreft_html, kreft_html5)
 ```
@@
 
+annual_profit
+balance_positive
+
 ## KPI ideas
@@
 ```
+
+```{r}
+
+```

### 2022-07-20 Shiny/Test/Test data.Rmd
Adding new variables to function. Returning to values in thousands. Plot experiments.

--- 2022-07-19 Shiny/Test/Test data.Rmd
@@
 
+read_proff8 <- function(file, start, end){
+    read_html(file, encoding = "UTF-8") |> 
+      html_node("#inner-frame") |> 
+      html_table() |> 
+      select( # removing empty column
+        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+      mutate( # removing whitespace
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |> 
+      filter( # removing duplicate table - no idea why it's necessary
+        !row_number() > 176
+      ) |>
+      filter( # removing duplicate rows
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |> 
+      pivot_longer( # tidying data
+        all_of(end):all_of(start), names_to = "year"
+        ) |> # using all_of()
+      mutate( # changing to real NAs and turning years into numbers
+        value = na_if(value, "-"),
+        year = as.integer(year), 
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |> 
+      fill( # filling the currency column
+        valutakode, .direction = "updown"
+      ) |> 
+      filter( # removing dates and redundant currency
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode") |> 
+      mutate( # removing whitespace in numbers
+        value = str_replace_all(value, "\\s", ""),
+        value = as.numeric(value),
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+      ) |> 
+      distinct() |> # removing any remaining duplicate rows
+      pivot_wider( # pivoting
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |> 
+      clean_names() |> # tidying names
+      select(-lederlonn_i_hele_1000,
+             -resultatregnskap_i_hele_1000,
+             -balanseregnskap_i_hele_1000) |> # removing headings
+      rename( # making it clear what lonn refers to
+        lederlonn = lonn
+      ) |> 
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |> 
+      arrange(desc(year)) # arranging by year
+}
+
 ```
@@
   read_excel(file) |> 
-    mutate(
-      Value = Value * 1000
-    ) |> 
   unite(
@@
 identical(kreft_html, kreft_html3)
+
+kreft_html5 <- read_proff8("Data/Kreftforeningen.html", "2002", "2021")
+
+identical(kreft_html, kreft_html5)
+
+kreft_html5 |> 
+  select(arsresultat, sum_egenkapital_og_gjeld, annual_profit, balance_positive)
 ```
@@
                          "Note: not all values are denominated in the same currency"))
+
+# New
+kreft_html5 |> 
+  select(
+          year, lederlonn, leder_annen_godtgjorelse
+         ) |> 
+        # Compensation in ordinary numbers
+        mutate(
+          lederlonn = lederlonn * 1000,
+          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
+        ) |> 
+        # renaming for the purposes of the plot
+        rename( 
+          Year = year,
+          Salary = lederlonn,
+          `Other compensation` = leder_annen_godtgjorelse
+        ) |> 
+        # pivoting to join the two values to use fill
+        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+        ggplot(aes(Year, value, fill = Compensation)) +
+        geom_col() +
+        scale_y_continuous(
+          # Checking for uniform currency code
+        labels = ifelse(length(unique(kreft_html5$valutakode)) == 1,
+                        # Using the currency code if uniform
+                    label_dollar(prefix = str_c(kreft_html5$valutakode[1], " ")),
+                        # alternative prefix if multiple currencies in the data
+                    label_dollar(prefix = "? "))
+         ) +
+        # pretty colours
+        scale_fill_met_d("Troy") + 
+        labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+             # Checking again for currency; including warning if not uniform
+        subtitle = ifelse(length(unique(kreft_html5$valutakode)) == 1,
+                         "",
+                         "Note: not all values are denominated in the same currency"))
+kreft_html5 |> 
+ggplot(aes(year, arsresultat, fill = annual_profit)) +
+      geom_col() +
+      scale_y_continuous(
+        labels = ifelse(length(unique(kreft_html5$valutakode)) == 1,
+                        # Using the currency code if uniform
+                    label_dollar(prefix = str_c(kreft_html5$valutakode[1], " ")),
+                        # alternative prefix if multiple currencies in the data
+                    label_dollar(prefix = "? "))
+        ) +
+        # slightly nicer colours + ensuring positive values are always green
+      scale_fill_met_d("Tsimshian",
+                   direction = ifelse(mean(kreft_html5$annual_profit, na.rm = TRUE) == 1,-1, 1)) +
+      theme(legend.position = "none") +
+      labs(x = "Year", y = "", title = "Annual result (in thousands)",
+             # Checking again for currency; including warning if not uniform
+        subtitle = ifelse(length(unique(kreft_html5$valutakode)) == 1,
+                         "",
+                         "Note: not all values are denominated in the same currency"))
+
+mean(kreft_html5$annual_profit)
+kreft_html5$annual_profit
 ```
@@
   theme(legend.position = "none") +
-  scale_fill_met_d("Tsimshian") +
+  scale_fill_met_d("Tsimshian",
+                   direction = ifelse(mean(kreft_html$ordinaert_resultat) == 1,-1, 1)) +
   theme(legend.position = "none") +
@@
 ```
+
+annual_profit
+balance_positive
 
@@
 ```
+
+```{r}
+
+```
+
+## More functions
+```{r}
+# mutate(
+#       KPI_admin =
+#         str_c( # Using str_c() to concatenate value and % sign
+#          format( # shortening(lengthening) the resulting value
+#            # calculating the value
+#           (expenses_administrative / expenses_program_services *100), digits = 3
+#           ), "%"
+#          )
+#       ) |> 
+#       filter(year == max(year)) |> # selecting latest available year
+#        select(KPI_admin) |> # dropping all other variables
+#         str_flatten() # turning into a single string
+
+kreft_excel2 |> 
+  filter(year == max(year)) |> 
+  transmute(
+    boxvalue = str_c(format(KPI_admin, digits = 3), "%")
+  ) |> 
+  str_flatten()
+```

### 2022-07-20 Shiny/Test/Flexdashboard rows.Rmd
Updating functions and plots.

@@
 
-### Plotly
+### Annual result
 ```{r}
-plotlyOutput("plotlyplot")
+plotlyOutput("arsresultat_plotly")
 ```
 
-### Annual result
+### Balance
 ```{r}
-plotlyOutput("arsresultat_plot")
+
 ```
@@
 ```{r}
-plotlyOutput("fancyplotly")
+plotlyOutput("CEO_plotly")
 ```
@@
     htmldata <- reactive({
-      read_proff6(input$html$datapath,
+      read_proff8(input$html$datapath,
                   as.character(input$start), as.character(input$end))
@@
 #---- Plots --------------------------------------------------------------------    
-# creating plots - CEO compensation
-
-    # ggplot as reactive function
-    lonnplot <- reactive({
-      ggplot(data = htmldata(), aes(year, lederlonn)) +
-        geom_col() +
-        scale_y_continuous(
-          labels = scales::label_dollar(prefix = ""))
-    })
-
-    # Creating plotly plot
-    output$plotlyplot <- renderPlotly(
-      ggplotly(lonnplot())
-    )
-
-    # Creating ggplot
-    output$htmlplot <- renderPlot(
-      lonnplot()
-    )
-
-    # fancier plot function
-    fancyplot <- reactive(
+#----CEO compensation-----------------------------------------------------------
+    CEO_plot <- reactive(
       htmldata() |> 
@@
          ) |> 
+        # Compensation in ordinary numbers
         mutate(
@@
   # Creating fancy plotly output  
-    output$fancyplotly <- renderPlotly(
-      ggplotly(fancyplot())
+    output$CEO_plotly <- renderPlotly(
+      ggplotly(CEO_plot())
     )
 
-# Creating plots - annual result
+#----Annual result---------------------------------------------
     arsresultat_plot <- reactive({
       htmldata() |>
-      mutate(
-        profit = ifelse(arsresultat >= 0, TRUE, FALSE)
-      ) |> 
-      ggplot(aes(year, arsresultat, fill = profit)) +
+      ggplot(aes(year, arsresultat, fill = annual_profit)) +
       geom_col() +
@@
         ) +
-      scale_fill_met_d("Tsimshian") +
+        # slightly nicer colours + ensuring positive values are always green
+      scale_fill_met_d("Tsimshian",
+                   direction = ifelse(mean(htmldata()$annual_profit, na.rm = TRUE) == 1,-1, 1)) +
       theme(legend.position = "none") +
@@
 
-    output$arsresultat_plot <- renderPlotly(
+    output$arsresultat_plotly <- renderPlotly(
       ggplotly(arsresultat_plot())
     )
+#----Balance--------------------------------------------------------------------
 
@@
 # Function to read html file
-read_proff6 <- function(file, start, end){
+read_proff8 <- function(file, start, end){
     read_html(file, encoding = "UTF-8") |> 
@@
       pivot_longer( # tidying data
-        all_of(end):all_of(start), names_to = "year") |> # using all_of()
+        all_of(end):all_of(start), names_to = "year"
+        ) |> # using all_of()
       mutate( # changing to real NAs and turning years into numbers
         value = na_if(value, "-"),
-        year = as.integer(year)) |> 
-      mutate( # adding currency column
+        year = as.integer(year), 
         valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
@@
       mutate( # removing whitespace in numbers
-        value = str_replace_all(value, "\\s", "")
-      ) |> 
-      mutate( # turning into numbers
-        value = as.numeric(value)
-      ) |> 
-      mutate( # removing years as values
+        value = str_replace_all(value, "\\s", ""),
+        value = as.numeric(value),
         value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
@@
         lederlonn = lonn
+      ) |> 
+      # Creating new variables for plots
+      mutate(
+        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
       ) |> 

### 2022-07-20 Shiny/Test/Flexdashboard rows.Rmd
Letting styler do it's thing to make everything more readable.

@@
 # Loading packages using pacman to ensure they're all installed
-if (!require("pacman")) install.packages("pacman"); library(pacman)
-p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-       flexdashboard, MetBrewer, fontawesome, htmltools)
+if (!require("pacman")) install.packages("pacman")
+library(pacman)
+p_load(
+  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
+  flexdashboard, MetBrewer, fontawesome, htmltools
+)
 # Using fontawesome for more up-to-date fontawesome icons
@@
 ```{r}
-
+plotlyOutput("balance_plotly")
 ```
@@
   htmldata <- reactive({
-      read_proff8(input$html$datapath,
-                  as.character(input$start), as.character(input$end))
+    read_proff8(
+      input$html$datapath,
+      as.character(input$start), as.character(input$end)
+    )
   })
@@
               # calculating the value
-          (expenses_administrative / expenses_program_services *100), digits = 3
+              (expenses_administrative / expenses_program_services * 100),
+              digits = 3
             ), "%"
@@
 
-    # Creating Fundraising ROI value
+  # Creating Fundraising ROI value -- transmute?
   KPI_fundraising_num <- reactive({
     exceldata() |>
-          mutate(
-            # Calculating the value as a percentage
+      # most recent year
+      filter(year == max(year)) |>
+      transmute(
+        # Calculating the value as a percentage; removing all other
         KPI_fundraising_num =
@@
             revenue_membership_fees - expenses_fundraising) /
-              expenses_fundraising * 100) |> 
-        # most recent year
-          filter(year == max(year)) |> 
+            expenses_fundraising * 100
+      ) |>
       # just want the single number
-          select(KPI_fundraising_num) |> 
       pull()
@@
       caption = "Fundraising ROI",
+
       # Colour depends on value
-        color = if (KPI_fundraising_num() < 100) {"danger"
-          } else if (KPI_fundraising_num() < 200) {"warning"
-            } else {"success"},
+      color = if (KPI_fundraising_num() < 100) {
+        "danger"
+      } else if (KPI_fundraising_num() < 200) {
+        "warning"
+      } else {
+        "success"
+      },
       icon = "fas fa-donate"
@@
       value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
-        caption = "success",
-        color = "success",
+      caption = "primary",
+      color = "primary",
       icon = "fas fa-coffee"
@@
       icon = "fas fa-skull-crossbones"
-    )})
+    )
+  })
 
@@
       icon = "fas fa-battery-empty"
-    )})
+    )
+  })
 
@@
   #----CEO compensation-----------------------------------------------------------
-    CEO_plot <- reactive(
+  CEO_plot <- reactive({
     htmldata() |>
+      # preparing for pivoting
       select(
@@
           # alternative prefix if multiple currencies in the data
-                    label_dollar(prefix = "? "))
+          label_dollar(prefix = "? ")
+        )
       ) +
@@
       scale_fill_met_d("Troy") +
-        labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+      labs(
+        x = "Year", y = "Total compensation", title = "Annual CEO compensation",
         # Checking again for currency; including warning if not uniform
@@
           "",
-                         "Note: not all values are denominated in the same currency"))
+          "Note: not all values are denominated in the same currency"
         )
+      )
+  })
 
@@
       scale_fill_met_d("Tsimshian",
-                   direction = ifelse(mean(htmldata()$annual_profit, na.rm = TRUE) == 1,-1, 1)) +
+        direction =
+          ifelse(mean(htmldata()$annual_profit,
+            na.rm = TRUE
+          ) == 1, -1, 1)
+      ) +
       theme(legend.position = "none") +
-      labs(x = "Year", y = "", title = "Annual result (in thousands)",
+      labs(
+        x = "Year", y = "", title = "Annual result (in thousands)",
         # Checking again for currency; including warning if not uniform
@@
           "",
-                         "Note: not all values are denominated in the same currency"))
+          "Note: not all values are denominated in the same currency"
+        )
+      )
   })
@@
   #----Balance--------------------------------------------------------------------    
+  balance_plot <- reactive({
+    htmldata() |>
+      ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
+      geom_col() +
+      scale_y_continuous(
+        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+          # Using the currency code if uniform
+          label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+          # alternative prefix if multiple currencies in the data
+          label_dollar(prefix = "? ")
+        )
+      ) +
+      # slightly nicer colours + ensuring positive values are always green
+      scale_fill_met_d("Tsimshian",
+        direction = ifelse(
+          mean(htmldata()$balance_positive,
+            na.rm = TRUE
+          ) == 1, -1, 1
+        )
+      ) +
+      theme(legend.position = "none") +
+      labs(
+        x = "Year", y = "", title = "Balance (in thousands)",
+        # Checking again for currency; including warning if not uniform
+        subtitle =
+          ifelse(length(unique(htmldata()$valutakode)) == 1,
+            "",
+            "Note: not all values are denominated in the same currency"
+          )
+      )
+  })
 
+  output$balance_plotly <- renderPlotly(
+    ggplotly(balance_plot())
+  )
 })
@@
 # Function to read html file
-read_proff8 <- function(file, start, end){
+read_proff8 <- function(file, start, end) {
   read_html(file, encoding = "UTF-8") |>
@@
     select( # removing empty column
-        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
+      REGNSKAPSPERIODE:all_of(start)
+    ) |> # using all_of()
     mutate( # removing whitespace
@@
     pivot_longer( # tidying data
-        all_of(end):all_of(start), names_to = "year"
+      all_of(end):all_of(start),
+      names_to = "year"
     ) |> # using all_of()
@@
     fill( # filling the currency column
-        valutakode, .direction = "updown"
+      valutakode,
+      .direction = "updown"
     ) |>
@@
       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-          REGNSKAPSPERIODE != "Valutakode") |> 
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
     mutate( # removing whitespace in numbers
@@
     clean_names() |> # tidying names
-      select(-lederlonn_i_hele_1000,
+    select(
+      -lederlonn_i_hele_1000,
       -resultatregnskap_i_hele_1000,
-             -balanseregnskap_i_hele_1000) |> # removing headings
+      -balanseregnskap_i_hele_1000
+    ) |> # removing headings
     rename( # making it clear what lonn refers to
@@
 # Function to read Excel file
-widen_excel <- function(file){
+widen_excel <- function(file) {
   read_excel(file) |>
     unite( # uniting Category and Name to deal with multiple 'Other'
-        entry, c(Category, Name), sep = "_"
+      entry, c(Category, Name),
+      sep = "_"
     ) |>

### 2022-07-20 Shiny/Test/Test data.Rmd
More plot and style experiments.

@@
       html_table() |> 
-      select( # removing empty column
-        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
-      mutate( # removing whitespace
+    # removing empty column
+      select( 
+        REGNSKAPSPERIODE:all_of(start)) |>
+    # removing whitespace
+      mutate( 
         REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
       ) |> 
-      filter( # removing duplicate table - no idea why it's necessary
+    # removing duplicate table - no idea why it's necessary
+      filter(
         !row_number() > 176
       ) |>
-      filter( # removing duplicate rows
+    # removing duplicate rows
+      filter( 
         !grepl("Lukk", REGNSKAPSPERIODE)
       ) |> 
-      pivot_longer( # tidying data
+    # tidying data
+      pivot_longer( 
         all_of(end):all_of(start), names_to = "year"
-        ) |> # using all_of()
-      mutate( # changing to real NAs and turning years into numbers
+        ) |> 
+    # changing to real NAs and turning years into numbers
+    # picking out the currency
+      mutate( 
         value = na_if(value, "-"),
@@
       ) |> 
-      fill( # filling the currency column
+    # filling the currency column
+      fill( 
         valutakode, .direction = "updown"
       ) |> 
-      filter( # removing dates and redundant currency
+    # removing dates and redundant currency
+      filter( 
         REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
           REGNSKAPSPERIODE != "Valutakode") |> 
-      mutate( # removing whitespace in numbers
+    # removing white space in numbers
+    # removing years as numbers
+      mutate( 
         value = str_replace_all(value, "\\s", ""),
@@
       ) |> 
-      distinct() |> # removing any remaining duplicate rows
-      pivot_wider( # pivoting
+    # removing any remaining duplicate rows
+      distinct() |> 
+    # pivoting
+      pivot_wider( 
         names_from = REGNSKAPSPERIODE, values_from = value
       ) |> 
-      clean_names() |> # tidying names
+    # tidying names
+      clean_names() |> 
+    # removing headings
       select(-lederlonn_i_hele_1000,
              -resultatregnskap_i_hele_1000,
-             -balanseregnskap_i_hele_1000) |> # removing headings
-      rename( # making it clear what lonn refers to
+             -balanseregnskap_i_hele_1000) |> 
+    # making it clear what lonn refers to
+      rename( 
         lederlonn = lonn
@@
     ) |> 
-      arrange(desc(year)) # arranging by year
+    # arranging by year
+      arrange(desc(year)) 
 }
@@
 
-kreft_html5 <- read_proff8("Data/Kreftforeningen.html", "2002", "2021")
+kreft_html_with_variables <- read_proff8("Data/Kreftforeningen.html", "2002", "2021")
 
@@
 
+kreft_html2 |> 
+  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = profit)) +
+  geom_col() +
+  scale_y_continuous(labels = label_dollar()) +
+  scale_fill_met_d("Tsimshian",
+                   direction = ifelse(mean(kreft_html2$profit) == 1,-1, 1))
+
 # Need to add profit variables
+
+kreft_html_with_variables |> 
+  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
+  geom_col() +
+  scale_y_continuous(
+        labels = ifelse(length(unique(kreft_html_with_variables$valutakode)) == 1,
+                        # Using the currency code if uniform
+                    label_dollar(prefix = str_c(kreft_html_with_variables$valutakode[1], " ")),
+                        # alternative prefix if multiple currencies in the data
+                    label_dollar(prefix = "? "))
+        ) +
+  scale_fill_met_d("Tsimshian",
+                   direction = ifelse(
+                     mean(kreft_html_with_variables$balance_positive,
+                          na.rm = TRUE) == 1,-1, 1)) +
+  theme(legend.position = "none") +
+  labs(x = "Year", y = "", title = "Balance (in thousands)",
+             # Checking again for currency; including warning if not uniform
+        subtitle = ifelse(length(unique(kreft_html_with_variables$valutakode)) == 1,
+                         "",
+                         "Note: not all values are denominated in the same currency"))
 

### 2022-07-20 Shiny/Test/Test data.Rmd
Stylering.

@@
 
-read_proff7 <- function(file, start, end){
+read_proff7 <- function(file, start, end) {
   read_html(file, encoding = "UTF-8") |>
@@
       value = na_if(value, "-"),
-        year = as.integer(year)) |> 
+      year = as.integer(year)
+    ) |>
     mutate( # adding currency column
@@
       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |> 
+    # arranging by year
+      arrange(desc(year)) 
+}
+
+read_proff_tidied <- function(file, start, end) {
+    read_html(file, encoding = "UTF-8") |>
+        html_node("#inner-frame") |>
+        html_table() |>
+        # removing empty column
+    select(REGNSKAPSPERIODE:all_of(start)) |>
+        # removing whitespace
+    mutate(REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)) |>
+        # removing duplicate table - no idea why it's necessary
+    filter(!row_number() > 176) |>
+        # removing duplicate rows
+    filter(!grepl("Lukk", REGNSKAPSPERIODE)) |>
+        # tidying data
+    pivot_longer(
+        all_of(end):all_of(start),
+        names_to = "year"
+    ) |>
+        # changing to real NAs and turning years into numbers picking
+        # out the currency
+    mutate(
+        value = na_if(value, "-"),
+        year = as.integer(year),
+        valutakode = ifelse(
+            REGNSKAPSPERIODE == "Valutakode", value,
+            NA
+        )
+    ) |>
+        # filling the currency column
+    fill(valutakode, .direction = "updown") |>
+        # removing dates and redundant currency
+    filter(
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE !=
+            "Startdato" & REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+        # removing white space in numbers and removing years as numbers
+    mutate(
+        value = str_replace_all(value, "\\s", ""),
+        value = as.numeric(value),
+        value = ifelse(
+            grepl("i hele 1000", REGNSKAPSPERIODE),
+            NA, value
+        )
+    ) |>
+        # removing any remaining duplicate rows
+    distinct() |>
+        # pivoting
+    pivot_wider(names_from = REGNSKAPSPERIODE, values_from = value) |>
+        # tidying names
+    clean_names() |>
+        # removing headings
+    select(
+        -lederlonn_i_hele_1000, -resultatregnskap_i_hele_1000,
+        -balanseregnskap_i_hele_1000
+    ) |>
+        # making it clear what lonn refers to
+    rename(lederlonn = lonn) |>
+        # Creating new variables for plots
+    mutate(
+        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+        balance_positive = ifelse(
+            sum_egenkapital_og_gjeld >= 0, TRUE,
+            FALSE
+        )
     ) |>
@@
       label_dollar(prefix = str_c(kreft_html$valutakode[1], " ")),
-                    label_dollar(prefix = "? "))
+      label_dollar(prefix = "? ")
+    )
   ) +
   scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
+  labs(
+    x = "Year", y = "Total compensation", title = "Annual CEO compensation",
     subtitle = ifelse(length(unique(kreft_html$valutakode)) == 1,
       "",
-                         "Note: not all values are denominated in the same currency"))
+      "Note: not all values are denominated in the same currency"
+    )
+  )
 

### 2022-07-20 Shiny/Test/Flexdashboard rows.Rmd
Testing themes.

@@
     vertical_layout: scroll
+    theme: lumen
 runtime: shiny
@@
 htmltools::tagList(fontawesome::fa_html_dependency())
+
+# Considering lumen or cerulean theme
 ```
@@
       caption = "Admin spend",
-      color = met.brewer("Wissing", n = 1),
+      color = met.brewer("Troy", n = 1),
       icon = "fa-random"
@@
 
-  # Creating fancy plotly output
+  # Creating plotly output
   output$CEO_plotly <- renderPlotly(

### 2022-07-20 Shiny/Test/Flexdashboard rows.Rmd
Renaming value boxes.

@@
 ```{r}
-valueBoxOutput("KPI_box")
+valueBoxOutput("KPI_box_admin")
 ```
@@
 ```{r}
-valueBoxOutput("KPI_box2")
+valueBoxOutput("KPI_box_fundraising")
 ```
@@
 
-  # Creating Fundraising ROI value -- transmute?
+  # Creating Fundraising ROI value
   KPI_fundraising_num <- reactive({
@@
   # Creating the value box for administrative expenses
-  output$KPI_box <- renderValueBox({
+  output$KPI_box_admin <- renderValueBox({
     valueBox(
@@
   # Unsure about colour levels
-  output$KPI_box2 <- renderValueBox({
+  output$KPI_box_fundraising <- renderValueBox({
     valueBox(

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
Changing a colour.

@@
       caption = "Admin spend",
-      color = met.brewer("Troy", n = 1),
+      color = met.brewer("Wissing", n = 1),
       icon = "fa-random"

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
KPIs.

--- 2022-07-20 Shiny/Test/Flexdashboard rows.Rmd
@@
 ```{r}
-valueBoxOutput("KPI_box3")
+valueBoxOutput("KPI_box_OPX")
 ```
@@
 plotlyOutput("balance_plotly")
+```
+
+Row 
+-----------------------------------------------------------------------
+
+### Expenses
+```{r}
+plotlyOutput("expenses_plotly")
 ```
@@
               # calculating the value
-              (expenses_administrative / expenses_program_services * 100),
+              (expenses_administrative / (expenses_program_services +
+                                            expenses_administrative +
+                                            expenses_fundraising +
+                                            expenses_other) * 100),
               digits = 3
@@
 
+  # Creating OPX KPI
+  KPI_OPX_num <- reactive({
+    exceldata() |> 
+      filter(year == max(year)) |> 
+      transmute(
+        KPI_OPX = (revenue_private_contributions + revenue_investments +
+                 assets_cash_in_bank_and_cash_equivalents) /
+      (liability_short_term_grants_payable +
+         liability_revocable_endowments +
+         liability_long_term_grants_payable) * 100
+      ) |> 
+      pull()
+  })
+
   # Creating the value box for administrative expenses
@@
       value = KPI_admin(),
-      caption = "Admin spend",
-      color = met.brewer("Troy", n = 1),
+      caption = "Administration as a % of total expenses",
+      color = met.brewer("Wissing", n = 1),
       icon = "fa-random"
@@
 
-  # Duplicate value boxes for testing
-  output$KPI_box3 <- renderValueBox({
+  # KPI_OPX
+  output$KPI_box_OPX <- renderValueBox({
     valueBox(
-      value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
-      caption = "primary",
-      color = "primary",
+      value = str_c(format(KPI_OPX_num(), digits = 3), "%"),
+      caption = "Liquid assets / obligations",
+      color = ifelse(KPI_OPX_num() <= 100, "success", "danger"),
       icon = "fas fa-coffee"
@@
   })
+
+  # Duplicate value boxes for testing
 
@@
   )
+
+  #----Expenses---------------------------------------
+  expenses_plot <- reactive({
+    exceldata() |> 
+  transmute(
+    Year = as.factor(year),
+    Other = expenses_other,
+    Administrative = expenses_administrative,
+    Fundraising = expenses_fundraising,
+    `Program services` = expenses_program_services
+  ) |> 
+  pivot_longer(c(Other:`Program services`), names_to = "Expenses") |> 
+  mutate(
+    Expenses = ordered(Expenses, levels =
+                         c("Program services", "Administrative",
+                         "Fundraising", "Other")) 
+  ) |> 
+  ggplot(aes(Year, value, fill = Expenses)) +
+  geom_col() +
+  coord_flip() +
+  scale_y_continuous(labels = label_dollar()) +
+  scale_fill_met_d("Troy")
+  })
+
+  output$expenses_plotly <- renderPlotly(
+    ggplotly(expenses_plot())
+  )
 })

### 2022-07-21 Shiny/Test/Test data.Rmd
KPIs.

@@
 ```{r view, eval=FALSE}
-View(kreft_html)
+View(kreft_html_with_variables)
 View(kreft_excel)
@@
 kreft_excel
+kreft_html_with_variables
+
 ```
@@
 
-annual_profit
-balance_positive
+### Expenses
+```{r}
+kreft_excel |> 
+  transmute(
+    Year = as.factor(year),
+    Other = expenses_other,
+    Administrative = expenses_administrative,
+    Fundraising = expenses_fundraising,
+    `Program services` = expenses_program_services
+  ) |> 
+  pivot_longer(c(Other:`Program services`), names_to = "Expenses") |> 
+  mutate(
+    Expenses = ordered(Expenses, levels =
+                         c("Program services", "Administrative",
+                         "Fundraising", "Other")) 
+  ) |> 
+  ggplot(aes(Year, value, fill = Expenses)) +
+  geom_col() +
+  coord_flip() +
+  scale_y_continuous(labels = label_dollar()) +
+  scale_fill_met_d("Troy")
+
+kreft_excel |> 
+  transmute(
+    Year = as.factor(year),
+    Administrative = expenses_administrative,
+    Fundraising = expenses_fundraising,
+    `Program services` = expenses_program_services,
+    Other = expenses_other
+  )
+?as_factor
+```
 
@@
 
-### Overhead
+### Private donation growth
 ```{r}
+kreft_excel |> 
+  select(year, revenue_private_contributions) |> 
+  pivot_wider(names_from = year, values_from = revenue_private_contributions) |> 
+  mutate(
+    growth = `2020` / `2019` * 100
+  )
 
+?pivot_wider
 ```
 
+### Their KPI
+
 ```{r}
+# private contributions, investments, cash in bank / short-term garnts payable, revocable endowments, long-term grants payable
 
+mutate(kreft_excel,
+    KPI_OPX = (revenue_private_contributions + revenue_investments +
+                 assets_cash_in_bank_and_cash_equivalents) /
+      (liability_short_term_grants_payable +
+         liability_revocable_endowments +
+         liability_long_term_grants_payable) * 100
+  ) |> select(year, KPI_OPX)
+
+kreft_excel |> 
+  filter(year == max(year)) |> 
+      transmute(
+        KPI_OPX = (revenue_private_contributions + revenue_investments +
+                 assets_cash_in_bank_and_cash_equivalents) /
+      (liability_short_term_grants_payable +
+         liability_revocable_endowments +
+         liability_long_term_grants_payable) * 100
+      ) |> 
+      pull()
 ```

### 2022-07-21 Shiny/Test/Test data.Rmd
Donation growth KPI.

@@
 kreft_excel |> 
-  select(year, revenue_private_contributions) |> 
-  pivot_wider(names_from = year, values_from = revenue_private_contributions) |> 
-  mutate(
-    growth = `2020` / `2019` * 100
-  )
+  transmute(
+    growth = revenue_private_contributions[1] / revenue_private_contributions[2] * 100
+  ) |> 
+  pull() |> 
+  first()
 
-?pivot_wider
+?pull
 ```

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
Plotly madness (fixing currency subtitle).

@@
   # Creating KPI_admin value
-  KPI_admin <- reactive({
+  KPI_admin_num <- reactive({
     exceldata() |>
-      mutate(
-        KPI_admin =
-          str_c( # Using str_c() to concatenate value and % sign
-            format( # shortening(lengthening) the resulting value
-              # calculating the value
-              (expenses_administrative / (expenses_program_services +
+      # most recent year
+      filter(year == max(year)) |>
+      # Calculating the value as a percentage; removing all other
+      transmute(
+        KPI_admin = expenses_administrative / (expenses_program_services +
           expenses_administrative +
           expenses_fundraising +
-                                            expenses_other) * 100),
-              digits = 3
-            ), "%"
-          )
+          expenses_other) * 100
       ) |>
-      filter(year == max(year)) |> # selecting latest available year
-      select(KPI_admin) |> # dropping all other variables
-      str_flatten() # turning into a single string
+      # just want the single number
+      pull()
   })
@@
     valueBox(
-      value = KPI_admin(),
+      value = KPI_to_string(KPI_admin_num()),
       caption = "Administration as a % of total expenses",
@@
     valueBox(
-      # shortening(lengthening) the value with format()
-      # Using str_c() to concatenate value and % sign;
-      value = str_c(format(KPI_fundraising_num(), digits = 3), "%"),
+      value = KPI_to_string(KPI_fundraising_num()),
       caption = "Fundraising ROI",
@@
     valueBox(
-      value = KPI_admin(),
-      caption = "info",
-      color = "info",
-      icon = "fas fa-skull-crossbones"
+      value = KPI_to_string(KPI_admin_num()),
+      caption = "Administration as a % of total expenses",
+      color = met.brewer("Wissing", n = 1),
+      icon = "fa-random"
     )
@@
     valueBox(
-      value = KPI_admin(),
-      caption = "warning",
-      color = "warning",
-      icon = "fas fa-battery-empty"
+      value = KPI_to_string(KPI_admin_num()),
+      caption = "Administration as a % of total expenses",
+      color = met.brewer("Wissing", n = 1),
+      icon = "fa-random"
     )
@@
       labs(
-        x = "Year", y = "Total compensation", title = "Annual CEO compensation",
-        # Checking again for currency; including warning if not uniform
-        subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
-          "",
-          "Note: not all values are denominated in the same currency"
-        )
+        x = "", y = "Total compensation", title = "CEO compensation"
       )
@@
   output$CEO_plotly <- renderPlotly(
-    ggplotly(CEO_plot())
+    ggplotly(CEO_plot()) |>
+      layout(
+        annotations =
+          list(
+            x = 1, y = -0.1, text = currency_caption(),
+            showarrow = F, xref = "paper", yref = "paper",
+            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+            font = list(size = 11)
+          )
+      )
   )
@@
       geom_col() +
-      scale_y_continuous(
-        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
-          # Using the currency code if uniform
-          label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
-          # alternative prefix if multiple currencies in the data
-          label_dollar(prefix = "? ")
-        )
-      ) +
+      currency_scale() +
       # slightly nicer colours + ensuring positive values are always green
@@
       labs(
-        x = "Year", y = "", title = "Annual result (in thousands)",
-        # Checking again for currency; including warning if not uniform
-        subtitle = ifelse(length(unique(htmldata()$valutakode)) == 1,
-          "",
-          "Note: not all values are denominated in the same currency"
-        )
+        x = "", y = "", title = "Annual result (in thousands)"
       )
@@
   output$arsresultat_plotly <- renderPlotly(
-    ggplotly(arsresultat_plot())
+    ggplotly(arsresultat_plot()) |>
+      layout(
+        annotations =
+          list(
+            x = 1, y = -0.1, text = currency_caption(),
+            showarrow = F, xref = "paper", yref = "paper",
+            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+            font = list(size = 11)
+          )
+      )
   )
@@
       geom_col() +
-    scale_y_continuous(
-      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
-        # Using the currency code if uniform
-        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
-        # alternative prefix if multiple currencies in the data
-        label_dollar(prefix = "? ")
-      )
-    ) +
+      currency_scale() +
       # slightly nicer colours + ensuring positive values are always green
@@
       labs(
-      x = "Year", y = "", title = "Balance (in thousands)",
-      # Checking again for currency; including warning if not uniform
-      subtitle =
-        ifelse(length(unique(htmldata()$valutakode)) == 1,
-          "",
-          "Note: not all values are denominated in the same currency"
-        )
+        x = "", y = "", title = "Balance (in thousands)",
       )
@@
   output$balance_plotly <- renderPlotly(
-    ggplotly(balance_plot())
+    ggplotly(balance_plot()) |>
+      layout(
+        annotations =
+          list(
+            x = 1, y = -0.1, text = currency_caption(),
+            showarrow = F, xref = "paper", yref = "paper",
+            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+            font = list(size = 11)
+          )
+      )
   )
@@
       mutate(
-    Expenses = ordered(Expenses, levels =
-                         c("Program services", "Administrative",
-                         "Fundraising", "Other")) 
+        Expenses = ordered(Expenses,
+          levels =
+            c(
+              "Program services", "Administrative",
+              "Fundraising", "Other"
+            )
+        )
       ) |>
@@
       coord_flip() +
-  scale_y_continuous(labels = label_dollar()) +
-  scale_fill_met_d("Troy")
+      currency_scale() +
+      scale_fill_met_d("Troy") +
+      labs(x = "", y = "", title = "Expenses (in thousands)")
   })
@@
   output$expenses_plotly <- renderPlotly(
-    ggplotly(expenses_plot())
+    ggplotly(expenses_plot()) |>
+      layout(
+        annotations =
+          list(
+            x = 1, y = -0.1, text = currency_caption(),
+            showarrow = F, xref = "paper", yref = "paper",
+            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+            font = list(size = 11)
+          )
+      )
+  )
+
+  #----Reactive functions-------------------------------
+  # Reactive function to check for currency and use currency code or ?
+  currency_scale <- reactive({
+    scale_y_continuous(
+      # Checking for uniform currency code
+      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+        # Using the currency code if uniform
+        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+        # alternative prefix if multiple currencies in the data
+        label_dollar(prefix = "? ")
+      )
     )
   })
+
+  # Reactive function to create currency caption
+  currency_caption <- reactive({
+    if (length(unique(htmldata()$valutakode)) == 1) {
+      ""
+    } else {
+      "Note: not all values are denominated in the same currency"
+    }
+  })
+})
+
 ```
 
-```{r functions, include=FALSE}
+```{r non-reactive functions, include=FALSE}
 # Function to read html file
@@
 }
+
+# Function to turn KPI values into NA or percentage string
+KPI_to_string <- function(variable){
+  if (variable == 0 | is.na(variable)){
+    NA
+  } else {
+    str_c(format(variable, digits = 3), "%")
+  }
+}
+
 ```

### 2022-07-21 Shiny/Test/Test data.Rmd
Plotly madness (fixing currency subtitle).

@@
       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-    ) |> 
-    # arranging by year
-      arrange(desc(year)) 
-}
-
-read_proff_tidied <- function(file, start, end) {
-    read_html(file, encoding = "UTF-8") |>
-        html_node("#inner-frame") |>
-        html_table() |>
-        # removing empty column
-    select(REGNSKAPSPERIODE:all_of(start)) |>
-        # removing whitespace
-    mutate(REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)) |>
-        # removing duplicate table - no idea why it's necessary
-    filter(!row_number() > 176) |>
-        # removing duplicate rows
-    filter(!grepl("Lukk", REGNSKAPSPERIODE)) |>
-        # tidying data
-    pivot_longer(
-        all_of(end):all_of(start),
-        names_to = "year"
-    ) |>
-        # changing to real NAs and turning years into numbers picking
-        # out the currency
-    mutate(
-        value = na_if(value, "-"),
-        year = as.integer(year),
-        valutakode = ifelse(
-            REGNSKAPSPERIODE == "Valutakode", value,
-            NA
-        )
-    ) |>
-        # filling the currency column
-    fill(valutakode, .direction = "updown") |>
-        # removing dates and redundant currency
-    filter(
-        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE !=
-            "Startdato" & REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-        # removing white space in numbers and removing years as numbers
-    mutate(
-        value = str_replace_all(value, "\\s", ""),
-        value = as.numeric(value),
-        value = ifelse(
-            grepl("i hele 1000", REGNSKAPSPERIODE),
-            NA, value
-        )
-    ) |>
-        # removing any remaining duplicate rows
-    distinct() |>
-        # pivoting
-    pivot_wider(names_from = REGNSKAPSPERIODE, values_from = value) |>
-        # tidying names
-    clean_names() |>
-        # removing headings
-    select(
-        -lederlonn_i_hele_1000, -resultatregnskap_i_hele_1000,
-        -balanseregnskap_i_hele_1000
-    ) |>
-        # making it clear what lonn refers to
-    rename(lederlonn = lonn) |>
-        # Creating new variables for plots
-    mutate(
-        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-        balance_positive = ifelse(
-            sum_egenkapital_og_gjeld >= 0, TRUE,
-            FALSE
-        )
     ) |> 
@@
                          "Note: not all values are denominated in the same currency"))
-kreft_html5 |> 
+testplot <- kreft_html5 |> 
 ggplot(aes(year, arsresultat, fill = annual_profit)) +
@@
              # Checking again for currency; including warning if not uniform
-        subtitle = ifelse(length(unique(kreft_html5$valutakode)) == 1,
-                         "",
-                         "Note: not all values are denominated in the same currency"))
+        subtitle = "test")
 
-mean(kreft_html5$annual_profit)
-kreft_html5$annual_profit
+testplot
+ggplotly(testplot) |> layout(title = list(text = str_c('US State Population and Life Expectancy',
+                                    '<br>',
+                                    '<sup>',
+                                    string,
+                                    '</sup>'),
+                                    y = 0.95, x = 0.5, xanchor = 'center', yanchor =  'top'))
+
+layout(title = list(text='Plot Title', y = 0.95, x = 0.5, xanchor = 'center', yanchor =  'top'))
+string <- "test"
+
+"Note: not all values are denominated in the same currency"
 ```
@@
 ggplotly(ordinaert_p_met)
+
+# function extracted
+
+  # Reactive function to check for currency and use an appropriate subtitle
+  currency_subtitle <- reactive({
+    if (length(unique(htmldata()$valutakode)) == 1){
+      ""
+    } else {
+      "Note: not all values are denominated in the same currency"
+    }
+  })
 ```
@@
   geom_col() +
-  scale_y_continuous(
-        labels = ifelse(length(unique(kreft_html_with_variables$valutakode)) == 1,
-                        # Using the currency code if uniform
-                    label_dollar(prefix = str_c(kreft_html_with_variables$valutakode[1], " ")),
-                        # alternative prefix if multiple currencies in the data
-                    label_dollar(prefix = "? "))
-        ) +
+  currency_scale() +
   scale_fill_met_d("Tsimshian",
@@
 
+currency_scale <- function()
+  {scale_y_continuous(
+        # Checking for uniform currency code
+        labels = ifelse(length(unique(kreft_html_with_variables$valutakode)) == 1,
+          # Using the currency code if uniform
+          label_dollar(prefix = str_c(kreft_html_with_variables$valutakode[1], " ")),
+          # alternative prefix if multiple currencies in the data
+          label_dollar(prefix = "? ")
+        )
+      )
+  }
 ```
@@
           select(KPI_admin)
+
+kreft_excel |>
+      mutate(
+        KPI_admin =
+          str_c( # Using str_c() to concatenate value and % sign
+            format( # shortening(lengthening) the resulting value
+              # calculating the value
+              (expenses_administrative / (expenses_program_services +
+                                            expenses_administrative +
+                                            expenses_fundraising +
+                                            expenses_other) * 100),
+              digits = 3
+            ), "%"
+          )
+      ) |>
+      filter(year == max(year)) |> # selecting latest available year
+      select(KPI_admin) |> # dropping all other variables
+      str_flatten() # turning into a single string
+
+kreft_excel |>
+      mutate(
+        KPI_admin =
+            format( # shortening(lengthening) the resulting value
+              # calculating the value
+              (expenses_administrative / (expenses_program_services +
+                                            expenses_administrative +
+                                            expenses_fundraising +
+                                            expenses_other) * 100),
+              digits = 3)
+      ) |>
+      filter(year == max(year)) |> # selecting latest available year
+      select(KPI_admin) |> # dropping all other variables
+      pull
+  class()
+
+KPI_admin_num <- kreft_excel |> 
+  filter(year == max(year)) |> 
+  transmute(
+    KPI_admin = expenses_administrative / (expenses_program_services +
+                                            expenses_administrative +
+                                            expenses_fundraising +
+                                            expenses_other) * 100
+  ) |> 
+  pull()
+
+KPI_to_string <- function(variable){
+  if (variable == 0 | is.na(variable)){
+    NA
+  } else {
+    str_c(format(variable, digits = 3), "%")
+  }
+}
+
+KPI_to_string(KPI_admin_num)
+
 ```

### 2022-07-21 Shiny/Test/Test data.Rmd
Plotly caption function test.

@@
 "Note: not all values are denominated in the same currency"
+
+# bad test
+    mutate(
+    valutakode = ifelse(year == 2019, "USD", valutakode)
+  ) |>
 ```
@@
 ```{r}
-kreft_excel |> 
+expenses_plot <- kreft_excel |> 
   transmute(
@@
   scale_fill_met_d("Troy")
+
+ggplotly(expenses_plot) |>
+      layout(
+        annotations =
+          list(
+            x = 1, y = -0.1, text = "currency_caption()",
+            showarrow = F, xref = "paper", yref = "paper",
+            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+            font = list(size = 11)
+          )
+      )
+
+plotlylayout <- function()({
+  list(
+            x = 1, y = -0.1, text = "currency_caption()",
+            showarrow = F, xref = "paper", yref = "paper",
+            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+            font = list(size = 11)
+          )
+})
+
+ggplotly(expenses_plot) |> 
+layout(annotations = plotlylayout())
 

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
Retitled to NGO analysis Plotly caption function. Green/Red scale function. Commenting. Fixing check for invalid value box variables. Commenting out real html function, adding erroneous one for testing.

@@
 ---
-title: "Charity analysis"
+title: "NGO analysis"
 output: 
@@
       geom_col() +
-      scale_y_continuous(
-        # Checking for uniform currency code
-        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
-          # Using the currency code if uniform
-          label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
-          # alternative prefix if multiple currencies in the data
-          label_dollar(prefix = "? ")
-        )
-      ) +
+      currency_scale() +
       # pretty colours
@@
         annotations =
-          list(
-            x = 1, y = -0.1, text = currency_caption(),
-            showarrow = F, xref = "paper", yref = "paper",
-            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
-            font = list(size = 11)
-          )
+          plotly_caption()
       )
@@
       currency_scale() +
-      # slightly nicer colours + ensuring positive values are always green
-      scale_fill_met_d("Tsimshian",
-        direction =
-          ifelse(mean(htmldata()$annual_profit,
-            na.rm = TRUE
-          ) == 1, -1, 1)
-      ) +
+      scale_fill_green() +
       theme(legend.position = "none") +
@@
       currency_scale() +
-      # slightly nicer colours + ensuring positive values are always green
-      scale_fill_met_d("Tsimshian",
-        direction = ifelse(
-          mean(htmldata()$balance_positive,
-            na.rm = TRUE
-          ) == 1, -1, 1
-        )
-      ) +
+      scale_fill_green() +
       theme(legend.position = "none") +
@@
     exceldata() |>
+      # Renaming for the purposes of the plot + dropping other variables
       transmute(
@@
       ) |>
+      # Pivoting to use fill
       pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
-      mutate(
-        Expenses = ordered(Expenses,
-          levels =
-            c(
-              "Program services", "Administrative",
-              "Fundraising", "Other"
-            )
-        )
-      ) |>
       ggplot(aes(Year, value, fill = Expenses)) +
       geom_col() +
+      # Flipping, because the Excel file only contains two years atm
       coord_flip() +
@@
   currency_caption <- reactive({
+    # Checking for uniform currency code
     if (length(unique(htmldata()$valutakode)) == 1) {
+      # Empty caption if uniform
       ""
+      # Warning if multiple currencies
     } else {
@@
     }
+  })
+
+  # Reactive function to create plotly caption
+  # (ggplot captions and subtitles are lost in the conversion)
+  # (shamelessly stolen from the Internet, with minor modifications)
+  plotly_caption <- reactive({
+    list(
+      x = 1, y = -0.1, text = currency_caption(),
+      showarrow = F, xref = "paper", yref = "paper",
+      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+      font = list(size = 11)
+    )
+  })
+
+  # Reactive function with slightly nicer colours + ensuring positive values are always green
+  scale_fill_green <- reactive({
+    scale_fill_met_d("Tsimshian",
+      direction = ifelse(
+        mean(htmldata()$balance_positive,
+          na.rm = TRUE
+        ) == 1, -1, 1
+      )
+    )
   })
@@
 # Function to read html file
+# read_proff8 <- function(file, start, end) {
+#   read_html(file, encoding = "UTF-8") |>
+#     html_node("#inner-frame") |>
+#     html_table() |>
+#     select( # removing empty column
+#       REGNSKAPSPERIODE:all_of(start)
+#     ) |> # using all_of()
+#     mutate( # removing whitespace
+#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+#     ) |>
+#     filter( # removing duplicate table - no idea why it's necessary
+#       !row_number() > 176
+#     ) |>
+#     filter( # removing duplicate rows
+#       !grepl("Lukk", REGNSKAPSPERIODE)
+#     ) |>
+#     pivot_longer( # tidying data
+#       all_of(end):all_of(start),
+#       names_to = "year"
+#     ) |> # using all_of()
+#     mutate( # changing to real NAs and turning years into numbers
+#       value = na_if(value, "-"),
+#       year = as.integer(year),
+#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+#     ) |>
+#     fill( # filling the currency column
+#       valutakode,
+#       .direction = "updown"
+#     ) |>
+#     filter( # removing dates and redundant currency
+#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+#         REGNSKAPSPERIODE != "Valutakode"
+#     ) |>
+#     mutate( # removing whitespace in numbers
+#       value = str_replace_all(value, "\\s", ""),
+#       value = as.numeric(value),
+#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+#     ) |>
+#     distinct() |> # removing any remaining duplicate rows
+#     pivot_wider( # pivoting
+#       names_from = REGNSKAPSPERIODE, values_from = value
+#     ) |>
+#     clean_names() |> # tidying names
+#     select(
+#       -lederlonn_i_hele_1000,
+#       -resultatregnskap_i_hele_1000,
+#       -balanseregnskap_i_hele_1000
+#     ) |> # removing headings
+#     rename( # making it clear what lonn refers to
+#       lederlonn = lonn
+#     ) |>
+#     # Creating new variables for plots
+#     mutate(
+#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+#     ) |>
+#     arrange(desc(year)) # arranging by year
+# }
+
+# Test version of function - multiple currencies
 read_proff8 <- function(file, start, end) {
@@
     ) |>
+    # bad test - REMOVE
+    mutate(
+    valutakode = ifelse(year == 2019, "USD", valutakode)
+  ) |>
     arrange(desc(year)) # arranging by year
@@
 KPI_to_string <- function(variable){
-  if (variable == 0 | is.na(variable)){
+  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
+  # layout-breaking errors, otherwise formatting
+  if (is.na(variable)){
+    NA
+  } else if (variable == Inf){
     NA
   } else {
+    # Shortening (lengthening) the value, adding %
     str_c(format(variable, digits = 3), "%")

### 2022-07-21 Shiny/Test/Colourful plots.Rmd
Do coders dream of art?

--- 2022-07-21 Shiny/Test/Test data.Rmd
@@
 ---
-title: "Test data"
+title: "Colourful plot tests"
 author: "Torbjørn Skinnemoen Ottersen"
-date: '2022-07-16'
-output: 
-  html_document: 
-    df_print: kable
+date: '2022-07-21'
+output: html_document
 ---
@@
 
-Testing ideas for the Shiny app/dashboard.
-
-# Background
-
-## Libraries
-```{r libraries}
+## R Markdown
+```{r}
 if (!require("pacman")) install.packages("pacman"); library(pacman)
@@
 
-## Functions
-
-### HTML
-```{r}
-read_proff6 <- function(file, start, end){
-    read_html(file, encoding = "UTF-8") |> 
-      html_node("#inner-frame") |> 
-      html_table() |> 
-      select( # removing empty column
-        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
-      mutate( # removing whitespace
-        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-      ) |> 
-      filter( # removing duplicate table - no idea why it's necessary
-        !row_number() > 176
-      ) |>
-      filter( # removing duplicate rows
-        !grepl("Lukk", REGNSKAPSPERIODE)
-      ) |> 
-      pivot_longer( # tidying data
-        all_of(end):all_of(start), names_to = "year") |> # using all_of()
-      mutate( # changing to real NAs and turning years into numbers
-        value = na_if(value, "-"),
-        year = as.integer(year)) |> 
-      mutate( # adding currency column
-        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-      ) |> 
-      fill( # filling the currency column
-        valutakode, .direction = "updown"
-      ) |> 
-      filter( # removing dates and redundant currency
-        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-          REGNSKAPSPERIODE != "Valutakode") |> 
-      mutate( # removing whitespace in numbers
-        value = str_replace_all(value, "\\s", "")
-      ) |> 
-      mutate( # turning into numbers
-        value = as.numeric(value)
-      ) |> 
-      mutate( # removing years as values
-        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-      ) |> 
-      mutate( # values no longer in 1000
-        value = value * 1000
-      ) |> 
-      distinct() |> # removing any remaining duplicate rows
-      pivot_wider( # pivoting
-        names_from = REGNSKAPSPERIODE, values_from = value
-      ) |> 
-      clean_names() |> # tidying names
-      select(-lederlonn_i_hele_1000,
-             -resultatregnskap_i_hele_1000,
-             -balanseregnskap_i_hele_1000) |> # removing headings
-      rename( # making it clear what lonn refers to
-        lederlonn = lonn
-      ) |> 
-      arrange(desc(year)) # arranging by year
-}
-
-read_proff7 <- function(file, start, end) {
-  read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame") |>
-    html_table() |>
-    select( # removing empty column
-      REGNSKAPSPERIODE:all_of(start)
-    ) |> # using all_of()
-    mutate( # removing whitespace
-      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-    ) |>
-    filter( # removing duplicate table - no idea why it's necessary
-      !row_number() > 176
-    ) |>
-    filter( # removing duplicate rows
-      !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |>
-    pivot_longer( # tidying data
-      all_of(end):all_of(start),
-      names_to = "year"
-    ) |> # using all_of()
-    mutate( # changing to real NAs and turning years into numbers
-      value = na_if(value, "-"),
-      year = as.integer(year)
-    ) |>
-    mutate( # adding currency column
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |>
-    fill( # filling the currency column
-      valutakode,
-      .direction = "updown"
-    ) |>
-    filter( # removing dates and redundant currency
-      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-        REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-    mutate( # removing whitespace in numbers
-      value = str_replace_all(value, "\\s", ""),
-      value = as.numeric(value),
-      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value),
-      value = value * 1000
-    ) |>
-    distinct() |> # removing any remaining duplicate rows
-    pivot_wider( # pivoting
-      names_from = REGNSKAPSPERIODE, values_from = value
-    ) |>
-    clean_names() |> # tidying names
-    select(
-      -lederlonn_i_hele_1000,
-      -resultatregnskap_i_hele_1000,
-      -balanseregnskap_i_hele_1000
-    ) |> # removing headings
-    rename( # making it clear what lonn refers to
-      lederlonn = lonn
-    ) |>
-    arrange(desc(year)) # arranging by year
-}
+```{r data, echo=FALSE}
 
@@
 
-read_proff_tidied <- function(file, start, end) {
-    read_html(file, encoding = "UTF-8") |>
-        html_node("#inner-frame") |>
-        html_table() |>
-        # removing empty column
-    select(REGNSKAPSPERIODE:all_of(start)) |>
-        # removing whitespace
-    mutate(REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)) |>
-        # removing duplicate table - no idea why it's necessary
-    filter(!row_number() > 176) |>
-        # removing duplicate rows
-    filter(!grepl("Lukk", REGNSKAPSPERIODE)) |>
-        # tidying data
-    pivot_longer(
-        all_of(end):all_of(start),
-        names_to = "year"
-    ) |>
-        # changing to real NAs and turning years into numbers picking
-        # out the currency
-    mutate(
-        value = na_if(value, "-"),
-        year = as.integer(year),
-        valutakode = ifelse(
-            REGNSKAPSPERIODE == "Valutakode", value,
-            NA
-        )
-    ) |>
-        # filling the currency column
-    fill(valutakode, .direction = "updown") |>
-        # removing dates and redundant currency
-    filter(
-        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE !=
-            "Startdato" & REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-        # removing white space in numbers and removing years as numbers
-    mutate(
-        value = str_replace_all(value, "\\s", ""),
-        value = as.numeric(value),
-        value = ifelse(
-            grepl("i hele 1000", REGNSKAPSPERIODE),
-            NA, value
-        )
-    ) |>
-        # removing any remaining duplicate rows
-    distinct() |>
-        # pivoting
-    pivot_wider(names_from = REGNSKAPSPERIODE, values_from = value) |>
-        # tidying names
-    clean_names() |>
-        # removing headings
-    select(
-        -lederlonn_i_hele_1000, -resultatregnskap_i_hele_1000,
-        -balanseregnskap_i_hele_1000
-    ) |>
-        # making it clear what lonn refers to
-    rename(lederlonn = lonn) |>
-        # Creating new variables for plots
-    mutate(
-        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-        balance_positive = ifelse(
-            sum_egenkapital_og_gjeld >= 0, TRUE,
-            FALSE
-        )
-    ) |>
-        # arranging by year
-    arrange(desc(year))
-}
-
-```
-
-### Excel wide
-```{r}
 widen_excel <- function(file){
@@
 }
-```
 
-## Loading data
-```{r}
-kreft_html <- read_proff6("Data/Kreftforeningen.html", "2002", "2021")
 kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
 
-kreft_html3 <- read_proff7("Data/Kreftforeningen.html", "2002", "2021")
-
-identical(kreft_html, kreft_html3)
-
-kreft_html_with_variables <- read_proff8("Data/Kreftforeningen.html", "2002", "2021")
-
-identical(kreft_html, kreft_html5)
-
-kreft_html5 |> 
-  select(arsresultat, sum_egenkapital_og_gjeld, annual_profit, balance_positive)
-```
-
-## Viewing data
-
-### `View()`
-```{r view, eval=FALSE}
-View(kreft_html_with_variables)
-View(kreft_excel)
-```
-
-### kabled
-```{r}
-kreft_html
-kreft_excel
-kreft_html_with_variables
+kreft_html <- read_proff8("Data/Kreftforeningen.html", "2002", "2021")
 
@@
 
-# Testing
+```{r}
+palettes <- colorblind_palettes
 
-## Data manipulation
-```{r}
-kreft_excel |> 
-  filter(year == max(year))
-
-kreft_html |> 
-  filter(year == max(kreft_excel$year))
-
-kreft_excel |> 
-  left_join(kreft_html, by = "year")
-```
-
-## Visualisation ideas
-
-### CEO compensation and currency testing
-```{r}
-compensation_p <- kreft_html |> 
-  group_by(year) |> 
-  mutate(
-    leder_total = sum(lederlonn, leder_annen_godtgjorelse, na.rm = TRUE)
-  ) |> 
-  ggplot(aes(year, leder_total, fill = leder_total)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
-ggplotly(compensation_p)
-
-compensation_p2 <- kreft_html |> 
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |> 
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |> 
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = scales::label_dollar(prefix =
-                                    str_c(kreft_html$valutakode[1], " "))) +
-  scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
-ggplotly(compensation_p2)
-
-# if else structure to check for uniform currency code?
-length(unique(kreft_html$valutakode)) == 1
-
-# Testing with uniform currency
-if (length(unique(kreft_html$valutakode)) == 1) {
-  kreft_html |> 
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |> 
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |> 
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = scales::label_dollar(prefix =
-                                    str_c(kreft_html$valutakode[1], " "))) +
-  scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
-} else {
-    kreft_html |> 
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |> 
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |> 
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = scales::label_dollar(prefix = "? ")) +
-  scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
-       subtitle = "Note: not all values are denominated in the same currency")
-}
-
-# Creating data frame with multiple currencies
-currency_test <- kreft_html |> 
-  mutate(
-    valutakode = ifelse(year == 2019, "USD", valutakode)
-  )
-
-# Testing new data frame
-if (length(unique(currency_test$valutakode)) == 1) {
-  currency_test |> 
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |> 
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |> 
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = scales::label_dollar(prefix =
-                                    str_c(kreft_html$valutakode[1], " "))) +
-  scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation")
-} else {
-    currency_test |> 
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |> 
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |> 
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = scales::label_dollar(prefix = "? ")) +
-  scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
-       subtitle = "Note: not all values are denominated in the same currency")
-}
-
-# Could probably be done with two internal ifelse() functions
-
-# ifelse() version
-
-# Testing
-currency_test |> 
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |> 
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |> 
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = ifelse(length(unique(currency_test$valutakode)) == 1,
-                    label_dollar(prefix = str_c(kreft_html$valutakode[1], " ")),
-                    label_dollar(prefix = "? "))
-      ) +
-  scale_fill_met_d("Troy") +
-  labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
-       subtitle = ifelse(length(unique(currency_test$valutakode)) == 1,
-                         element_blank(),
-                         "Note: not all values are denominated in the same currency"))
-
-# Original
-kreft_html |>
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |>
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |>
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  scale_y_continuous(
-    labels = ifelse(length(unique(kreft_html$valutakode)) == 1,
-      label_dollar(prefix = str_c(kreft_html$valutakode[1], " ")),
-      label_dollar(prefix = "? ")
-    )
-  ) +
-  scale_fill_met_d("Troy") +
-  labs(
-    x = "Year", y = "Total compensation", title = "Annual CEO compensation",
-    subtitle = ifelse(length(unique(kreft_html$valutakode)) == 1,
-      "",
-      "Note: not all values are denominated in the same currency"
-    )
-  )
-
-# New
-kreft_html5 |> 
-  select(
-          year, lederlonn, leder_annen_godtgjorelse
-         ) |> 
-        # Compensation in ordinary numbers
-        mutate(
-          lederlonn = lederlonn * 1000,
-          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-        ) |> 
-        # renaming for the purposes of the plot
-        rename( 
-          Year = year,
-          Salary = lederlonn,
-          `Other compensation` = leder_annen_godtgjorelse
-        ) |> 
-        # pivoting to join the two values to use fill
-        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
-        ggplot(aes(Year, value, fill = Compensation)) +
-        geom_col() +
-        scale_y_continuous(
-          # Checking for uniform currency code
-        labels = ifelse(length(unique(kreft_html5$valutakode)) == 1,
-                        # Using the currency code if uniform
-                    label_dollar(prefix = str_c(kreft_html5$valutakode[1], " ")),
-                        # alternative prefix if multiple currencies in the data
-                    label_dollar(prefix = "? "))
-         ) +
-        # pretty colours
-        scale_fill_met_d("Troy") + 
-        labs(x = "Year", y = "Total compensation", title = "Annual CEO compensation",
-             # Checking again for currency; including warning if not uniform
-        subtitle = ifelse(length(unique(kreft_html5$valutakode)) == 1,
-                         "",
-                         "Note: not all values are denominated in the same currency"))
-kreft_html5 |> 
-ggplot(aes(year, arsresultat, fill = annual_profit)) +
-      geom_col() +
-      scale_y_continuous(
-        labels = ifelse(length(unique(kreft_html5$valutakode)) == 1,
-                        # Using the currency code if uniform
-                    label_dollar(prefix = str_c(kreft_html5$valutakode[1], " ")),
-                        # alternative prefix if multiple currencies in the data
-                    label_dollar(prefix = "? "))
-        ) +
-        # slightly nicer colours + ensuring positive values are always green
-      scale_fill_met_d("Tsimshian",
-                   direction = ifelse(mean(kreft_html5$annual_profit, na.rm = TRUE) == 1,-1, 1)) +
-      theme(legend.position = "none") +
-      labs(x = "Year", y = "", title = "Annual result (in thousands)",
-             # Checking again for currency; including warning if not uniform
-        subtitle = ifelse(length(unique(kreft_html5$valutakode)) == 1,
-                         "",
-                         "Note: not all values are denominated in the same currency"))
-
-mean(kreft_html5$annual_profit)
-kreft_html5$annual_profit
-```
-
-### Result
-```{r}
-arsresultat_p <- kreft_html |> 
-  mutate(
-    profit = ifelse(arsresultat >= 0, TRUE, FALSE)
-  ) |> 
-  ggplot(aes(year, arsresultat, fill = profit)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = ""))
-ggplotly(arsresultat_p)
-
-arsresultat_p_met <- kreft_html |> 
-  mutate(
-    profit = ifelse(arsresultat >= 0, TRUE, FALSE)
-  ) |> 
-  ggplot(aes(year, arsresultat, fill = profit)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
-  scale_fill_met_d("Tsimshian") +
-  theme(legend.position = "none")
-ggplotly(arsresultat_p_met)
-
-ordinaert_p <- kreft_html |> 
-  mutate(
-    profit = ifelse(ordinaert_resultat >= 0, TRUE, FALSE)
-  ) |> 
-  ggplot(aes(year, ordinaert_resultat, fill = profit)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
-  theme(legend.position = "none")
-ggplotly(ordinaert_p)
-
-ordinaert_p_met <- kreft_html |> 
-  mutate(
-    profit = ifelse(ordinaert_resultat >= 0, TRUE, FALSE)
-  ) |> 
-  ggplot(aes(year, ordinaert_resultat, fill = profit)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = "")) +
-  theme(legend.position = "none") +
-  scale_fill_met_d("Tsimshian",
-                   direction = ifelse(mean(kreft_html$ordinaert_resultat) == 1,-1, 1)) +
-  theme(legend.position = "none") +
-  labs(x = "", y = "")
-ggplotly(ordinaert_p_met)
-```
-
-### Egenkapital og gjeld
-```{r}
-kreft_html2 <- kreft_html |> 
-  mutate(
-    profit = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-  )
-kreft_html2 |> 
-  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = profit)) +
-  geom_col() +
-  scale_y_continuous(labels = label_dollar()) +
-  scale_fill_met_d("Tsimshian",
-                   direction = if (mean(kreft_html2$profit) == 1){-1} else {1})
-
-kreft_html2 |> 
-  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = profit)) +
-  geom_col() +
-  scale_y_continuous(labels = label_dollar()) +
-  scale_fill_met_d("Tsimshian",
-                   direction = ifelse(mean(kreft_html2$profit) == 1,-1, 1))
-
-kreft_html2 |> 
-  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = profit)) +
-  geom_col() +
-  scale_y_continuous(labels = label_dollar()) +
-  scale_fill_met_d("Tsimshian",
-                   direction = ifelse(mean(kreft_html2$profit) == 1,-1, 1))
-
-# Need to add profit variables
-
-kreft_html_with_variables |> 
-  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
-  geom_col() +
-  scale_y_continuous(
-        labels = ifelse(length(unique(kreft_html_with_variables$valutakode)) == 1,
-                        # Using the currency code if uniform
-                    label_dollar(prefix = str_c(kreft_html_with_variables$valutakode[1], " ")),
-                        # alternative prefix if multiple currencies in the data
-                    label_dollar(prefix = "? "))
-        ) +
-  scale_fill_met_d("Tsimshian",
-                   direction = ifelse(
-                     mean(kreft_html_with_variables$balance_positive,
-                          na.rm = TRUE) == 1,-1, 1)) +
-  theme(legend.position = "none") +
-  labs(x = "Year", y = "", title = "Balance (in thousands)",
-             # Checking again for currency; including warning if not uniform
-        subtitle = ifelse(length(unique(kreft_html_with_variables$valutakode)) == 1,
-                         "",
-                         "Note: not all values are denominated in the same currency"))
-
-```
-
-### Expenses
-```{r}
-kreft_excel |> 
+for (i in 1:length(palettes)) {
+  print(kreft_excel |> 
     transmute(
@@
     scale_y_continuous(labels = label_dollar()) +
-  scale_fill_met_d("Troy")
-
-kreft_excel |> 
-  transmute(
-    Year = as.factor(year),
-    Administrative = expenses_administrative,
-    Fundraising = expenses_fundraising,
-    `Program services` = expenses_program_services,
-    Other = expenses_other
-  )
-?as_factor
+    scale_fill_met_d(palettes[i]) +
+    labs(x = "Year", y = "", title = palettes[i]))
+}
 ```
 
-## KPI ideas
-
-### Fundraising ROI
 ```{r}
-kreft_excel |> 
-  mutate(
-    KPI_ROI = expenses_fundraising / (revenue_private_contributions + revenue_membership_fees)
+for (i in 1:length(palettes)) {
+  print(kreft_html |> 
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
   ) |> 
-  select(year, KPI_ROI)
-
-kreft_excel |> 
-          mutate(
-            KPI_fundraising =
-              str_c(format((revenue_private_contributions + revenue_membership_fees) / expenses_fundraising) * 100, digits = 3
-          ), "%"
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
   ) |> 
-  filter(year == max(year)) |> # selecting latest available year
-             select(KPI_fundraising) |> 
-        str_flatten() # turning into a single string
-
-# Numeric, for further transformation
-
-KPI_fundraising_num <- kreft_excel |> 
-          mutate(
-            KPI_fundraising_num =
-              (revenue_private_contributions +
-                 revenue_membership_fees - expenses_fundraising) /
-              expenses_fundraising * 100) |> 
-  filter(year == max(year)) |> 
-  select(KPI_fundraising_num) |> 
-  pull()
-
-# To turn into percentage
-str_c(format(KPI_fundraising_num, digits = 3), "%")
-
-KPI_fundraising_num
-
-if (KPI_fundraising_num < 100) {"danger"} else if (KPI_fundraising_num < 200) {"warning"} else {"success"}
-
-kreft_excel |> 
-  mutate(
-    KPI_admin =
-      str_c(
-        format(
-        ((revenue_private_contributions + revenue_membership_fees - expenses_fundraising) / expenses_fundraising * 100), digits = 3
-        ), "%"
-        )
-  ) |> 
-  filter(year == max(year)) |> 
-  select(KPI_admin) |> 
-  str_flatten()
-
-(kreft_excel$revenue_private_contributions + kreft_excel$revenue_membership_fees) / kreft_excel$expenses_fundraising
-
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |> 
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix =
+                                    str_c(kreft_html$valutakode[1], " "))) +
+  scale_fill_met_d(palettes[i]) +
+  labs(x = "Year", y = "Total compensation", title = palettes[i]))
+}
 ```
-
-### Administrative expenses
-```{r}
-kreft_excel |> 
-  mutate(
-    KPI_admin =
-      str_c(
-        format(
-        (expenses_administrative / expenses_program_services *100), digits = 3
-        ), "%"
-        )
-  ) |> 
-  select(year, KPI_admin)
-
-kreft_excel |> 
-  mutate(
-    KPI_admin =
-      str_c(
-        format(
-        (expenses_administrative / expenses_program_services *100), digits = 3
-        ), "%"
-        )
-  ) |> 
-  filter(year == max(year)) |> 
-  select(KPI_admin) |> 
-  str_flatten()
-
-kreft_excel |> mutate(
-            KPI_admin =
-              expenses_administrative / expenses_program_services *100) |> 
-           filter(year == max(year)) |> 
-          select(KPI_admin)
-```
-
-### Private donation growth
-```{r}
-kreft_excel |> 
-  transmute(
-    growth = revenue_private_contributions[1] / revenue_private_contributions[2] * 100
-  ) |> 
-  pull() |> 
-  first()
-
-?pull
-```
-
-### Their KPI
-
-```{r}
-# private contributions, investments, cash in bank / short-term garnts payable, revocable endowments, long-term grants payable
-
-mutate(kreft_excel,
-    KPI_OPX = (revenue_private_contributions + revenue_investments +
-                 assets_cash_in_bank_and_cash_equivalents) /
-      (liability_short_term_grants_payable +
-         liability_revocable_endowments +
-         liability_long_term_grants_payable) * 100
-  ) |> select(year, KPI_OPX)
-
-kreft_excel |> 
-  filter(year == max(year)) |> 
-      transmute(
-        KPI_OPX = (revenue_private_contributions + revenue_investments +
-                 assets_cash_in_bank_and_cash_equivalents) /
-      (liability_short_term_grants_payable +
-         liability_revocable_endowments +
-         liability_long_term_grants_payable) * 100
-      ) |> 
-      pull()
-```
-
-## More functions
-```{r}
-# mutate(
-#       KPI_admin =
-#         str_c( # Using str_c() to concatenate value and % sign
-#          format( # shortening(lengthening) the resulting value
-#            # calculating the value
-#           (expenses_administrative / expenses_program_services *100), digits = 3
-#           ), "%"
-#          )
-#       ) |> 
-#       filter(year == max(year)) |> # selecting latest available year
-#        select(KPI_admin) |> # dropping all other variables
-#         str_flatten() # turning into a single string
-
-kreft_excel2 |> 
-  filter(year == max(year)) |> 
-  transmute(
-    boxvalue = str_c(format(KPI_admin, digits = 3), "%")
-  ) |> 
-  str_flatten()
-```

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
Do coders dream of art?

@@
       currency_scale() +
-      scale_fill_met_d("Troy") +
+      scale_fill_met_d("VanGogh3") +
       labs(x = "", y = "", title = "Expenses (in thousands)")

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
Creating KPIs private donations, public grants.

--- 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
@@
 ```{r}
-valueBoxOutput("KPI_box4")
+valueBoxOutput("KPI_box_private")
 ```
@@
 
+  # Creating private donations growth
+  KPI_private_num <- reactive({
+    exceldata() |>
+      transmute(
+        growth = revenue_private_contributions[1] / revenue_private_contributions[2] * 100
+      ) |>
+      pull() |>
+      first()
+  })
+
+  # Creating public grants growth
+  KPI_public_num <- reactive({
+    exceldata() |>
+      transmute(
+        growth = revenue_public_grants[1] / revenue_public_grants[2] * 100
+      ) |>
+      pull() |>
+      first()
+  })
+
   # Creating the value box for administrative expenses
@@
       caption = "Administration as a % of total expenses",
-      color = met.brewer("Wissing", n = 1),
+      color = "primary",
       icon = "fa-random"
@@
       color = ifelse(KPI_OPX_num() <= 100, "success", "danger"),
-      icon = "fas fa-coffee"
+      icon = "fas fa-seedling"
     )
@@
 
-  # Duplicate value boxes for testing
-
-  output$KPI_box4 <- renderValueBox({
+  # Creating value box for public grants growth
+  output$KPI_box_private <- renderValueBox({
     valueBox(
-      value = KPI_to_string(KPI_admin_num()),
-      caption = "Administration as a % of total expenses",
-      color = met.brewer("Wissing", n = 1),
-      icon = "fa-random"
+      value = KPI_to_string(KPI_private_num()),
+      caption = "Private donation growth",
+      color = ifelse(KPI_private_num() < 100, "warning", "success"),
+      icon = "fas fa-hands-helping"
     )
@@
     valueBox(
-      value = KPI_to_string(KPI_admin_num()),
-      caption = "Administration as a % of total expenses",
-      color = met.brewer("Wissing", n = 1),
-      icon = "fa-random"
+      value = KPI_to_string(KPI_public_num()),
+      caption = "Public grants growth",
+      color = ifelse(KPI_public_num() < 100, "warning", "success"),
+      icon = "fas fa-hand-holding-usd"
     )
@@
       # pretty colours
-      scale_fill_met_d("Troy") +
+      scale_fill_met_d("Hokusai2", direction = -1) +
       labs(
@@
       pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
+      # Ordering for the plot
+      mutate(
+        Expenses = ordered(Expenses,
+          levels =
+            c("Program services", "Other", "Fundraising", "Administrative")
+        )
+      ) |>
       ggplot(aes(Year, value, fill = Expenses)) +
@@
       currency_scale() +
-      scale_fill_met_d("Troy") +
+      scale_fill_met_d("Hokusai2", direction = -1) +
       labs(x = "", y = "", title = "Expenses (in thousands)")
@@
 
-  # Reactive function with slightly nicer colours + ensuring positive values are always green
+  # Reactive function with slightly nicer colours + ensuring positive values are
+  # always green
   scale_fill_green <- reactive({

### 2022-07-21 Shiny/Test/Test data.Rmd
Public grants KPI, proposed private/public KPI.

@@
 
+### Public grants growth
+```{r}
+kreft_excel |> 
+  transmute(
+    growth = revenue_public_grants[1] / revenue_public_grants[2] * 100
+  ) |> 
+  pull() |> 
+  first()
+```
+
+### Private/public
+```{r}
+kreft_excel |> 
+  transmute(
+    proportion = revenue_private_contributions / revenue_public_grants * 100
+  ) |> 
+  pull() |> 
+  first()
+```
+
 ### Their KPI

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
Fixing public grants value box reference.

@@
 
-### KPI
+### KPI 1
 ```{r}
@@
 ```{r}
-valueBoxOutput("KPI_box5")
+valueBoxOutput("KPI_box_public")
 ```
@@
 
-  output$KPI_box5 <- renderValueBox({
+  output$KPI_box_public <- renderValueBox({
     valueBox(

### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Facet test. Plot and KPI fixes.

--- 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
@@
 )
-# Using fontawesome for more up-to-date fontawesome icons
+# Using fontawesome for more up-to-date (and currently supported) fontawesome icons
 htmltools::tagList(fontawesome::fa_html_dependency())
@@
 
-### KPI
-```{r}
-valueBoxOutput("KPI_box_admin")
-```
-
-### KPI 2
+### KPI Fundraising
 ```{r}
@@
 
-### KPI 3
+### KPI OPX
 ```{r}
@@
 
-### KPI 4
+### KPI Private growth
 ```{r}
@@
 
-### KPI 5
+### KPI Public growth
 ```{r}
-valueBoxOutput("KPI_box5")
+valueBoxOutput("KPI_box_public")
+```
+
+### KPI Private/public proportion
+```{r}
+valueBoxOutput("KPI_box_proportion")
 ```
@@
 plotlyOutput("expenses_plotly")
+```
+
+### Revenues
+```{r}
+plotlyOutput("revenue_plotly")
 ```
@@
 
-  # Creating KPI_admin value
-  KPI_admin_num <- reactive({
-    exceldata() |>
-      # most recent year
-      filter(year == max(year)) |>
-      # Calculating the value as a percentage; removing all other
-      transmute(
-        KPI_admin = expenses_administrative / (expenses_program_services +
-          expenses_administrative +
-          expenses_fundraising +
-          expenses_other) * 100
-      ) |>
-      # just want the single number
-      pull()
-  })
-
   # Creating Fundraising ROI value
@@
     exceldata() |>
+      # most recent year
       filter(year == max(year)) |>
+      # Calculating the value as a %
       transmute(
@@
       ) |>
-      pull()
+      # Just the single number
+      pull() |> 
+      first()
   })
@@
     exceldata() |>
+      # Calculating the two most recent years as a %
       transmute(
@@
       ) |>
+      # Pulling the first instance of the number
       pull() |>
@@
 
-  # Creating the value box for administrative expenses
-  output$KPI_box_admin <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_admin_num()),
-      caption = "Administration as a % of total expenses",
-      color = "primary",
-      icon = "fa-random"
-    )
+  # Creating public/private proportion
+  KPI_proportion <- reactive({
+    exceldata() |> 
+    transmute(
+    proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
+  ) |> 
+  pull() |> 
+  first()
   })
@@
       caption = "Fundraising ROI",
-
       # Colour depends on value
-      color = if (KPI_fundraising_num() < 100) {
+      color = if (is.na(KPI_fundraising_num())) {
+        "primary"
+      } else if (KPI_fundraising_num() < 100) {
         "danger"
@@
     valueBox(
-      value = str_c(format(KPI_OPX_num(), digits = 3), "%"),
+      value = KPI_to_string(KPI_OPX_num()),
       caption = "Liquid assets / obligations",
-      color = ifelse(KPI_OPX_num() <= 100, "success", "danger"),
+      color = if (is.na(KPI_OPX_num())) {
+        "primary"
+      } else if (KPI_OPX_num() < 100) {
+        "warning"
+      } else {
+        "success"
+      },
       icon = "fas fa-seedling"
@@
 
-  # Creating value box for public grants growth
+  # Creating value box for private donations growth
   output$KPI_box_private <- renderValueBox({
@@
       caption = "Private donation growth",
-      color = ifelse(KPI_private_num() < 100, "warning", "success"),
+      color = if (is.na(KPI_private_num())){
+        "primary"
+      } else if (KPI_private_num() < 100){
+        "warning"
+      } else {
+        "success"
+      },
       icon = "fas fa-hands-helping"
@@
 
-  output$KPI_box5 <- renderValueBox({
+  # Creating value box for public grants growth
+  output$KPI_box_public <- renderValueBox({
     valueBox(
@@
       caption = "Public grants growth",
-      color = ifelse(KPI_public_num() < 100, "warning", "success"),
+      color = if (is.na(KPI_public_num())){
+        "primary"
+      } else if (KPI_public_num() < 100){
+        "warning"
+      } else {
+        "success"
+      },
       icon = "fas fa-hand-holding-usd"
+    )
+  })
+
+  # Creating value box for private/public proportion
+  output$KPI_box_proportion <- renderValueBox({
+    valueBox(
+      value = KPI_to_string(KPI_proportion()),
+      caption = "Private donations",
+      color = "info",
+      icon = "fas fa-percentage"
     )
@@
       geom_col() +
-      # Flipping, because the Excel file only contains two years atm
-      coord_flip() +
-      currency_scale() +
-      scale_fill_met_d("Hokusai2", direction = -1) +
+      scale_y_continuous(labels = label_dollar(prefix = "NOK ")) +
+      scale_fill_met_d("Hokusai2") +
       labs(x = "", y = "", title = "Expenses (in thousands)")
@@
   output$expenses_plotly <- renderPlotly(
-    ggplotly(expenses_plot()) |>
-      layout(
-        annotations =
-          plotly_caption()
+    ggplotly(expenses_plot())
   )
+
+  #----Revenue------------------------------------------
+  revenue_plot <- reactive({
+    exceldata() |> 
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year                    = as.factor(year),
+        `Private contributions` = revenue_private_contributions,
+        `Membership fees`       = revenue_membership_fees,
+        `Public grants`         = revenue_public_grants,
+        Investments             = revenue_investments,
+        `Operational activities`= revenue_operational_activities,
+        `Commercial activities` = revenue_commercial_activities,
+        Other                   = revenue_other
+      ) |> 
+      # Pivoting to use fill
+      pivot_longer(
+        c(`Private contributions`:Other), names_to = "Revenue source"
+      ) |> 
+      ggplot(aes(Year, value, fill = `Revenue source`)) +
+      geom_col() +
+      scale_y_continuous(labels = label_dollar(prefix = "NOK ")) +
+      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
+      labs(x = "", y = "", title = "Revenue (in thousands)")
+  })
+
+  output$revenue_plotly <- renderPlotly(
+    ggplotly(revenue_plot())
   )
@@
   # layout-breaking errors, otherwise formatting
-  if (is.na(variable)){
+  if (is.na(variable)) {
     NA
-  } else if (variable == Inf){
+  } else if (variable == Inf) {
     NA

### 2022-07-22 Shiny/Test/Test data.Rmd
Facet test. Plot and KPI fixes.

@@
 
+kreft_excel |> 
+      filter(year == max(year)) |>
+      transmute(
+        # Calculating the value as a percentage; removing all other
+        KPI_fundraising_num =
+          (revenue_private_contributions +
+            revenue_membership_fees - expenses_fundraising) /
+            expenses_fundraising * 100
+      ) |>
+      # just want the single number
+      pull()
+
 ?pull
@@
   transmute(
-    proportion = revenue_private_contributions / revenue_public_grants * 100
+    proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
   ) |> 
@@
   first()
+
+kreft_excel
 ```
@@
       pull()
+kreft_excel |> 
+filter(year == max(year)) |>
+      # Calculating the value as a %
+      transmute(
+        KPI_OPX = (revenue_private_contributions + revenue_investments +
+          assets_cash_in_bank_and_cash_equivalents) /
+          (liability_short_term_grants_payable +
+            liability_revocable_endowments +
+            liability_long_term_grants_payable) * 100
+      ) |>
+      # Just the single number
+      pull()
+```
+
+### Revenue plot
+```{r}
+kreft_excel |> 
+  transmute(
+        Year                    = as.factor(year),
+        `Private contributions` = revenue_private_contributions,
+        `Membership fees`       = revenue_membership_fees,
+        `Public grants`         = revenue_public_grants,
+        Investments             = revenue_investments,
+        `Operational activities`= revenue_operational_activities,
+        `Commercial activities` = revenue_commercial_activities,
+        Other                   = revenue_other
+      ) |> 
+      # Pivoting to use fill
+      pivot_longer(
+        c(`Private contributions`:Other), names_to = "Revenue source"
+      ) |> 
+      ggplot(aes(Year, value, fill = `Revenue source`)) +
+      geom_col() +
+      # Flipping, because the Excel file only contains two years atm
+      coord_flip() +
+      scale_fill_met_d("Hokusai2", direction = -1) +
+      labs(x = "", y = "", title = "Revenue (in thousands)")
 ```

### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Updating to use observe() |> bindEvent(). Added function to create a maximum value for the revenue and expenses plots. Adding this to the y limits for these plots.
Many tests in the Facet test document.

@@
 
-### Information 1
+### Key Performance Indicators
 
@@
 
-### Information 2
+### Plots
 ```{r}
@@
 # Shiny actions
-observeEvent(input$click, { # wait for button click
+observe({ # wait for button click
 
-  #---- Data frames --------------------------------------------------------------      
+  #---- Data frames ------------------------------------------------------------      
   # creating data frame from html file
@@
 
-  #---- Value Boxes --------------------------------------------------------------
+  #---- Value Boxes ------------------------------------------------------------
 
@@
       # just want the single number
-      pull()
+      pull() |> 
+      first()
   })
@@
       geom_col() +
-      scale_y_continuous(labels = label_dollar(prefix = "NOK ")) +
+      scale_y_continuous(labels = label_dollar(prefix = "NOK "),
+                         limits = c(NA, maximum())) +
       scale_fill_met_d("Hokusai2") +
@@
       geom_col() +
-      scale_y_continuous(labels = label_dollar(prefix = "NOK ")) +
+      scale_y_continuous(labels = label_dollar(prefix = "NOK "),
+                         limits = c(NA, maximum())) +
       scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
@@
   })
+
+  # Reactive function to create the maximum value for revenue and expenses plots
+  maximum <- reactive({
+    read_excel(input$excel$datapath) |>
+      select(-Firm, -`Sub-category`) |>
+      filter(Category == "Revenue" | Category == "Expenses") |>
+      group_by(Year, Category) |>
+      summarise(
+        sum = sum(Value, na.rm = TRUE)
+      ) |>
+      pull() |>
+      max()
   })
+}) |> bindEvent(input$click)
 
@@
 }
+
+?observeEvent
 ```

### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Letting styler do its thing.

@@
       caption = "Private donation growth",
-      color = if (is.na(KPI_private_num())){
+      color = if (is.na(KPI_private_num())) {
         "primary"
-      } else if (KPI_private_num() < 100){
+      } else if (KPI_private_num() < 100) {
         "warning"
@@
       caption = "Public grants growth",
-      color = if (is.na(KPI_public_num())){
+      color = if (is.na(KPI_public_num())) {
         "primary"
-      } else if (KPI_public_num() < 100){
+      } else if (KPI_public_num() < 100) {
         "warning"
@@
       geom_col() +
-      scale_y_continuous(labels = label_dollar(prefix = "NOK "),
-                         limits = c(NA, maximum())) +
+      scale_y_continuous(
+        labels = label_dollar(prefix = "NOK "),
+        limits = c(NA, maximum())
+      ) +
       scale_fill_met_d("Hokusai2") +
@@
         Investments = revenue_investments,
-        `Operational activities`= revenue_operational_activities,
+        `Operational activities` = revenue_operational_activities,
         `Commercial activities` = revenue_commercial_activities,
@@
       pivot_longer(
-        c(`Private contributions`:Other), names_to = "Revenue source"
+        c(`Private contributions`:Other),
+        names_to = "Revenue source"
       ) |>
@@
       geom_col() +
-      scale_y_continuous(labels = label_dollar(prefix = "NOK "),
-                         limits = c(NA, maximum())) +
+      scale_y_continuous(
+        labels = label_dollar(prefix = "NOK "),
+        limits = c(NA, maximum())
+      ) +
       scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
@@
 # Function to turn KPI values into NA or percentage string
-KPI_to_string <- function(variable){
+KPI_to_string <- function(variable) {
   # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
@@
 }
-
-?observeEvent
 ```

### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Adding data tests for plots based on html data. Rewriting erroneous scale_fill_green() function. Commenting and stylering.

@@
 )
-# Using fontawesome for more up-to-date (and currently supported) fontawesome icons
+# Using fontawesome for more up-to-date (and currently supported) fontawesome
+# icons
 htmltools::tagList(fontawesome::fa_html_dependency())
-
-# Considering lumen or cerulean theme
 ```
@@
 
-### Transparency
+### Transparency: CEO compensation
 ```{r}
@@
   htmldata <- reactive({
-    read_proff8(
+    read_proff(
       input$html$datapath,
@@
   CEO_plot <- reactive({
+    # Testing for the presence of data
+    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
+      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
+      # Printing 'no information' plot if no data
+      no_information()
+    } else {
+      # Creating plot if any data exist
+      # Note that the plot will still look empty, but have a legend and scales,
+      # if all values are explicit zeroes
       htmldata() |>
@@
         geom_col() +
+        # using currency scale function to add correct currency code or ?
         currency_scale() +
@@
         )
+    }
   })
@@
     ggplotly(CEO_plot()) |>
+      ## ggplot subtitles and captions are lost when using ggplotly(), a
+      ## currency warning caption is therefore added in plotly, if needed, using
+      ## a separate function
       layout(
@@
   arsresultat_plot <- reactive({
+    # testing for the presence of data
+    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
+      # Printing 'no information' plot if no data
+      no_information()
+    } else {
       htmldata() |>
@@
         geom_col() +
+        # using currency scale function to add correct currency code or ?
         currency_scale() +
-      scale_fill_green() +
+        # using a function to ensure uniform colours irrespective of data
+        scale_fill_green(htmldata()$annual_profit) +
         theme(legend.position = "none") +
@@
     ggplotly(arsresultat_plot()) |>
+      ## ggplot subtitles and captions are lost when using ggplotly(), a
+      ## currency warning caption is therefore added in plotly, if needed, using
+      ## a separate function
       layout(
@@
   )
+
   #----Balance--------------------------------------------------------------------    
   balance_plot <- reactive({
+    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
+      no_information()
+    } else {
       htmldata() |>
@@
         geom_col() +
+        # using currency scale function to add correct currency code or ?
         currency_scale() +
-      scale_fill_green() +
+        # using a function to ensure uniform colours irrespective of data
+        scale_fill_green(htmldata()$balance_positive) +
         theme(legend.position = "none") +
@@
     ggplotly(balance_plot()) |>
+      ## ggplot subtitles and captions are lost when using ggplotly(), a
+      ## currency warning caption is therefore added in plotly, if needed, using
+      ## a separate function
       layout(
@@
   expenses_plot <- reactive({
+    # No testing with user-supplied Excel data
     exceldata() |>
@@
       scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
         labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
         limits = c(NA, maximum())
@@
       scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
         labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
         limits = c(NA, maximum())
       ) +
+      # Specifying the number of values, as the base palette only includes 6
       scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
@@
 
-  # Reactive function with slightly nicer colours + ensuring positive values are
-  # always green
-  scale_fill_green <- reactive({
-    scale_fill_met_d("Tsimshian",
-      direction = ifelse(
-        mean(htmldata()$balance_positive,
-          na.rm = TRUE
-        ) == 1, -1, 1
-      )
-    )
-  })
-
   # Reactive function to create the maximum value for revenue and expenses plots
   maximum <- reactive({
+    # reading the Excel sheet anew
     read_excel(input$excel$datapath) |>
+      # removing the firm and sub-category columns
       select(-Firm, -`Sub-category`) |>
+      # filtering to retain only revenue and expenses entries
       filter(Category == "Revenue" | Category == "Expenses") |>
+      # grouping and summarising to calculate maximum value
       group_by(Year, Category) |>
@@
       ) |>
+      # Creating list of sums
       pull() |>
+      # selecting the highest value
       max()
   })
+  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
 }) |> bindEvent(input$click)
@@
 # Function to read html file
-# read_proff8 <- function(file, start, end) {
+# read_proff <- function(file, start, end) {
 #   read_html(file, encoding = "UTF-8") |>
@@
 
-# Test version of function - multiple currencies
-read_proff8 <- function(file, start, end) {
+# Test version of function - multiple currencies and no CEO compensation
+read_proff <- function(file, start, end) {
   read_html(file, encoding = "UTF-8") |>
@@
     mutate(
-      valutakode = ifelse(year == 2019, "USD", valutakode)
+      valutakode = ifelse(year == 2019, "USD", valutakode),
+      leder_annen_godtgjorelse = NA,
+      lederlonn = NA,
+      arsresultat = NA,
+      sum_egenkapital_og_gjeld = NA
     ) |>
@@
 }
+
+# Function with slightly nicer colours than default; ensuring uniformity; +
+# ensuring positive values are always green through a test for the presence of
+# FALSE values
+scale_fill_green <- function(variable) {
+    scale_fill_met_d("Tsimshian",
+      direction = ifelse(
+        mean(variable,
+          na.rm = TRUE
+        ) == 1, -1, 1
+      )
+    )
+  }
+
+# Function to print a 'no information' plot, if necessary
+no_information <- function() {
+  ggplot() +
+    # text geom with a space to create the plot, but otherwise show nothing
+    geom_text(aes(0, 0, label = " ")) +
+    # as blank a slate as possible
+    theme_void() +
+    annotate(
+      "text",
+      label = "No information available",
+      # Using the same palette as the other non-profit/loss plots
+      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+    )
+}
 ```

### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Excel data quality tests and warnings.

@@
   KPI_fundraising_num <- reactive({
+    # testing for data quality, returning NA if bad
+    if (datatest() == TRUE) {
+      NA
+    } else {
       exceldata() |>
@@
         first()
+    }
   })
@@
   KPI_OPX_num <- reactive({
+    # testing for data quality, returning NA if bad
+    if (datatest() == TRUE) {
+      NA
+    } else {
     exceldata() |>
@@
       first()
+    }
   })
@@
   KPI_private_num <- reactive({
+    # testing for data quality, returning NA if bad
+    if (datatest() == TRUE) {
+      NA
+    } else {
     exceldata() |>
@@
   KPI_public_num <- reactive({
+    # testing for data quality, returning NA if bad
+    if (datatest() == TRUE) {
+      NA
+    } else {
     exceldata() |>
@@
   KPI_proportion <- reactive({
+    # testing for data quality, returning NA if bad
+    if (datatest() == TRUE) {
+      NA
+    } else {
       exceldata() |>
@@
   balance_plot <- reactive({
+    # testing for the presence of data
     if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
+      # Printing 'no information' plot if no data
       no_information()
@@
   expenses_plot <- reactive({
-    # No testing with user-supplied Excel data
+    # testing for data quality
+    if (datatest() == TRUE) {
+      # printing 'bad data' plot if data are bad
+      bad_data()
+    } else {
     exceldata() |>
@@
       labs(x = "", y = "", title = "Expenses (in thousands)")
+    }
   })
@@
   revenue_plot <- reactive({
+    # testing for data quality
+    if (datatest() == TRUE) {
+      # printing 'bad data' plot if data are bad
+      bad_data()
+    } else {
     exceldata() |>
@@
       labs(x = "", y = "", title = "Revenue (in thousands)")
+    }
   })
@@
   #----Reactive functions-------------------------------
+
+  # Reactive function to test for Excel data quality (basic)
+  datatest <- reactive({
+    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
+      TRUE
+    } else {
+      FALSE
+    }
+  })
+
   # Reactive function to check for currency and use currency code or ?
@@
 # Function to read html file
+read_proff <- function(file, start, end) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    select( # removing empty column
+      REGNSKAPSPERIODE:all_of(start)
+    ) |> # using all_of()
+    mutate( # removing whitespace
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    filter( # removing duplicate table - no idea why it's necessary
+      !row_number() > 176
+    ) |>
+    filter( # removing duplicate rows
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    pivot_longer( # tidying data
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |> # using all_of()
+    mutate( # changing to real NAs and turning years into numbers
+      value = na_if(value, "-"),
+      year = as.integer(year),
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    fill( # filling the currency column
+      valutakode,
+      .direction = "updown"
+    ) |>
+    filter( # removing dates and redundant currency
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate( # removing whitespace in numbers
+      value = str_replace_all(value, "\\s", ""),
+      value = as.numeric(value),
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    distinct() |> # removing any remaining duplicate rows
+    pivot_wider( # pivoting
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    clean_names() |> # tidying names
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |> # removing headings
+    rename( # making it clear what lonn refers to
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    arrange(desc(year)) # arranging by year
+}
+
+# Test version of function that creates multiple data problems
 # read_proff <- function(file, start, end) {
@@
 #     ) |>
+#     # bad test - REMOVE
+#     mutate(
+#       valutakode = ifelse(year == 2019, "USD", valutakode),
+#       leder_annen_godtgjorelse = NA,
+#       lederlonn = NA,
+#       arsresultat = NA,
+#       sum_egenkapital_og_gjeld = NA
+#     ) |>
 #     arrange(desc(year)) # arranging by year
 # }
-
-# Test version of function - multiple currencies and no CEO compensation
-read_proff <- function(file, start, end) {
-  read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame") |>
-    html_table() |>
-    select( # removing empty column
-      REGNSKAPSPERIODE:all_of(start)
-    ) |> # using all_of()
-    mutate( # removing whitespace
-      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-    ) |>
-    filter( # removing duplicate table - no idea why it's necessary
-      !row_number() > 176
-    ) |>
-    filter( # removing duplicate rows
-      !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |>
-    pivot_longer( # tidying data
-      all_of(end):all_of(start),
-      names_to = "year"
-    ) |> # using all_of()
-    mutate( # changing to real NAs and turning years into numbers
-      value = na_if(value, "-"),
-      year = as.integer(year),
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |>
-    fill( # filling the currency column
-      valutakode,
-      .direction = "updown"
-    ) |>
-    filter( # removing dates and redundant currency
-      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-        REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-    mutate( # removing whitespace in numbers
-      value = str_replace_all(value, "\\s", ""),
-      value = as.numeric(value),
-      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-    ) |>
-    distinct() |> # removing any remaining duplicate rows
-    pivot_wider( # pivoting
-      names_from = REGNSKAPSPERIODE, values_from = value
-    ) |>
-    clean_names() |> # tidying names
-    select(
-      -lederlonn_i_hele_1000,
-      -resultatregnskap_i_hele_1000,
-      -balanseregnskap_i_hele_1000
-    ) |> # removing headings
-    rename( # making it clear what lonn refers to
-      lederlonn = lonn
-    ) |>
-    # Creating new variables for plots
-    mutate(
-      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-    ) |>
-    # bad test - REMOVE
-    mutate(
-      valutakode = ifelse(year == 2019, "USD", valutakode),
-      leder_annen_godtgjorelse = NA,
-      lederlonn = NA,
-      arsresultat = NA,
-      sum_egenkapital_og_gjeld = NA
-    ) |>
-    arrange(desc(year)) # arranging by year
-}
 
@@
     ) |>
-    select(-`Sub-category`) |> # deleting Sub-category
+    select(-`Sub-category`, -Firm) |> # deleting Sub-category and Firm
     # pivoting
@@
 }
+
+# Function to print a 'bad data' plot, if necessary
+bad_data <- function() {
+  ggplot() +
+    # text geom with a space to create the plot, but otherwise show nothing
+    geom_text(aes(0, 0, label = " ")) +
+    # as blank a slate as possible
+    theme_void() +
+    annotate(
+      "text",
+      label = "Something is wrong\nwith the Excel data.",
+      # Using the same palette as the other non-profit/loss plots
+      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+    )
+}
 ```

### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Merging currency caption functions. Commenting.

@@
 
+  # Creating plotly output
   output$arsresultat_plotly <- renderPlotly(
@@
 
+  # Creating plotly output
   output$balance_plotly <- renderPlotly(
@@
 
+  # Creating plotly output
   output$expenses_plotly <- renderPlotly(
@@
 
+  # Creating plotly output
   output$revenue_plotly <- renderPlotly(
@@
   datatest <- reactive({
+    # let me know if anyone has a better idea for testing
     if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
@@
 
-  # Reactive function to create currency caption
-  currency_caption <- reactive({
-    # Checking for uniform currency code
-    if (length(unique(htmldata()$valutakode)) == 1) {
-      # Empty caption if uniform
-      ""
-      # Warning if multiple currencies
-    } else {
-      "Note: not all values are denominated in the same currency"
-    }
-  })
-
-  # Reactive function to create plotly caption
-  # (ggplot captions and subtitles are lost in the conversion)
-  # (shamelessly stolen from the Internet, with minor modifications)
+  # Reactive function to create plotly caption (ggplot captions and subtitles
+  # are lost in the conversion) (shamelessly stolen from the Internet, with
+  # minor modifications).
+  # Checks for uniform currency code and prints a warning
+  # if there are multiple currency codes.
   plotly_caption <- reactive({
     list(
-      x = 1, y = -0.1, text = currency_caption(),
+      x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
+        "", "Note: not all values are denominated in the same currency"
+      ),
       showarrow = F, xref = "paper", yref = "paper",
@@
 # Function to read html file
+# read_proff <- function(file, start, end) {
+#   read_html(file, encoding = "UTF-8") |>
+#     html_node("#inner-frame") |>
+#     html_table() |>
+#     select( # removing empty column
+#       REGNSKAPSPERIODE:all_of(start)
+#     ) |> # using all_of()
+#     mutate( # removing whitespace
+#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+#     ) |>
+#     filter( # removing duplicate table - no idea why it's necessary
+#       !row_number() > 176
+#     ) |>
+#     filter( # removing duplicate rows
+#       !grepl("Lukk", REGNSKAPSPERIODE)
+#     ) |>
+#     pivot_longer( # tidying data
+#       all_of(end):all_of(start),
+#       names_to = "year"
+#     ) |> # using all_of()
+#     mutate( # changing to real NAs and turning years into numbers
+#       value = na_if(value, "-"),
+#       year = as.integer(year),
+#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+#     ) |>
+#     fill( # filling the currency column
+#       valutakode,
+#       .direction = "updown"
+#     ) |>
+#     filter( # removing dates and redundant currency
+#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+#         REGNSKAPSPERIODE != "Valutakode"
+#     ) |>
+#     mutate( # removing whitespace in numbers
+#       value = str_replace_all(value, "\\s", ""),
+#       value = as.numeric(value),
+#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+#     ) |>
+#     distinct() |> # removing any remaining duplicate rows
+#     pivot_wider( # pivoting
+#       names_from = REGNSKAPSPERIODE, values_from = value
+#     ) |>
+#     clean_names() |> # tidying names
+#     select(
+#       -lederlonn_i_hele_1000,
+#       -resultatregnskap_i_hele_1000,
+#       -balanseregnskap_i_hele_1000
+#     ) |> # removing headings
+#     rename( # making it clear what lonn refers to
+#       lederlonn = lonn
+#     ) |>
+#     # Creating new variables for plots
+#     mutate(
+#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+#     ) |>
+#     arrange(desc(year)) # arranging by year
+# }
+
+# Test version of function that creates multiple data problems
 read_proff <- function(file, start, end) {
@@
     ) |>
+    # bad test - REMOVE
+    mutate(
+      valutakode = ifelse(year == 2019, "USD", valutakode)
+    ) |>
     arrange(desc(year)) # arranging by year
@@
 
-# Test version of function that creates multiple data problems
-# read_proff <- function(file, start, end) {
-#   read_html(file, encoding = "UTF-8") |>
-#     html_node("#inner-frame") |>
-#     html_table() |>
-#     select( # removing empty column
-#       REGNSKAPSPERIODE:all_of(start)
-#     ) |> # using all_of()
-#     mutate( # removing whitespace
-#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-#     ) |>
-#     filter( # removing duplicate table - no idea why it's necessary
-#       !row_number() > 176
-#     ) |>
-#     filter( # removing duplicate rows
-#       !grepl("Lukk", REGNSKAPSPERIODE)
-#     ) |>
-#     pivot_longer( # tidying data
-#       all_of(end):all_of(start),
-#       names_to = "year"
-#     ) |> # using all_of()
-#     mutate( # changing to real NAs and turning years into numbers
-#       value = na_if(value, "-"),
-#       year = as.integer(year),
-#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-#     ) |>
-#     fill( # filling the currency column
-#       valutakode,
-#       .direction = "updown"
-#     ) |>
-#     filter( # removing dates and redundant currency
-#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-#         REGNSKAPSPERIODE != "Valutakode"
-#     ) |>
-#     mutate( # removing whitespace in numbers
-#       value = str_replace_all(value, "\\s", ""),
-#       value = as.numeric(value),
-#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-#     ) |>
-#     distinct() |> # removing any remaining duplicate rows
-#     pivot_wider( # pivoting
-#       names_from = REGNSKAPSPERIODE, values_from = value
-#     ) |>
-#     clean_names() |> # tidying names
-#     select(
-#       -lederlonn_i_hele_1000,
-#       -resultatregnskap_i_hele_1000,
-#       -balanseregnskap_i_hele_1000
-#     ) |> # removing headings
-#     rename( # making it clear what lonn refers to
-#       lederlonn = lonn
-#     ) |>
-#     # Creating new variables for plots
-#     mutate(
-#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-#     ) |>
-#     # bad test - REMOVE
-#     mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
+# ,
 #       leder_annen_godtgjorelse = NA,
@@
 #       sum_egenkapital_og_gjeld = NA
-#     ) |>
-#     arrange(desc(year)) # arranging by year
-# }
 

### 2022-07-22 Shiny/Test/Flexdashboard rows.Rmd
Fixing errors in calculating growth rates. Fixing levels in Revenue. Switching Revenues and Expenses in dashboard. Moving functions to more readable positions.

@@
 
+### Revenues
+```{r}
+plotlyOutput("revenue_plotly")
+```
+
 ### Expenses
@@
 plotlyOutput("expenses_plotly")
-```
-
-### Revenues
-```{r}
-plotlyOutput("revenue_plotly")
 ```
@@
 
+  # Creating value box for fundraising ROI
+  # Unsure about colour levels
+  output$KPI_box_fundraising <- renderValueBox({
+    valueBox(
+      value = KPI_to_string(KPI_fundraising_num()),
+      caption = "Fundraising ROI",
+      # Colour depends on value
+      color = if (is.na(KPI_fundraising_num())) {
+        "primary"
+      } else if (KPI_fundraising_num() < 100) {
+        "danger"
+      } else if (KPI_fundraising_num() < 200) {
+        "warning"
+      } else {
+        "success"
+      },
+      icon = "fas fa-donate"
+    )
+  })
+
   # Creating OPX KPI
@@
 
+  # Creating value box for OPX KPI
+  output$KPI_box_OPX <- renderValueBox({
+    valueBox(
+      value = KPI_to_string(KPI_OPX_num()),
+      caption = "Liquid assets / obligations",
+      color = if (is.na(KPI_OPX_num())) {
+        "primary"
+      } else if (KPI_OPX_num() < 100) {
+        "warning"
+      } else {
+        "success"
+      },
+      icon = "fas fa-seedling"
+    )
+  })
+
   # Creating private donations growth
@@
         transmute(
-        growth = revenue_private_contributions[1] / revenue_private_contributions[2] * 100
+          growth = 
+            (revenue_private_contributions[1] - revenue_private_contributions[2]) /
+            revenue_private_contributions[2] * 100
         ) |>
@@
     }
+  })
+
+  # Creating value box for private donations growth
+  output$KPI_box_private <- renderValueBox({
+    valueBox(
+      value = KPI_to_string(KPI_private_num()),
+      caption = "Private donation growth",
+      color = if (is.na(KPI_private_num())) {
+        "primary"
+      } else if (KPI_private_num() < 0) {
+        "warning"
+      } else {
+        "success"
+      },
+      icon = "fas fa-hands-helping"
+    )
   })
@@
         transmute(
-        growth = revenue_public_grants[1] / revenue_public_grants[2] * 100
+          growth =
+            (revenue_public_grants[1] - revenue_public_grants[2]) / 
+            revenue_public_grants[2] * 100
         ) |>
@@
     }
+  })
+
+  # Creating value box for public grants growth
+  output$KPI_box_public <- renderValueBox({
+    valueBox(
+      value = KPI_to_string(KPI_public_num()),
+      caption = "Public grants growth",
+      color = if (is.na(KPI_public_num())) {
+        "primary"
+      } else if (KPI_public_num() < 0) {
+        "warning"
+      } else {
+        "success"
+      },
+      icon = "fas fa-hand-holding-usd"
+    )
   })
@@
 
-  # Creating value box for fundraising ROI
-  # Unsure about colour levels
-  output$KPI_box_fundraising <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_fundraising_num()),
-      caption = "Fundraising ROI",
-      # Colour depends on value
-      color = if (is.na(KPI_fundraising_num())) {
-        "primary"
-      } else if (KPI_fundraising_num() < 100) {
-        "danger"
-      } else if (KPI_fundraising_num() < 200) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-donate"
-    )
-  })
-
-  # KPI_OPX
-  output$KPI_box_OPX <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_OPX_num()),
-      caption = "Liquid assets / obligations",
-      color = if (is.na(KPI_OPX_num())) {
-        "primary"
-      } else if (KPI_OPX_num() < 100) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-seedling"
-    )
-  })
-
-  # Creating value box for private donations growth
-  output$KPI_box_private <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_private_num()),
-      caption = "Private donation growth",
-      color = if (is.na(KPI_private_num())) {
-        "primary"
-      } else if (KPI_private_num() < 100) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-hands-helping"
-    )
-  })
-
-  # Creating value box for public grants growth
-  output$KPI_box_public <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_public_num()),
-      caption = "Public grants growth",
-      color = if (is.na(KPI_public_num())) {
-        "primary"
-      } else if (KPI_public_num() < 100) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-hand-holding-usd"
-    )
-  })
-
   # Creating value box for private/public proportion
@@
   #---- Plots --------------------------------------------------------------------
-  #----CEO compensation-----------------------------------------------------------
-  CEO_plot <- reactive({
-    # Testing for the presence of data
-    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
-      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      # Creating plot if any data exist
-      # Note that the plot will still look empty, but have a legend and scales,
-      # if all values are explicit zeroes
-      htmldata() |>
-        # preparing for pivoting
-        select(
-          year, lederlonn, leder_annen_godtgjorelse
-        ) |>
-        # Compensation in ordinary numbers
-        mutate(
-          lederlonn = lederlonn * 1000,
-          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-        ) |>
-        # renaming for the purposes of the plot
-        rename(
-          Year = year,
-          Salary = lederlonn,
-          `Other compensation` = leder_annen_godtgjorelse
-        ) |>
-        # pivoting to join the two values to use fill
-        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-        ggplot(aes(Year, value, fill = Compensation)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # pretty colours
-        scale_fill_met_d("Hokusai2", direction = -1) +
-        labs(
-          x = "", y = "Total compensation", title = "CEO compensation"
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$CEO_plotly <- renderPlotly(
-    ggplotly(CEO_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
 
@@
         ) |>
+        mutate(
+          `Revenue source` =
+            ordered(`Revenue source`, levels =
+                      c("Other", "Commercial activities",
+                        "Operational activities", "Investments",
+                        "Membership fees", "Private contributions",
+                        "Public grants"))
+        ) |> 
         ggplot(aes(Year, value, fill = `Revenue source`)) +
@@
     ggplotly(revenue_plot())
+  )
+
+  #----CEO compensation---------------------------------------------------------
+  CEO_plot <- reactive({
+    # Testing for the presence of data
+    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
+      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
+      # Printing 'no information' plot if no data
+      no_information()
+    } else {
+      # Creating plot if any data exist
+      # Note that the plot will still look empty, but have a legend and scales,
+      # if all values are explicit zeroes
+      htmldata() |>
+        # preparing for pivoting
+        select(
+          year, lederlonn, leder_annen_godtgjorelse
+        ) |>
+        # Compensation in ordinary numbers
+        mutate(
+          lederlonn = lederlonn * 1000,
+          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
+        ) |>
+        # renaming for the purposes of the plot
+        rename(
+          Year = year,
+          Salary = lederlonn,
+          `Other compensation` = leder_annen_godtgjorelse
+        ) |>
+        # pivoting to join the two values to use fill
+        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
+        ggplot(aes(Year, value, fill = Compensation)) +
+        geom_col() +
+        # using currency scale function to add correct currency code or ?
+        currency_scale() +
+        # pretty colours
+        scale_fill_met_d("Hokusai2", direction = -1) +
+        labs(
+          x = "", y = "Total compensation", title = "CEO compensation"
+        )
+    }
+  })
+
+  # Creating plotly output
+  output$CEO_plotly <- renderPlotly(
+    ggplotly(CEO_plot()) |>
+      ## ggplot subtitles and captions are lost when using ggplotly(), a
+      ## currency warning caption is therefore added in plotly, if needed, using
+      ## a separate function
+      layout(
+        annotations =
+          plotly_caption()
+      )
   )

### 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
File input validation. Adding KPI equations.

@@
 
-The KPIs are based on the following calculations:
+The KPIs are based on the following calculations, using data derived from the Excel sheet:
 
-- **admin** is based on dividing - expenses_administrative / expenses_program_services *100
-- **Fundraising ROI** - (revenue_private_contributions +
-                 revenue_membership_fees - expenses_fundraising) /
-              expenses_fundraising * 100
-```{r}
-
-```
+- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
+- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
+- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
+- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
 
@@
 ### Plots
-```{r}
-
-```
 
@@
   htmldata <- reactive({
+    validate(need(tools::file_ext(input$html$datapath) == "html",
+                  "Please upload an html file."))
     read_proff(
@@
   exceldata <- reactive({
+    validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
+                  "Please upload an Excel file."))
     widen_excel(input$excel$datapath)
@@
         ) |>
+        # Ordering for the plot
         mutate(

### 2022-07-23 Shiny/Test/Colourful plots.Rmd
Code hiding as default.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 ```
 
-## R Markdown
 ```{r}
@@
 
-```{r data, echo=FALSE}
+```{r data}
 
@@
 
+## Four colours
 ```{r}
@@
 
+## Two colours
 ```{r}

### 2022-07-23 Report/KPIs and Visualisations.Rmd
Creating Report folder. Uploading a document about KPIs and visualisations (with code). Adding info about the KPIs and plots in the dashboard.

---
title: "KPIs and Visualisations"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-23'
output: 
  html_document: 
    toc: yes
    toc_depth: 5
    code_folding: show
---

```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
```

>Sannsynligvis mest oversiktlig hvis man kjører "Knit".

Laster opp dette dokumentet som informasjon om KPIer og visualiseringer -- samt muligens som inspirasjon for visualiseringer og utregninger i rapporten.

Legg merke til at innenfor Shiny så er de fleste objekter ikke objekter men funksjoner, dvs. at objektet `htmldata` her heter `htmldata()` i Shiny-koden. I tillegg er mesteparten av elementene kodet som "reaktive funksjoner", dvs. pakket inn i `reactive({})`. Den delen av koden kan fjernes i et vanlig R Markdown-dokument. De forskjellige datatestene er nok heller ikke nødvendige.

En idé om hvordan pakken [MetBrewer](https://github.com/BlakeRMills/MetBrewer), som jeg har brukt for farger, virker i praksis kan man se hvis man kjører "Knit" på dokumentet "Colourful Plots" i Shiny/Test.

Legg merke til at ingen av koden, unntatt setup-chunken kjøres når man knitter, men koden vises. (`knitr::opts_chunk$set(echo = TRUE, eval = FALSE)`)

*Si ifra hvis du/dere har spørsmål!*

## KPIs

### Key Performance Indicators

>Her har jeg brukt enkle LaTeX mattefunsksjoner for å vise utregningene.

The KPIs are based on the following calculations, using data derived from the Excel sheet:

- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
  - A basic measure of the ROI of direct fundraising expenses.
- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
  - A measure of the relationship between the organisation's liquid assets and its obligations.
- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
  - A calculation of the percentage change in private contributions based on the two most recent years.
- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
  - A calculation of the percentage change in public grants received based on the two most recent years.
- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.

### Code

#### Fundraising ROI
```{r}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (revenue_private_contributions +
            revenue_membership_fees - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

#### OPX KPI
```{r}
# Creating OPX KPI
KPI_OPX_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      # Calculating the value as a %
      transmute(
        KPI_OPX = (revenue_private_contributions + revenue_investments +
          assets_cash_in_bank_and_cash_equivalents) /
          (liability_short_term_grants_payable +
            liability_revocable_endowments +
            liability_long_term_grants_payable) * 100
      ) |>
      # Just the single number
      pull() |>
      first()
  }
})
```

#### private donations growth
```{r}
# Creating private donations growth
KPI_private_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_private_contributions[1] - revenue_private_contributions[2]) /
            revenue_private_contributions[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first()
  }
})
```

#### public grants growth
```{r}
# Creating public grants growth
KPI_public_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_public_grants[1] - revenue_public_grants[2]) /
            revenue_public_grants[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first()
  }
})
```

#### public/private proportion
```{r}
# Creating public/private proportion
KPI_proportion <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      transmute(
        proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
      ) |>
      pull() |>
      first()
  }
})
```

## Plots

The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

- **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
- **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

### Code

#### Årsresultat
```{r}
arsresultat_plot <- reactive({
  # testing for the presence of data
  if (sum(!is.na(htmldata()$arsresultat)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    htmldata() |>
      ggplot(aes(year, arsresultat, fill = annual_profit)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale() +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata()$annual_profit) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Annual result (in thousands)"
      )
  }
})

# Creating plotly output
output$arsresultat_plotly <- renderPlotly(
  ggplotly(arsresultat_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)
```

#### Balance
```{r}
#----Balance--------------------------------------------------------------------
balance_plot <- reactive({
  # testing for the presence of data
  if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    htmldata() |>
      ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale() +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata()$balance_positive) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Balance (in thousands)",
      )
  }
})

# Creating plotly output
output$balance_plotly <- renderPlotly(
  ggplotly(balance_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)
```

#### Expenses
```{r}
expenses_plot <- reactive({
  # testing for data quality
  if (datatest() == TRUE) {
    # printing 'bad data' plot if data are bad
    bad_data()
  } else {
    exceldata() |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum())
      ) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)")
  }
})

# Creating plotly output
output$expenses_plotly <- renderPlotly(
  ggplotly(expenses_plot())
)
```

#### Revenue
```{r}
revenue_plot <- reactive({
  # testing for data quality
  if (datatest() == TRUE) {
    # printing 'bad data' plot if data are bad
    bad_data()
  } else {
    exceldata() |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum())
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)")
  }
})

# Creating plotly output
output$revenue_plotly <- renderPlotly(
  ggplotly(revenue_plot())
)
```

#### CEO compensation
```{r}
#----CEO compensation---------------------------------------------------------
CEO_plot <- reactive({
  # Testing for the presence of data
  if (sum(!is.na(htmldata()$lederlonn)) == 0 &
    sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
    # Printing 'no information' plot if no data
    no_information()
  } else {
    # Creating plot if any data exist
    # Note that the plot will still look empty, but have a legend and scales,
    # if all values are explicit zeroes
    htmldata() |>
      # preparing for pivoting
      select(
        year, lederlonn, leder_annen_godtgjorelse
      ) |>
      # Compensation in ordinary numbers
      mutate(
        lederlonn = lederlonn * 1000,
        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
      ) |>
      # renaming for the purposes of the plot
      rename(
        Year = year,
        Salary = lederlonn,
        `Other compensation` = leder_annen_godtgjorelse
      ) |>
      # pivoting to join the two values to use fill
      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
      ggplot(aes(Year, value, fill = Compensation)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale() +
      # pretty colours
      scale_fill_met_d("Hokusai2", direction = -1) +
      labs(
        x = "", y = "Total compensation", title = "CEO compensation"
      )
  }
})

# Creating plotly output
output$CEO_plotly <- renderPlotly(
  ggplotly(CEO_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)
```

## Background functions

### Code

#### Reactive functions

##### datatest
```{r}
# Reactive function to test for Excel data quality (basic)
datatest <- reactive({
  # let me know if anyone has a better idea for testing
  if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
    TRUE
  } else {
    FALSE
  }
})
```

##### currency_scale
```{r}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

##### plotly_caption
```{r}
# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

##### maximum
```{r}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  # reading the Excel sheet anew
  read_excel(input$excel$datapath) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
})
```

#### Static functions

##### Creating the necessary data
```{r data}
# Function to read html file
read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    arrange(desc(year)) # arranging by year
}

# Creating data frame
htmldata <- read_proff("Data/Kreftforeningen.html", "2002", "2021")

# Function to read Excel file
widen_excel <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`, -Firm) |> # deleting Sub-category and Firm
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() # cleaning names
}

# Creating data frame
exeldata <- widen_excel("Data/OPXfinans_final.xlsx")
```

##### KPI_to_string
```{r}
# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}
```

##### scale_fill_green
```{r}
# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}
```

##### no_information
```{r}
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "No information available",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

##### bad_data
```{r}
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "Something is wrong\nwith the Excel data.",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

### 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
Creating Report folder. Uploading a document about KPIs and visualisations (with code). Adding info about the KPIs and plots in the dashboard.

@@
 - **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
+  - A basic measure of the ROI of direct fundraising expenses.
 - **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
+  - A measure of the relationship between the organisation's liquid assets and its obligations.
 - **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
+  - A calculation of the percentage change in private contributions based on the two most recent years.
 - **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+  - A calculation of the percentage change in public grants received based on the two most recent years.
 - **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
+  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.
 
@@
 ### Plots
+
+The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.
+
+- **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
+- **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
+- **Revenues**: Based on the category *Revenue* in the Excel sheet.
+- **Expenses**: Based on the category *Expenses* in the Excel sheet.
+- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
+  - A common measure of transparency, but should normally be evaluated in context.
 

### 2022-07-23 Report/Draft Dashboard section.Rmd
Uploading a draft of a possible report dashboard section. Commenting.

---
title: "Dashboard section"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-23'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

>Ca. 6 sider.

# The Dashboard

One of the outputs we have created is an example dashboard that calculates some KPIs and creates a few visualisations based on user-uploaded data from Proff.no and the Excel template we developed. The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'uner the hood' was created using Shiny.

The dashboard consists of three main parts: a sidebar, the dashboard proper, and a 'Details' tab that provides basic information about the KPIs and plots.

## The Sidebar

The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data, and an 'Analyse' button which is connected to a Shiny `observe() |> bindEvent()` structure, ensuring that most of the code is not executed until the user has finished entering data.
```{r dashboard_sidebar, eval=FALSE}
# HTML
fileInput("html", label = "Upload html file", accept = ".html")

# Year inputs
numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Button
actionButton("click", "Analyse")
```

Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded:
```{r dashboard_file_validation, eval=FALSE}
validate(need(tools::file_ext(input$html$datapath) == "html",
                  "Please upload an html file."))
validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
                  "Please upload an Excel file."))
```

If the validation is successful the files are read using the functions developed for this purpose.

## The dashboard proper

### Key Performance Indicators

The main section of the dashboard has a row-based layout, with a top row containing five 'value boxes' displaying KPIs derived from data in the Excel sheet, calculated and formatted in separate reactive functions. As Flexdashboard fails particularly gracelessly when errors occur with the value boxes, each function relies on some basic checks carried out by separate functions created for this purpose. In the case of both this is a simple function that checks if any of the numeric columns contain characters:
```{r dashboard_datatest, eval=FALSE}
datatest <- reactive({
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
```

If there are any characters, the calculation outputs `NA`, otherwise it carries out the calculation. One example:
```{r dashboard_KPI_calculation}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (revenue_private_contributions +
            revenue_membership_fees - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

The value box then also relies on a function that converts the number into a percentage string, checking into process if the number is `NA` or `Inf` (indicating a division by zero, and thus either a data problem or data not suited to this calculation).
```{r}
# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}
```

In the case of some value boxes, such as the Fundraising ROI one, the function includes logic that changes the colour of the box based on the calculated number. The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
```{r}
# Creating value box for fundraising ROI
# Unsure about colour levels
# testing for data quality, returning NA if bad
if (datatest() == TRUE) {
    NA
  } else {
    output$KPI_box_fundraising <- renderValueBox({
      valueBox(
        value = KPI_to_string(KPI_fundraising_num()),
        caption = "Fundraising ROI",
        # Colour depends on value
        color = if (is.na(KPI_fundraising_num())) {
          "primary"
        } else if (KPI_fundraising_num() < 100) {
          "danger"
        } else if (KPI_fundraising_num() < 200) {
          "warning"
        } else {
          "success"
        },
        icon = "fas fa-donate"
      )
    })
  }
```

All the boxes also include a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1.

### Plots

Below the row of KPI boxes, the dashboard contains three more rows containing five plots, three based on Proff.no data, and two based on Excel data. The html-based plots include a basic test of whether the necessary data are available, as some or all may sometimes be missing, printing a 'no information' plot using a separate function if the data is missing.
```{r}
# testing for the presence of data
if (sum(!is.na(htmldata()$arsresultat)) == 0) {
  # Printing 'no information' plot if no data
  no_information()
} else {
  ggplot() # create plot
}
```

The 'no information' plot is an essentially empty plot displaying a notice to the user:
```{r}
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "No information available",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

The top row contains two plots showing the annual result and balance data for the years covered by Proff.no. These are fairly simple bar plots, including a `fill` parameter colouring the bars depending on whether the numbers are positive or negative. They do, however, rely on two simple additional function, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, using `scales::label_dollar()`, or, otherwise, uses a question mark instead of a currency code.
```{r}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

A second function is used for the `fill` parameter in order to (1) ensure that positive numbers are always green, irrespective of the presence of negative numbers, and vice versa, and (2) to use slightly nicer colours than the default ggplot2 ones, making use of the 'Tsimshian' palette from Blake R. Mills' creative [MetBrewer](https://github.com/BlakeRMills/MetBrewer) package.

The plots are converted to plotly using `ggplotly()`. As subtitles and captions do not survive the conversion, a separate function outputs a plotly annotation if the data contains multiple currency codes:
```{r}
# Creating plotly output
output$arsresultat_plotly <- renderPlotly(
  ggplotly(arsresultat_plot()) |>
    ## ggplot subtitles and captions are lost when using ggplotly(), a
    ## currency warning caption is therefore added in plotly, if needed, using
    ## a separate function
    layout(
      annotations =
        plotly_caption()
    )
)

# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

Below these plots are two plots based on Excel data. These both use the same check for data quality as the value box, outputting a 'bad data' plot if the data are bad using a separate function:
```{r}
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "Something is wrong\nwith the Excel data.",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting - in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order. Excel data are presumed to be supplied in NOK.
```{r}
exceldata() |>
  # Renaming for the purposes of the plot + dropping other variables
  transmute(
    Year = as.factor(year),
    `Private contributions` = revenue_private_contributions,
    `Membership fees` = revenue_membership_fees,
    `Public grants` = revenue_public_grants,
    Investments = revenue_investments,
    `Operational activities` = revenue_operational_activities,
    `Commercial activities` = revenue_commercial_activities,
    Other = revenue_other
  ) |>
  # Pivoting to use fill
  pivot_longer(
    c(`Private contributions`:Other),
    names_to = "Revenue source"
  ) |>
  # Ordering for the plot
  mutate(
    `Revenue source` =
      ordered(`Revenue source`,
        levels =
          c(
            "Other", "Commercial activities",
            "Operational activities", "Investments",
            "Membership fees", "Private contributions",
            "Public grants"
          )
      )
  ) |>
  ggplot(aes(Year, value, fill = `Revenue source`)) +
  geom_col() +
  scale_y_continuous(
    # Excel data are assumed to be in NOK; if this changes, code based on
    # the currency_code() function may be adapted
    labels = label_dollar(prefix = "NOK "),
    # Basing lower limit on the data (not suppressing erroneous negative
    # numbers), and the upper limit on the maximum value in revenues and
    # expenses, to match the two plots
    limits = c(NA, maximum())
  ) +
  # Specifying the number of values, as the base palette only includes 6
  scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
  labs(x = "", y = "", title = "Revenue (in thousands)")
```

The revenue and expenses plots rely on a separate 'maximum' function to create to y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data. The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals. (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
```{r}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  # reading the Excel sheet anew
  read_excel(input$excel$datapath) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
})
```

All the plots based on the Excel data make use of the "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()`, as the base palette only contains six colours, but the plot has seven. The same function is used to generate the text colour in the 'no information' and 'bad data' plots.

Below these two plots is a plot showing the total reported CEO compensation in the years covered by the html data, where this information is available.

## Details tab

Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots.

### 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
Uploading a draft of a possible report dashboard section. Commenting.

@@
 actionButton("click", "Analyse")
-
 ```
@@
   htmldata <- reactive({
+    # validating file
     validate(need(tools::file_ext(input$html$datapath) == "html",
@@
   exceldata <- reactive({
+    # validating file
     validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
@@
   # Unsure about colour levels
+  # testing for data quality, returning NA if bad
+  if (datatest() == TRUE) {
+    NA
+  } else {
     output$KPI_box_fundraising <- renderValueBox({
@@
     })
+  }
 

### 2022-07-23 Report/Draft Dashboard section.Rmd
Minor fixes, naming chunks.

@@
 One of the outputs we have created is an example dashboard that calculates some KPIs and creates a few visualisations based on user-uploaded data from Proff.no and the Excel template we developed.
-The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'uner the hood' was created using Shiny.
+The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'under the hood' was created using Shiny.
+
+As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed:
+```{r dashboard_packages, eval=FALSE}
+if (!require("pacman")) install.packages("pacman")
+library(pacman)
+p_load(
+  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
+  flexdashboard, MetBrewer, fontawesome, htmltools
+)
+```
 
@@
 As Flexdashboard fails particularly gracelessly when errors occur with the value boxes, each function relies on some basic checks carried out by separate functions created for this purpose.
-In the case of both this is a simple function that checks if any of the numeric columns contain characters:
+In the case of both this is a simple function that checks if any of the numeric columns contain characters (in which case the data frame is probably hopelessly corrupted):
 ```{r dashboard_datatest, eval=FALSE}
@@
 
-The value box then also relies on a function that converts the number into a percentage string, checking into process if the number is `NA` or `Inf` (indicating a division by zero, and thus either a data problem or data not suited to this calculation).
-```{r}
+The value box then also relies on a function that converts the number into a percentage string, checking in the process if the number is `NA` or `Inf` (indicating a division by zero, and thus either a data problem or data not suited to this calculation).
+```{r dashboard_KPI_string, eval=FALSE}
 # Function to turn KPI values into NA or percentage string
@@
   # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
-  # layout-breaking errors, otherwise formatting
+  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
+  # avoid passing an NA to subsequent operations.)
   if (is.na(variable)) {
@@
 The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
-```{r}
+```{r dashboard_valuebox, eval=FALSE}
 # Creating value box for fundraising ROI
@@
 All the boxes also include a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1.
+```{r dashboard_fontawesome, eval=FALSE}
+# Using fontawesome for more up-to-date (and currently supported) fontawesome
+# icons
+htmltools::tagList(fontawesome::fa_html_dependency())
+```
 
@@
 The html-based plots include a basic test of whether the necessary data are available, as some or all may sometimes be missing, printing a 'no information' plot using a separate function if the data is missing.
-```{r}
+```{r html_datatest, eval=FALSE}
 # testing for the presence of data
@@
 The 'no information' plot is an essentially empty plot displaying a notice to the user:
-```{r}
+```{r dashboard_no_information, eval=FALSE}
 # Function to print a 'no information' plot, if necessary
@@
 They do, however, rely on two simple additional function, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, using `scales::label_dollar()`, or, otherwise, uses a question mark instead of a currency code.
-```{r}
+```{r dashboard_currency, eval=FALSE}
 # Reactive function to check for currency and use currency code or ?
@@
 As subtitles and captions do not survive the conversion, a separate function outputs a plotly annotation if the data contains multiple currency codes:
-```{r}
+```{r dashboard_plotly, eval=FALSE}
 # Creating plotly output
@@
 Below these plots are two plots based on Excel data. These both use the same check for data quality as the value box, outputting a 'bad data' plot if the data are bad using a separate function:
-```{r}
+```{r dashboard_excel_datatest, eval=FALSE}
 # Function to print a 'bad data' plot, if necessary
@@
 Excel data are presumed to be supplied in NOK.
-```{r}
+```{r dashboard_excelplot, eval=FALSE}
 exceldata() |>
@@
 (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
-```{r}
+```{r dashboard_maximum, eval=FALSE}
 # Reactive function to create the maximum value for revenue and expenses plots

### 2022-07-23 Report/KPIs and Visualisations.Rmd
Copying over chang. Spelling mistake.

@@
 
->Her har jeg brukt enkle LaTeX mattefunsksjoner for å vise utregningene.
+>Her har jeg brukt enkle LaTeX mattefunksjoner for å vise utregningene.
 

### 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
Comment headers. Stylering.

--- 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
@@
 
-  #---- Data frames ------------------------------------------------------------      
+  # Data frames ================================================================
+
   # creating data frame from html file
   htmldata <- reactive({
-    validate(need(tools::file_ext(input$html$datapath) == "html",
-                  "Please upload an html file."))
+    # validating file
+    validate(need(
+      tools::file_ext(input$html$datapath) == "html",
+      "Please upload an html file."
+    ))
     read_proff(
@@
   exceldata <- reactive({
-    validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
-                  "Please upload an Excel file."))
+    # validating file
+    validate(need(
+      tools::file_ext(input$excel$datapath) == "xlsx",
+      "Please upload an Excel file."
+    ))
     widen_excel(input$excel$datapath)
@@
 
-  #---- Value Boxes ------------------------------------------------------------
+  # Value Boxes ================================================================
 
+  ## Fundraising ROI------------------------------------------------------------
   # Creating Fundraising ROI value
@@
   # Unsure about colour levels
+  # testing for data quality, returning NA if bad
   output$KPI_box_fundraising <- renderValueBox({
@@
 
+  ## OPX KPI--------------------------------------------------------------------
   # Creating OPX KPI
@@
 
+  ## Private donations growth---------------------------------------------------
   # Creating private donations growth
@@
 
+  ## Public grants growth-------------------------------------------------------
   # Creating public grants growth
@@
 
+  ## Public/private proportion--------------------------------------------------
   # Creating public/private proportion
@@
 
-  #---- Plots --------------------------------------------------------------------
+  # Plots ======================================================================
 
-  #----Annual result---------------------------------------------
+  ## Annual result--------------------------------------------------------------
   arsresultat_plot <- reactive({
@@
 
-  #----Balance--------------------------------------------------------------------    
+  ## Balance--------------------------------------------------------------------
   balance_plot <- reactive({
@@
 
-  #----Expenses---------------------------------------
+  ## Expenses-------------------------------------------------------------------
   expenses_plot <- reactive({
@@
 
-  #----Revenue------------------------------------------
+  ## Revenue--------------------------------------------------------------------
   revenue_plot <- reactive({
@@
           `Revenue source` =
-            ordered(`Revenue source`, levels =
-                      c("Other", "Commercial activities",
+            ordered(`Revenue source`,
+              levels =
+                c(
+                  "Other", "Commercial activities",
                   "Operational activities", "Investments",
                   "Membership fees", "Private contributions",
-                        "Public grants"))
+                  "Public grants"
+                )
+            )
         ) |>
@@
 
-  #----CEO compensation---------------------------------------------------------
+  ## CEO compensation-----------------------------------------------------------
   CEO_plot <- reactive({
@@
 
-  #----Reactive functions-------------------------------
+  # Reactive functions =========================================================
 
+  ## Excel data test------------------------------------------------------------
   # Reactive function to test for Excel data quality (basic)
@@
 
+  ## Currency scale-------------------------------------------------------------
   # Reactive function to check for currency and use currency code or ?
@@
 
+  ## Plotly caption-------------------------------------------------------------
   # Reactive function to create plotly caption (ggplot captions and subtitles
@@
 
+  ## scale maximum--------------------------------------------------------------
   # Reactive function to create the maximum value for revenue and expenses plots
@@
 ```{r non-reactive functions, include=FALSE}
-# Function to read html file
+# Data reading functions =======================================================
+
+## Function to read html file --------------------------------------------------
 # read_proff <- function(file, start, end) {
@@
 
-# Test version of function that creates multiple data problems
+### Test version of function that creates multiple data problems----------------
 read_proff <- function(file, start, end) {
@@
 
-# Function to read Excel file
+## Function to read Excel file -------------------------------------------------
 widen_excel <- function(file) {
@@
 
+# Support functions ============================================================
+
+## KPI_to_string ---------------------------------------------------------------
 # Function to turn KPI values into NA or percentage string
@@
 
+## scale_fill_green ------------------------------------------------------------
 # Function with slightly nicer colours than default; ensuring uniformity; +
@@
 
+## no_information plot ---------------------------------------------------------
 # Function to print a 'no information' plot, if necessary
@@
 
+## bad_data plot ---------------------------------------------------------------
 # Function to print a 'bad data' plot, if necessary

### 2022-07-24 Shiny/Test/Flexdashboard rows tests.Rmd
Turns out it's necessary to split the server logic into two observe() events to deal with R's limits on variable name length. Thus a new version.

--- 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
@@
 
-```{r shiny, include=FALSE}
+```{r value_boxes, include=FALSE}
 # Shiny actions
@@
           KPI_fundraising_num =
-            (revenue_private_contributions +
-              revenue_membership_fees - expenses_fundraising) /
+            (sum(revenue_private_contributions,
+              revenue_membership_fees,
+              na.rm = TRUE
+            ) - expenses_fundraising) /
               expenses_fundraising * 100
@@
         transmute(
-          KPI_OPX = (revenue_private_contributions + revenue_investments +
-            assets_cash_in_bank_and_cash_equivalents) /
-            (liability_short_term_grants_payable +
-              liability_revocable_endowments +
-              liability_long_term_grants_payable) * 100
+          KPI_OPX =
+            sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents, na.rm = TRUE) /
+              sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable, na.rm = TRUE) * 100
         ) |>
@@
 
+  ## Excel data test------------------------------------------------------------
+  # Reactive function to test for Excel data quality (basic)
+  datatest <- reactive({
+    # let me know if anyone has a better idea for testing
+    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
+      TRUE
+    } else {
+      FALSE
+    }
+  })
+}) |> bindEvent(input$click)
+```
+
+```{r plots, include=FALSE}
+# It's necessary to split the observers into separate parts, with all necessary
+# code duplicated.
+
+observe({
+
   # Plots ======================================================================
+
+  # Data frames ================================================================
+
+  # creating data frame from html file
+  htmldata <- reactive({
+    # validating file
+    validate(need(
+      tools::file_ext(input$html$datapath) == "html",
+      "Please upload an html file."
+    ))
+    read_proff(
+      input$html$datapath,
+      as.character(input$start), as.character(input$end)
+    )
+  })
+
+  # Creating data frame from Excel file
+  exceldata <- reactive({
+    # validating file
+    validate(need(
+      tools::file_ext(input$excel$datapath) == "xlsx",
+      "Please upload an Excel file."
+    ))
+    widen_excel(input$excel$datapath)
+  })
 
@@
 
-  ## Excel data test------------------------------------------------------------
-  # Reactive function to test for Excel data quality (basic)
-  datatest <- reactive({
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
-    }
-  })
-
   ## Currency scale-------------------------------------------------------------
@@
   })
+
+  ## Excel data test------------------------------------------------------------
+  # Reactive function to test for Excel data quality (basic)
+  datatest <- reactive({
+    # let me know if anyone has a better idea for testing
+    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
+      TRUE
+    } else {
+      FALSE
+    }
+  })
   # Let the magic/disaster happen (i.e. execute when the user clicks the button)
@@
 ## Function to read html file --------------------------------------------------
+read_proff <- function(file, start, end) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    select( # removing empty column
+      REGNSKAPSPERIODE:all_of(start)
+    ) |> # using all_of()
+    mutate( # removing whitespace
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    filter( # removing duplicate table - no idea why it's necessary
+      !row_number() > 176
+    ) |>
+    filter( # removing duplicate rows
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    pivot_longer( # tidying data
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |> # using all_of()
+    mutate( # changing to real NAs and turning years into numbers
+      value = na_if(value, "-"),
+      year = as.integer(year),
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    fill( # filling the currency column
+      valutakode,
+      .direction = "updown"
+    ) |>
+    filter( # removing dates and redundant currency
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate( # removing whitespace in numbers
+      value = str_replace_all(value, "\\s", ""),
+      value = as.numeric(value),
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    distinct() |> # removing any remaining duplicate rows
+    pivot_wider( # pivoting
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    clean_names() |> # tidying names
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |> # removing headings
+    rename( # making it clear what lonn refers to
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    arrange(desc(year)) # arranging by year
+}
+
+### Test version of function that creates multiple data problems----------------
 # read_proff <- function(file, start, end) {
@@
 #     ) |>
+#     # bad test - REMOVE
+#     mutate(
+#       valutakode = ifelse(year == 2019, "USD", valutakode)
+#     ) |>
 #     arrange(desc(year)) # arranging by year
 # }
-
-### Test version of function that creates multiple data problems----------------
-read_proff <- function(file, start, end) {
-  read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame") |>
-    html_table() |>
-    select( # removing empty column
-      REGNSKAPSPERIODE:all_of(start)
-    ) |> # using all_of()
-    mutate( # removing whitespace
-      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-    ) |>
-    filter( # removing duplicate table - no idea why it's necessary
-      !row_number() > 176
-    ) |>
-    filter( # removing duplicate rows
-      !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |>
-    pivot_longer( # tidying data
-      all_of(end):all_of(start),
-      names_to = "year"
-    ) |> # using all_of()
-    mutate( # changing to real NAs and turning years into numbers
-      value = na_if(value, "-"),
-      year = as.integer(year),
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |>
-    fill( # filling the currency column
-      valutakode,
-      .direction = "updown"
-    ) |>
-    filter( # removing dates and redundant currency
-      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-        REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-    mutate( # removing whitespace in numbers
-      value = str_replace_all(value, "\\s", ""),
-      value = as.numeric(value),
-      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-    ) |>
-    distinct() |> # removing any remaining duplicate rows
-    pivot_wider( # pivoting
-      names_from = REGNSKAPSPERIODE, values_from = value
-    ) |>
-    clean_names() |> # tidying names
-    select(
-      -lederlonn_i_hele_1000,
-      -resultatregnskap_i_hele_1000,
-      -balanseregnskap_i_hele_1000
-    ) |> # removing headings
-    rename( # making it clear what lonn refers to
-      lederlonn = lonn
-    ) |>
-    # Creating new variables for plots
-    mutate(
-      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-    ) |>
-    # bad test - REMOVE
-    mutate(
-      valutakode = ifelse(year == 2019, "USD", valutakode)
-    ) |>
-    arrange(desc(year)) # arranging by year
-}
-
-# ,
+# 
 #       leder_annen_godtgjorelse = NA,
@@
   # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
-  # layout-breaking errors, otherwise formatting
+  # layout-breaking errors, otherwise formatting- (The check for NA is needed to
+  # avoid passing an NA to subsequent operations.)
   if (is.na(variable)) {
@@
 # Function to print a 'no information' plot, if necessary
+# no_information <- function() {
+#   ggplot() +
+#     # text geom with a space to create the plot, but otherwise show nothing
+#     geom_text(aes(0, 0, label = " ")) +
+#     # as blank a slate as possible
+#     theme_void() +
+#     annotate(
+#       "text",
+#       label = "No information available",
+#       # Using the same palette as the other non-profit/loss plots
+#       x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+#     )
+# }
+
+## bad_data plot ---------------------------------------------------------------
+# Function to print a 'bad data' plot, if necessary
+# bad_data <- function() {
+#   ggplot() +
+#     # text geom with a space to create the plot, but otherwise show nothing
+#     geom_text(aes(0, 0, label = " ")) +
+#     # as blank a slate as possible
+#     theme_void() +
+#     annotate(
+#       "text",
+#       label = "Something is wrong\nwith the Excel data.",
+#       # Using the same palette as the other non-profit/loss plots
+#       x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+#     )
+# }
+
+## no_information plot ---------------------------------------------------------
+# Function to print a 'no information' plot, if necessary
 no_information <- function() {
+  empty_plot("No information available")
+}
+
+## bad_data plot ---------------------------------------------------------------
+# Function to print a 'bad data' plot, if necessary
+bad_data <- function() {
+  empty_plot("Something is wrong\nwith the Excel data.")
+}
+
+empty_plot <- function(warning) {
   ggplot() +
@@
       "text",
-      label = "No information available",
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
-}
-
-## bad_data plot ---------------------------------------------------------------
-# Function to print a 'bad data' plot, if necessary
-bad_data <- function() {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    annotate(
-      "text",
-      label = "Something is wrong\nwith the Excel data.",
+      label = warning,
       # Using the same palette as the other non-profit/loss plots

### 2022-07-24 Report/Draft Dashboard section.Rmd
Spelling.

@@
 
-Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded:
+Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded:
 ```{r dashboard_file_validation, eval=FALSE}
@@
 
-Below the row of KPI boxes, the dashboard contains three more rows containing five plots, three based on Proff.no data, and two based on Excel data.
+Below the row of KPI boxes, the dashboard contains three more rows containing five plots, three based on Proff.no data and two based on Excel data.
 The html-based plots include a basic test of whether the necessary data are available, as some or all may sometimes be missing, printing a 'no information' plot using a separate function if the data is missing.
@@
 These are fairly simple bar plots, including a `fill` parameter colouring the bars depending on whether the numbers are positive or negative.
-They do, however, rely on two simple additional function, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, using `scales::label_dollar()`, or, otherwise, uses a question mark instead of a currency code.
+They do, however, rely on two simple additional functions, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, or, otherwise, uses a question mark instead of a currency code, both using `scales::label_dollar()`.
 ```{r dashboard_currency, eval=FALSE}

### 2022-07-24 Report/Draft Dashboard section.Rmd
Updating functions. Adding information about two 'observe()' functions.

@@
 
->Ca. 6 sider.
+>Ca. 6,5 sider.
 
@@
 
-The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data, and an 'Analyse' button which is connected to a Shiny `observe() |> bindEvent()` structure, ensuring that most of the code is not executed until the user has finished entering data.
+The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data, and an 'Analyse' button which is connected to two Shiny `observe() |> bindEvent()` structures, ensuring that most of the code is not executed until the user has finished entering data.
+(Though both structures await the same button click, the division was necessary to remain within R's 10,000 byte limit.)
 ```{r dashboard_sidebar, eval=FALSE}
@@
 ```{r dashboard_valuebox, eval=FALSE}
-# Creating value box for fundraising ROI
-# Unsure about colour levels
+# Creating Fundraising ROI value
+KPI_fundraising_num <- reactive({
   # testing for data quality, returning NA if bad
@@
   } else {
-    output$KPI_box_fundraising <- renderValueBox({
-      valueBox(
-        value = KPI_to_string(KPI_fundraising_num()),
-        caption = "Fundraising ROI",
-        # Colour depends on value
-        color = if (is.na(KPI_fundraising_num())) {
-          "primary"
-        } else if (KPI_fundraising_num() < 100) {
-          "danger"
-        } else if (KPI_fundraising_num() < 200) {
-          "warning"
-        } else {
-          "success"
-        },
-        icon = "fas fa-donate"
-      )
+    exceldata() |>
+      # most recent year
+      filter(year == max(year)) |>
+      transmute(
+        # Calculating the value as a percentage; removing all other
+        KPI_fundraising_num =
+          (sum(revenue_private_contributions,
+            revenue_membership_fees,
+            na.rm = TRUE
+          ) - expenses_fundraising) /
+            expenses_fundraising * 100
+      ) |>
+      # just want the single number
+      pull() |>
+      first()
+  }
 })
-  }
 ```
@@
 
-The 'no information' plot is an essentially empty plot displaying a notice to the user:
+The 'no information' plot is an essentially empty plot displaying a notice to the user, and is created through a function that passes label information into a more general blank plot function:
 ```{r dashboard_no_information, eval=FALSE}
+## no_information plot ---------------------------------------------------------
 # Function to print a 'no information' plot, if necessary
 no_information <- function() {
+  empty_plot("No information available")
+}
+
+## empty_plot-------------------------------------------------------------------
+# Function to print an empty plot with a supplied message
+empty_plot <- function(warning) {
   ggplot() +
@@
       "text",
-      label = "No information available",
+      label = warning,
       # Using the same palette as the other non-profit/loss plots
@@
 
-Below these plots are two plots based on Excel data. These both use the same check for data quality as the value box, outputting a 'bad data' plot if the data are bad using a separate function:
+Below these plots are two plots based on Excel data.
+These both use the same check for data quality as the value box, outputting a 'bad data' plot if the data are bad using a function that passes label information to the `empty_plot()` function:
 ```{r dashboard_excel_datatest, eval=FALSE}
+## bad_data plot ---------------------------------------------------------------
 # Function to print a 'bad data' plot, if necessary
 bad_data <- function() {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    annotate(
-      "text",
-      label = "Something is wrong\nwith the Excel data.",
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
+  empty_plot("Something is wrong\nwith the Excel data.")
 }

### 2022-07-24 Report/KPIs and Visualisations.Rmd
Copying over updated function.

@@
 ```{r data}
-# Function to read html file
+## Function to read html file --------------------------------------------------
 read_proff <- function(file, start, end) {
+  # Specifying encoding to deal with occasional errors
   read_html(file, encoding = "UTF-8") |>
@@
     html_table() |>
-    select( # removing empty column
+    # removing empty column, using all_of(argument)
+    select( 
       REGNSKAPSPERIODE:all_of(start)
-    ) |> # using all_of()
-    mutate( # removing whitespace
+    ) |> 
+    # removing white space
+    mutate( 
       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
     ) |>
-    filter( # removing duplicate table - no idea why it's necessary
+    # removing duplicate table sometimes created by read_html()
+    filter( 
       !row_number() > 176
     ) |>
-    filter( # removing duplicate rows
+    # removing duplicate rows
+    filter( 
       !grepl("Lukk", REGNSKAPSPERIODE)
     ) |>
-    pivot_longer( # tidying data
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer( 
       all_of(end):all_of(start),
       names_to = "year"
-    ) |> # using all_of()
-    mutate( # changing to real NAs and turning years into numbers
+    ) |> 
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate( 
+      # changing to real NAs
       value = na_if(value, "-"),
+      # turning years into numbers
       year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
     ) |>
-    fill( # filling the currency column
+    # filling the currency column
+    fill(
       valutakode,
@@
     ) |>
-    filter( # removing dates and redundant currency
+    # removing dates and redundant currency
+    filter( 
       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
@@
     ) |>
-    mutate( # removing whitespace in numbers
+    mutate( 
+      # removing white space in numbers
       value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
       value = as.numeric(value),
+      # removing values that are years
       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
     ) |>
-    distinct() |> # removing any remaining duplicate rows
-    pivot_wider( # pivoting
+    # removing any remaining duplicate rows
+    distinct() |> 
+    # pivoting to create 'tidy' variables
+    pivot_wider( 
       names_from = REGNSKAPSPERIODE, values_from = value
     ) |>
-    clean_names() |> # tidying names
+    # tidying names
+    clean_names() |> 
+    # removing headings from the table
     select(
@@
       -balanseregnskap_i_hele_1000
-    ) |> # removing headings
-    rename( # making it clear what lonn refers to
+    ) |> 
+    # making it clear what lonn refers to
+    rename( 
       lederlonn = lonn
@@
     ) |>
-    arrange(desc(year)) # arranging by year
+    # arranging by year for practical display
+    arrange(desc(year)) 
 }

### 2022-07-24 Shiny/Test/Flexdashboard rows tests.Rmd
Using sum() in private calculation. Adding NA check for matching colours in private box. Commenting. Removing unused code. Adding testing arguments section.

@@
 )
-# Using fontawesome for more up-to-date (and currently supported) fontawesome
-# icons
+# Using the fontawesome package for more up-to-date (and currently supported)
+# Fontawesome icons
 htmltools::tagList(fontawesome::fa_html_dependency())
@@
   # Unsure about colour levels
-  # testing for data quality, returning NA if bad
   output$KPI_box_fundraising <- renderValueBox({
@@
         transmute(
-          proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
+          proportion = sum(revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) / sum(revenue_public_grants, revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) * 100
         ) |>
@@
       caption = "Private donations",
-      color = "info",
+      # Adding in the same check as elsewhere - colours match if all are NA
+      color = if (is.na(KPI_proportion())) {
+        "primary"
+      } else {
+        "info"
+      },
       icon = "fas fa-percentage"
@@
 observe({
-
-  # Plots ======================================================================
 
@@
   })
+
+  # Plots ======================================================================
 
@@
 read_proff <- function(file, start, end) {
+  # Specifying encoding to deal with occasional errors
   read_html(file, encoding = "UTF-8") |>
@@
 #       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE),
 #     ) |>
@@
 #     mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode)
-#     ) |>
-#     arrange(desc(year)) # arranging by year
-# }
-# 
+#       valutakode = ifelse(year == 2019, "USD", valutakode),
 #       leder_annen_godtgjorelse = NA,
@@
 #       sum_egenkapital_og_gjeld = NA
+#     ) |>
+#     arrange(desc(year)) # arranging by year
+# }
 
@@
   read_excel(file) |>
-    unite( # uniting Category and Name to deal with multiple 'Other'
+    # uniting Category and Name to deal with multiple 'Other'
+    unite(
       entry, c(Category, Name),
@@
     ) |>
-    select(-`Sub-category`, -Firm) |> # deleting Sub-category and Firm
-    # pivoting
+    # deleting Sub-category and Firm
+    select(-`Sub-category`, -Firm) |> 
+    # pivoting to create 'tidy' variables
     pivot_wider(names_from = entry, values_from = Value) |>
-    clean_names() # cleaning names
+    # cleaning names
+    clean_names() 
 }
@@
   # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
-  # layout-breaking errors, otherwise formatting- (The check for NA is needed to
+  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
   # avoid passing an NA to subsequent operations.)
@@
 # Function to print a 'no information' plot, if necessary
-# no_information <- function() {
-#   ggplot() +
-#     # text geom with a space to create the plot, but otherwise show nothing
-#     geom_text(aes(0, 0, label = " ")) +
-#     # as blank a slate as possible
-#     theme_void() +
-#     annotate(
-#       "text",
-#       label = "No information available",
-#       # Using the same palette as the other non-profit/loss plots
-#       x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-#     )
-# }
-
-## bad_data plot ---------------------------------------------------------------
-# Function to print a 'bad data' plot, if necessary
-# bad_data <- function() {
-#   ggplot() +
-#     # text geom with a space to create the plot, but otherwise show nothing
-#     geom_text(aes(0, 0, label = " ")) +
-#     # as blank a slate as possible
-#     theme_void() +
-#     annotate(
-#       "text",
-#       label = "Something is wrong\nwith the Excel data.",
-#       # Using the same palette as the other non-profit/loss plots
-#       x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-#     )
-# }
-
-## no_information plot ---------------------------------------------------------
-# Function to print a 'no information' plot, if necessary
 no_information <- function() {
@@
 
+## empty_plot-------------------------------------------------------------------
+# Function to print an empty plot with a supplied message
 empty_plot <- function(warning) {
@@
 }
+
+# Testing arguments ============================================================
+# One or more of the elements of this mutate() operation may be inserted at the
+# end of the read_proff() function to create various data problems for some
+# quick and dirty testing
+
+# mutate(
+#       valutakode = ifelse(year == 2019, "USD", valutakode),
+#       leder_annen_godtgjorelse = NA,
+#       lederlonn = NA,
+#       arsresultat = NA,
+#       sum_egenkapital_og_gjeld = NA
+#     ) |>
 ```

### 2022-07-24 Report/KPIs and Visualisations.Rmd
Updating with code from the dashboard.

@@
         KPI_fundraising_num =
-          (revenue_private_contributions +
-            revenue_membership_fees - expenses_fundraising) /
+          (sum(revenue_private_contributions,
+            revenue_membership_fees,
+            na.rm = TRUE
+          ) - expenses_fundraising) /
             expenses_fundraising * 100
@@
       transmute(
-        KPI_OPX = (revenue_private_contributions + revenue_investments +
-          assets_cash_in_bank_and_cash_equivalents) /
-          (liability_short_term_grants_payable +
-            liability_revocable_endowments +
-            liability_long_term_grants_payable) * 100
+        KPI_OPX =
+          sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents, na.rm = TRUE) /
+            sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable, na.rm = TRUE) * 100
       ) |>
@@
       transmute(
-        proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
+        proportion = sum(revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) / sum(revenue_public_grants, revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) * 100
       ) |>
@@
 
-# Function to read Excel file
+## Function to read Excel file -------------------------------------------------
 widen_excel <- function(file) {
   read_excel(file) |>
-    unite( # uniting Category and Name to deal with multiple 'Other'
+    # uniting Category and Name to deal with multiple 'Other'
+    unite(
       entry, c(Category, Name),
@@
 ```{r}
+## no_information plot ---------------------------------------------------------
 # Function to print a 'no information' plot, if necessary
 no_information <- function() {
+  empty_plot("No information available")
+}
+```
+
+##### bad_data
+```{r}
+## bad_data plot ---------------------------------------------------------------
+# Function to print a 'bad data' plot, if necessary
+bad_data <- function() {
+  empty_plot("Something is wrong\nwith the Excel data.")
+}
+```
+
+#### empty_plot
+```{r}
+## empty_plot-------------------------------------------------------------------
+# Function to print an empty plot with a supplied message
+empty_plot <- function(warning) {
   ggplot() +
@@
 ```
-
-##### bad_data
-```{r}
-# Function to print a 'bad data' plot, if necessary
-bad_data <- function() {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    annotate(
-      "text",
-      label = "Something is wrong\nwith the Excel data.",
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
-}
-```

### 2022-07-24 Report/Draft Dashboard section.Rmd
Updating; spelling.

@@
 
-The revenue and expenses plots rely on a separate 'maximum' function to create to y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data.
+The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data.
 The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals.
@@
 
-All the plots based on the Excel data make use of the "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()`, as the base palette only contains six colours, but the plot has seven.
+All the plots based on the Excel data make use of the "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories.
 The same function is used to generate the text colour in the 'no information' and 'bad data' plots.

### 2022-07-24 Report/KPIs and Visualisations.Rmd
Adding information on visual editing, outline view, and commenting.

@@
 
->Sannsynligvis mest oversiktlig hvis man kjører "Knit".
+> Sannsynligvis mest oversiktlig hvis man kjører "Knit". (Et annet lite triks er å klikke på "Visual" øverst til venstre, eller "Outline" øverst til høyre.)
+
+<!--# I "Visual" kan man også lett finne koden for å legge til en kommentar som ikke blir del av det ferdige dokumentet. -->
 
@@
 
->Her har jeg brukt enkle LaTeX mattefunksjoner for å vise utregningene.
+> Her har jeg brukt enkle LaTeX mattefunksjoner for å vise utregningene.
 
@@
 #### Fundraising ROI
+
 ```{r}
@@
 #### OPX KPI
+
 ```{r}
@@
 #### private donations growth
+
 ```{r}
@@
 #### public grants growth
+
 ```{r}
@@
 #### public/private proportion
+
 ```{r}
@@
 #### Årsresultat
+
 ```{r}
@@
 #### Balance
+
 ```{r}
@@
 #### Expenses
+
 ```{r}
@@
 #### Revenue
+
 ```{r}
@@
 #### CEO compensation
+
 ```{r}
@@
 ##### datatest
+
 ```{r}
@@
 ##### currency_scale
+
 ```{r}
@@
 ##### plotly_caption
+
 ```{r}
@@
 ##### maximum
+
 ```{r}
@@
 ##### Creating the necessary data
+
 ```{r data}
@@
 ##### KPI_to_string
+
 ```{r}
@@
 ##### scale_fill_green
+
 ```{r}
@@
 ##### no_information
+
 ```{r}
@@
 ##### bad_data
+
 ```{r}
@@
 #### empty_plot
+
 ```{r}

### 2022-07-24 Shiny/Test/Dashboard downloads test.Rmd
Experiments with generating a downloadable report.

--- 2022-07-24 Shiny/Test/Flexdashboard rows tests.Rmd
@@
 actionButton("click", "Analyse")
+```
+
+If desired, a copy of the analysis in html format may then be downloaded.
+
+```{r sidebar_button}
+# download
+uiOutput("downloadUI")
 ```
@@
 
-- **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
-- **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
+- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
+- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
 - **Revenues**: Based on the category *Revenue* in the Excel sheet.
@@
 # Shiny actions
+
+  # Report Output ==============================================================
+
+  output$downloadUI <- renderUI( {
+  downloadButton("report", "Generate report")
+})
+
+  output$report <- downloadHandler(
+    filename = "report.html",
+    content = function(file) {
+      tempReport <- file.path(tempdir(), "report.Rmd")
+      file.copy("report.Rmd", tempReport, overwrite = TRUE)
+
+      params <- list(htmlfile = input$html$datapath,
+                     excelfile = input$excel$datapath,
+                     start = input$start,
+                     end = input$end)
+
+      rmarkdown::render(tempReport, output_file = file,
+                        params = params,
+                        envir = new.env(parent = globalenv()))
+    }
+  )
+
 observe({ # wait for button click
@@
           growth =
-            (revenue_private_contributions[1] - revenue_private_contributions[2]) /
+            (revenue_private_contributions[1] -
+              revenue_private_contributions[2]) /
               revenue_private_contributions[2] * 100
@@
         transmute(
-          proportion = sum(revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) / sum(revenue_public_grants, revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) * 100
+          proportion =
+            sum(revenue_private_contributions,
+              revenue_membership_fees,
+              na.rm = TRUE
+            ) /
+              sum(revenue_public_grants, revenue_private_contributions,
+                revenue_membership_fees,
+                na.rm = TRUE
+              ) * 100
         ) |>
@@
       htmldata() |>
-        ggplot(aes(year, arsresultat, fill = annual_profit)) +
+        # renaming for plotly hover
+        rename(
+          Year = year,
+          `Net income` = arsresultat,
+          `Net positive` = annual_profit
+        ) |> 
+        ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
         geom_col() +
@@
         labs(
-          x = "", y = "", title = "Annual result (in thousands)"
+          x = "", y = "", title = "Net income (in thousands)"
         )
@@
       htmldata() |>
-        ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
+        # rename for plotly hover
+        rename(
+          Year = year,
+          `Total equity/liabilities` = sum_egenkapital_og_gjeld,
+          `Positive equity` = balance_positive
+        ) |> 
+        ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
         geom_col() +
@@
         labs(
-          x = "", y = "", title = "Balance (in thousands)",
+          x = "", y = "", title = "Total equity/liabilities (in thousands)",
         )

### 2022-07-24 Shiny/Test/report.Rmd
Experiments with generating a downloadable report.

--- 2022-07-24 Shiny/Test/Flexdashboard rows tests.Rmd
@@
 ---
-title: "NGO analysis"
-output: 
-  flexdashboard::flex_dashboard:
-    orientation: rows
-    vertical_layout: scroll
-    theme: lumen
-runtime: shiny
+title: "Report"
+author: "Torbjørn Skinnemoen Ottersen"
+date: '2022-07-24'
+output: html_document
+params:
+  htmlfile: NA
+  excelfile: NA
+  start: NA
+  end: NA
 ---
@@
 ```{r setup, include=FALSE}
-# Loading packages using pacman to ensure they're all installed
+knitr::opts_chunk$set(echo = TRUE)
+
 if (!require("pacman")) install.packages("pacman")
@@
 
-Sidebar {.sidebar}
-=======================================================================
-
-### File upload
-
-Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.
-
-Then upload an Excel file in the set format.
-
-Click 'Analyse' to analyse the data.
-
-```{r sidebar}
-# HTML
-fileInput("html", label = "Upload html file", accept = ".html")
-
-# Year inputs
-numericInput("start", label = "Start", value = 2002)
-numericInput("end", label = "End", value = 2021)
-
-# Excel
-fileInput("excel", "Upload Excel file", accept = ".xlsx")
-
-# Button
-actionButton("click", "Analyse")
-```
-
-Dashboard
-=======================================================================
-
-Row 
------------------------------------------------------------------------
-
-### KPI Fundraising
+# Functions
 ```{r}
-valueBoxOutput("KPI_box_fundraising")
-```
-
-### KPI OPX
-```{r}
-valueBoxOutput("KPI_box_OPX")
-```
-
-### KPI Private growth
-```{r}
-valueBoxOutput("KPI_box_private")
-```
-
-### KPI Public growth
-```{r}
-valueBoxOutput("KPI_box_public")
-```
-
-### KPI Private/public proportion
-```{r}
-valueBoxOutput("KPI_box_proportion")
-```
-
-Row
------------------------------------------------------------------------
-
-### Annual result
-```{r}
-plotlyOutput("arsresultat_plotly")
-```
-
-### Balance
-```{r}
-plotlyOutput("balance_plotly")
-```
-
-Row 
------------------------------------------------------------------------
-
-### Revenues
-```{r}
-plotlyOutput("revenue_plotly")
-```
-
-### Expenses
-```{r}
-plotlyOutput("expenses_plotly")
-```
-
-Row 
------------------------------------------------------------------------
-
-### Transparency: CEO compensation
-```{r}
-plotlyOutput("CEO_plotly")
-```
-
-Details
-=======================================================================
-
-Row 
------------------------------------------------------------------------
-
-### Key Performance Indicators
-
-The KPIs are based on the following calculations, using data derived from the Excel sheet:
-
-- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
-  - A basic measure of the ROI of direct fundraising expenses.
-- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
-  - A measure of the relationship between the organisation's liquid assets and its obligations.
-- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
-  - A calculation of the percentage change in private contributions based on the two most recent years.
-- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
-  - A calculation of the percentage change in public grants received based on the two most recent years.
-- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
-  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.
-
-Row 
------------------------------------------------------------------------
-
-### Plots
-
-The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.
-
-- **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
-- **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
-- **Revenues**: Based on the category *Revenue* in the Excel sheet.
-- **Expenses**: Based on the category *Expenses* in the Excel sheet.
-- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
-  - A common measure of transparency, but should normally be evaluated in context.
-
-```{r value_boxes, include=FALSE}
-# Shiny actions
-observe({ # wait for button click
-
-  # Data frames ================================================================
-
-  # creating data frame from html file
-  htmldata <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$html$datapath) == "html",
-      "Please upload an html file."
-    ))
-    read_proff(
-      input$html$datapath,
-      as.character(input$start), as.character(input$end)
-    )
-  })
-
-  # Creating data frame from Excel file
-  exceldata <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
-    widen_excel(input$excel$datapath)
-  })
-
-  # Value Boxes ================================================================
-
-  ## Fundraising ROI------------------------------------------------------------
-  # Creating Fundraising ROI value
-  KPI_fundraising_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # most recent year
-        filter(year == max(year)) |>
-        transmute(
-          # Calculating the value as a percentage; removing all other
-          KPI_fundraising_num =
-            (sum(revenue_private_contributions,
-              revenue_membership_fees,
-              na.rm = TRUE
-            ) - expenses_fundraising) /
-              expenses_fundraising * 100
-        ) |>
-        # just want the single number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for fundraising ROI
-  # Unsure about colour levels
-  output$KPI_box_fundraising <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_fundraising_num()),
-      caption = "Fundraising ROI",
-      # Colour depends on value
-      color = if (is.na(KPI_fundraising_num())) {
-        "primary"
-      } else if (KPI_fundraising_num() < 100) {
-        "danger"
-      } else if (KPI_fundraising_num() < 200) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-donate"
-    )
-  })
-
-  ## OPX KPI--------------------------------------------------------------------
-  # Creating OPX KPI
-  KPI_OPX_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # most recent year
-        filter(year == max(year)) |>
-        # Calculating the value as a %
-        transmute(
-          KPI_OPX =
-            sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents, na.rm = TRUE) /
-              sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable, na.rm = TRUE) * 100
-        ) |>
-        # Just the single number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for OPX KPI
-  output$KPI_box_OPX <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_OPX_num()),
-      caption = "Liquid assets / obligations",
-      color = if (is.na(KPI_OPX_num())) {
-        "primary"
-      } else if (KPI_OPX_num() < 100) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-seedling"
-    )
-  })
-
-  ## Private donations growth---------------------------------------------------
-  # Creating private donations growth
-  KPI_private_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # Calculating the two most recent years as a %
-        transmute(
-          growth =
-            (revenue_private_contributions[1] - revenue_private_contributions[2]) /
-              revenue_private_contributions[2] * 100
-        ) |>
-        # Pulling the first instance of the number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for private donations growth
-  output$KPI_box_private <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_private_num()),
-      caption = "Private donation growth",
-      color = if (is.na(KPI_private_num())) {
-        "primary"
-      } else if (KPI_private_num() < 0) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-hands-helping"
-    )
-  })
-
-  ## Public grants growth-------------------------------------------------------
-  # Creating public grants growth
-  KPI_public_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # Calculating the two most recent years as a %
-        transmute(
-          growth =
-            (revenue_public_grants[1] - revenue_public_grants[2]) /
-              revenue_public_grants[2] * 100
-        ) |>
-        # Pulling the first instance of the number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for public grants growth
-  output$KPI_box_public <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_public_num()),
-      caption = "Public grants growth",
-      color = if (is.na(KPI_public_num())) {
-        "primary"
-      } else if (KPI_public_num() < 0) {
-        "warning"
-      } else {
-        "success"
-      },
-      icon = "fas fa-hand-holding-usd"
-    )
-  })
-
-  ## Public/private proportion--------------------------------------------------
-  # Creating public/private proportion
-  KPI_proportion <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        transmute(
-          proportion = sum(revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) / sum(revenue_public_grants, revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) * 100
-        ) |>
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for private/public proportion
-  output$KPI_box_proportion <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_proportion()),
-      caption = "Private donations",
-      # Adding in the same check as elsewhere - colours match if all are NA
-      color = if (is.na(KPI_proportion())) {
-        "primary"
-      } else {
-        "info"
-      },
-      icon = "fas fa-percentage"
-    )
-  })
-
-  ## Excel data test------------------------------------------------------------
-  # Reactive function to test for Excel data quality (basic)
-  datatest <- reactive({
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
-    }
-  })
-}) |> bindEvent(input$click)
-```
-
-```{r plots, include=FALSE}
-# It's necessary to split the observers into separate parts, with all necessary
-# code duplicated.
-
-observe({
-
-  # Data frames ================================================================
-
-  # creating data frame from html file
-  htmldata <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$html$datapath) == "html",
-      "Please upload an html file."
-    ))
-    read_proff(
-      input$html$datapath,
-      as.character(input$start), as.character(input$end)
-    )
-  })
-
-  # Creating data frame from Excel file
-  exceldata <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
-    widen_excel(input$excel$datapath)
-  })
-
-  # Plots ======================================================================
-
-  ## Annual result--------------------------------------------------------------
-  arsresultat_plot <- reactive({
-    # testing for the presence of data
-    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      htmldata() |>
-        ggplot(aes(year, arsresultat, fill = annual_profit)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # using a function to ensure uniform colours irrespective of data
-        scale_fill_green(htmldata()$annual_profit) +
-        theme(legend.position = "none") +
-        labs(
-          x = "", y = "", title = "Annual result (in thousands)"
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$arsresultat_plotly <- renderPlotly(
-    ggplotly(arsresultat_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  ## Balance--------------------------------------------------------------------
-  balance_plot <- reactive({
-    # testing for the presence of data
-    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      htmldata() |>
-        ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # using a function to ensure uniform colours irrespective of data
-        scale_fill_green(htmldata()$balance_positive) +
-        theme(legend.position = "none") +
-        labs(
-          x = "", y = "", title = "Balance (in thousands)",
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$balance_plotly <- renderPlotly(
-    ggplotly(balance_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  ## Expenses-------------------------------------------------------------------
-  expenses_plot <- reactive({
-    # testing for data quality
-    if (datatest() == TRUE) {
-      # printing 'bad data' plot if data are bad
-      bad_data()
-    } else {
-      exceldata() |>
-        # Renaming for the purposes of the plot + dropping other variables
-        transmute(
-          Year = as.factor(year),
-          Other = expenses_other,
-          Administrative = expenses_administrative,
-          Fundraising = expenses_fundraising,
-          `Program services` = expenses_program_services
-        ) |>
-        # Pivoting to use fill
-        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
-        # Ordering for the plot
-        mutate(
-          Expenses = ordered(Expenses,
-            levels =
-              c("Program services", "Other", "Fundraising", "Administrative")
-          )
-        ) |>
-        ggplot(aes(Year, value, fill = Expenses)) +
-        geom_col() +
-        scale_y_continuous(
-          # Excel data are assumed to be in NOK; if this changes, code based on
-          # the currency_code() function may be adapted
-          labels = label_dollar(prefix = "NOK "),
-          # Basing lower limit on the data (not suppressing erroneous negative
-          # numbers), and the upper limit on the maximum value in revenues and
-          # expenses, to match the two plots
-          limits = c(NA, maximum())
-        ) +
-        scale_fill_met_d("Hokusai2") +
-        labs(x = "", y = "", title = "Expenses (in thousands)")
-    }
-  })
-
-  # Creating plotly output
-  output$expenses_plotly <- renderPlotly(
-    ggplotly(expenses_plot())
-  )
-
-  ## Revenue--------------------------------------------------------------------
-  revenue_plot <- reactive({
-    # testing for data quality
-    if (datatest() == TRUE) {
-      # printing 'bad data' plot if data are bad
-      bad_data()
-    } else {
-      exceldata() |>
-        # Renaming for the purposes of the plot + dropping other variables
-        transmute(
-          Year = as.factor(year),
-          `Private contributions` = revenue_private_contributions,
-          `Membership fees` = revenue_membership_fees,
-          `Public grants` = revenue_public_grants,
-          Investments = revenue_investments,
-          `Operational activities` = revenue_operational_activities,
-          `Commercial activities` = revenue_commercial_activities,
-          Other = revenue_other
-        ) |>
-        # Pivoting to use fill
-        pivot_longer(
-          c(`Private contributions`:Other),
-          names_to = "Revenue source"
-        ) |>
-        # Ordering for the plot
-        mutate(
-          `Revenue source` =
-            ordered(`Revenue source`,
-              levels =
-                c(
-                  "Other", "Commercial activities",
-                  "Operational activities", "Investments",
-                  "Membership fees", "Private contributions",
-                  "Public grants"
-                )
-            )
-        ) |>
-        ggplot(aes(Year, value, fill = `Revenue source`)) +
-        geom_col() +
-        scale_y_continuous(
-          # Excel data are assumed to be in NOK; if this changes, code based on
-          # the currency_code() function may be adapted
-          labels = label_dollar(prefix = "NOK "),
-          # Basing lower limit on the data (not suppressing erroneous negative
-          # numbers), and the upper limit on the maximum value in revenues and
-          # expenses, to match the two plots
-          limits = c(NA, maximum())
-        ) +
-        # Specifying the number of values, as the base palette only includes 6
-        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
-        labs(x = "", y = "", title = "Revenue (in thousands)")
-    }
-  })
-
-  # Creating plotly output
-  output$revenue_plotly <- renderPlotly(
-    ggplotly(revenue_plot())
-  )
-
-  ## CEO compensation-----------------------------------------------------------
-  CEO_plot <- reactive({
-    # Testing for the presence of data
-    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
-      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      # Creating plot if any data exist
-      # Note that the plot will still look empty, but have a legend and scales,
-      # if all values are explicit zeroes
-      htmldata() |>
-        # preparing for pivoting
-        select(
-          year, lederlonn, leder_annen_godtgjorelse
-        ) |>
-        # Compensation in ordinary numbers
-        mutate(
-          lederlonn = lederlonn * 1000,
-          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-        ) |>
-        # renaming for the purposes of the plot
-        rename(
-          Year = year,
-          Salary = lederlonn,
-          `Other compensation` = leder_annen_godtgjorelse
-        ) |>
-        # pivoting to join the two values to use fill
-        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-        ggplot(aes(Year, value, fill = Compensation)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # pretty colours
-        scale_fill_met_d("Hokusai2", direction = -1) +
-        labs(
-          x = "", y = "Total compensation", title = "CEO compensation"
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$CEO_plotly <- renderPlotly(
-    ggplotly(CEO_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  # Reactive functions =========================================================
-
-  ## Currency scale-------------------------------------------------------------
-  # Reactive function to check for currency and use currency code or ?
-  currency_scale <- reactive({
-    scale_y_continuous(
-      # Checking for uniform currency code
-      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
-        # Using the currency code if uniform
-        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
-        # alternative prefix if multiple currencies in the data
-        label_dollar(prefix = "? ")
-      )
-    )
-  })
-
-  ## Plotly caption-------------------------------------------------------------
-  # Reactive function to create plotly caption (ggplot captions and subtitles
-  # are lost in the conversion) (shamelessly stolen from the Internet, with
-  # minor modifications).
-  # Checks for uniform currency code and prints a warning
-  # if there are multiple currency codes.
-  plotly_caption <- reactive({
-    list(
-      x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
-        "", "Note: not all values are denominated in the same currency"
-      ),
-      showarrow = F, xref = "paper", yref = "paper",
-      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
-      font = list(size = 11)
-    )
-  })
-
-  ## scale maximum--------------------------------------------------------------
-  # Reactive function to create the maximum value for revenue and expenses plots
-  maximum <- reactive({
-    # reading the Excel sheet anew
-    read_excel(input$excel$datapath) |>
-      # removing the firm and sub-category columns
-      select(-Firm, -`Sub-category`) |>
-      # filtering to retain only revenue and expenses entries
-      filter(Category == "Revenue" | Category == "Expenses") |>
-      # grouping and summarising to calculate maximum value
-      group_by(Year, Category) |>
-      summarise(
-        sum = sum(Value, na.rm = TRUE)
-      ) |>
-      # Creating list of sums
-      pull() |>
-      # selecting the highest value
-      max()
-  })
-
-  ## Excel data test------------------------------------------------------------
-  # Reactive function to test for Excel data quality (basic)
-  datatest <- reactive({
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
-    }
-  })
-  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
-}) |> bindEvent(input$click)
-
-```
-
-```{r non-reactive functions, include=FALSE}
 # Data reading functions =======================================================
@@
 
-### Test version of function that creates multiple data problems----------------
-# read_proff <- function(file, start, end) {
-#   read_html(file, encoding = "UTF-8") |>
-#     html_node("#inner-frame") |>
-#     html_table() |>
-#     select( # removing empty column
-#       REGNSKAPSPERIODE:all_of(start)
-#     ) |> # using all_of()
-#     mutate( # removing whitespace
-#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-#     ) |>
-#     filter( # removing duplicate table - no idea why it's necessary
-#       !row_number() > 176
-#     ) |>
-#     filter( # removing duplicate rows
-#       !grepl("Lukk", REGNSKAPSPERIODE)
-#     ) |>
-#     pivot_longer( # tidying data
-#       all_of(end):all_of(start),
-#       names_to = "year"
-#     ) |> # using all_of()
-#     mutate( # changing to real NAs and turning years into numbers
-#       value = na_if(value, "-"),
-#       year = as.integer(year),
-#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-#     ) |>
-#     fill( # filling the currency column
-#       valutakode,
-#       .direction = "updown"
-#     ) |>
-#     filter( # removing dates and redundant currency
-#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-#         REGNSKAPSPERIODE != "Valutakode"
-#     ) |>
-#     mutate( # removing whitespace in numbers
-#       value = str_replace_all(value, "\\s", ""),
-#       value = as.numeric(value),
-#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-#     ) |>
-#     distinct() |> # removing any remaining duplicate rows
-#     pivot_wider( # pivoting
-#       names_from = REGNSKAPSPERIODE, values_from = value
-#     ) |>
-#     clean_names() |> # tidying names
-#     select(
-#       -lederlonn_i_hele_1000,
-#       -resultatregnskap_i_hele_1000,
-#       -balanseregnskap_i_hele_1000
-#     ) |> # removing headings
-#     rename( # making it clear what lonn refers to
-#       lederlonn = lonn
-#     ) |>
-#     # Creating new variables for plots
-#     mutate(
-#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE),
-#     ) |>
-#     # bad test - REMOVE
-#     mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
-#       leder_annen_godtgjorelse = NA,
-#       lederlonn = NA,
-#       arsresultat = NA,
-#       sum_egenkapital_og_gjeld = NA
-#     ) |>
-#     arrange(desc(year)) # arranging by year
-# }
-
 ## Function to read Excel file -------------------------------------------------
@@
     clean_names()
+}
+
+htmldata <- read_proff(
+      params$htmlfile,
+      as.character(params$start), as.character(params$end)
+    )
+
+exceldata <- widen_excel(params$excelfile)
+
+  ## Excel data test------------------------------------------------------------
+  # Reactive function to test for Excel data quality (basic)
+datatest <- 
+    # let me know if anyone has a better idea for testing
+    if (mean(unlist(lapply(exceldata, is.character))) != 0) {
+      TRUE
+    } else {
+      FALSE
     }
@@
 
-# Testing arguments ============================================================
-# One or more of the elements of this mutate() operation may be inserted at the
-# end of the read_proff() function to create various data problems for some
-# quick and dirty testing
+## Fundraising ROI------------------------------------------------------------
+  # Creating Fundraising ROI value
+    # testing for data quality, returning NA if bad
+fundraising <- if (datatest == TRUE) {
+      NA
+    } else {
+      exceldata |>
+        # most recent year
+        filter(year == max(year)) |>
+        transmute(
+          # Calculating the value as a percentage; removing all other
+          KPI_fundraising_num =
+            (sum(revenue_private_contributions,
+              revenue_membership_fees,
+              na.rm = TRUE
+            ) - expenses_fundraising) /
+              expenses_fundraising * 100
+        ) |>
+        # just want the single number
+        pull() |>
+        first()
+    } |> KPI_to_string()
 
-# mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
-#       leder_annen_godtgjorelse = NA,
-#       lederlonn = NA,
-#       arsresultat = NA,
-#       sum_egenkapital_og_gjeld = NA
-#     ) |>
 ```
+
+```
+
+```{r data, include=FALSE}
+
+```
+
+# Key Performance Indicators
+
+The KPIs are based on the following calculations, using data derived from the Excel sheet:
+
+- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
+  - A basic measure of the ROI of direct fundraising expenses.
+- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
+  - A measure of the relationship between the organisation's liquid assets and its obligations.
+- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
+  - A calculation of the percentage change in private contributions based on the two most recent years.
+- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+  - A calculation of the percentage change in public grants received based on the two most recent years.
+- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
+  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.
+
+### Fundraising ROI: `r fundraising`

### 2022-07-24 Report/KPIs and Visualisations.Rmd
Updating translations.

@@
         KPI_OPX =
-          sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents, na.rm = TRUE) /
-            sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable, na.rm = TRUE) * 100
+          sum(revenue_private_contributions,
+            revenue_investments, assets_cash_in_bank_and_cash_equivalents,
+            na.rm = TRUE
+          ) /
+            sum(liability_short_term_grants_payable,
+              liability_revocable_endowments,
+              liability_long_term_grants_payable,
+              na.rm = TRUE
+            ) * 100
       ) |>
@@
         growth =
-          (revenue_private_contributions[1] - revenue_private_contributions[2]) /
+          (revenue_private_contributions[1] -
+             revenue_private_contributions[2]) /
             revenue_private_contributions[2] * 100
@@
       transmute(
-        proportion = sum(revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) / sum(revenue_public_grants, revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) * 100
+        proportion =
+          sum(revenue_private_contributions,
+            revenue_membership_fees,
+            na.rm = TRUE
+          ) /
+            sum(revenue_public_grants, revenue_private_contributions,
+              revenue_membership_fees,
+              na.rm = TRUE
+            ) * 100
       ) |>
@@
 
--   **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
--   **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
+- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
+- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
 - **Revenues**: Based on the category *Revenue* in the Excel sheet.
@@
     htmldata() |>
-      ggplot(aes(year, arsresultat, fill = annual_profit)) +
+      # renaming for plotly hover
+      rename(
+        Year = year,
+        `Net income` = arsresultat,
+        `Net positive` = annual_profit
+      ) |>
+      ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
       geom_col() +
@@
       labs(
-        x = "", y = "", title = "Annual result (in thousands)"
+        x = "", y = "", title = "Net income (in thousands)"
       )
@@
 ```{r}
-#----Balance--------------------------------------------------------------------
 balance_plot <- reactive({
@@
     htmldata() |>
-      ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
+      # rename for plotly hover
+      rename(
+        Year = year,
+        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
+        `Positive equity` = balance_positive
+      ) |>
+      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
       geom_col() +
@@
       labs(
-        x = "", y = "", title = "Balance (in thousands)",
+        x = "", y = "", title = "Total equity/liabilities (in thousands)",
       )

### 2022-07-24 Dashboard/NGO Analysis dashboard alpha.Rmd
KPI_OPX changing line breaks.

--- 2022-07-24 Shiny/Test/Dashboard downloads test.Rmd
@@
 actionButton("click", "Analyse")
-```
-
-If desired, a copy of the analysis in html format may then be downloaded.
-
-```{r sidebar_button}
-# download
-uiOutput("downloadUI")
 ```
@@
 # Shiny actions
-
-  # Report Output ==============================================================
-
-  output$downloadUI <- renderUI( {
-  downloadButton("report", "Generate report")
-})
-
-  output$report <- downloadHandler(
-    filename = "report.html",
-    content = function(file) {
-      tempReport <- file.path(tempdir(), "report.Rmd")
-      file.copy("report.Rmd", tempReport, overwrite = TRUE)
-
-      params <- list(htmlfile = input$html$datapath,
-                     excelfile = input$excel$datapath,
-                     start = input$start,
-                     end = input$end)
-
-      rmarkdown::render(tempReport, output_file = file,
-                        params = params,
-                        envir = new.env(parent = globalenv()))
-    }
-  )
-
 observe({ # wait for button click
@@
           KPI_OPX =
-            sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents, na.rm = TRUE) /
-              sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable, na.rm = TRUE) * 100
+            sum(revenue_private_contributions,
+                revenue_investments, assets_cash_in_bank_and_cash_equivalents,
+                na.rm = TRUE) /
+              sum(liability_short_term_grants_payable,
+                  liability_revocable_endowments,
+                  liability_long_term_grants_payable, na.rm = TRUE) * 100
         ) |>

### 2022-07-24 Shiny/Test/Dashboard downloads test.Rmd
Including link to Excel template.

@@
 
-Then upload an Excel file in the set format.
+Then upload an Excel file in the set format (available [here](https://github.com/amaliehc/ISSSV1337/blob/5c0a8a2beae499b3d2524a3965ceebf3b3ebad0e/Dashboard/Data/NGO_template.xltx)).
 
@@
 
-  output$downloadUI <- renderUI( {
+output$downloadUI <- renderUI({
   downloadButton("report", "Generate report")
@@
 
-      params <- list(htmlfile = input$html$datapath,
+    params <- list(
+      htmlfile = input$html$datapath,
       excelfile = input$excel$datapath,
       start = input$start,
-                     end = input$end)
-
-      rmarkdown::render(tempReport, output_file = file,
+      end = input$end
+    )
+
+    rmarkdown::render(tempReport,
+      output_file = file,
       params = params,
-                        envir = new.env(parent = globalenv()))
+      envir = new.env(parent = globalenv())
+    )
   }

### 2022-07-24 Shiny/Test/Dashboard downloads test.Rmd
Testing automatic report generation.

@@
 
-If desired, a copy of the analysis in html format may then be downloaded.
+If desired, a copy of the analysis in PDF format may then be downloaded.
+
+Enter the name of the NGO in the text box to include it in the report.
 
 ```{r sidebar_button}
+textInput("name", "NGO name", value = "NGO name")
 # download
@@
 output$downloadUI <- renderUI({
-  downloadButton("report", "Generate report")
+  downloadButton("report", "Download report")
 })
@@
 output$report <- downloadHandler(
-  filename = "report.html",
+  filename = "report.pdf",
   content = function(file) {
@@
       start = input$start,
-      end = input$end
+      end = input$end,
+      title = input$name
     )

### 2022-07-24 Shiny/Test/report.Rmd
Testing automatic report generation.

@@
 ---
-title: "Report"
-author: "Torbjørn Skinnemoen Ottersen"
-date: '2022-07-24'
-output: html_document
+title: "NGO Analysis: `r params$title`"
+date: "`r Sys.Date()`"
+output: pdf_document
 params:
@@
   end: NA
+  title: NA
 ---
@@
 )
-# Using the fontawesome package for more up-to-date (and currently supported)
-# Fontawesome icons
-htmltools::tagList(fontawesome::fa_html_dependency())
 ```
 
-# Functions
-```{r}
+```{r data, echo=FALSE, message=FALSE, warning=FALSE}
 # Data reading functions =======================================================
@@
 exceldata <- widen_excel(params$excelfile)
+```
 
-  ## Excel data test------------------------------------------------------------
-  # Reactive function to test for Excel data quality (basic)
-datatest <- 
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata, is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
+```{r static_functions, echo=FALSE, message=FALSE, warning=FALSE}
+## empty_plot-------------------------------------------------------------------
+# Function to print an empty plot with a supplied message
+empty_plot <- function(warning) {
+  ggplot() +
+    # text geom with a space to create the plot, but otherwise show nothing
+    geom_text(aes(0, 0, label = " ")) +
+    # as blank a slate as possible
+    theme_void() +
+    annotate(
+      "text",
+      label = warning,
+      # Using the same palette as the other non-profit/loss plots
+      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+    )
 }
 
-# Support functions ============================================================
+## no_information plot ---------------------------------------------------------
+# Function to print a 'no information' plot, if necessary
+no_information <- function() {
+  empty_plot("No information available")
+}
+
+## bad_data plot ---------------------------------------------------------------
+# Function to print a 'bad data' plot, if necessary
+bad_data <- function() {
+  empty_plot("Something is wrong\nwith the Excel data.")
+}
 
@@
 }
+```
 
-## no_information plot ---------------------------------------------------------
-# Function to print a 'no information' plot, if necessary
-no_information <- function() {
-  empty_plot("No information available")
+```{r functions_reliant, echo=FALSE, message=FALSE, warning=FALSE}
+## scale maximum--------------------------------------------------------------
+# Reactive function to create the maximum value for revenue and expenses plots
+maximum <- 
+    # reading the Excel sheet anew
+    read_excel(params$excelfile) |>
+      # removing the firm and sub-category columns
+      select(-Firm, -`Sub-category`) |>
+      # filtering to retain only revenue and expenses entries
+      filter(Category == "Revenue" | Category == "Expenses") |>
+      # grouping and summarising to calculate maximum value
+      group_by(Year, Category) |>
+      summarise(
+        sum = sum(Value, na.rm = TRUE)
+      ) |>
+      # Creating list of sums
+      pull() |>
+      # selecting the highest value
+      max()
+
+## Excel data test------------------------------------------------------------
+# Reactive function to test for Excel data quality (basic)
+datatest <-
+  # let me know if anyone has a better idea for testing
+  if (mean(unlist(lapply(exceldata, is.character))) != 0) {
+    TRUE
+  } else {
+    FALSE
   }
 
-## bad_data plot ---------------------------------------------------------------
-# Function to print a 'bad data' plot, if necessary
-bad_data <- function() {
-  empty_plot("Something is wrong\nwith the Excel data.")
-}
+## Currency scale-------------------------------------------------------------
+# Reactive function to check for currency and use currency code or ?
+currency_scale <-
+  scale_y_continuous(
+    # Checking for uniform currency code
+    labels = ifelse(length(unique(htmldata$valutakode)) == 1,
+      # Using the currency code if uniform
+      label_dollar(prefix = str_c(htmldata$valutakode[1], " ")),
+      # alternative prefix if multiple currencies in the data
+      label_dollar(prefix = "? ")
+    )
+  )
 
-## empty_plot-------------------------------------------------------------------
-# Function to print an empty plot with a supplied message
-empty_plot <- function(warning) {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    annotate(
-      "text",
-      label = warning,
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+## Plotly caption-------------------------------------------------------------
+# Reactive function to create plotly caption (ggplot captions and subtitles
+# are lost in the conversion) (shamelessly stolen from the Internet, with
+# minor modifications).
+# Checks for uniform currency code and prints a warning
+# if there are multiple currency codes.
+# plotly_caption <-
+#   list(
+#     x = 1, y = -0.1, text = ifelse(length(unique(htmldata$valutakode)) == 1,
+#       "", "Note: not all values are denominated in the same currency"
+#     ),
+#     showarrow = F, xref = "paper", yref = "paper",
+#     xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+#     font = list(size = 11)
+#   )
+
+currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
+      "", "Note: not all values are denominated in the same currency"
     )
-}
+```
 
+```{r KPIs, echo=FALSE, message=FALSE, warning=FALSE}
 ## Fundraising ROI------------------------------------------------------------
@@
 } else {
+  {
     exceldata |>
@@
   } |> KPI_to_string()
+}
 
-```
+## OPX KPI--------------------------------------------------------------------
+# Creating OPX KPI
+KPI_OPX_num <-
+  # testing for data quality, returning NA if bad
+  if (datatest == TRUE) {
+    NA
+  } else {
+    exceldata |>
+      # most recent year
+      filter(year == max(year)) |>
+      # Calculating the value as a %
+      transmute(
+        KPI_OPX =
+          sum(revenue_private_contributions,
+            revenue_investments, assets_cash_in_bank_and_cash_equivalents,
+            na.rm = TRUE
+          ) /
+            sum(liability_short_term_grants_payable,
+              liability_revocable_endowments,
+              liability_long_term_grants_payable,
+              na.rm = TRUE
+            ) * 100
+      ) |>
+      # Just the single number
+      pull() |>
+      first() |>
+      KPI_to_string()
+  }
 
-```
+# Creating private donations growth
+KPI_private_num <-
+  # testing for data quality, returning NA if bad
+  if (datatest == TRUE) {
+    NA
+  } else {
+    exceldata |>
+      # Calculating the two most recent years as a %
+      transmute(
+        growth =
+          (revenue_private_contributions[1] -
+            revenue_private_contributions[2]) /
+            revenue_private_contributions[2] * 100
+      ) |>
+      # Pulling the first instance of the number
+      pull() |>
+      first() |> 
+      KPI_to_string()
+  }
 
-```{r data, include=FALSE}
+## Public grants growth-------------------------------------------------------
+# Creating public grants growth
+KPI_public_num <-
+  # testing for data quality, returning NA if bad
+  if (datatest == TRUE) {
+    NA
+  } else {
+    exceldata |>
+      # Calculating the two most recent years as a %
+      transmute(
+        growth =
+          (revenue_public_grants[1] - revenue_public_grants[2]) /
+            revenue_public_grants[2] * 100
+      ) |>
+      # Pulling the first instance of the number
+      pull() |>
+      first() |> 
+      KPI_to_string()
+  }
 
+## Public/private proportion--------------------------------------------------
+# Creating public/private proportion
+KPI_proportion <-
+  # testing for data quality, returning NA if bad
+  if (datatest == TRUE) {
+    NA
+  } else {
+    exceldata |>
+      transmute(
+        proportion =
+          sum(revenue_private_contributions,
+            revenue_membership_fees,
+            na.rm = TRUE
+          ) /
+            sum(revenue_public_grants, revenue_private_contributions,
+              revenue_membership_fees,
+              na.rm = TRUE
+            ) * 100
+      ) |>
+      pull() |>
+      first() |>
+      KPI_to_string()
+  }
 ```
@@
 ### Fundraising ROI: `r fundraising`
+
+```{r plots, echo=FALSE, message=FALSE, warning=FALSE}
+# Plots ======================================================================
+
+## Annual result--------------------------------------------------------------
+arsresultat_plot <-
+  # testing for the presence of data
+  if (sum(!is.na(htmldata$arsresultat)) == 0) {
+    # Printing 'no information' plot if no data
+    no_information()
+  } else {
+    htmldata |>
+      # renaming for plotly hover
+      rename(
+        Year = year,
+        `Net income` = arsresultat,
+        `Net positive` = annual_profit
+      ) |>
+      ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
+      geom_col() +
+      # using currency scale function to add correct currency code or ?
+      currency_scale +
+      # using a function to ensure uniform colours irrespective of data
+      scale_fill_green(htmldata$annual_profit) +
+      theme(legend.position = "none") +
+      labs(
+        x = "", y = "", title = "Net income (in thousands)",
+        caption = currency_caption
+      )
+  }
+
+arsresultat_plot
+
+## Balance--------------------------------------------------------------------
+balance_plot <- 
+  # testing for the presence of data
+  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
+    # Printing 'no information' plot if no data
+    no_information()
+  } else {
+    htmldata |>
+      # rename for plotly hover
+      rename(
+        Year = year,
+        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
+        `Positive equity` = balance_positive
+      ) |>
+      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
+      geom_col() +
+      # using currency scale function to add correct currency code or ?
+      currency_scale +
+      # using a function to ensure uniform colours irrespective of data
+      scale_fill_green(htmldata$balance_positive) +
+      theme(legend.position = "none") +
+      labs(
+        x = "", y = "", title = "Total equity/liabilities (in thousands)",
+        caption = currency_caption
+      )
+  }
+
+balance_plot
+
+# Creating plotly output
+#ggplotly(balance_plot) |>
+  ## ggplot subtitles and captions are lost when using ggplotly(), a
+  ## currency warning caption is therefore added in plotly, if needed, using
+  ## a separate function
+  # layout(
+  #   annotations =
+  #     plotly_caption
+  # )
+
+## Expenses-------------------------------------------------------------------
+expenses_plot <- 
+    # testing for data quality
+    if (datatest == TRUE) {
+      # printing 'bad data' plot if data are bad
+      bad_data()
+    } else {
+      exceldata |>
+        # Renaming for the purposes of the plot + dropping other variables
+        transmute(
+          Year = as.factor(year),
+          Other = expenses_other,
+          Administrative = expenses_administrative,
+          Fundraising = expenses_fundraising,
+          `Program services` = expenses_program_services
+        ) |>
+        # Pivoting to use fill
+        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
+        # Ordering for the plot
+        mutate(
+          Expenses = ordered(Expenses,
+            levels =
+              c("Program services", "Other", "Fundraising", "Administrative")
+          )
+        ) |>
+        ggplot(aes(Year, value, fill = Expenses)) +
+        geom_col() +
+        scale_y_continuous(
+          # Excel data are assumed to be in NOK; if this changes, code based on
+          # the currency_code() function may be adapted
+          labels = label_dollar(prefix = "NOK "),
+          # Basing lower limit on the data (not suppressing erroneous negative
+          # numbers), and the upper limit on the maximum value in revenues and
+          # expenses, to match the two plots
+          limits = c(NA, maximum)
+        ) +
+        scale_fill_met_d("Hokusai2") +
+        labs(x = "", y = "", title = "Expenses (in thousands)")
+    }
+
+expenses_plot
+
+# Creating plotly output
+# ggplotly(expenses_plot)
+
+## Revenue--------------------------------------------------------------------
+revenue_plot <-
+  # testing for data quality
+  if (datatest == TRUE) {
+    # printing 'bad data' plot if data are bad
+    bad_data()
+  } else {
+    exceldata |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        `Private contributions` = revenue_private_contributions,
+        `Membership fees` = revenue_membership_fees,
+        `Public grants` = revenue_public_grants,
+        Investments = revenue_investments,
+        `Operational activities` = revenue_operational_activities,
+        `Commercial activities` = revenue_commercial_activities,
+        Other = revenue_other
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(
+        c(`Private contributions`:Other),
+        names_to = "Revenue source"
+      ) |>
+      # Ordering for the plot
+      mutate(
+        `Revenue source` =
+          ordered(`Revenue source`,
+            levels =
+              c(
+                "Other", "Commercial activities",
+                "Operational activities", "Investments",
+                "Membership fees", "Private contributions",
+                "Public grants"
+              )
+          )
+      ) |>
+      ggplot(aes(Year, value, fill = `Revenue source`)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum)
+      ) +
+      # Specifying the number of values, as the base palette only includes 6
+      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
+      labs(x = "", y = "", title = "Revenue (in thousands)")
+  }
+
+revenue_plot
+
+# Creating plotly output
+# ggplotly(revenue_plot)
+
+## CEO compensation-----------------------------------------------------------
+CEO_plot <-
+  # Testing for the presence of data
+  if (sum(!is.na(htmldata$lederlonn)) == 0 &
+    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
+    # Printing 'no information' plot if no data
+    no_information()
+  } else {
+    # Creating plot if any data exist
+    # Note that the plot will still look empty, but have a legend and scales,
+    # if all values are explicit zeroes
+    htmldata |>
+      # preparing for pivoting
+      select(
+        year, lederlonn, leder_annen_godtgjorelse
+      ) |>
+      # Compensation in ordinary numbers
+      mutate(
+        lederlonn = lederlonn * 1000,
+        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
+      ) |>
+      # renaming for the purposes of the plot
+      rename(
+        Year = year,
+        Salary = lederlonn,
+        `Other compensation` = leder_annen_godtgjorelse
+      ) |>
+      # pivoting to join the two values to use fill
+      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
+      ggplot(aes(Year, value, fill = Compensation)) +
+      geom_col() +
+      # using currency scale function to add correct currency code or ?
+      currency_scale +
+      # pretty colours
+      scale_fill_met_d("Hokusai2", direction = -1) +
+      labs(
+        x = "", y = "Total compensation", title = "CEO compensation"
+      )
+  }
+
+CEO_plot
+
+# Creating plotly output
+# ggplotly(CEO_plot) |>
+  ## ggplot subtitles and captions are lost when using ggplotly(), a
+  ## currency warning caption is therefore added in plotly, if needed, using
+  ## a separate function
+  # layout(
+  #   annotations =
+  #     plotly_caption
+  # )
+```

### 2022-07-24 Shiny/Test/report.Rmd
Getting close to finally having a report download feature.

@@
 
-```{r plots, echo=FALSE, message=FALSE, warning=FALSE}
+# Plots
+
+### Plots
+
+The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.
+
+- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
+
+```{r arsresultat_plot, echo=FALSE, message=FALSE, warning=FALSE}
 # Plots ======================================================================
@@
     # Printing 'no information' plot if no data
-    no_information()
+    NA
   } else {
@@
   }
+if (sum(!is.na(htmldata$arsresultat)) != 0) {
+  arsresultat_plot
+}
+```
 
-arsresultat_plot
+- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
 
+```{r balance_plot, echo=FALSE, message=FALSE, warning=FALSE}
 ## Balance--------------------------------------------------------------------
@@
   }
-
+if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) != 0) {
   balance_plot
-
+}
 # Creating plotly output
@@
   # )
+```
 
+- **Expenses**: Based on the category *Expenses* in the Excel sheet.
+
+```{r expenses_plot, echo=FALSE, message=FALSE, warning=FALSE}
 ## Expenses-------------------------------------------------------------------
@@
       # printing 'bad data' plot if data are bad
-      bad_data()
+      NA
     } else {
@@
     }
-
+ if (datatest == FALSE) {
    expenses_plot
-
+ }
 # Creating plotly output
 # ggplotly(expenses_plot)
+```
 
+- **Revenues**: Based on the category *Revenue* in the Excel sheet.
+
+```{r revenue_plot, echo=FALSE, message=FALSE, warning=FALSE}
 ## Revenue--------------------------------------------------------------------
@@
     # printing 'bad data' plot if data are bad
-    bad_data()
+    NA
   } else {
@@
   }
-
+if (datatest == FALSE) {
   revenue_plot
-
+}
 # Creating plotly output
 # ggplotly(revenue_plot)
+```
 
+- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
+  - A common measure of transparency, but should normally be evaluated in context.
+
+```{r compensation_plot, echo=FALSE, message=FALSE, warning=FALSE}
 ## CEO compensation-----------------------------------------------------------
@@
     sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
-    # Printing 'no information' plot if no data
-    no_information()
+    # Returning NA if no data
+    NA
   } else {
@@
   }
-
+if (sum(!is.na(htmldata$lederlonn)) != 0 &
+  sum(!is.na(htmldata$leder_annen_godtgjorelse)) != 0) {
   CEO_plot
+}
 

### 2022-07-25 Shiny/Test/Fontawesome.Rmd
Adding info on Fontawesome update.
Ideer til ikoner

@@
 ```
+
+Fontawesome er nå oppdatert til v6.1.1.
 
@@
 
+new icons
+private donations:  https://fontawesome.com/icons/wallet?s=solid
+
+På feks liquid assets/obligations:
+Scale balanced når det er grønt: https://fontawesome.com/icons/scale-balanced?s=solid
+Scale unbalanced når det er rødt: https://fontawesome.com/icons/scale-unbalanced?s=solid
+
+På private donations: https://fontawesome.com/icons/chart-line?s=solid
+
+Andre forslag generelt:
+Globe: https://fontawesome.com/icons/globe?s=solid
+Magnifying glass + chart: https://fontawesome.com/icons/magnifying-glass-chart?s=solid
+
 #### Checked

### 2022-07-25 Shiny/Test/report.Rmd
So many changes.

@@
 # Reactive function to create the maximum value for revenue and expenses plots
-maximum <- 
+maximum <- if (datatest == TRUE) {
+  NA
+} else {
     # reading the Excel sheet anew
@@
       max()
+  }
 
@@
   if (sum(!is.na(htmldata$arsresultat)) == 0) {
-    # Printing 'no information' plot if no data
     NA
@@
 }
+
 ```
+
+`r ifelse(!is.na(arsresultat_plot), "", "Data are bad")`
 
@@
 
-# Creating plotly output
-# ggplotly(CEO_plot) |>
-  ## ggplot subtitles and captions are lost when using ggplotly(), a
-  ## currency warning caption is therefore added in plotly, if needed, using
-  ## a separate function
-  # layout(
-  #   annotations =
-  #     plotly_caption
-  # )
 ```

### 2022-07-25 Shiny/Test/Fontawesome.Rmd
Adding some more ideas, styling links

--- 2022-07-19 Shiny/Test/Fontawesome.Rmd
@@
 
+new icons
+
+[private donations](https://fontawesome.com/icons/wallet?s=solid)
+
+På feks liquid assets/obligations:
+
+[Scale balanced når det er grønt](https://fontawesome.com/icons/scale-balanced?s=solid): fa-solid fa-scale-balanced
+
+[Scale unbalanced når det er rødt](https://fontawesome.com/icons/scale-unbalanced?s=solid)
+
+[På private donations](https://fontawesome.com/icons/chart-line?s=solid)
+
+Andre forslag generelt:
+
+[Globe](https://fontawesome.com/icons/globe?s=solid)
+
+[Magnifying glass + chart](https://fontawesome.com/icons/magnifying-glass-chart?s=solid)
+
+https://fontawesome.com/icons/arrow-up-right-dots?s=solid
+
+https://fontawesome.com/icons/chart-pie?s=solid
+
+https://fontawesome.com/icons/landmark?s=solid
+
+https://fontawesome.com/icons/building-columns?s=solid
+
+https://fontawesome.com/icons/money-bill-trend-up?s=solid
+
+https://fontawesome.com/icons/circle-xmark?s=solid
+
 #### Checked

### 2022-07-25 Dashboard/NGO Analysis dashboard alpha.Rmd
Changing icons. Adding a data test to maximum(). Updating translations in headings.

@@
 
-### Annual result
+### Net income
 ```{r}
@@
 
-### Balance
+### Total equity/liabilities
 ```{r}
@@
       },
-      icon = "fas fa-donate"
+      icon = if (is.na(KPI_fundraising_num())) {
+        "fa-solid fa-circle-xmark"
+      } else {
+        "fas fa-donate"
+      }
     )
@@
       },
-      icon = "fas fa-seedling"
+      icon = if (is.na(KPI_OPX_num())) {
+        "fa-solid fa-circle-xmark"
+      } else if (KPI_OPX_num() < 100) {
+        "fa-solid fa-scale-unbalanced"
+      } else {
+        "fa-solid fa-scale-balanced"
+      }
     )
@@
       },
-      icon = "fas fa-hands-helping"
+      icon = if (is.na(KPI_private_num())) {
+        "fa-solid fa-circle-xmark"
+      } else {
+        "fas fa-hands-helping"
+      }
     )
@@
       },
-      icon = "fas fa-hand-holding-usd"
+      icon = if (is.na(KPI_public_num())) {
+        "fa-solid fa-circle-xmark"
+      } else {
+        "fa-solid fa-building-columns"
+      }
     )
@@
       },
-      icon = "fas fa-percentage"
+      icon = if (is.na(KPI_proportion())) {
+        "fa-solid fa-circle-xmark"
+      } else {
+        "fa-solid fa-chart-pie"
+      }
     )
@@
 
-  ## Excel data test------------------------------------------------------------
+  ## Excel data test (also below)-----------------------------------------------
   # Reactive function to test for Excel data quality (basic)
@@
   maximum <- reactive({
+    if (datatest() == TRUE) {
+      NA
+    } else {
     # reading the Excel sheet anew
@@
       max()
+    }
   })
 
-  ## Excel data test------------------------------------------------------------
+  ## Excel data test (also above)-----------------------------------------------
   # Reactive function to test for Excel data quality (basic)

### 2022-07-25 Shiny/Test/Dashboard downloads test.Rmd
So many changes.
Adding NGO title. Requiring purrr. Reformatting KPI details.

@@
   shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-  flexdashboard, MetBrewer, fontawesome, htmltools
+  flexdashboard, MetBrewer, fontawesome, htmltools, purrr
 )
@@
 
-Enter the name of the NGO in the text box to include it in the report.
-
 ```{r sidebar_button}
-textInput("name", "NGO name", value = "NGO name")
 # download
@@
 
+Row {data-height=60}
+-----------------------------------------------------------------------
+
+### NGO analysed
+
+```{r}
+# Using htmlOutput() to use html styling
+htmlOutput("title")
+```
+
 Row 
@@
 
-- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
-  - A basic measure of the ROI of direct fundraising expenses.
-- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
-  - A measure of the relationship between the organisation's liquid assets and its obligations.
-- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
-  - A calculation of the percentage change in private contributions based on the two most recent years.
-- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
-  - A calculation of the percentage change in public grants received based on the two most recent years.
-- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
-  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.
+- **Fundraising ROI**:
+  - A basic measure of the return on investment of direct fundraising expenses:
+
+  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
+
+- **Liquid assets/obligations**:
+  - A measure of the relationship between the organisation's liquid assets and its obligations:
+
+  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
+
+- **Private donations growth**:
+  - A calculation of the percentage change in private contributions based on the two most recent years:
+
+  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
+
+- **Public grants growth**:
+  - A calculation of the percentage change in public grants received based on the two most recent years:
+
+$$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+
+- **Private/public proportion**:
+  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
+
+$$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
 
@@
       start = input$start,
-      end = input$end,
-      title = input$name
+      end = input$end
     )
@@
     widen_excel(input$excel$datapath)
+  })
+
+  # NGO title ==================================================================
+  NGO <- reactive({
+    read_excel(input$excel$datapath) |>
+      # Turning the needed column into a list
+      pull(Firm) |>
+      # discarding all NAs in case the name is somewhere
+      purrr::discard(is.na) |>
+      # choosing the first available name
+      first()
+  })
+
+  title <- reactive({
+    # testing if NGO() is NA
+    if (is.na(NGO())) {
+      # Printing warning if NA
+      "Name of NGO missing"
+      # Printing name of NGO
+    } else {
+      NGO()
+    }
+  })
+
+  # Creating NGO title using html formatting (colour based on MetBrewer
+  # "Hokusai2")
+  output$title <- renderText({
+    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
   })

### 2022-07-25 Shiny/Test/report.Rmd
Moving plots around, reconsidering layout of pdf.

--- 2022-07-24 Shiny/Test/report.Rmd
@@
 ---
-title: "NGO Analysis: `r params$title`"
 date: "`r Sys.Date()`"
-output: pdf_document
+output: 
+  pdf_document:
+    fig_crop: false
 params:
@@
   end: NA
-  title: NA
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.asp=0.618, out.width="70%", fig.align='center')
 
+# Repeating loading the libraries, in case the information isn't passed on from
+# the dashboard
 if (!require("pacman")) install.packages("pacman")
@@
 
-```{r data, echo=FALSE, message=FALSE, warning=FALSE}
-# Data reading functions =======================================================
+<!--# The title element of the YAML block is further down, as it relies on data from the Excel sheet. -->
+
+```{r data}
+# Data reading =================================================================
 
@@
 }
-
+### Reading html ---------------------
 htmldata <- read_proff(
@@
 
+### Reading excel ---------------------
 exceldata <- widen_excel(params$excelfile)
+
+# NGO title ==================================================================
+
+NGO <- read_excel(params$excelfile) |>
+  # Turning the needed column into a list
+  pull(Firm) |>
+  # discarding all NAs in case the name is somewhere
+  purrr::discard(is.na) |>
+  # choosing the first available name
+  first()
+
+title <- # testing if NGO() is NA
+  if (is.na(NGO)) {
+    # Printing warning if NA
+    "Name of NGO missing"
+    # Printing name of NGO
+  } else {
+    NGO
+  }
 ```
 
-```{r static_functions, echo=FALSE, message=FALSE, warning=FALSE}
-## empty_plot-------------------------------------------------------------------
-# Function to print an empty plot with a supplied message
-empty_plot <- function(warning) {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    annotate(
-      "text",
-      label = warning,
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
+<!--# Report title-->
+
+---
+title: "NGO Analysis: `r title`"
+---
+
+```{r data_warnings}
+# Excel data test =============================================================
+# Function to test for Excel data quality (basic)
+datatest <-
+  # let me know if anyone has a better idea for testing
+  if (mean(unlist(lapply(exceldata, is.character))) != 0) {
+    TRUE
+  } else {
+    FALSE
   }
 
-## no_information plot ---------------------------------------------------------
-# Function to print a 'no information' plot, if necessary
-no_information <- function() {
-  empty_plot("No information available")
+## Creating the Excel warning text --------------------------------------------
+excel_warning <- if (datatest == TRUE) {
+  "**    Warning: something is wrong with the Excel data.**"
+} else {
+  ""
 }
 
-## bad_data plot ---------------------------------------------------------------
-# Function to print a 'bad data' plot, if necessary
-bad_data <- function() {
-  empty_plot("Something is wrong\nwith the Excel data.")
+# HTML warnings ===============================================================
+# Creating the arsresultat warning text
+arsresultat_plot_warning <- if (sum(!is.na(htmldata$arsresultat)) == 0) {
+  "*    No information about net income/loss is available.*"
+} else {
+  ""
 }
 
+# Creating the balance warning text
+balance_plot_warning <-  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
+  "*    No information about total equity/liabilities is available.*"
+} else {
+  ""
+}
+
+# Creating the CEO_plot warning text
+CEO_plot_warning <- if (sum(!is.na(htmldata$lederlonn)) == 0 &
+    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
+  "*    No information about CEO compensation is available.*"
+} else {
+  ""
+}
+
+# Creating general warning
+general_warning <- if (datatest == TRUE | sum(!is.na(htmldata$arsresultat)) == 0 | sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 | (sum(!is.na(htmldata$lederlonn)) == 0 &
+    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0)) {
+  "## Note:"
+} else {
+  ""
+}
+```
+
+<!--# Functions that don't rely on any of the parameters passed from the dashboard -->
+
+```{r static_functions}
 ## KPI_to_string ---------------------------------------------------------------
@@
 
-```{r functions_reliant, echo=FALSE, message=FALSE, warning=FALSE}
+<!--# Functions that rely on parameters from the dashboard -->
+
+```{r functions_reliant}
+
 ## scale maximum--------------------------------------------------------------
-# Reactive function to create the maximum value for revenue and expenses plots
-maximum <- 
+# Function to create the maximum value for revenue and expenses plots
+maximum <- if (datatest == TRUE) {
+  NA
+} else {
     # reading the Excel sheet anew
@@
       max()
-
-## Excel data test------------------------------------------------------------
-# Reactive function to test for Excel data quality (basic)
-datatest <-
-  # let me know if anyone has a better idea for testing
-  if (mean(unlist(lapply(exceldata, is.character))) != 0) {
-    TRUE
-  } else {
-    FALSE
   }
@@
 ## Currency scale-------------------------------------------------------------
-# Reactive function to check for currency and use currency code or ?
+# Function to check for currency and use currency code or ?
 currency_scale <-
@@
 
-## Plotly caption-------------------------------------------------------------
-# Reactive function to create plotly caption (ggplot captions and subtitles
-# are lost in the conversion) (shamelessly stolen from the Internet, with
-# minor modifications).
-# Checks for uniform currency code and prints a warning
-# if there are multiple currency codes.
-# plotly_caption <-
-#   list(
-#     x = 1, y = -0.1, text = ifelse(length(unique(htmldata$valutakode)) == 1,
-#       "", "Note: not all values are denominated in the same currency"
-#     ),
-#     showarrow = F, xref = "paper", yref = "paper",
-#     xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
-#     font = list(size = 11)
+currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
+      "Note: not all values are denominated in the same currency.", ""
+    )
+# currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
+#       "", "Note: not all values are denominated in the same currency.\n"
 #     )
-
-currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
-      "", "Note: not all values are denominated in the same currency"
-    )
 ```
 
-```{r KPIs, echo=FALSE, message=FALSE, warning=FALSE}
+<!--# Creating KPIs -->
+
+```{r KPIs}
 ## Fundraising ROI------------------------------------------------------------
@@
 # Creating OPX KPI
-KPI_OPX_num <-
+KPI_OPX <-
   # testing for data quality, returning NA if bad
@@
 
+## Private donations ---------------------------------------------------------
 # Creating private donations growth
-KPI_private_num <-
+KPI_private <-
   # testing for data quality, returning NA if bad
@@
 # Creating public grants growth
-KPI_public_num <-
+KPI_public <-
   # testing for data quality, returning NA if bad
@@
   }
+
+# Testing for the presence of NAs; printing warning if there are any
+KPI_warning <- if (is.na(fundraising) | is.na(KPI_OPX) | is.na(KPI_private) |
+  is.na(KPI_public) | is.na(KPI_proportion)) {
+  "*A value of 'NA' indicates that there is something wrong with the Excel data, that the necessary data are not available, or that the data are unsuited to an automatic calculation of the KPI.*"
+} else {
+  ""
+}
 ```
@@
 
-- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
-  - A basic measure of the ROI of direct fundraising expenses.
-- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
-  - A measure of the relationship between the organisation's liquid assets and its obligations.
-- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
-  - A calculation of the percentage change in private contributions based on the two most recent years.
-- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
-  - A calculation of the percentage change in public grants received based on the two most recent years.
-- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
-  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.
+<!--# Need to decide whether to underline the KPIs, as in the first one -->
 
-### Fundraising ROI: `r fundraising`
+- **Fundraising ROI**: [`r fundraising`]{.underline}
+  - A basic measure of the return on investment of direct fundraising expenses:
+  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
 
-```{r plots, echo=FALSE, message=FALSE, warning=FALSE}
-# Plots ======================================================================
+- **Liquid assets/obligations**: `r KPI_OPX`
+  - A measure of the relationship between the organisation's liquid assets and its obligations:
+  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
 
+- **Private donations growth**: `r KPI_private`
+  - A calculation of the percentage change in private contributions based on the two most recent years:
+  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
+
+- **Public grants growth**: `r KPI_public`
+  - A calculation of the percentage change in public grants received based on the two most recent years:
+  $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+
+- **Private/public proportion**: `r KPI_proportion`
+  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
+  $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
+
+`r KPI_warning`
+
+<!--# Latex command to force new page - ensures the plot heading doesn't move up to the first page-->
+
+\newpage
+
+# Plots
+
+<!--# Warning header, if necessary -->
+
+`r general_warning`
+
+<!--# specific alerts -->
+
+`r excel_warning`
+
+`r arsresultat_plot_warning`
+
+`r balance_plot_warning`
+
+`r CEO_plot_warning`
+
+<!-- But of course a bit of raw LaTeX is necessary to print a left quotation mark-->
+
+```{r arsresultat_plot, fig.cap="Based on the entry \`Årsresultat' in the Proff.no data."}
 ## Annual result--------------------------------------------------------------
@@
   if (sum(!is.na(htmldata$arsresultat)) == 0) {
-    # Printing 'no information' plot if no data
-    no_information()
+    NA
   } else {
@@
       labs(
-        x = "", y = "", title = "Net income (in thousands)",
-        caption = currency_caption
+        x = "", y = "", caption = currency_caption
       )
   }
+# Printing plot if the data are available
+if (sum(!is.na(htmldata$arsresultat)) != 0) {
+  arsresultat_plot
+}
 
-arsresultat_plot
+```
 
+```{r balance_plot, fig.cap="Based on the entry \`Sum egenkapital og gjeld' in the Proff.no data."}
 ## Balance--------------------------------------------------------------------
@@
       labs(
-        x = "", y = "", title = "Total equity/liabilities (in thousands)",
+        x = "", y = "",
         caption = currency_caption
@@
 
+# Printing the plot if the data are available
+if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) != 0) {
   balance_plot
+}
+```
 
-# Creating plotly output
-#ggplotly(balance_plot) |>
-  ## ggplot subtitles and captions are lost when using ggplotly(), a
-  ## currency warning caption is therefore added in plotly, if needed, using
-  ## a separate function
-  # layout(
-  #   annotations =
-  #     plotly_caption
-  # )
-
-## Expenses-------------------------------------------------------------------
-expenses_plot <- 
-    # testing for data quality
-    if (datatest == TRUE) {
-      # printing 'bad data' plot if data are bad
-      bad_data()
-    } else {
-      exceldata |>
-        # Renaming for the purposes of the plot + dropping other variables
-        transmute(
-          Year = as.factor(year),
-          Other = expenses_other,
-          Administrative = expenses_administrative,
-          Fundraising = expenses_fundraising,
-          `Program services` = expenses_program_services
-        ) |>
-        # Pivoting to use fill
-        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
-        # Ordering for the plot
-        mutate(
-          Expenses = ordered(Expenses,
-            levels =
-              c("Program services", "Other", "Fundraising", "Administrative")
-          )
-        ) |>
-        ggplot(aes(Year, value, fill = Expenses)) +
-        geom_col() +
-        scale_y_continuous(
-          # Excel data are assumed to be in NOK; if this changes, code based on
-          # the currency_code() function may be adapted
-          labels = label_dollar(prefix = "NOK "),
-          # Basing lower limit on the data (not suppressing erroneous negative
-          # numbers), and the upper limit on the maximum value in revenues and
-          # expenses, to match the two plots
-          limits = c(NA, maximum)
-        ) +
-        scale_fill_met_d("Hokusai2") +
-        labs(x = "", y = "", title = "Expenses (in thousands)")
-    }
-
-expenses_plot
-
-# Creating plotly output
-# ggplotly(expenses_plot)
-
+```{r revenue_plot, fig.cap="Based on the category \`Revenue' in the Excel sheet."}
 ## Revenue--------------------------------------------------------------------
@@
       scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
-      labs(x = "", y = "", title = "Revenue (in thousands)")
+      labs(x = "", y = "")
   }
 
+# Printing the plot if the data are OK
+if (datatest == FALSE) {
   revenue_plot
+}
+```
 
-# Creating plotly output
-# ggplotly(revenue_plot)
+```{r expenses_plot, fig.cap="Based on the category \`Expenses' in the Excel sheet."}
+expenses_plot <- 
+    # testing for data quality
+    if (datatest == TRUE) {
+      # printing 'bad data' plot if data are bad
+      NA
+    } else {
+      exceldata |>
+        # Renaming for the purposes of the plot + dropping other variables
+        transmute(
+          Year = as.factor(year),
+          Other = expenses_other,
+          Administrative = expenses_administrative,
+          Fundraising = expenses_fundraising,
+          `Program services` = expenses_program_services
+        ) |>
+        # Pivoting to use fill
+        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
+        # Ordering for the plot
+        mutate(
+          Expenses = ordered(Expenses,
+            levels =
+              c("Program services", "Other", "Fundraising", "Administrative")
+          )
+        ) |>
+        ggplot(aes(Year, value, fill = Expenses)) +
+        geom_col() +
+        scale_y_continuous(
+          # Excel data are assumed to be in NOK; if this changes, code based on
+          # the currency_code() function may be adapted
+          labels = label_dollar(prefix = "NOK "),
+          # Basing lower limit on the data (not suppressing erroneous negative
+          # numbers), and the upper limit on the maximum value in revenues and
+          # expenses, to match the two plots
+          limits = c(NA, maximum)
+        ) +
+        scale_fill_met_d("Hokusai2") +
+        labs(x = "", y = "")
+    }
 
+# Printing the plot if the data are OK
+ if (datatest == FALSE) {
+   expenses_plot
+ }
+```
+
+```{r compensation_plot, out.width="100%", fig.cap="Based on the category \`Lederlønn' in the Proff.no data.
+Not in thousands.
+Includes the entries \`Lønn' and \`Leder annen godtgjørelse'.
+A common measure of transparency, but should normally be evaluated in context."}
 ## CEO compensation-----------------------------------------------------------
@@
       labs(
-        x = "", y = "Total compensation", title = "CEO compensation"
-      )
+        x = "", y = "Total compensation"
+      ) +
+      theme(legend.position = "bottom")
   }
 
+# Printing the plot if the data are available
+if (sum(!is.na(htmldata$lederlonn)) != 0 |
+    sum(!is.na(htmldata$leder_annen_godtgjorelse)) != 0) {
   CEO_plot
-
-# Creating plotly output
-# ggplotly(CEO_plot) |>
-  ## ggplot subtitles and captions are lost when using ggplotly(), a
-  ## currency warning caption is therefore added in plotly, if needed, using
-  ## a separate function
-  # layout(
-  #   annotations =
-  #     plotly_caption
-  # )
+} 
 ```

### 2022-07-25 Shiny/Test/report.Rmd
Abandoning all hope of dealing with cropped axes.
Commenting. Stylering.

@@
 date: "`r Sys.Date()`"
-output: 
-  pdf_document:
-    fig_crop: false
+output: pdf_document
 params:
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.asp=0.618, out.width="70%", fig.align='center')
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 0.618, out.width = "70%", fig.align = "center")
 
@@
 <!--# The title element of the YAML block is further down, as it relies on data from the Excel sheet. -->
+
+<!-- # Need plot titles -->
 
@@
 ```{r functions_reliant}
-
 ## scale maximum--------------------------------------------------------------
@@
 
+## Currency caption-----------------------------------------------------------s
+# Fake currency caption
 currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
@@
 )
+# Real currency caption
 # currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
@@
 
+### NA warning-----------------------------------------------------------------
 # Testing for the presence of NAs; printing warning if there are any
@@
 
-<!-- But of course a bit of raw LaTeX is necessary to print a left quotation mark-->
+<!--# Creating plots -->
+
+<!-- But of course a bit of raw LaTeX is necessary to print a left quotation mark -->
 
@@
 ```{r expenses_plot, fig.cap="Based on the category \`Expenses' in the Excel sheet."}
+## Expenses -------------------------------------------------------------------
 expenses_plot <-

### 2022-07-25 Shiny/Test/report.Rmd
Errant character.

@@
 
-## Currency caption-----------------------------------------------------------s
+## Currency caption-----------------------------------------------------------
 # Fake currency caption

### 2022-07-26 Report/Draft Dashboard section.Rmd
Addin an eval = FALSE.

@@
 If there are any characters, the calculation outputs `NA`, otherwise it carries out the calculation. One example:
-```{r dashboard_KPI_calculation}
+```{r dashboard_KPI_calculation, eval=FALSE}
 # Creating Fundraising ROI value

### 2022-07-26 Shiny/Test/Dashboard downloads test.Rmd
Adding a download notification.

@@
 output$report <- downloadHandler(
+
   filename = "report.pdf",
@@
     )
+
+    id <- showNotification(
+      "Generating report...",
+      duration = NULL,
+      closeButton = FALSE
+    )
+    on.exit(removeNotification(id), add = TRUE)
 

### 2022-07-26 Shiny/Test/report.Rmd
Changing figure size; bolding KPIS.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 0.618, out.width = "70%", fig.align = "center")
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 0.618, out.width = "75%", fig.align = "center")
 
@@
 
-- **Fundraising ROI**: [`r fundraising`]{.underline}
+- **Fundraising ROI: `r fundraising`**
   - A basic measure of the return on investment of direct fundraising expenses:
@@
 
-- **Liquid assets/obligations**: `r KPI_OPX`
+- **Liquid assets/obligations: `r KPI_OPX`**
   - A measure of the relationship between the organisation's liquid assets and its obligations:
@@
 
-- **Private donations growth**: `r KPI_private`
+- **Private donations growth: `r KPI_private`**
   - A calculation of the percentage change in private contributions based on the two most recent years:
@@
 
-- **Public grants growth**: `r KPI_public`
+- **Public grants growth: `r KPI_public`**
   - A calculation of the percentage change in public grants received based on the two most recent years:
@@
 
-- **Private/public proportion**: `r KPI_proportion`
+- **Private/public proportion: `r KPI_proportion`**
   - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
@@
 
-# Plots
+# Graphs
 
@@
       labs(
-        x = "", y = "", caption = currency_caption
+        x = "", y = "", caption = currency_caption, title = "Net income (in thousands)"
       )
@@
         x = "", y = "",
-        caption = currency_caption
+        caption = currency_caption, title = "Total equity/liabilities (in thousands)"
       )
@@
       scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
-      labs(x = "", y = "")
+      labs(x = "", y = "", title = "Revenue (in thousands)")
   }
@@
       scale_fill_met_d("Hokusai2") +
-      labs(x = "", y = "")
+      labs(x = "", y = "", title = "Expenses (in thousands)")
   }
@@
       labs(
-        x = "", y = "Total compensation"
+        x = "", y = "Total compensation", title = "CEO compensation"
       ) +

### 2022-07-26 Dashboard/NGO Analysis dashboard alpha.Rmd
Copying over report generation andv title box.

@@
 
+If desired, a copy of the analysis in PDF format may then be downloaded.
+
+```{r sidebar_button}
+# download button
+uiOutput("downloadUI")
+```
+
 Dashboard
@@
 
-### Plots
-
-The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.
+### Graphs
+
+The dashboard includes five graphs, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.
 
@@
 
-```{r value_boxes, include=FALSE}
-# Shiny actions
-observe({ # wait for button click
+```{r report, include=FALSE}
+# Shiny actions begin here =====================================================
+
+# Report Output ================================================================
+
+output$downloadUI <- renderUI({
+  downloadButton("report", "Download report")
+})
+
+## Download and parameters -----------------------------------------------------
+output$report <- downloadHandler(
+
+  filename = "report.pdf",
+  content = function(file) {
+    tempReport <- file.path(tempdir(), "report.Rmd")
+    file.copy("report.Rmd", tempReport, overwrite = TRUE)
+
+    params <- list(
+      htmlfile = input$html$datapath,
+      excelfile = input$excel$datapath,
+      start = input$start,
+      end = input$end
+    )
+
+    id <- showNotification(
+      "Generating report...",
+      duration = NULL,
+      closeButton = FALSE
+    )
+    on.exit(removeNotification(id), add = TRUE)
+
+    rmarkdown::render(tempReport,
+      output_file = file,
+      params = params,
+      envir = new.env(parent = globalenv())
+    )
+  }
+)
+```
+
+```{r reactive_title_value_boxes, include=FALSE}
+# First reactive chunk =========================================================
+
+# wait for button click
+observe({
 
@@
     ))
+    # Creating data frame
     read_proff(
@@
     ))
+    # Creating data frame
     widen_excel(input$excel$datapath)
@@
 
+  # NGO title ==================================================================
+  NGO <- reactive({
+    read_excel(input$excel$datapath) |>
+      # Turning the needed column into a list
+      pull(Firm) |>
+      # discarding all NAs in case the name is somewhere
+      purrr::discard(is.na) |>
+      # choosing the first available name
+      first()
+  })
+
+  title <- reactive({
+    # testing if NGO() is NA
+    if (is.na(NGO())) {
+      # Printing warning if NA
+      "Name of NGO missing"
+      # Printing name of NGO
+    } else {
+      NGO()
+    }
+  })
+
+  # Creating NGO title using html formatting (colour based on MetBrewer
+  # "Hokusai2")
+  output$title <- renderText({
+    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
+  })
+
   # Value Boxes ================================================================
+
+  # All boxes use the colour "primary" and the icon fa-solid fa-circle-xmark" if
+  # the value is NA
 
@@
       },
+      # icon depends on whether or not the value is NA
       icon = if (is.na(KPI_fundraising_num())) {
@@
       caption = "Liquid assets / obligations",
+      # colour depends on value
       color = if (is.na(KPI_OPX_num())) {
@@
       },
+      # icon depends on value
       icon = if (is.na(KPI_OPX_num())) {
@@
       caption = "Private donation growth",
+      # colour depends on value
       color = if (is.na(KPI_private_num())) {
@@
       },
+      # icon depends on whether or not the value is NA
       icon = if (is.na(KPI_private_num())) {
@@
       caption = "Public grants growth",
+      # colour depends on value
       color = if (is.na(KPI_public_num())) {
@@
       },
+      # icon depends on whether or not the value is NA
       icon = if (is.na(KPI_public_num())) {
@@
       exceldata() |>
+        # calculating the value as a %
         transmute(
@@
         pull() |>
+        # getting the most recent year
         first()
@@
       },
+      # icon depends on whether or not the value is NA
       icon = if (is.na(KPI_proportion())) {
@@
 
-  ## Excel data test (also below)-----------------------------------------------
+  ## Excel data test -----------------------------------------------------------
+  # (also below, in the second chunk)
   # Reactive function to test for Excel data quality (basic)
@@
   })
+  # the button click
 }) |> bindEvent(input$click)
@@
 
-```{r plots, include=FALSE}
+```{r reactive_plots, include=FALSE}
+
+# Beginning of the second reactive chunk =======================================
 # It's necessary to split the observers into separate parts, with all necessary
-# code duplicated.
+# code duplicated, because of R's limitations.
 
@@
   # Data frames ================================================================
+  # Recreating, to make available to this chunk as well
 

### 2022-07-26 Shiny/Test/report.Rmd
Moving KPI styling to function.

@@
 <!--# The title element of the YAML block is further down, as it relies on data from the Excel sheet. -->
-
-<!-- # Need plot titles -->
 
@@
   } else {
-    # Shortening (lengthening) the value, adding %
-    str_c(format(variable, digits = 3), "%")
+    # Shortening (lengthening) the value, adding %; making the text bold
+    str_c("**", format(variable, digits = 3), "%**")
   }
@@
 
-- **Fundraising ROI: `r fundraising`**
+- **Fundraising ROI:** `r fundraising`
   - A basic measure of the return on investment of direct fundraising expenses:
@@
 
-- **Liquid assets/obligations: `r KPI_OPX`**
+- **Liquid assets/obligations:** `r KPI_OPX`
   - A measure of the relationship between the organisation's liquid assets and its obligations:
@@
 
-- **Private donations growth: `r KPI_private`**
+- **Private donations growth:** `r KPI_private`
   - A calculation of the percentage change in private contributions based on the two most recent years:
@@
 
-- **Public grants growth: `r KPI_public`**
+- **Public grants growth:** `r KPI_public`
   - A calculation of the percentage change in public grants received based on the two most recent years:
@@
 
-- **Private/public proportion: `r KPI_proportion`**
+- **Private/public proportion:** `r KPI_proportion`
   - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:

### 2022-07-26 Dashboard/NGO Analysis dashboard alpha.Rmd
OrOr? AndAnd. (Programming logicals).

@@
     # Testing for the presence of data
-    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
+    if (sum(!is.na(htmldata()$lederlonn)) == 0 &&
       sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {

### 2022-07-26 Dashboard/report.Rmd
OrOr? AndAnd. (Programming logicals).

--- 2022-07-26 Shiny/Test/report.Rmd
@@
 excel_warning <- if (datatest == TRUE) {
-  "**    Warning: something is wrong with the Excel data.**"
+  "***    Warning: something is wrong with the Excel data.***"
 } else {
@@
 # Creating the CEO_plot warning text
-CEO_plot_warning <- if (sum(!is.na(htmldata$lederlonn)) == 0 &
+CEO_plot_warning <- if (sum(!is.na(htmldata$lederlonn)) == 0 &&
     sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
@@
 # Creating general warning
-general_warning <- if (datatest == TRUE | sum(!is.na(htmldata$arsresultat)) == 0 | sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 | (sum(!is.na(htmldata$lederlonn)) == 0 &
+general_warning <- if (datatest == TRUE || sum(!is.na(htmldata$arsresultat)) == 0 || sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 || (sum(!is.na(htmldata$lederlonn)) == 0 &&
     sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0)) {
@@
 # Testing for the presence of NAs; printing warning if there are any
-KPI_warning <- if (is.na(fundraising) | is.na(KPI_OPX) | is.na(KPI_private) |
-  is.na(KPI_public) | is.na(KPI_proportion)) {
+KPI_warning <- if (is.na(fundraising) || is.na(KPI_OPX) || is.na( KPI_private) ||
+  is.na(KPI_public) || is.na(KPI_proportion)) {
   "*A value of 'NA' indicates that there is something wrong with the Excel data, that the necessary data are not available, or that the data are unsuited to an automatic calculation of the KPI.*"
@@
   # Testing for the presence of data
-  if (sum(!is.na(htmldata$lederlonn)) == 0 &
+  if (sum(!is.na(htmldata$lederlonn)) == 0 &&
     sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {

### 2022-07-26 Dashboard/report.Rmd
Theme_light(), returning to real currency caption, joining revenue and expenses.

@@
 ## Currency caption-----------------------------------------------------------
-# Fake currency caption
+# Real currency caption
 currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
-  "Note: not all values are denominated in the same currency.", ""
+      "", "Note: not all values are denominated in the same currency.\n"
     )
-# Real currency caption
+
+# Fake currency caption - turns the values around
 # currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
-#       "", "Note: not all values are denominated in the same currency.\n"
+#   "Note: not all values are denominated in the same currency.", ""
 # )
@@
       scale_fill_green(htmldata$annual_profit) +
-      theme(legend.position = "none") +
       labs(
         x = "", y = "", caption = currency_caption, title = "Net income (in thousands)"
-      )
+      ) +
+      theme_light() +
+      theme(legend.position = "none")
   }
@@
       scale_fill_green(htmldata$balance_positive) +
-      theme(legend.position = "none") +
       labs(
@@
         caption = currency_caption, title = "Total equity/liabilities (in thousands)"
-      )
+      ) +
+      theme_light() +
+      theme(legend.position = "none")
   }
@@
 
-```{r revenue_plot, fig.cap="Based on the category \`Revenue' in the Excel sheet."}
+```{r revenue_expenses_plot, fig.asp=1, fig.show="hold", out.width="45%", fig.cap="Based on the categories \`Revenue' and \`Expenses' in the Excel sheet."}
 ## Revenue--------------------------------------------------------------------
@@
       scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
-      labs(x = "", y = "", title = "Revenue (in thousands)")
+      labs(x = "", y = "", title = "Revenue (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
   }
@@
 }
-```
 
-```{r expenses_plot, fig.cap="Based on the category \`Expenses' in the Excel sheet."}
 ## Expenses -------------------------------------------------------------------
@@
       scale_fill_met_d("Hokusai2") +
-      labs(x = "", y = "", title = "Expenses (in thousands)")
+      labs(x = "", y = "", title = "Expenses (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
   }
@@
       labs(
-        x = "", y = "Total compensation", title = "CEO compensation"
+        x = "", y = "", title = "CEO compensation"
       ) +
-      theme(legend.position = "bottom")
+      theme(legend.position = "bottom") +
+      theme_light()
   }

### 2022-07-26 Dashboard/NGO Analysis dashboard alpha.Rmd
Adding validation to title function.

@@
   NGO <- reactive({
+    # validating the file again
+    validate(need(
+      tools::file_ext(input$excel$datapath) == "xlsx",
+      "Please upload an Excel file."
+    ))
+    # fetching the name
     read_excel(input$excel$datapath) |>

### 2022-07-26 Dashboard/report.Rmd
Moving CEO legend to bottom.

@@
       ) +
-      theme(legend.position = "bottom") +
-      theme_light()
+      theme_light() +
+      theme(legend.position = "bottom")
   }

### 2022-07-26 Dashboard/report.Rmd
OrOr in CEO plot.

@@
 # Printing the plot if the data are available
-if (sum(!is.na(htmldata$lederlonn)) != 0 |
+if (sum(!is.na(htmldata$lederlonn)) != 0 ||
   sum(!is.na(htmldata$leder_annen_godtgjorelse)) != 0) {

### 2022-07-26 Old/Dag 3 heck.Rmd
Cleanup.

---
title: "Data analysis and data transformation"
author: "Henrik Rui"
date: '2022-06-29'
output:
  pdf_document: default
  html_document: default
---

#Tinytex
#Setter eval til FALSE for aa ikke evaluere kode når man lager dokumentet

```{r, eval=FALSE}
install.packages("tinytex")
library(tinytex)
install_tinytex()

```

### 2022-07-26 Old/Shiny/Test/Facet test.Rmd
Cleanup.

---
title: "Facet test"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny,tidyverse, scales, rvest, janitor, plotly, readxl, MetBrewer, cowplot)
```

```{r}
revenue <- kreft_excel |> 
  # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year                    = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees`       = revenue_membership_fees,
        `Public grants`         = revenue_public_grants,
        Investments             = revenue_investments,
        `Operational activities`= revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other                   = revenue_other
      ) |> 
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other), names_to = "Revenue source"
      ) |> 
  ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
scale_y_continuous(labels = label_dollar(prefix = "NOK "),
                         limits = c(NA, maximum)) +
  scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)")
revenue
expenses <- kreft_excel |> 
  transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |> 
  ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(labels = label_dollar(prefix = "NOK "),
                         limits = c(NA, maximum)) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)")

expenses

maximum

?scale_y_continuous()

expenses

maximum <- read_excel("Data/OPXfinans_final.xlsx") |> 
  select(-Firm, -`Sub-category`) |> 
  filter(Category == "Revenue" | Category == "Expenses") |> 
  group_by(Year, Category) |> 
  summarise(
    sum = sum(Value, na.rm = TRUE)
  ) |> 
  pull() |> 
  max()

maximum

excel_long

bare
filtered <- bare |> 
  filter(Category == "Revenue" | Category == "Expenses")

filtered
filtered <- filtered |>
  select(-Firm, -`Sub-category`)

filtered |> 
  ggplot(aes(as.factor(Year), Value, fill = Name)) +
  geom_col() +
  facet_wrap(~Category) +
  scale_y_continuous(labels = label_dollar()) +
  scale_fill_manual(values = met.brewer("Hokusai3", 10))

maximum <- filtered |> 
  group_by(Year, Category) |> 
  summarise(
    sum = sum(Value, na.rm = TRUE)
  ) |> 
  pull() |> 
  max()

filtered |> 
  group_by(Year, Category) |> 
  summarise(
    sum = sum(Value, na.rm = TRUE)
  )

```

# Data
```{r}
widen_excel <- function(file){
  read_excel(file) |> 
  unite(
    entry, c(Category, Name), sep = "_"
  ) |> 
  select(-`Sub-category`) |> 
  pivot_wider(names_from = entry, values_from = Value) |> 
  clean_names()
}

kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
```

### 2022-07-26 Old/Shiny/Test/Flexdashboard rows function tests.Rmd
Cleanup.

---
title: "NGO analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
runtime: shiny
---

```{r setup, include=FALSE}
# Loading packages using pacman to ensure they're all installed
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
# Using fontawesome for more up-to-date (and currently supported) fontawesome
# icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

Sidebar {.sidebar}
=======================================================================

### File upload

Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.

Then upload an Excel file in the set format.

Click 'Analyse' to analyse the data.

```{r sidebar}
# HTML
fileInput("html", label = "Upload html file", accept = ".html")

# Year inputs
numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Button
actionButton("click", "Analyse")
```

Dashboard
=======================================================================

Row 
-----------------------------------------------------------------------

### KPI Fundraising
```{r}
valueBoxOutput("KPI_box_fundraising")
```

### KPI OPX
```{r}
valueBoxOutput("KPI_box_OPX")
```

### KPI Private growth
```{r}
valueBoxOutput("KPI_box_private")
```

### KPI Public growth
```{r}
valueBoxOutput("KPI_box_public")
```

### KPI Private/public proportion
```{r}
valueBoxOutput("KPI_box_proportion")
```

Row
-----------------------------------------------------------------------

### Annual result
```{r}
plotlyOutput("arsresultat_plotly")
```

### Balance
```{r}
plotlyOutput("balance_plotly")
```

Row 
-----------------------------------------------------------------------

### Revenues
```{r}
plotlyOutput("revenue_plotly")
```

### Expenses
```{r}
plotlyOutput("expenses_plotly")
```

Row 
-----------------------------------------------------------------------

### Transparency: CEO compensation
```{r}
plotlyOutput("CEO_plotly")
```

Details
=======================================================================

Row 
-----------------------------------------------------------------------

### Key Performance Indicators

The KPIs are based on the following calculations, using data derived from the Excel sheet:

- **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
  - A basic measure of the ROI of direct fundraising expenses.
- **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
  - A measure of the relationship between the organisation's liquid assets and its obligations.
- **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
  - A calculation of the percentage change in private contributions based on the two most recent years.
- **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
  - A calculation of the percentage change in public grants received based on the two most recent years.
- **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.

Row 
-----------------------------------------------------------------------

### Plots

The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

- **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
- **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

```{r shiny, include=FALSE}
# Shiny actions
observe({ # wait for button click

  #---- Data frames ------------------------------------------------------------      
  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(tools::file_ext(input$html$datapath) == "html",
                  "Please upload an html file."))
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
                  "Please upload an Excel file."))
    widen_excel(input$excel$datapath)
  })

  #---- Value Boxes ------------------------------------------------------------

  # Creating Fundraising ROI value
  KPI_fundraising_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        transmute(
          # Calculating the value as a percentage; removing all other
          KPI_fundraising_num =
            (revenue_private_contributions +
              revenue_membership_fees - expenses_fundraising) /
              expenses_fundraising * 100
        ) |>
        # just want the single number
        pull() |>
        first()
    }
  })

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  # testing for data quality, returning NA if bad
    output$KPI_box_fundraising <- renderValueBox({
      valueBox(
        value = KPI_to_string(KPI_fundraising_num()),
        caption = "Fundraising ROI",
        # Colour depends on value
        color = if (is.na(KPI_fundraising_num())) {
          "primary"
        } else if (KPI_fundraising_num() < 100) {
          "danger"
        } else if (KPI_fundraising_num() < 200) {
          "warning"
        } else {
          "success"
        },
        icon = "fas fa-donate"
      )
    })

  # Creating OPX KPI
  KPI_OPX_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        # Calculating the value as a %
        transmute(
          KPI_OPX = (revenue_private_contributions + revenue_investments +
            assets_cash_in_bank_and_cash_equivalents) /
            (liability_short_term_grants_payable +
              liability_revocable_endowments +
              liability_long_term_grants_payable) * 100
        ) |>
        # Just the single number
        pull() |>
        first()
    }
  })

  # Creating value box for OPX KPI
  output$KPI_box_OPX <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_OPX_num()),
      caption = "Liquid assets / obligations",
      color = if (is.na(KPI_OPX_num())) {
        "primary"
      } else if (KPI_OPX_num() < 100) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-seedling"
    )
  })

  # Creating private donations growth
  KPI_private_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth = 
            (revenue_private_contributions[1] - revenue_private_contributions[2]) /
            revenue_private_contributions[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for private donations growth
  output$KPI_box_private <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_private_num()),
      caption = "Private donation growth",
      color = if (is.na(KPI_private_num())) {
        "primary"
      } else if (KPI_private_num() < 0) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-hands-helping"
    )
  })

  # Creating public grants growth
  KPI_public_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth =
            (revenue_public_grants[1] - revenue_public_grants[2]) / 
            revenue_public_grants[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for public grants growth
  output$KPI_box_public <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_public_num()),
      caption = "Public grants growth",
      color = if (is.na(KPI_public_num())) {
        "primary"
      } else if (KPI_public_num() < 0) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-hand-holding-usd"
    )
  })

  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        transmute(
          proportion = (revenue_private_contributions + revenue_membership_fees) / (revenue_public_grants + revenue_private_contributions + revenue_membership_fees) * 100
        ) |>
        pull() |>
        first()
    }
  })

  # Creating value box for private/public proportion
  output$KPI_box_proportion <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_proportion()),
      caption = "Private donations",
      color = "info",
      icon = "fas fa-percentage"
    )
  })

  #---- Plots --------------------------------------------------------------------

  #----Annual result---------------------------------------------
  arsresultat_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        ggplot(aes(year, arsresultat, fill = annual_profit)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$annual_profit) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Annual result (in thousands)"
        )
    }
  })

  # Creating plotly output
  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Balance--------------------------------------------------------------------    
  balance_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$balance_positive) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Balance (in thousands)",
        )
    }
  })

  # Creating plotly output
  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Expenses---------------------------------------
  expenses_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          Other = expenses_other,
          Administrative = expenses_administrative,
          Fundraising = expenses_fundraising,
          `Program services` = expenses_program_services
        ) |>
        # Pivoting to use fill
        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
        # Ordering for the plot
        mutate(
          Expenses = ordered(Expenses,
            levels =
              c("Program services", "Other", "Fundraising", "Administrative")
          )
        ) |>
        ggplot(aes(Year, value, fill = Expenses)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        scale_fill_met_d("Hokusai2") +
        labs(x = "", y = "", title = "Expenses (in thousands)")
    }
  })

  # Creating plotly output
  output$expenses_plotly <- renderPlotly(
    ggplotly(expenses_plot())
  )

  #----Revenue------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        # Ordering for the plot
        mutate(
          `Revenue source` =
            ordered(`Revenue source`, levels =
                      c("Other", "Commercial activities",
                        "Operational activities", "Investments",
                        "Membership fees", "Private contributions",
                        "Public grants"))
        ) |> 
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

  # Creating plotly output
  output$revenue_plotly <- renderPlotly(
    ggplotly(revenue_plot())
  )

  #----CEO compensation---------------------------------------------------------
  CEO_plot <- reactive({
    # Testing for the presence of data
    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      # Creating plot if any data exist
      # Note that the plot will still look empty, but have a legend and scales,
      # if all values are explicit zeroes
      create_CEO_plot(htmldata(), currency_scale())
    }
  })

  # Creating plotly output
  output$CEO_plotly <- renderPlotly(
    ggplotly(CEO_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Reactive functions-------------------------------

  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })

  # Reactive function to check for currency and use currency code or ?
  currency_scale <- reactive({
    scale_y_continuous(
      # Checking for uniform currency code
      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
        # Using the currency code if uniform
        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
        # alternative prefix if multiple currencies in the data
        label_dollar(prefix = "? ")
      )
    )
  })

  # Reactive function to create plotly caption (ggplot captions and subtitles
  # are lost in the conversion) (shamelessly stolen from the Internet, with
  # minor modifications).
  # Checks for uniform currency code and prints a warning
  # if there are multiple currency codes.
  plotly_caption <- reactive({
    list(
      x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
        "", "Note: not all values are denominated in the same currency"
      ),
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 11)
    )
  })

  # Reactive function to create the maximum value for revenue and expenses plots
  maximum <- reactive({
    # reading the Excel sheet anew
    read_excel(input$excel$datapath) |>
      # removing the firm and sub-category columns
      select(-Firm, -`Sub-category`) |>
      # filtering to retain only revenue and expenses entries
      filter(Category == "Revenue" | Category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(Year, Category) |>
      summarise(
        sum = sum(Value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()
  })
  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
}) |> bindEvent(input$click)

```

```{r non-reactive functions, include=FALSE}
# Function to read html file
# read_proff <- function(file, start, end) {
#   read_html(file, encoding = "UTF-8") |>
#     html_node("#inner-frame") |>
#     html_table() |>
#     select( # removing empty column
#       REGNSKAPSPERIODE:all_of(start)
#     ) |> # using all_of()
#     mutate( # removing whitespace
#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
#     ) |>
#     filter( # removing duplicate table - no idea why it's necessary
#       !row_number() > 176
#     ) |>
#     filter( # removing duplicate rows
#       !grepl("Lukk", REGNSKAPSPERIODE)
#     ) |>
#     pivot_longer( # tidying data
#       all_of(end):all_of(start),
#       names_to = "year"
#     ) |> # using all_of()
#     mutate( # changing to real NAs and turning years into numbers
#       value = na_if(value, "-"),
#       year = as.integer(year),
#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
#     ) |>
#     fill( # filling the currency column
#       valutakode,
#       .direction = "updown"
#     ) |>
#     filter( # removing dates and redundant currency
#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
#         REGNSKAPSPERIODE != "Valutakode"
#     ) |>
#     mutate( # removing whitespace in numbers
#       value = str_replace_all(value, "\\s", ""),
#       value = as.numeric(value),
#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
#     ) |>
#     distinct() |> # removing any remaining duplicate rows
#     pivot_wider( # pivoting
#       names_from = REGNSKAPSPERIODE, values_from = value
#     ) |>
#     clean_names() |> # tidying names
#     select(
#       -lederlonn_i_hele_1000,
#       -resultatregnskap_i_hele_1000,
#       -balanseregnskap_i_hele_1000
#     ) |> # removing headings
#     rename( # making it clear what lonn refers to
#       lederlonn = lonn
#     ) |>
#     # Creating new variables for plots
#     mutate(
#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
#     ) |>
#     arrange(desc(year)) # arranging by year
# }

# Test version of function that creates multiple data problems
read_proff <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # bad test - REMOVE
    mutate(
      valutakode = ifelse(year == 2019, "USD", valutakode)
    ) |>
    arrange(desc(year)) # arranging by year
}

# ,
#       leder_annen_godtgjorelse = NA,
#       lederlonn = NA,
#       arsresultat = NA,
#       sum_egenkapital_og_gjeld = NA

# Function to read Excel file
widen_excel <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`, -Firm) |> # deleting Sub-category and Firm
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() # cleaning names
}

# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}

# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
    scale_fill_met_d("Tsimshian",
      direction = ifelse(
        mean(variable,
          na.rm = TRUE
        ) == 1, -1, 1
      )
    )
  }

# Function to print a 'no information' plot, if necessary
no_information <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "No information available",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}

# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = "Something is wrong\nwith the Excel data.",
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}

# Function to create CEO plot
create_CEO_plot <- function(dataframe, scale){
  dataframe |>
        # preparing for pivoting
        select(
          year, lederlonn, leder_annen_godtgjorelse
        ) |>
        # Compensation in ordinary numbers
        mutate(
          lederlonn = lederlonn * 1000,
          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
        ) |>
        # renaming for the purposes of the plot
        rename(
          Year = year,
          Salary = lederlonn,
          `Other compensation` = leder_annen_godtgjorelse
        ) |>
        # pivoting to join the two values to use fill
        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
        ggplot(aes(Year, value, fill = Compensation)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        scale +
        # pretty colours
        scale_fill_met_d("Hokusai2", direction = -1) +
        labs(
          x = "", y = "Total compensation", title = "CEO compensation"
        )
}
```

### 2022-07-26 Old/Shiny/Test/Flexdashboard rows.Rmd
Cleanup.

--- 2022-07-23 Shiny/Test/Flexdashboard rows.Rmd
@@
           KPI_fundraising_num =
-            (revenue_private_contributions +
-              revenue_membership_fees - expenses_fundraising) /
+            (sum(revenue_private_contributions,
+              revenue_membership_fees, na.rm = TRUE) - expenses_fundraising) /
               expenses_fundraising * 100
@@
   # Creating OPX KPI
-  KPI_OPX_num <- reactive({
+
+  # For some bizarre reason I cannot do na.rm = TRUE in these sum
+
+  OPX_num <- reactive({
     # testing for data quality, returning NA if bad
@@
         transmute(
-          KPI_OPX = (revenue_private_contributions + revenue_investments +
-            assets_cash_in_bank_and_cash_equivalents) /
-            (liability_short_term_grants_payable +
-              liability_revocable_endowments +
-              liability_long_term_grants_payable) * 100
+          KPI_OPX =
+            sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents) /
+            sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable) * 100
         ) |>
@@
     valueBox(
-      value = KPI_to_string(KPI_OPX_num()),
+      value = KPI_to_string(OPX_num()),
       caption = "Liquid assets / obligations",
-      color = if (is.na(KPI_OPX_num())) {
+      color = if (is.na(OPX_num())) {
         "primary"
-      } else if (KPI_OPX_num() < 100) {
+      } else if (OPX_num() < 100) {
         "warning"

### 2022-07-26 Old/Shiny/Test/Reading OPX table.Rmd
Cleanup.

---
title: "Reading OPX table"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
care <- read_html("Shiny/Test/Data/NGO Profile Page - OPX.webarchive") |> 
  html_element("body > opx-root > opx-ngo-profile-sidenav > opx-sidenav > div > mat-drawer-container > mat-drawer-content > div > opx-ngo-profile > opx-ngo-financial > div") |> 
  html_table()

care
care2 <- care |> 
  rename(
    Name = X1,
    `2020` = X2,
    `2019` = X3,
    `2018` = X4,
    `2017` = X5
  ) |> 
  pivot_longer(`2020`:`2017`, names_to = "Year")
care2 |> 
  mutate(
    value = (str_replace_all(value, "\\s", "")),
    value = as.numeric(value),
    Year = as.integer(Year)
  ) |> 
  arrange(desc(Year)) |> 
  write_excel_csv("Shiny/Test/Data/care.csv")
```

### 2022-07-26 Old/Test parsing of Proff table.Rmd
Cleanup.

--- 2022-07-15 Test parsing of Proff table.Rmd
@@
 
-read_proff6("Links/Kors.html", "2001", "2020")
+rodekors <- read_proff6("Links/Kors.html", "2001", "2020")
 opera <- read_proff6("Links/opera.html", "1998", "2021")
@@
 ggplotly(opera_lonn)
+
+rodekors
+
+ggplot(rodekors, aes(year, sum_egenkapital_og_gjeld)) +
+  geom_col() +
+  theme_minimal() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = ""))
 ```

### 2022-07-26 Shiny/Test/app.R

@@ -1,167 +1,0 @@
-# Using pacman to load libraries because so many are needed
-
-if (!require("pacman")) install.packages("pacman"); library(pacman)
-p_load(shiny, tidyverse, scales, rvest, janitor, plotly, readxl)
-
-# Creating ui
-ui <- fluidPage(
-
-    # Application title
-    titlePanel("Charity analysis"),
-    
-    # Explanation of inputs
-    htmlOutput("explanation"),
-    
-    # Inputs
-      # need more strict validation for files?
-    # html
-    fileInput("html", label = "Upload html file", accept = ".html"),
-    numericInput("start", label = "Start", value = 2002),
-    numericInput("end", label = "End", value = 2021),
-    # Excel
-    fileInput("excel", "Upload Excel file", accept = ".xlsx"),
-    # Button
-    actionButton("click", "Analyse"),
-    
-    # Outputs (Planning on removing the tables once testing is done)
-      # need to work on how to display KPIs
-    
-    tableOutput("htmltable"),
-    tableOutput("exceltable"),
-    plotOutput("htmlplot"),
-    plotlyOutput("plotlyplot")
-)
-
-# Define server logic
-server <- function(input, output) {
-  
-  # function to read html file
-  
-  read_proff6 <- function(file, start, end){
-    read_html(file, encoding = "UTF-8") |> 
-      html_node("#inner-frame") |> 
-      html_table() |> 
-      select( # removing empty column
-        REGNSKAPSPERIODE:all_of(start)) |> # using all_of()
-      mutate( # removing whitespace
-        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-      ) |> 
-      filter( # removing duplicate table - no idea why it's necessary
-        !row_number() > 176
-      ) |>
-      filter( # removing duplicate rows
-        !grepl("Lukk", REGNSKAPSPERIODE)
-      ) |> 
-      pivot_longer( # tidying data
-        all_of(end):all_of(start), names_to = "year") |> # using all_of()
-      mutate( # changing to real NAs and turning years into numbers
-        value = na_if(value, "-"),
-        year = as.integer(year)) |> 
-      mutate( # adding currency column
-        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-      ) |> 
-      fill( # filling the currency column
-        valutakode, .direction = "updown"
-      ) |> 
-      filter( # removing dates and redundant currency
[…]

### 2022-07-26 Report/Draft Dashboard section.Rmd
Updating with info on report; updating some of the code as well.

--- 2022-07-24 Report/Draft Dashboard section.Rmd
@@
 
->Ca. 6,5 sider.
+>Ca. 8 sider.
 
@@
 The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'under the hood' was created using Shiny.
+Additionally, the dashboard provides the user with the option of downloading a pdf version of the analysis, using a parameterised R Markdown document.
 
@@
 
-In the case of some value boxes, such as the Fundraising ROI one, the function includes logic that changes the colour of the box based on the calculated number.
+In the case of some value boxes, such as the Fundraising ROI one, the function includes logic that changes the colour and icon of the box based on the calculated number.
 The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
 ```{r dashboard_valuebox, eval=FALSE}
-# Creating Fundraising ROI value
-KPI_fundraising_num <- reactive({
-  # testing for data quality, returning NA if bad
-  if (datatest() == TRUE) {
-    NA
+# Creating value box for fundraising ROI
+output$KPI_box_fundraising <- renderValueBox({
+  valueBox(
+    value = KPI_to_string(KPI_fundraising_num()),
+    caption = "Fundraising ROI",
+    # Colour depends on value
+    color = if (is.na(KPI_fundraising_num())) {
+      "primary"
+    } else if (KPI_fundraising_num() < 100) {
+      "danger"
+    } else if (KPI_fundraising_num() < 200) {
+      "warning"
     } else {
-    exceldata() |>
-      # most recent year
-      filter(year == max(year)) |>
-      transmute(
-        # Calculating the value as a percentage; removing all other
-        KPI_fundraising_num =
-          (sum(revenue_private_contributions,
-            revenue_membership_fees,
-            na.rm = TRUE
-          ) - expenses_fundraising) /
-            expenses_fundraising * 100
-      ) |>
-      # just want the single number
-      pull() |>
-      first()
+      "success"
+    },
+    # icon depends on whether or not the value is NA
+    icon = if (is.na(KPI_fundraising_num())) {
+      "fa-solid fa-circle-xmark"
+    } else {
+      "fas fa-donate"
     }
+  )
 })
@@
 
-All the boxes also include a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1.
+All the boxes also include, as mentioned, a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1.
+All boxes use the theme colour 'primary' and the icon 'fa-solid fa-circle-xmark' if the value is NA.
 ```{r dashboard_fontawesome, eval=FALSE}
@@
 
-Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots.
+Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots. At the top it includes a heading containing the name of the NGO as written in the Excel file:
+```{r dashboard_title_creation, eval=FALSE}
+# NGO title ==================================================================
+NGO <- reactive({
+  # validating the file again
+  validate(need(
+    tools::file_ext(input$excel$datapath) == "xlsx",
+    "Please upload an Excel file."
+  ))
+  # fetching the name
+  read_excel(input$excel$datapath) |>
+    # Turning the needed column into a list
+    pull(Firm) |>
+    # discarding all NAs in case the name is somewhere
+    purrr::discard(is.na) |>
+    # choosing the first available name
+    first()
+})
+
+title <- reactive({
+  # testing if NGO() is NA
+  if (is.na(NGO())) {
+    # Printing warning if NA
+    "Name of NGO missing"
+    # Printing name of NGO
+  } else {
+    NGO()
+  }
+})
+```
+
+The title is output using `htmlOutput()` to use some basic html styling:
+```{r dashboard_title, eval=FALSE}
+# Creating NGO title using html formatting (colour based on MetBrewer
+# "Hokusai2")
+output$title <- renderText({
+  paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
+})
+```
+
+## The parameterised R Markdown report
+
+The dashboard also allows the user to download a pdf version of the analysis through clicking on a button.
+A `downloadHandler()` function passes the necessary parameters to the R Markdown document and displays a temporary notification while the file is being created.
+The function includes temporary file creation arguments that should enable it to function on e.g.
+shinyapps.io and similar server contexts, but this has not been tested.
+```{r}
+## Download and parameters -----------------------------------------------------
+output$report <- downloadHandler(
+
+  # Based on code in Mastering Shiny
+  filename = "report.pdf",
+  content = function(file) {
+    tempReport <- file.path(tempdir(), "report.Rmd")
+    file.copy("report.Rmd", tempReport, overwrite = TRUE)
+
+    # parameters
+    params <- list(
+      htmlfile = input$html$datapath,
+      excelfile = input$excel$datapath,
+      start = input$start,
+      end = input$end
+    )
+
+    # notification
+    id <- showNotification(
+      "Generating report...",
+      duration = NULL,
+      closeButton = FALSE
+    )
+    on.exit(removeNotification(id), add = TRUE)
+
+    # R Markdown render
+    rmarkdown::render(tempReport,
+      output_file = file,
+      params = params,
+      envir = new.env(parent = globalenv())
+    )
+  }
+)
+```
+
+The code in the R Markdown document is very similar to that of the dashboard, but structured in a more conventional way.
+Data warnings are here text rather than blank plots, and the KPIs are integrated into the text instead of being presented in value boxes.

### 2022-07-27 Old/Shiny/Test/Data tests.Rmd
Data tests.

---
title: "Data tests"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-07-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rvest)

read_html("Dashboard/Data/")
```

```{r}
read_proff <- function(file, start, end) {
  # Specifying encoding to deal with occasional errors
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column
    fill(
      valutakode,
      .direction = "updown"
    ) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}
```

### 2022-07-27 Dashboard/NGO Analysis dashboard alpha.Rmd

@@
 
+  # Based on code in Mastering Shiny
   filename = "report.pdf",
@@
 
+    # parameters
     params <- list(
@@
 
+    # notification
     id <- showNotification(
@@
 
+    # R Markdown render
     rmarkdown::render(tempReport,
@@
 
-  # All boxes use the colour "primary" and the icon fa-solid fa-circle-xmark" if
+  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark" if
   # the value is NA

### 2022-07-27 Old/Shiny/Test/Flexdashboard rows tests.Rmd

--- 2022-07-24 Shiny/Test/Dashboard downloads test.Rmd
@@
 
-Then upload an Excel file in the set format (available [here](https://github.com/amaliehc/ISSSV1337/blob/5c0a8a2beae499b3d2524a3965ceebf3b3ebad0e/Dashboard/Data/NGO_template.xltx)).
+Then upload an Excel file in the set format.
 
@@
 actionButton("click", "Analyse")
-```
-
-If desired, a copy of the analysis in PDF format may then be downloaded.
-
-Enter the name of the NGO in the text box to include it in the report.
-
-```{r sidebar_button}
-textInput("name", "NGO name", value = "NGO name")
-# download
-uiOutput("downloadUI")
 ```
@@
 # Shiny actions
-
-# Report Output ==============================================================
-
-output$downloadUI <- renderUI({
-  downloadButton("report", "Download report")
-})
-
-output$report <- downloadHandler(
-  filename = "report.pdf",
-  content = function(file) {
-    tempReport <- file.path(tempdir(), "report.Rmd")
-    file.copy("report.Rmd", tempReport, overwrite = TRUE)
-
-    params <- list(
-      htmlfile = input$html$datapath,
-      excelfile = input$excel$datapath,
-      start = input$start,
-      end = input$end,
-      title = input$name
-    )
-
-    rmarkdown::render(tempReport,
-      output_file = file,
-      params = params,
-      envir = new.env(parent = globalenv())
-    )
-  }
-)
-
 observe({ # wait for button click

### 2022-07-27 Old/Shiny/Test/Test data.Rmd

--- 2022-07-22 Shiny/Test/Test data.Rmd
@@
 ```
+
+```{r}
+NGO <- read_excel("Data/OPXfinans_final.xlsx") |> 
+  pull(Firm) |>
+  first()
+
+title <- if (is.na(NGO)) {
+  "Name of NGO missing"
+} else {
+  NGO
+}
+
+title
+
+NGO_blank <- read_excel("Data/Bad_data_no_name.xlsx") |> 
+  pull(Firm) |>
+  first()
+
+NGO_blank <- read_excel("Data/Bad_data_no_name.xlsx") |> 
+  pull(Firm) |> 
+  purrr::discard(is.na) |> 
+  first()
+
+length(NGO_blank)
+
+```
+
+```{r}
+summary(kreft_excel)
+```
+{.tabset-fade}
+
+{.no-title}

### 2022-07-27 Old/Shiny/Test/Testing bad data.Rmd

---
title: "Testing bad data"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman"); library(pacman)
p_load(shiny,tidyverse, scales, rvest, janitor, plotly, readxl, MetBrewer)

```

## Functions and data
```{r}
read_proff8 <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    arrange(desc(year)) # arranging by year
}

riksmaal <- read_proff8("Shiny/Test/Data/Riksmaal2011_21.html", "2011", "2021")

print(sum(riksmaal$lederlonn, na.rm = TRUE))

sum(is.na(riksmaal$lederlonn)) == nrow(riksmaal)
nrow(riksmaal)

sum(!is.na(riksmaal$lederlonn))

sum

if (sum(is.na(riksmaal$lederlonn)) == nrow(riksmaal)) {
  "yes"
} else {
  "no"
}

testplot <- if (sum(!is.na(riksmaal$lederlonn)) == 0 &
    sum(!is.na(riksmaal$leder_annen_godtgjorelse)) == 0) {
    ggplot() +
    geom_text(aes(0, 0, label = "No information available")) +
    theme_void()
} else {
  "yes"
}

ggplotly(testplot)

riksmaal$leder_annen_godtgjorelse

sum(!is.na(riksmaal$lederlonn)) == 0

riksmaal |> 
    ggplot() +
    annotate("No information available") +
    theme_void()

ggplot() +
    theme_void() +
    geom_text(aes(0,0,label='N/A')) +
    xlab(NULL)

ggplot() +
    geom_text(aes(0, 0, label = "No information available")) +
    theme_void()

ggplot() +
    geom_text(aes(0, 0, label = " ")) +
    theme_void() +
  annotate(
    "text", label = "No information available",
    x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
  )

widen_excel_bad <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`) |> # deleting Sub-category
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() |>  # cleaning names
    mutate(
      leder_annen_godtgjorelse = 1000
    )
}

widen_excel <- function(file) {
  read_excel(file) |>
    unite( # uniting Category and Name to deal with multiple 'Other'
      entry, c(Category, Name),
      sep = "_"
    ) |>
    select(-`Sub-category`, -Firm) |> # deleting Sub-category
    # pivoting
    pivot_wider(names_from = entry, values_from = Value) |>
    clean_names() # cleaning names
}
test <- widen_excel("Shiny/Test/Data/Bad_data.xlsx")

test
good <- widen_excel("Shiny/Test/Data/OPXfinans_final.xlsx")
good
testnumber <- NA
ifelse(testnumber >= 0, TRUE, FALSE)

is.character(testnumber)

?is.numeric

is.numeric(kreft_excel$assets_property_plant_equipment_and_intangible_assets)
is.numeric(c())

exceltest <- read_excel("Shiny/Test/Data/Bad_data.xlsx")

is.numeric()

if (mean(unlist(lapply(exceldata(), is.numeric))) != 0) {
  TRUE
} else {
  FALSE
}

is.numeric(NA)
is.character(NA)

class(good)
?if_any

?lapply
```

### 2022-07-27 Old/Dag 3 heck.Rmd

@@
 ```
+
+```{r}
+library(tidyverse)
+costofliving <- read_csv("https://raw.githubusercontent.com/sbjorkholt/ISSSV1337/main/Week%201/Day%203%20-%20Introduction%20to%20data%20analysis%20and%20data%20transformation/Tasks/costofliving.csv")
+```
+
+```{r}
+install.packages("janitor")
+library(janitor)
+
+costofliving_clean<-costofliving %>%
+  clean_names()
+
+```
+
+## Tidyverse functions
+
+-select
+-filter
+-mutate
+-rename
+-group_by
+-summarise
+
+### Select
+
+```{r}
+costofliving_subset <- costofliving_clean %>%
+  select(cost_of_living_index,groceries_index)
+```
+
+```{r}
+costofliving_clean_subset <- costofliving_clean %>%
+  select(cost_of_living_index)
+```
+```
+
+```{r}

### 2022-07-27 Old/Shiny/Test/Data tests.Rmd
Data tests.

@@
 knitr::opts_chunk$set(echo = TRUE)
+if (!require("pacman")) install.packages("pacman")
+library(pacman)
+p_load(
+  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
+  flexdashboard, MetBrewer, fontawesome, htmltools
+)
 ```
@@
 
-read_html("Dashboard/Data/")
+read_html("Data/Kreftforeningen.html")
 ```
@@
 ```
+
+```{r}
+getwd()
+kreft <- read_proff("Data/Kreftforeningen.html", "2002", "2021")
+```
+
+```{r}
+widen_excel <- function(file) {
+  read_excel(file) |>
+    # uniting Category and Name to deal with multiple 'Other'
+    unite(
+      entry, c(Category, Name),
+      sep = "_"
+    ) |>
+    # deleting Sub-category and Firm
+    select(-`Sub-category`, -Firm) |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(names_from = entry, values_from = Value) |>
+    # cleaning names
+    clean_names()
+}
+```
+
+```{r}
+kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
+```
+
+```{r}
+excel_base <- read_excel("Data/OPXfinans_final.xlsx")
+
+if (grepl("Category", names(excel_base))) {
+  TRUE
+}
+
+?grepl
+
+grepl("Category", names(excel_base))
+
+if ("Category" %in% names(excel_base)) {
+  TRUE
+}
+
+if ("2022" %in% names(kreft) && "2002" %in% names(kreft)) {
+  TRUE
+}
+getwd()
+blank <- read_excel("Data/Blank2.xlsx")
+
+if ("Year" %in% names(blank) && "Category" %in% names(blank)) {
+  TRUE
+} else {
+  FALSE
+}
+
+if ("Year" %in% names(excel_base) && "Category" %in% names(excel_base)) {
+  TRUE
+} else {
+  FALSE
+}
+
+if ("Year" %in% names(excel_base)) {
+  TRUE
+} else {
+  FALSE
+}
+
+?missing
+```
+
+```{r}
+read_proff_test <- function(file) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+}
+
+read_proff_complete <- function(file, start, end) {
+    if (as.character(all_of(start)) %in% names(file) && as.character(all_of(end)) %in% names(file)) {
+    # removing empty column, using all_of(argument)
+    file |> 
+      select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column
+    fill(
+      valutakode,
+      .direction = "updown"
+    ) |>
+    # removing dates and redundant currency
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+    } else {
+      message("Please ensure you have entered the correct start and end years")
+    }
+  }
+```
+
+```{r}
+html_test <- read_proff_test("Data/Kreftforeningen.html")
+
+html_test
+
+testing <- read_proff_complete(html_test, "2002", "2021")
+
+names(html_test)
+
+testing
+```

### 2022-07-27 Old/Shiny/Test/Data tests.Rmd
Testing for years and correct excel format tests.

--- 2022-07-27 Old/Shiny/Test/Data tests.Rmd
@@
 ```
+
+```{r}
+getwd()
+kreft <- read_proff("Data/Kreftforeningen.html", "2002", "2021")
+```
+
+```{r}
+widen_excel <- function(file) {
+  read_excel(file) |>
+    # uniting Category and Name to deal with multiple 'Other'
+    unite(
+      entry, c(Category, Name),
+      sep = "_"
+    ) |>
+    # deleting Sub-category and Firm
+    select(-`Sub-category`, -Firm) |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(names_from = entry, values_from = Value) |>
+    # cleaning names
+    clean_names()
+}
+```
+
+```{r}
+kreft_excel <- widen_excel("Data/OPXfinans_final.xlsx")
+```
+
+```{r}
+excel_base <- read_excel("Data/OPXfinans_final.xlsx")
+
+if (grepl("Category", names(excel_base))) {
+  TRUE
+}
+
+?grepl
+
+grepl("Category", names(excel_base))
+
+if ("Category" %in% names(excel_base)) {
+  TRUE
+}
+
+if ("2022" %in% names(kreft) && "2002" %in% names(kreft)) {
+  TRUE
+}
+getwd()
+blank <- read_excel("Data/Blank2.xlsx")
+
+if ("Year" %in% names(blank) && "Category" %in% names(blank)) {
+  TRUE
+} else {
+  FALSE
+}
+
+if ("Year" %in% names(excel_base) && "Category" %in% names(excel_base)) {
+  TRUE
+} else {
+  FALSE
+}
+
+if ("Year" %in% names(excel_base)) {
+  TRUE
+} else {
+  FALSE
+}
+
+?missing
+```
+
+```{r}
+read_proff_test <- function(file) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+}
+
+read_proff_complete <- function(object, start, end) {
+    if (as.character(all_of(start)) %in% names(object) && as.character(all_of(end)) %in% names(object)) {
+    # removing empty column, using all_of(argument)
+    object |> 
+      select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column
+    fill(
+      valutakode,
+      .direction = "updown"
+    ) |>
+    # removing dates and redundant currency
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+    } else {
+      message("Please ensure you have entered the correct start and end years")
+    }
+  }
+```
+
+```{r}
+html_test <- read_proff_test("Data/Kreftforeningen.html")
+
+html_test
+
+testing <- read_proff_complete(html_test, "2002", "2021")
+
+names(html_test)
+
+testing
+```

### 2022-07-27 Dashboard/Dashboard data tests.Rmd
Maybe a functional function?

--- 2022-07-27 Dashboard/NGO Analysis dashboard alpha.Rmd
@@
     ))
+
+    # Testing
+    html_test <- read_proff_test(input$html$datapath)
+
     # Creating data frame
-    read_proff(
-      input$html$datapath,
+    if (as.character(all_of(input$start)) %in% names(html_test) && as.character(all_of(input$end)) %in% names(html_test)) {
+    read_proff_complete(
+      html_test,
       as.character(input$start), as.character(input$end)
     )
+    } else {
+      validate("Please ensure you have entered the correct start and end years.")
+    }
   })
@@
 ```
+
+```{r new_data_tests}
+read_proff_test <- function(file) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+}
+
+read_proff_complete <- function(object, start, end) {
+    # removing empty column, using all_of(argument)
+    object |> 
+      select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column
+    fill(
+      valutakode,
+      .direction = "updown"
+    ) |>
+    # removing dates and redundant currency
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+  }
+```

### 2022-07-27 Dashboard/Dashboard data tests.Rmd
Deleting unneeded data creation.
More tests and testing data.

--- 2022-07-27 Dashboard/NGO Analysis dashboard alpha.Rmd
@@
 
-  # Data frames ================================================================
-
-  # creating data frame from html file
-  htmldata <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$html$datapath) == "html",
-      "Please upload an html file."
-    ))
-    # Creating data frame
-    read_proff(
-      input$html$datapath,
-      as.character(input$start), as.character(input$end)
-    )
-  })
+  # Data frame ================================================================
 
   # Creating data frame from Excel file
-  exceldata <- reactive({
+  excel_test <- reactive({
     # validating file
@@
     ))
-    # Creating data frame
-    widen_excel(input$excel$datapath)
+
+    widen_excel_test(input$excel$datapath)
+  })
+
+  exceldata <- reactive({
+    widen_excel(excel_test())
   })
@@
   NGO <- reactive({
-    # validating the file again
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
+    # checking for the required column again
+    if ("firm" %in% names(excel_test())) {
       # fetching the name
-    read_excel(input$excel$datapath) |>
+      excel_test() |>
         # Turning the needed column into a list
-      pull(Firm) |>
+        pull(firm) |>
         # discarding all NAs in case the name is somewhere
@@
         first()
+    } else {
+      validate("Please upload data using the specified Excel template.")
+    }
   })
@@
   # Data frames ================================================================
-  # Recreating, to make available to this chunk as well
 
@@
     ))
+
+    # Testing
+    html_test <- read_proff_test(input$html$datapath)
+
     # Creating data frame
-    read_proff(
-      input$html$datapath,
+
+    read_proff_complete(
+      html_test,
       as.character(input$start), as.character(input$end)
@@
 
+  # Recreating, to make available to this chunk as well
   # Creating data frame from Excel file
-  exceldata <- reactive({
+  excel_test <- reactive({
     # validating file
@@
       NA
-    } else {
-    # reading the Excel sheet anew
-    read_excel(input$excel$datapath) |>
+    } else if ("year" %in% names(excel_test()) && "category" %in% names(excel_test())) {
+      # starting with the unpivoted excel data
+      excel_test() |>
         # removing the firm and sub-category columns
-      select(-Firm, -`Sub-category`) |>
+        select(-firm, -sub_category) |>
         # filtering to retain only revenue and expenses entries
-      filter(Category == "Revenue" | Category == "Expenses") |>
+        filter(category == "Revenue" | category == "Expenses") |>
         # grouping and summarising to calculate maximum value
-      group_by(Year, Category) |>
+        group_by(year, category) |>
         summarise(
-        sum = sum(Value, na.rm = TRUE)
+          sum = sum(value, na.rm = TRUE)
         ) |>
@@
         max()
+    } else {
+      NA
     }
@@
 ## Function to read html file --------------------------------------------------
-read_proff <- function(file, start, end) {
-  # Specifying encoding to deal with occasional errors
-  read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame") |>
-    html_table() |>
-    # removing empty column, using all_of(argument)
-    select(
-      REGNSKAPSPERIODE:all_of(start)
-    ) |>
-    # removing white space
-    mutate(
-      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-    ) |>
-    # removing duplicate table sometimes created by read_html()
-    filter(
-      !row_number() > 176
-    ) |>
-    # removing duplicate rows
-    filter(
-      !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |>
-    # tidying data using pivot_longer() and all_of(argument)
-    pivot_longer(
-      all_of(end):all_of(start),
-      names_to = "year"
-    ) |>
-    # creating a currency code column and turning current currency codes into
-    # NAs
-    mutate(
-      # changing to real NAs
-      value = na_if(value, "-"),
-      # turning years into numbers
-      year = as.integer(year),
-      # creating a currency code column and turning current currency codes into
-      # NAs
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |>
-    # filling the currency column
-    fill(
-      valutakode,
-      .direction = "updown"
-    ) |>
-    # removing dates and redundant currency
-    filter(
-      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-        REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-    mutate(
-      # removing white space in numbers
-      value = str_replace_all(value, "\\s", ""),
-      # turning numbers into true numbers
-      value = as.numeric(value),
-      # removing values that are years
-      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-    ) |>
-    # removing any remaining duplicate rows
-    distinct() |>
-    # pivoting to create 'tidy' variables
-    pivot_wider(
-      names_from = REGNSKAPSPERIODE, values_from = value
-    ) |>
-    # tidying names
-    clean_names() |>
-    # removing headings from the table
-    select(
-      -lederlonn_i_hele_1000,
-      -resultatregnskap_i_hele_1000,
-      -balanseregnskap_i_hele_1000
-    ) |>
-    # making it clear what lonn refers to
-    rename(
-      lederlonn = lonn
-    ) |>
-    # Creating new variables for plots
-    mutate(
-      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-    ) |>
-    # arranging by year for practical display
-    arrange(desc(year))
-}
+# read_proff <- function(file, start, end) {
+#   # Specifying encoding to deal with occasional errors
+#   read_html(file, encoding = "UTF-8") |>
+#     html_node("#inner-frame") |>
+#     html_table() |>
+#     # removing empty column, using all_of(argument)
+#     select(
+#       REGNSKAPSPERIODE:all_of(start)
+#     ) |>
+#     # removing white space
+#     mutate(
+#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+#     ) |>
+#     # removing duplicate table sometimes created by read_html()
+#     filter(
+#       !row_number() > 176
+#     ) |>
+#     # removing duplicate rows
+#     filter(
+#       !grepl("Lukk", REGNSKAPSPERIODE)
+#     ) |>
+#     # tidying data using pivot_longer() and all_of(argument)
+#     pivot_longer(
+#       all_of(end):all_of(start),
+#       names_to = "year"
+#     ) |>
+#     # creating a currency code column and turning current currency codes into
+#     # NAs
+#     mutate(
+#       # changing to real NAs
+#       value = na_if(value, "-"),
+#       # turning years into numbers
+#       year = as.integer(year),
+#       # creating a currency code column and turning current currency codes into
+#       # NAs
+#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+#     ) |>
+#     # filling the currency column
+#     fill(
+#       valutakode,
+#       .direction = "updown"
+#     ) |>
+#     # removing dates and redundant currency
+#     filter(
+#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+#         REGNSKAPSPERIODE != "Valutakode"
+#     ) |>
+#     mutate(
+#       # removing white space in numbers
+#       value = str_replace_all(value, "\\s", ""),
+#       # turning numbers into true numbers
+#       value = as.numeric(value),
+#       # removing values that are years
+#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+#     ) |>
+#     # removing any remaining duplicate rows
+#     distinct() |>
+#     # pivoting to create 'tidy' variables
+#     pivot_wider(
+#       names_from = REGNSKAPSPERIODE, values_from = value
+#     ) |>
+#     # tidying names
+#     clean_names() |>
+#     # removing headings from the table
+#     select(
+#       -lederlonn_i_hele_1000,
+#       -resultatregnskap_i_hele_1000,
+#       -balanseregnskap_i_hele_1000
+#     ) |>
+#     # making it clear what lonn refers to
+#     rename(
+#       lederlonn = lonn
+#     ) |>
+#     # Creating new variables for plots
+#     mutate(
+#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+#     ) |>
+#     # arranging by year for practical display
+#     arrange(desc(year))
+# }
 
@@
 ## Function to read Excel file -------------------------------------------------
-widen_excel <- function(file) {
+widen_excel_test <- function(file) {
   read_excel(file) |> 
-    # uniting Category and Name to deal with multiple 'Other'
+    # cleaning names to enable later operations
+    clean_names()
+}
+
+widen_excel <- function(object) {
+  if ("year" %in% names(object) && "firm" %in% names(object) &&
+    "category" %in% names(object) && "sub_category" %in% names(object) &&
+    "name" %in% names(object) && "value" %in% names(object)) {
+    object |>
+      # uniting category and name to deal with multiple 'Other'
       unite(
-      entry, c(Category, Name),
+        entry, c(category, name),
         sep = "_"
@@
       # deleting Sub-category and Firm
-    select(-`Sub-category`, -Firm) |>
+      select(-sub_category, -firm) |>
       # pivoting to create 'tidy' variables
-    pivot_wider(names_from = entry, values_from = Value) |>
+      pivot_wider(names_from = entry, values_from = value) |>
       # cleaning names
       clean_names()
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
 }
@@
 ```
+
+```{r new_data_tests}
+# Part 1
+read_proff_test <- function(file) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+}
+
+# Part 2 - version with integrated test
+read_proff_complete <- function(object, start, end) {
+  if (as.character(all_of(start)) %in% names(object) &&
+    as.character(all_of(end)) %in% names(object)) {
+    object |>
+      # removing empty column, using all_of(argument)
+      select(
+        REGNSKAPSPERIODE:all_of(start)
+      ) |>
+      # removing white space
+      mutate(
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |>
+      # removing duplicate table sometimes created by read_html()
+      filter(
+        !row_number() > 176
+      ) |>
+      # removing duplicate rows
+      filter(
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |>
+      # tidying data using pivot_longer() and all_of(argument)
+      pivot_longer(
+        all_of(end):all_of(start),
+        names_to = "year"
+      ) |>
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      mutate(
+        # changing to real NAs
+        value = na_if(value, "-"),
+        # turning years into numbers
+        year = as.integer(year),
+        # creating a currency code column and turning current currency codes into
+        # NAs
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |>
+      # filling the currency column
+      fill(
+        valutakode,
+        .direction = "updown"
+      ) |>
+      # removing dates and redundant currency
+      filter(
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode"
+      ) |>
+      mutate(
+        # removing white space in numbers
+        value = str_replace_all(value, "\\s", ""),
+        # turning numbers into true numbers
+        value = as.numeric(value),
+        # removing values that are years
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+      ) |>
+      # removing any remaining duplicate rows
+      distinct() |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |>
+      # tidying names
+      clean_names() |>
+      # removing headings from the table
+      select(
+        -lederlonn_i_hele_1000,
+        -resultatregnskap_i_hele_1000,
+        -balanseregnskap_i_hele_1000
+      ) |>
+      # making it clear what lonn refers to
+      rename(
+        lederlonn = lonn
+      ) |>
+      # Creating new variables for plots
+      mutate(
+        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+      ) |>
+      # arranging by year for practical display
+      arrange(desc(year))
+  } else {
+    validate("Please ensure you have entered the correct start and end years")
+  }
+}
+```

### 2022-07-27 Old/Shiny/Test/Data tests.Rmd
Maybe a functional function?
More tests and testing data.

@@
 
-?missing
+names(excel_base)
+
+excel_base$`Sub-category`
+
+excel_tested <- read_excel("Data/Blank2.xlsx") |> 
+  clean_names()
+
+widen_excel_testing <- function(object) {
+  if ("Year" %in% names(object) && "Firm" %in% names(object) &&
+    "Category" %in% names(object) && "'Sub-category'" %in% names(object) &&
+    "Name" %in% names(object) && "Value" %in% names(object)) {
+    object |>
+      # uniting Category and Name to deal with multiple 'Other'
+      unite(
+        entry, c(Category, Name),
+        sep = "_"
+      ) |>
+      # deleting Sub-category and Firm
+      select(-`Sub-category`, -Firm) |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(names_from = entry, values_from = Value) |>
+      # cleaning names
+      clean_names()
+  } else {
+    message("Bad")
+  }
+}
+
+getwd()
+
+widen_excel_testing(excel_base)
+
+exceltest |> clean_names()
+getwd()
+blank <- read_excel("Data/Blank2.xlsx")
+names(blank)
+
+clean_names(blank)
 ```
@@
 ```
+```{r}
+?validate
+```
+
+```{r}
+read_proff_names <- function(file, start, end) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    names()
+}
+
+read_proff_names("Data/Kreftforeningen.html", "2002")
+
+nrk <- read_html("Data/NRK.html")
+nrk_node <- nrk |> html_node("#inner-frame")
+
+class(nrk_node)
+
+wrong <- read_html("Data/Kreft_wrong.html") |> html_node("#inner-frame") |> html_table()
+
+wrong
+
+right <- read_html("Data/Kreftforeningen.html", encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+
+right
+
+"REGNSKAPSPERIODE" %in% names(wrong)
+
+if (class(nrk_node) == "xml_missing") {
+  TRUE
+} else {
+  FALSE
+}
+
+read_proff_read <- function(file) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+}
+
+testing <- read_html("Data/Kreftforeningen.html", encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+class(testing)
+
+read_proff_table <- function(object) {
+  html_table()
+}
+
+empty <- read_proff_read("Data/Kreft_wrong.html") |> html_table()
+
+"REGNSKAPSPERIODE" %in% names(empty)
+
+read_proff_testing(testing, "2002", "2021")
+
+exceldata <- reactive({
+
+})
+
+```
+
+- test if html
+- test of node is present
+- test if correct table
+- go for it
+
+```{r}
+read_proff_test <- function(file) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+}
+
+# Part 2 - version with integrated test
+read_proff_complete <- function(object, start, end) {
+  if (as.character(all_of(start)) %in% names(object) &&
+    as.character(all_of(end)) %in% names(object)) {
+    object |>
+      # removing empty column, using all_of(argument)
+      select(
+        REGNSKAPSPERIODE:all_of(start)
+      ) |>
+      # removing white space
+      mutate(
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |>
+      # removing duplicate table sometimes created by read_html()
+      filter(
+        !row_number() > 176
+      ) |>
+      # removing duplicate rows
+      filter(
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |>
+      # tidying data using pivot_longer() and all_of(argument)
+      pivot_longer(
+        all_of(end):all_of(start),
+        names_to = "year"
+      ) |>
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      mutate(
+        # changing to real NAs
+        value = na_if(value, "-"),
+        # turning years into numbers
+        year = as.integer(year),
+        # creating a currency code column and turning current currency codes into
+        # NAs
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |>
+      # filling the currency column
+      fill(
+        valutakode,
+        .direction = "updown"
+      ) |>
+      # removing dates and redundant currency
+      filter(
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode"
+      ) |>
+      mutate(
+        # removing white space in numbers
+        value = str_replace_all(value, "\\s", ""),
+        # turning numbers into true numbers
+        value = as.numeric(value),
+        # removing values that are years
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+      ) |>
+      # removing any remaining duplicate rows
+      distinct() |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |>
+      # tidying names
+      clean_names() |>
+      # removing headings from the table
+      select(
+        -lederlonn_i_hele_1000,
+        -resultatregnskap_i_hele_1000,
+        -balanseregnskap_i_hele_1000
+      ) |>
+      # making it clear what lonn refers to
+      rename(
+        lederlonn = lonn
+      ) |>
+      # Creating new variables for plots
+      mutate(
+        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+      ) |>
+      # arranging by year for practical display
+      arrange(desc(year))
+  } else {
+    validate("Please ensure you have entered the correct start and end years")
+  }
+}
+```

### 2022-07-27 Dashboard/NGO Analysis dashboard alpha.Rmd
Commenting. Moving testing function.

@@
 
-# Button
+# Analyse button
 actionButton("click", "Analyse")
@@
 
-  # creating data frame from html file
+  # creating data frame from html file <- Can be deleted
   htmldata <- reactive({
@@
 }
+
+## Function to read Excel file -------------------------------------------------
+widen_excel <- function(file) {
+  read_excel(file) |>
+    # uniting Category and Name to deal with multiple 'Other'
+    unite(
+      entry, c(Category, Name),
+      sep = "_"
+    ) |>
+    # deleting Sub-category and Firm
+    select(-`Sub-category`, -Firm) |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(names_from = entry, values_from = Value) |>
+    # cleaning names
+    clean_names()
+}
+
+# Support functions ============================================================
+
+## KPI_to_string ---------------------------------------------------------------
+# Function to turn KPI values into NA or percentage string
+KPI_to_string <- function(variable) {
+  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
+  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
+  # avoid passing an NA to subsequent operations.)
+  if (is.na(variable)) {
+    NA
+  } else if (variable == Inf) {
+    NA
+  } else {
+    # Shortening (lengthening) the value, adding %
+    str_c(format(variable, digits = 3), "%")
+  }
+}
+
+## scale_fill_green ------------------------------------------------------------
+# Function with slightly nicer colours than default; ensuring uniformity; +
+# ensuring positive values are always green through a test for the presence of
+# FALSE values
+scale_fill_green <- function(variable) {
+  scale_fill_met_d("Tsimshian",
+    direction = ifelse(
+      mean(variable,
+        na.rm = TRUE
+      ) == 1, -1, 1
+    )
+  )
+}
+
+## no_information plot ---------------------------------------------------------
+# Function to print a 'no information' plot, if necessary
+no_information <- function() {
+  empty_plot("No information available")
+}
+
+## bad_data plot ---------------------------------------------------------------
+# Function to print a 'bad data' plot, if necessary
+bad_data <- function() {
+  empty_plot("Something is wrong\nwith the Excel data.")
+}
+
+## empty_plot-------------------------------------------------------------------
+# Function to print an empty plot with a supplied message
+empty_plot <- function(warning) {
+  ggplot() +
+    # text geom with a space to create the plot, but otherwise show nothing
+    geom_text(aes(0, 0, label = " ")) +
+    # as blank a slate as possible
+    theme_void() +
+    annotate(
+      "text",
+      label = warning,
+      # Using the same palette as the other non-profit/loss plots
+      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+    )
+}
+
+# Testing arguments ============================================================
+# One or more of the elements of this mutate() operation may be inserted at the
+# end of the read_proff() function to create various data problems for some
+# quick and dirty testing
+
+# mutate(
+#       valutakode = ifelse(year == 2019, "USD", valutakode),
+#       leder_annen_godtgjorelse = NA,
+#       lederlonn = NA,
+#       arsresultat = NA,
+#       sum_egenkapital_og_gjeld = NA
+#     ) |>
 
@@
 # }
-
-## Function to read Excel file -------------------------------------------------
-widen_excel <- function(file) {
-  read_excel(file) |>
-    # uniting Category and Name to deal with multiple 'Other'
-    unite(
-      entry, c(Category, Name),
-      sep = "_"
-    ) |>
-    # deleting Sub-category and Firm
-    select(-`Sub-category`, -Firm) |>
-    # pivoting to create 'tidy' variables
-    pivot_wider(names_from = entry, values_from = Value) |>
-    # cleaning names
-    clean_names()
-}
-
-# Support functions ============================================================
-
-## KPI_to_string ---------------------------------------------------------------
-# Function to turn KPI values into NA or percentage string
-KPI_to_string <- function(variable) {
-  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
-  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
-  # avoid passing an NA to subsequent operations.)
-  if (is.na(variable)) {
-    NA
-  } else if (variable == Inf) {
-    NA
-  } else {
-    # Shortening (lengthening) the value, adding %
-    str_c(format(variable, digits = 3), "%")
-  }
-}
-
-## scale_fill_green ------------------------------------------------------------
-# Function with slightly nicer colours than default; ensuring uniformity; +
-# ensuring positive values are always green through a test for the presence of
-# FALSE values
-scale_fill_green <- function(variable) {
-  scale_fill_met_d("Tsimshian",
-    direction = ifelse(
-      mean(variable,
-        na.rm = TRUE
-      ) == 1, -1, 1
-    )
-  )
-}
-
-## no_information plot ---------------------------------------------------------
-# Function to print a 'no information' plot, if necessary
-no_information <- function() {
-  empty_plot("No information available")
-}
-
-## bad_data plot ---------------------------------------------------------------
-# Function to print a 'bad data' plot, if necessary
-bad_data <- function() {
-  empty_plot("Something is wrong\nwith the Excel data.")
-}
-
-## empty_plot-------------------------------------------------------------------
-# Function to print an empty plot with a supplied message
-empty_plot <- function(warning) {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    annotate(
-      "text",
-      label = warning,
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
-}
-
-# Testing arguments ============================================================
-# One or more of the elements of this mutate() operation may be inserted at the
-# end of the read_proff() function to create various data problems for some
-# quick and dirty testing
-
-# mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
-#       leder_annen_godtgjorelse = NA,
-#       lederlonn = NA,
-#       arsresultat = NA,
-#       sum_egenkapital_og_gjeld = NA
-#     ) |>
-```
+```

### 2022-07-27 Old/Shiny/Test/Data tests.Rmd
Spelling.
Maybe found the function.

--- 2022-07-27 Old/Shiny/Test/Data tests.Rmd
@@
 ```
+```{r}
+?validate
+```
+
+```{r}
+read_proff_names <- function(file, start, end) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    names()
+}
+
+read_proff_names("Data/Kreftforeningen.html", "2002")
+
+nrk <- read_html("Data/NRK.html")
+nrk_node <- nrk |> html_node("#inner-frame")
+
+class(nrk_node)
+
+wrong <- read_html("Data/Kreft_wrong.html") |> html_node("#inner-frame") |> html_table()
+
+wrong
+
+right <- read_html("Data/Kreftforeningen.html", encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+
+right
+
+"REGNSKAPSPERIODE" %in% names(wrong)
+
+if (class(nrk_node) == "xml_missing") {
+  TRUE
+} else {
+  FALSE
+}
+
+read_proff_read <- function(file) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+}
+
+testing <- read_html("Data/Kreftforeningen.html", encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+class(testing)
+
+read_proff_table <- function(object) {
+  html_table()
+}
+
+empty <- read_proff_read("Data/Kreft_wrong.html") |> html_table()
+
+"REGNSKAPSPERIODE" %in% names(empty)
+
+read_proff_testing(testing, "2002", "2021")
+
+htmldata <- function(file, start, end) {
+  node <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  if (class(node) == "xml_node") {
+    table <- node |> html_table()
+  } else {
+    message("Table error.")
+  }
+
+  if ("REGNSKAPSPERIODE" %in% names(table)) {
+    if (as.character(all_of(start)) %in% names(table) &&
+    as.character(all_of(end)) %in% names(table)) {
+      read_proff_notest("Data/Kreftforeningen.html", start, end)
+    } else {
+      message("character error")
+    }
+  } else {
+    message("Regnskap error.")
+  }
+}
+
+htmldata("Data/Kreftforeningen.html", "2002", "2021")
+
+weird <- read_html("Data/Kreftforeningen.html", encoding = "UTF-8") |> 
+  html_element("#inner-frame") |> html_table()
+
+names(weird)
+
+"2002" %in% names(weird)
+
+read_test <- function(object, start, end) {
+  if (as.character(all_of(start)) %in% names(object) &&
+    as.character(all_of(end)) %in% names(object)) {
+    TRUE
+  } else {
+    FALSE
+  }
+}
+
+html_first_test <- function(file) {
+
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |> html_table()
+  } else {
+    "table error"
+  }
+
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    "regnskap error"
+  }
+}
+
+html_first_test("Data/Kreftforeningen.html")
+
+html_second_test <- function(object, start, end) {
+  if (as.character(all_of(start)) %in% names(object) &&
+    as.character(all_of(end)) %in% names(object)) {
+    read_proff_notest(object, start, end)
+  } else {
+    FALSE
+  }
+}
+
+html_second_test(htmlready, "2002", "2021")
+
+read_test(weird, "2002", "2021")
+
+html_testing <- function(file, start, end) {
+
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |> html_table()
+  } else {
+    "table error"
+  }
+
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    "regnskap error"
+  }
+
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    read_proff_notest(file, start, end)
+  } else {
+    "year error"
+  }
+}
+
+html_testing("Data/Kreftforeningen.html", "2002", "2021")
+
+check
+
+if (check == TRUE) {
+  read_proff_notest("Data/Kreftforeningen.html", "2002", "2021")
+} else {
+  "what's wrong"
+}
+```
+
+- test if html
+- test if node is present
+- test if correct table
+- go for it
+
+```{r}
+read_proff_test <- function(file) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table()
+}
+
+# Part 2 - version with integrated test
+read_proff_complete <- function(object, start, end) {
+  if (as.character(all_of(start)) %in% names(object) &&
+    as.character(all_of(end)) %in% names(object)) {
+    object |>
+      # removing empty column, using all_of(argument)
+      select(
+        REGNSKAPSPERIODE:all_of(start)
+      ) |>
+      # removing white space
+      mutate(
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |>
+      # removing duplicate table sometimes created by read_html()
+      filter(
+        !row_number() > 176
+      ) |>
+      # removing duplicate rows
+      filter(
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |>
+      # tidying data using pivot_longer() and all_of(argument)
+      pivot_longer(
+        all_of(end):all_of(start),
+        names_to = "year"
+      ) |>
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      mutate(
+        # changing to real NAs
+        value = na_if(value, "-"),
+        # turning years into numbers
+        year = as.integer(year),
+        # creating a currency code column and turning current currency codes into
+        # NAs
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |>
+      # filling the currency column
+      fill(
+        valutakode,
+        .direction = "updown"
+      ) |>
+      # removing dates and redundant currency
+      filter(
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode"
+      ) |>
+      mutate(
+        # removing white space in numbers
+        value = str_replace_all(value, "\\s", ""),
+        # turning numbers into true numbers
+        value = as.numeric(value),
+        # removing values that are years
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+      ) |>
+      # removing any remaining duplicate rows
+      distinct() |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |>
+      # tidying names
+      clean_names() |>
+      # removing headings from the table
+      select(
+        -lederlonn_i_hele_1000,
+        -resultatregnskap_i_hele_1000,
+        -balanseregnskap_i_hele_1000
+      ) |>
+      # making it clear what lonn refers to
+      rename(
+        lederlonn = lonn
+      ) |>
+      # Creating new variables for plots
+      mutate(
+        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+      ) |>
+      # arranging by year for practical display
+      arrange(desc(year))
+  } else {
+    validate("Please ensure you have entered the correct start and end years")
+  }
+}
+
+read_proff_notest <- function(file, start, end) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |> 
+      # removing empty column, using all_of(argument)
+      select(
+        REGNSKAPSPERIODE:all_of(start)
+      ) |>
+      # removing white space
+      mutate(
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |>
+      # removing duplicate table sometimes created by read_html()
+      filter(
+        !row_number() > 176
+      ) |>
+      # removing duplicate rows
+      filter(
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |>
+      # tidying data using pivot_longer() and all_of(argument)
+      pivot_longer(
+        all_of(end):all_of(start),
+        names_to = "year"
+      ) |>
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      mutate(
+        # changing to real NAs
+        value = na_if(value, "-"),
+        # turning years into numbers
+        year = as.integer(year),
+        # creating a currency code column and turning current currency codes into
+        # NAs
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |>
+      # filling the currency column
+      fill(
+        valutakode,
+        .direction = "updown"
+      ) |>
+      # removing dates and redundant currency
+      filter(
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode"
+      ) |>
+      mutate(
+        # removing white space in numbers
+        value = str_replace_all(value, "\\s", ""),
+        # turning numbers into true numbers
+        value = as.numeric(value),
+        # removing values that are years
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+      ) |>
+      # removing any remaining duplicate rows
+      distinct() |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |>
+      # tidying names
+      clean_names() |>
+      # removing headings from the table
+      select(
+        -lederlonn_i_hele_1000,
+        -resultatregnskap_i_hele_1000,
+        -balanseregnskap_i_hele_1000
+      ) |>
+      # making it clear what lonn refers to
+      rename(
+        lederlonn = lonn
+      ) |>
+      # Creating new variables for plots
+      mutate(
+        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+      ) |>
+      # arranging by year for practical display
+      arrange(desc(year))
+}
+```
+
+```{r}
+# This may be the function =====================================================
+```

### 2022-07-28 Dashboard/Dashboard data tests.Rmd
Commenting, deleting sep in Excel function in data tests.

@@
 
-## Function to read Excel file -------------------------------------------------
+## Functions to read Excel file ------------------------------------------------
 widen_excel_test <- function(file) {
@@
       unite(
-        entry, c(category, name),
-        sep = "_"
+        entry, c(category, name)
       ) |>

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha.Rmd
Commenting, deleting sep in Excel function in data tests.

@@
 ```{r setup, include=FALSE}
-# Loading packages using pacman to ensure they're all installed
+# Loading packages using pacman to ensure they're all installed and avoiding
+# catastrophic dashboard errors if they're not installed
 if (!require("pacman")) install.packages("pacman")
@@
     ) |>
-    # removing white space
+    # removing text white space
     mutate(

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha.Rmd
Adding further Excel testing.

@@
 ```{r sidebar_button}
-# download button
+# download button - using uiOutput() because of an apparent bug
 uiOutput("downloadUI")
@@
 
+## Download button -------------------------------------------------------------
 output$downloadUI <- renderUI({
@@
 
-## Download and parameters -----------------------------------------------------
+## Report download and parameters ----------------------------------------------s
 output$report <- downloadHandler(
@@
 
-  # Data frames ================================================================
-
-  # creating data frame from html file <- Can be deleted
-  htmldata <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$html$datapath) == "html",
-      "Please upload an html file."
-    ))
-    # Creating data frame
-    read_proff(
-      input$html$datapath,
-      as.character(input$start), as.character(input$end)
-    )
-  })
-
-  # Creating data frame from Excel file
-  exceldata <- reactive({
+  # Excel data frame ===========================================================
+
+  ## Validating and creating long version of Excel file, if possible------------
+  excel_test <- reactive({
     # validating file
@@
     ))
-    # Creating data frame
-    widen_excel(input$excel$datapath)
-  })
+
+    widen_excel_test(input$excel$datapath)
+  })
+
+  ## Creating the primary Excel data frame--------------------------------------
+  exceldata <- reactive({
+    widen_excel(excel_test())
+  })
+
+  # # Creating data frame from Excel file
+  # exceldata <- reactive({
+  #   # validating file
+  #   validate(need(
+  #     tools::file_ext(input$excel$datapath) == "xlsx",
+  #     "Please upload an Excel file."
+  #   ))
+  #   # Creating data frame
+  #   widen_excel(input$excel$datapath)
+  # })
 
@@
   })
+
+  # NGO <- reactive({
+  #   # validating the file again
+  #   validate(need(
+  #     tools::file_ext(input$excel$datapath) == "xlsx",
+  #     "Please upload an Excel file."
+  #   ))
+  #   # fetching the name
+  #   read_excel(input$excel$datapath) |>
+  #     # Turning the needed column into a list
+  #     pull(Firm) |>
+  #     # discarding all NAs in case the name is somewhere
+  #     purrr::discard(is.na) |>
+  #     # choosing the first available name
+  #     first()
+  # })
+  # 
+  # title <- reactive({
+  #   # testing if NGO() is NA
+  #   if (is.na(NGO())) {
+  #     # Printing warning if NA
+  #     "Name of NGO missing"
+  #     # Printing name of NGO
+  #   } else {
+  #     NGO()
+  #   }
+  # })
 
@@
   # Data frames ================================================================
-  # Recreating, to make available to this chunk as well
-
-  # creating data frame from html file
+  ## creating data frame from html file-----------------------------------------
   htmldata <- reactive({
@@
 
-  # Creating data frame from Excel file
-  exceldata <- reactive({
+  # Recreating, to make available to this chunk as well
+  ## Validating and creating long version of Excel file, if possible------------
+  excel_test <- reactive({
     # validating file
@@
     ))
-    # Creating data frame
-    widen_excel(input$excel$datapath)
+
+    widen_excel_test(input$excel$datapath)
+  })
+
+  ## Creating the primary Excel data frame--------------------------------------
+  exceldata <- reactive({
+    widen_excel(excel_test())
   })
@@
 
-## Function to read Excel file -------------------------------------------------
-widen_excel <- function(file) {
+## Functions to read Excel file ------------------------------------------------
+widen_excel_test <- function(file) {
+  # Reading Excel file
   read_excel(file) |> 
-    # uniting Category and Name to deal with multiple 'Other'
+    # cleaning names to enable later operations
+    clean_names()
+}
+
+# Function to create primary Excel data frame
+widen_excel <- function(object) {
+  # testing if all columns are present
+  if (setequal(
+    names(object),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  )) {
+    object |>
+      # uniting category and name to deal with multiple 'Other'
       unite(
-      entry, c(Category, Name),
-      sep = "_"
+        entry, c(category, name)
       ) |>
       # deleting Sub-category and Firm
-    select(-`Sub-category`, -Firm) |>
+      select(-sub_category, -firm) |>
       # pivoting to create 'tidy' variables
-    pivot_wider(names_from = entry, values_from = Value) |>
+      pivot_wider(names_from = entry, values_from = value) |>
       # cleaning names
       clean_names()
-}
+    # Notfification if all columns are not present
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
+}
+
+# widen_excel <- function(file) {
+#   read_excel(file) |>
+#     # uniting Category and Name to deal with multiple 'Other'
+#     unite(
+#       entry, c(Category, Name),
+#       sep = "_"
+#     ) |>
+#     # deleting Sub-category and Firm
+#     select(-`Sub-category`, -Firm) |>
+#     # pivoting to create 'tidy' variables
+#     pivot_wider(names_from = entry, values_from = Value) |>
+#     # cleaning names
+#     clean_names()
+# }
 
@@
 #     ) |>
-
-### Test version of function that creates multiple data problems----------------
-# read_proff <- function(file, start, end) {
-#   read_html(file, encoding = "UTF-8") |>
-#     html_node("#inner-frame") |>
-#     html_table() |>
-#     select( # removing empty column
-#       REGNSKAPSPERIODE:all_of(start)
-#     ) |> # using all_of()
-#     mutate( # removing whitespace
-#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-#     ) |>
-#     filter( # removing duplicate table - no idea why it's necessary
-#       !row_number() > 176
-#     ) |>
-#     filter( # removing duplicate rows
-#       !grepl("Lukk", REGNSKAPSPERIODE)
-#     ) |>
-#     pivot_longer( # tidying data
-#       all_of(end):all_of(start),
-#       names_to = "year"
-#     ) |> # using all_of()
-#     mutate( # changing to real NAs and turning years into numbers
-#       value = na_if(value, "-"),
-#       year = as.integer(year),
-#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-#     ) |>
-#     fill( # filling the currency column
-#       valutakode,
-#       .direction = "updown"
-#     ) |>
-#     filter( # removing dates and redundant currency
-#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-#         REGNSKAPSPERIODE != "Valutakode"
-#     ) |>
-#     mutate( # removing whitespace in numbers
-#       value = str_replace_all(value, "\\s", ""),
-#       value = as.numeric(value),
-#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-#     ) |>
-#     distinct() |> # removing any remaining duplicate rows
-#     pivot_wider( # pivoting
-#       names_from = REGNSKAPSPERIODE, values_from = value
-#     ) |>
-#     clean_names() |> # tidying names
-#     select(
-#       -lederlonn_i_hele_1000,
-#       -resultatregnskap_i_hele_1000,
-#       -balanseregnskap_i_hele_1000
-#     ) |> # removing headings
-#     rename( # making it clear what lonn refers to
-#       lederlonn = lonn
-#     ) |>
-#     # Creating new variables for plots
-#     mutate(
-#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE),
-#     ) |>
-#     # bad test - REMOVE
-#     mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
-#       leder_annen_godtgjorelse = NA,
-#       lederlonn = NA,
-#       arsresultat = NA,
-#       sum_egenkapital_og_gjeld = NA
-#     ) |>
-#     arrange(desc(year)) # arranging by year
-# }
-```
+```

### 2022-07-28 Dashboard/Dashboard data tests.Rmd
Discovered setequal().

@@
 widen_excel_test <- function(file) {
+  # Reading Excel file
   read_excel(file) |> 
@@
 widen_excel <- function(object) {
-  if ("year" %in% names(object) && "firm" %in% names(object) &&
-    "category" %in% names(object) && "sub_category" %in% names(object) &&
-    "name" %in% names(object) && "value" %in% names(object)) {
+  # testing if all columns are present
+  if (setequal(names(object),
+               c("year", "firm", "category", "sub_category", "name", "value"))) {
     object |>

### 2022-07-28 Old/Shiny/Test/Data tests.Rmd
Actually writing the function.
setequal() testing.

@@
 clean_names(blank)
+
+getwd()
+excel_logic <- read_excel("Data/OPXfinans_final.xlsx") |> clean_names()
+
+excel_logic
+
+excel_names <- names(excel_logic)
+
+excel_names == c("year", "firm", "category", "sub_category", "name", "value")
+
+class(excel_names)
+
+if (excel_names == c("year", "firm", "category", "sub_category", "name", "value")) {
+  "yes"
+} else {
+  "no"
+}
+
+setequal(excel_names, c("year", "firm", "category", "sub_category", "name", "value"))
 ```
@@
 # This may be the function =====================================================
+
+html_testing <- function(file, start, end) {
+
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |> html_table()
+  } else {
+    "table error"
+  }
+
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    "regnskap error"
+  }
+
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    read_proff_notest(file, start, end)
+  } else {
+    "year error"
+  }
+}
 ```

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha.Rmd
Spelling.

@@
   # Data frames ================================================================
-  ## creating data frame from html file-----------------------------------------
+  ## Creating data frame from html file-----------------------------------------
   htmldata <- reactive({

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002.Rmd
New version.

--- 2022-07-28 Dashboard/NGO Analysis dashboard alpha.Rmd
@@
 # Loading packages using pacman to ensure they're all installed and avoiding
-# catastrophic dashboard errors if they're not installed
+# catastrophic dashboard errors if they're not available
 if (!require("pacman")) install.packages("pacman")
@@
   })
-
-  # # Creating data frame from Excel file
-  # exceldata <- reactive({
-  #   # validating file
-  #   validate(need(
-  #     tools::file_ext(input$excel$datapath) == "xlsx",
-  #     "Please upload an Excel file."
-  #   ))
-  #   # Creating data frame
-  #   widen_excel(input$excel$datapath)
-  # })
 
@@
   })
-
-  # NGO <- reactive({
-  #   # validating the file again
-  #   validate(need(
-  #     tools::file_ext(input$excel$datapath) == "xlsx",
-  #     "Please upload an Excel file."
-  #   ))
-  #   # fetching the name
-  #   read_excel(input$excel$datapath) |>
-  #     # Turning the needed column into a list
-  #     pull(Firm) |>
-  #     # discarding all NAs in case the name is somewhere
-  #     purrr::discard(is.na) |>
-  #     # choosing the first available name
-  #     first()
-  # })
-  # 
-  # title <- reactive({
-  #   # testing if NGO() is NA
-  #   if (is.na(NGO())) {
-  #     # Printing warning if NA
-  #     "Name of NGO missing"
-  #     # Printing name of NGO
-  #   } else {
-  #     NGO()
-  #   }
-  # })
 
@@
     ))
-    # Creating data frame
-    read_proff(
-      input$html$datapath,
-      as.character(input$start), as.character(input$end)
-    )
+
+    # Testing and creating data frame
+    html_test_create(input$html$datapath,
+                 as.character(input$start), as.character(input$end))
   })
@@
 
-## Function to read html file --------------------------------------------------
+## Functions to read html file -------------------------------------------------
+html_test_create <- function(file, start, end) {
+  # Reading file and node
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  # Checking if node exists
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |>
+      html_table()
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Basic check if the table is correct
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Testing if the correct years have been entered
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    # Creating the data frame
+    read_proff(file, start, end)
+  } else {
+    validate("Please ensure you have entered the correct start and end years.")
+  }
+}
+
 read_proff <- function(file, start, end) {
-  # Specifying encoding to deal with occasional errors
   read_html(file, encoding = "UTF-8") |>
@@
       ) |>
-    # removing text white space
+      # removing white space
       mutate(
@@
 }
-
-# widen_excel <- function(file) {
-#   read_excel(file) |>
-#     # uniting Category and Name to deal with multiple 'Other'
-#     unite(
-#       entry, c(Category, Name),
-#       sep = "_"
-#     ) |>
-#     # deleting Sub-category and Firm
-#     select(-`Sub-category`, -Firm) |>
-#     # pivoting to create 'tidy' variables
-#     pivot_wider(names_from = entry, values_from = Value) |>
-#     # cleaning names
-#     clean_names()
-# }
 

### 2022-07-28 Old/Shiny/Test/Dashboard data tests.Rmd
Moving old and testing documents.

--- 2022-07-28 Dashboard/Dashboard data tests.Rmd
@@
 
-    # Testing
-    html_test <- read_proff_test(input$html$datapath)
-
-    # Creating data frame
-
-    read_proff_complete(
-      html_test,
-      as.character(input$start), as.character(input$end)
-    )
+    # Testing and creating data frame
+    html_test_create(input$html$datapath,
+                 as.character(input$start), as.character(input$end))
   })
@@
 ```{r new_data_tests}
-# Part 1
-read_proff_test <- function(file) {
-  # Specifying encoding to deal with occasional errors
+html_test_create <- function(file, start, end) {
+  # Reading file and node
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  # Checking if node exists
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |>
+      html_table()
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Basic check if the table is correct
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Testing if the correct years have been entered
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    # Creating the data frame
+    read_proff(file, start, end)
+  } else {
+    validate("Please ensure you have entered the correct start and end years.")
+  }
+}
+
+read_proff <- function(file, start, end) {
   read_html(file, encoding = "UTF-8") |>
     html_node("#inner-frame") |>
-    html_table()
-}
-
-# Part 2 - version with integrated test
-read_proff_complete <- function(object, start, end) {
-  if (as.character(all_of(start)) %in% names(object) &&
-    as.character(all_of(end)) %in% names(object)) {
-    object |>
+    html_table() |> 
       # removing empty column, using all_of(argument)
@@
       arrange(desc(year))
-  } else {
-    validate("Please ensure you have entered the correct start and end years")
-  }
-}
-```
+}
+```

### 2022-07-28 Old/Shiny/Test/HTML tests.Rmd
HTML function testing.

--- 2022-07-28 Old/Shiny/Test/Dashboard data tests.Rmd
@@
 ---
-title: "NGO analysis"
-output: 
-  flexdashboard::flex_dashboard:
-    orientation: rows
-    vertical_layout: scroll
-    theme: lumen
-runtime: shiny
+title: "HTML tests"
+author: "Torbjørn Skinnemoen Ottersen"
+date: "2022-07-28"
+output: html_document
 ---
@@
 ```{r setup, include=FALSE}
-# Loading packages using pacman to ensure they're all installed
-if (!require("pacman")) install.packages("pacman")
-library(pacman)
-p_load(
-  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-  flexdashboard, MetBrewer, fontawesome, htmltools
-)
-# Using the fontawesome package for more up-to-date (and currently supported)
-# Fontawesome icons
-htmltools::tagList(fontawesome::fa_html_dependency())
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-Sidebar {.sidebar}
-=======================================================================
+```{r}
+# Part 1
+read_proff_read <- function(file) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+}
 
-### File upload
-
-Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.
-
-Then upload an Excel file in the set format.
-
-Click 'Analyse' to analyse the data.
-
-```{r sidebar}
-# HTML
-fileInput("html", label = "Upload html file", accept = ".html")
-
-# Year inputs
-numericInput("start", label = "Start", value = 2002)
-numericInput("end", label = "End", value = 2021)
-
-# Excel
-fileInput("excel", "Upload Excel file", accept = ".xlsx")
-
-# Button
-actionButton("click", "Analyse")
-```
-
-If desired, a copy of the analysis in PDF format may then be downloaded.
-
-```{r sidebar_button}
-# download button
-uiOutput("downloadUI")
-```
-
-Dashboard
-=======================================================================
-
-Row 
------------------------------------------------------------------------
-
-### KPI Fundraising
-```{r}
-valueBoxOutput("KPI_box_fundraising")
-```
-
-### KPI OPX
-```{r}
-valueBoxOutput("KPI_box_OPX")
-```
-
-### KPI Private growth
-```{r}
-valueBoxOutput("KPI_box_private")
-```
-
-### KPI Public growth
-```{r}
-valueBoxOutput("KPI_box_public")
-```
-
-### KPI Private/public proportion
-```{r}
-valueBoxOutput("KPI_box_proportion")
-```
-
-Row
------------------------------------------------------------------------
-
-### Net income
-```{r}
-plotlyOutput("arsresultat_plotly")
-```
-
-### Total equity/liabilities
-```{r}
-plotlyOutput("balance_plotly")
-```
-
-Row 
------------------------------------------------------------------------
-
-### Revenues
-```{r}
-plotlyOutput("revenue_plotly")
-```
-
-### Expenses
-```{r}
-plotlyOutput("expenses_plotly")
-```
-
-Row 
------------------------------------------------------------------------
-
-### Transparency: CEO compensation
-```{r}
-plotlyOutput("CEO_plotly")
-```
-
-Details
-=======================================================================
-
-Row {data-height=60}
------------------------------------------------------------------------
-
-### NGO analysed
-
-```{r}
-# Using htmlOutput() to use html styling
-htmlOutput("title")
-```
-
-Row 
------------------------------------------------------------------------
-
-### Key Performance Indicators
-
-The KPIs are based on the following calculations, using data derived from the Excel sheet:
-
-- **Fundraising ROI**:
-  - A basic measure of the return on investment of direct fundraising expenses:
-
-  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
-
-- **Liquid assets/obligations**:
-  - A measure of the relationship between the organisation's liquid assets and its obligations:
-
-  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
-
-- **Private donations growth**:
-  - A calculation of the percentage change in private contributions based on the two most recent years:
-
-  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
-
-- **Public grants growth**:
-  - A calculation of the percentage change in public grants received based on the two most recent years:
-
-$$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
-
-- **Private/public proportion**:
-  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
-
-$$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
-
-Row 
------------------------------------------------------------------------
-
-### Graphs
-
-The dashboard includes five graphs, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.
-
-- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
-- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
-- **Revenues**: Based on the category *Revenue* in the Excel sheet.
-- **Expenses**: Based on the category *Expenses* in the Excel sheet.
-- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
-  - A common measure of transparency, but should normally be evaluated in context.
-
-```{r report, include=FALSE}
-# Shiny actions begin here =====================================================
-
-# Report Output ================================================================
-
-output$downloadUI <- renderUI({
-  downloadButton("report", "Download report")
-})
-
-## Download and parameters -----------------------------------------------------
-output$report <- downloadHandler(
-
-  # Based on code in Mastering Shiny
-  filename = "report.pdf",
-  content = function(file) {
-    tempReport <- file.path(tempdir(), "report.Rmd")
-    file.copy("report.Rmd", tempReport, overwrite = TRUE)
-
-    # parameters
-    params <- list(
-      htmlfile = input$html$datapath,
-      excelfile = input$excel$datapath,
-      start = input$start,
-      end = input$end
-    )
-
-    # notification
-    id <- showNotification(
-      "Generating report...",
-      duration = NULL,
-      closeButton = FALSE
-    )
-    on.exit(removeNotification(id), add = TRUE)
-
-    # R Markdown render
-    rmarkdown::render(tempReport,
-      output_file = file,
-      params = params,
-      envir = new.env(parent = globalenv())
-    )
+# Part 2 - version with integrated test
+read_proff_complete <- function(object, start, end) {
+  if (as.character(all_of(start)) %in% names(object) &&
+    as.character(all_of(end)) %in% names(object)) {
+    object |>
+      # removing empty column, using all_of(argument)
+      select(
+        REGNSKAPSPERIODE:all_of(start)
+      ) |>
+      # removing white space
+      mutate(
+        REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+      ) |>
+      # removing duplicate table sometimes created by read_html()
+      filter(
+        !row_number() > 176
+      ) |>
+      # removing duplicate rows
+      filter(
+        !grepl("Lukk", REGNSKAPSPERIODE)
+      ) |>
+      # tidying data using pivot_longer() and all_of(argument)
+      pivot_longer(
+        all_of(end):all_of(start),
+        names_to = "year"
+      ) |>
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      mutate(
+        # changing to real NAs
+        value = na_if(value, "-"),
+        # turning years into numbers
+        year = as.integer(year),
+        # creating a currency code column and turning current currency codes into
+        # NAs
+        valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+      ) |>
+      # filling the currency column
+      fill(
+        valutakode,
+        .direction = "updown"
+      ) |>
+      # removing dates and redundant currency
+      filter(
+        REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+          REGNSKAPSPERIODE != "Valutakode"
+      ) |>
+      mutate(
+        # removing white space in numbers
+        value = str_replace_all(value, "\\s", ""),
+        # turning numbers into true numbers
+        value = as.numeric(value),
+        # removing values that are years
+        value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+      ) |>
+      # removing any remaining duplicate rows
+      distinct() |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(
+        names_from = REGNSKAPSPERIODE, values_from = value
+      ) |>
+      # tidying names
+      clean_names() |>
+      # removing headings from the table
+      select(
+        -lederlonn_i_hele_1000,
+        -resultatregnskap_i_hele_1000,
+        -balanseregnskap_i_hele_1000
+      ) |>
+      # making it clear what lonn refers to
+      rename(
+        lederlonn = lonn
+      ) |>
+      # Creating new variables for plots
+      mutate(
+        annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+        balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+      ) |>
+      # arranging by year for practical display
+      arrange(desc(year))
+  } else {
+    validate("Please ensure you have entered the correct start and end years")
   }
-)
-```
-
-```{r reactive_title_value_boxes, include=FALSE}
-# First reactive chunk =========================================================
-
-# wait for button click
-observe({
-
-  # Data frame ================================================================
-
-  # Creating data frame from Excel file
-  excel_test <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
-
-    widen_excel_test(input$excel$datapath)
-  })
-
-  exceldata <- reactive({
-    widen_excel(excel_test())
-  })
-
-  # NGO title ==================================================================
-  NGO <- reactive({
-    # checking for the required column again
-    if ("firm" %in% names(excel_test())) {
-      # fetching the name
-      excel_test() |>
-        # Turning the needed column into a list
-        pull(firm) |>
-        # discarding all NAs in case the name is somewhere
-        purrr::discard(is.na) |>
-        # choosing the first available name
-        first()
-    } else {
-      validate("Please upload data using the specified Excel template.")
 }
-  })
-
-  title <- reactive({
-    # testing if NGO() is NA
-    if (is.na(NGO())) {
-      # Printing warning if NA
-      "Name of NGO missing"
-      # Printing name of NGO
-    } else {
-      NGO()
-    }
-  })
-
-  # Creating NGO title using html formatting (colour based on MetBrewer
-  # "Hokusai2")
-  output$title <- renderText({
-    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
-  })
-
-  # Value Boxes ================================================================
-
-  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark" if
-  # the value is NA
-
-  ## Fundraising ROI------------------------------------------------------------
-  # Creating Fundraising ROI value
-  KPI_fundraising_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # most recent year
-        filter(year == max(year)) |>
-        transmute(
-          # Calculating the value as a percentage; removing all other
-          KPI_fundraising_num =
-            (sum(revenue_private_contributions,
-              revenue_membership_fees,
-              na.rm = TRUE
-            ) - expenses_fundraising) /
-              expenses_fundraising * 100
-        ) |>
-        # just want the single number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for fundraising ROI
-  # Unsure about colour levels
-  output$KPI_box_fundraising <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_fundraising_num()),
-      caption = "Fundraising ROI",
-      # Colour depends on value
-      color = if (is.na(KPI_fundraising_num())) {
-        "primary"
-      } else if (KPI_fundraising_num() < 100) {
-        "danger"
-      } else if (KPI_fundraising_num() < 200) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_fundraising_num())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fas fa-donate"
-      }
-    )
-  })
-
-  ## OPX KPI--------------------------------------------------------------------
-  # Creating OPX KPI
-  KPI_OPX_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # most recent year
-        filter(year == max(year)) |>
-        # Calculating the value as a %
-        transmute(
-          KPI_OPX =
-            sum(revenue_private_contributions,
-                revenue_investments, assets_cash_in_bank_and_cash_equivalents,
-                na.rm = TRUE) /
-              sum(liability_short_term_grants_payable,
-                  liability_revocable_endowments,
-                  liability_long_term_grants_payable, na.rm = TRUE) * 100
-        ) |>
-        # Just the single number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for OPX KPI
-  output$KPI_box_OPX <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_OPX_num()),
-      caption = "Liquid assets / obligations",
-      # colour depends on value
-      color = if (is.na(KPI_OPX_num())) {
-        "primary"
-      } else if (KPI_OPX_num() < 100) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on value
-      icon = if (is.na(KPI_OPX_num())) {
-        "fa-solid fa-circle-xmark"
-      } else if (KPI_OPX_num() < 100) {
-        "fa-solid fa-scale-unbalanced"
-      } else {
-        "fa-solid fa-scale-balanced"
-      }
-    )
-  })
-
-  ## Private donations growth---------------------------------------------------
-  # Creating private donations growth
-  KPI_private_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # Calculating the two most recent years as a %
-        transmute(
-          growth =
-            (revenue_private_contributions[1] -
-              revenue_private_contributions[2]) /
-              revenue_private_contributions[2] * 100
-        ) |>
-        # Pulling the first instance of the number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for private donations growth
-  output$KPI_box_private <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_private_num()),
-      caption = "Private donation growth",
-      # colour depends on value
-      color = if (is.na(KPI_private_num())) {
-        "primary"
-      } else if (KPI_private_num() < 0) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_private_num())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fas fa-hands-helping"
-      }
-    )
-  })
-
-  ## Public grants growth-------------------------------------------------------
-  # Creating public grants growth
-  KPI_public_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # Calculating the two most recent years as a %
-        transmute(
-          growth =
-            (revenue_public_grants[1] - revenue_public_grants[2]) /
-              revenue_public_grants[2] * 100
-        ) |>
-        # Pulling the first instance of the number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for public grants growth
-  output$KPI_box_public <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_public_num()),
-      caption = "Public grants growth",
-      # colour depends on value
-      color = if (is.na(KPI_public_num())) {
-        "primary"
-      } else if (KPI_public_num() < 0) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_public_num())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fa-solid fa-building-columns"
-      }
-    )
-  })
-
-  ## Public/private proportion--------------------------------------------------
-  # Creating public/private proportion
-  KPI_proportion <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # calculating the value as a %
-        transmute(
-          proportion =
-            sum(revenue_private_contributions,
-              revenue_membership_fees,
-              na.rm = TRUE
-            ) /
-              sum(revenue_public_grants, revenue_private_contributions,
-                revenue_membership_fees,
-                na.rm = TRUE
-              ) * 100
-        ) |>
-        pull() |>
-        # getting the most recent year
-        first()
-    }
-  })
-
-  # Creating value box for private/public proportion
-  output$KPI_box_proportion <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_proportion()),
-      caption = "Private donations",
-      # Adding in the same check as elsewhere - colours match if all are NA
-      color = if (is.na(KPI_proportion())) {
-        "primary"
-      } else {
-        "info"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_proportion())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fa-solid fa-chart-pie"
-      }
-    )
-  })
-
-  ## Excel data test -----------------------------------------------------------
-  # (also below, in the second chunk)
-  # Reactive function to test for Excel data quality (basic)
-  datatest <- reactive({
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
-    }
-  })
-  # the button click
-}) |> bindEvent(input$click)
-```
-
-```{r reactive_plots, include=FALSE}
-
-# Beginning of the second reactive chunk =======================================
-# It's necessary to split the observers into separate parts, with all necessary
-# code duplicated, because of R's limitations.
-
-observe({
-
-  # Data frames ================================================================
 
@@
 
-    # Testing and creating data frame
-    html_test_create(input$html$datapath,
-                 as.character(input$start), as.character(input$end))
-  })
+    # Testing
+    html_test <- read_proff_test(input$html$datapath)
 
-  # Recreating, to make available to this chunk as well
-  # Creating data frame from Excel file
-  excel_test <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
+    # Creating data frame
 
-    widen_excel_test(input$excel$datapath)
-  })
-
-  exceldata <- reactive({
-    widen_excel(excel_test())
-  })
-
-  # Plots ======================================================================
-
-  ## Annual result--------------------------------------------------------------
-  arsresultat_plot <- reactive({
-    # testing for the presence of data
-    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      htmldata() |>
-        # renaming for plotly hover
-        rename(
-          Year = year,
-          `Net income` = arsresultat,
-          `Net positive` = annual_profit
-        ) |> 
-        ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # using a function to ensure uniform colours irrespective of data
-        scale_fill_green(htmldata()$annual_profit) +
-        theme(legend.position = "none") +
-        labs(
-          x = "", y = "", title = "Net income (in thousands)"
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$arsresultat_plotly <- renderPlotly(
-    ggplotly(arsresultat_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  ## Balance--------------------------------------------------------------------
-  balance_plot <- reactive({
-    # testing for the presence of data
-    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      htmldata() |>
-        # rename for plotly hover
-        rename(
-          Year = year,
-          `Total equity/liabilities` = sum_egenkapital_og_gjeld,
-          `Positive equity` = balance_positive
-        ) |> 
-        ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # using a function to ensure uniform colours irrespective of data
-        scale_fill_green(htmldata()$balance_positive) +
-        theme(legend.position = "none") +
-        labs(
-          x = "", y = "", title = "Total equity/liabilities (in thousands)",
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$balance_plotly <- renderPlotly(
-    ggplotly(balance_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  ## Expenses-------------------------------------------------------------------
-  expenses_plot <- reactive({
-    # testing for data quality
-    if (datatest() == TRUE) {
-      # printing 'bad data' plot if data are bad
-      bad_data()
-    } else {
-      exceldata() |>
-        # Renaming for the purposes of the plot + dropping other variables
-        transmute(
-          Year = as.factor(year),
-          Other = expenses_other,
-          Administrative = expenses_administrative,
-          Fundraising = expenses_fundraising,
-          `Program services` = expenses_program_services
-        ) |>
-        # Pivoting to use fill
-        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
-        # Ordering for the plot
-        mutate(
-          Expenses = ordered(Expenses,
-            levels =
-              c("Program services", "Other", "Fundraising", "Administrative")
-          )
-        ) |>
-        ggplot(aes(Year, value, fill = Expenses)) +
-        geom_col() +
-        scale_y_continuous(
-          # Excel data are assumed to be in NOK; if this changes, code based on
-          # the currency_code() function may be adapted
-          labels = label_dollar(prefix = "NOK "),
-          # Basing lower limit on the data (not suppressing erroneous negative
-          # numbers), and the upper limit on the maximum value in revenues and
-          # expenses, to match the two plots
-          limits = c(NA, maximum())
-        ) +
-        scale_fill_met_d("Hokusai2") +
-        labs(x = "", y = "", title = "Expenses (in thousands)")
-    }
-  })
-
-  # Creating plotly output
-  output$expenses_plotly <- renderPlotly(
-    ggplotly(expenses_plot())
-  )
-
-  ## Revenue--------------------------------------------------------------------
-  revenue_plot <- reactive({
-    # testing for data quality
-    if (datatest() == TRUE) {
-      # printing 'bad data' plot if data are bad
-      bad_data()
-    } else {
-      exceldata() |>
-        # Renaming for the purposes of the plot + dropping other variables
-        transmute(
-          Year = as.factor(year),
-          `Private contributions` = revenue_private_contributions,
-          `Membership fees` = revenue_membership_fees,
-          `Public grants` = revenue_public_grants,
-          Investments = revenue_investments,
-          `Operational activities` = revenue_operational_activities,
-          `Commercial activities` = revenue_commercial_activities,
-          Other = revenue_other
-        ) |>
-        # Pivoting to use fill
-        pivot_longer(
-          c(`Private contributions`:Other),
-          names_to = "Revenue source"
-        ) |>
-        # Ordering for the plot
-        mutate(
-          `Revenue source` =
-            ordered(`Revenue source`,
-              levels =
-                c(
-                  "Other", "Commercial activities",
-                  "Operational activities", "Investments",
-                  "Membership fees", "Private contributions",
-                  "Public grants"
-                )
-            )
-        ) |>
-        ggplot(aes(Year, value, fill = `Revenue source`)) +
-        geom_col() +
-        scale_y_continuous(
-          # Excel data are assumed to be in NOK; if this changes, code based on
-          # the currency_code() function may be adapted
-          labels = label_dollar(prefix = "NOK "),
-          # Basing lower limit on the data (not suppressing erroneous negative
-          # numbers), and the upper limit on the maximum value in revenues and
-          # expenses, to match the two plots
-          limits = c(NA, maximum())
-        ) +
-        # Specifying the number of values, as the base palette only includes 6
-        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
-        labs(x = "", y = "", title = "Revenue (in thousands)")
-    }
-  })
-
-  # Creating plotly output
-  output$revenue_plotly <- renderPlotly(
-    ggplotly(revenue_plot())
-  )
-
-  ## CEO compensation-----------------------------------------------------------
-  CEO_plot <- reactive({
-    # Testing for the presence of data
-    if (sum(!is.na(htmldata()$lederlonn)) == 0 &&
-      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      # Creating plot if any data exist
-      # Note that the plot will still look empty, but have a legend and scales,
-      # if all values are explicit zeroes
-      htmldata() |>
-        # preparing for pivoting
-        select(
-          year, lederlonn, leder_annen_godtgjorelse
-        ) |>
-        # Compensation in ordinary numbers
-        mutate(
-          lederlonn = lederlonn * 1000,
-          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-        ) |>
-        # renaming for the purposes of the plot
-        rename(
-          Year = year,
-          Salary = lederlonn,
-          `Other compensation` = leder_annen_godtgjorelse
-        ) |>
-        # pivoting to join the two values to use fill
-        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-        ggplot(aes(Year, value, fill = Compensation)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # pretty colours
-        scale_fill_met_d("Hokusai2", direction = -1) +
-        labs(
-          x = "", y = "Total compensation", title = "CEO compensation"
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$CEO_plotly <- renderPlotly(
-    ggplotly(CEO_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  # Reactive functions =========================================================
-
-  ## Currency scale-------------------------------------------------------------
-  # Reactive function to check for currency and use currency code or ?
-  currency_scale <- reactive({
-    scale_y_continuous(
-      # Checking for uniform currency code
-      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
-        # Using the currency code if uniform
-        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
-        # alternative prefix if multiple currencies in the data
-        label_dollar(prefix = "? ")
-      )
+    read_proff_complete(
+      html_test,
+      as.character(input$start), as.character(input$end)
     )
   })
-
-  ## Plotly caption-------------------------------------------------------------
-  # Reactive function to create plotly caption (ggplot captions and subtitles
-  # are lost in the conversion) (shamelessly stolen from the Internet, with
-  # minor modifications).
-  # Checks for uniform currency code and prints a warning
-  # if there are multiple currency codes.
-  plotly_caption <- reactive({
-    list(
-      x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
-        "", "Note: not all values are denominated in the same currency"
-      ),
-      showarrow = F, xref = "paper", yref = "paper",
-      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
-      font = list(size = 11)
-    )
-  })
-
-  ## scale maximum--------------------------------------------------------------
-  # Reactive function to create the maximum value for revenue and expenses plots
-  maximum <- reactive({
-    if (datatest() == TRUE) {
-      NA
-    } else if ("year" %in% names(excel_test()) && "category" %in% names(excel_test())) {
-      # starting with the unpivoted excel data
-      excel_test() |>
-        # removing the firm and sub-category columns
-        select(-firm, -sub_category) |>
-        # filtering to retain only revenue and expenses entries
-        filter(category == "Revenue" | category == "Expenses") |>
-        # grouping and summarising to calculate maximum value
-        group_by(year, category) |>
-        summarise(
-          sum = sum(value, na.rm = TRUE)
-        ) |>
-        # Creating list of sums
-        pull() |>
-        # selecting the highest value
-        max()
-    } else {
-      NA
-    }
-  })
-
-  ## Excel data test (also above)-----------------------------------------------
-  # Reactive function to test for Excel data quality (basic)
-  datatest <- reactive({
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
-    }
-  })
-  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
-}) |> bindEvent(input$click)
-
 ```
 
-```{r non-reactive functions, include=FALSE}
-# Data reading functions =======================================================
-
-## Function to read html file --------------------------------------------------
-# read_proff <- function(file, start, end) {
-#   # Specifying encoding to deal with occasional errors
-#   read_html(file, encoding = "UTF-8") |>
-#     html_node("#inner-frame") |>
-#     html_table() |>
-#     # removing empty column, using all_of(argument)
-#     select(
-#       REGNSKAPSPERIODE:all_of(start)
-#     ) |>
-#     # removing white space
-#     mutate(
-#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-#     ) |>
-#     # removing duplicate table sometimes created by read_html()
-#     filter(
-#       !row_number() > 176
-#     ) |>
-#     # removing duplicate rows
-#     filter(
-#       !grepl("Lukk", REGNSKAPSPERIODE)
-#     ) |>
-#     # tidying data using pivot_longer() and all_of(argument)
-#     pivot_longer(
-#       all_of(end):all_of(start),
-#       names_to = "year"
-#     ) |>
-#     # creating a currency code column and turning current currency codes into
-#     # NAs
-#     mutate(
-#       # changing to real NAs
-#       value = na_if(value, "-"),
-#       # turning years into numbers
-#       year = as.integer(year),
-#       # creating a currency code column and turning current currency codes into
-#       # NAs
-#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-#     ) |>
-#     # filling the currency column
-#     fill(
-#       valutakode,
-#       .direction = "updown"
-#     ) |>
-#     # removing dates and redundant currency
-#     filter(
-#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-#         REGNSKAPSPERIODE != "Valutakode"
-#     ) |>
-#     mutate(
-#       # removing white space in numbers
-#       value = str_replace_all(value, "\\s", ""),
-#       # turning numbers into true numbers
-#       value = as.numeric(value),
-#       # removing values that are years
-#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-#     ) |>
-#     # removing any remaining duplicate rows
-#     distinct() |>
-#     # pivoting to create 'tidy' variables
-#     pivot_wider(
-#       names_from = REGNSKAPSPERIODE, values_from = value
-#     ) |>
-#     # tidying names
-#     clean_names() |>
-#     # removing headings from the table
-#     select(
-#       -lederlonn_i_hele_1000,
-#       -resultatregnskap_i_hele_1000,
-#       -balanseregnskap_i_hele_1000
-#     ) |>
-#     # making it clear what lonn refers to
-#     rename(
-#       lederlonn = lonn
-#     ) |>
-#     # Creating new variables for plots
-#     mutate(
-#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-#     ) |>
-#     # arranging by year for practical display
-#     arrange(desc(year))
-# }
-
-### Test version of function that creates multiple data problems----------------
-# read_proff <- function(file, start, end) {
-#   read_html(file, encoding = "UTF-8") |>
-#     html_node("#inner-frame") |>
-#     html_table() |>
-#     select( # removing empty column
-#       REGNSKAPSPERIODE:all_of(start)
-#     ) |> # using all_of()
-#     mutate( # removing whitespace
-#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-#     ) |>
-#     filter( # removing duplicate table - no idea why it's necessary
-#       !row_number() > 176
-#     ) |>
-#     filter( # removing duplicate rows
-#       !grepl("Lukk", REGNSKAPSPERIODE)
-#     ) |>
-#     pivot_longer( # tidying data
-#       all_of(end):all_of(start),
-#       names_to = "year"
-#     ) |> # using all_of()
-#     mutate( # changing to real NAs and turning years into numbers
-#       value = na_if(value, "-"),
-#       year = as.integer(year),
-#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-#     ) |>
-#     fill( # filling the currency column
-#       valutakode,
-#       .direction = "updown"
-#     ) |>
-#     filter( # removing dates and redundant currency
-#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-#         REGNSKAPSPERIODE != "Valutakode"
-#     ) |>
-#     mutate( # removing whitespace in numbers
-#       value = str_replace_all(value, "\\s", ""),
-#       value = as.numeric(value),
-#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-#     ) |>
-#     distinct() |> # removing any remaining duplicate rows
-#     pivot_wider( # pivoting
-#       names_from = REGNSKAPSPERIODE, values_from = value
-#     ) |>
-#     clean_names() |> # tidying names
-#     select(
-#       -lederlonn_i_hele_1000,
-#       -resultatregnskap_i_hele_1000,
-#       -balanseregnskap_i_hele_1000
-#     ) |> # removing headings
-#     rename( # making it clear what lonn refers to
-#       lederlonn = lonn
-#     ) |>
-#     # Creating new variables for plots
-#     mutate(
-#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE),
-#     ) |>
-#     # bad test - REMOVE
-#     mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
-#       leder_annen_godtgjorelse = NA,
-#       lederlonn = NA,
-#       arsresultat = NA,
-#       sum_egenkapital_og_gjeld = NA
-#     ) |>
-#     arrange(desc(year)) # arranging by year
-# }
-
-## Functions to read Excel file ------------------------------------------------
-widen_excel_test <- function(file) {
-  # Reading Excel file
-  read_excel(file) |> 
-    # cleaning names to enable later operations
-    clean_names()
-}
-
-widen_excel <- function(object) {
-  # testing if all columns are present
-  if (setequal(names(object),
-               c("year", "firm", "category", "sub_category", "name", "value"))) {
-    object |>
-      # uniting category and name to deal with multiple 'Other'
-      unite(
-        entry, c(category, name)
-      ) |>
-      # deleting Sub-category and Firm
-      select(-sub_category, -firm) |>
-      # pivoting to create 'tidy' variables
-      pivot_wider(names_from = entry, values_from = value) |>
-      # cleaning names
-      clean_names()
-  } else {
-    validate("Please upload data using the specified Excel template.")
-  }
-}
-
-# Support functions ============================================================
-
-## KPI_to_string ---------------------------------------------------------------
-# Function to turn KPI values into NA or percentage string
-KPI_to_string <- function(variable) {
-  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
-  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
-  # avoid passing an NA to subsequent operations.)
-  if (is.na(variable)) {
-    NA
-  } else if (variable == Inf) {
-    NA
-  } else {
-    # Shortening (lengthening) the value, adding %
-    str_c(format(variable, digits = 3), "%")
-  }
-}
-
-## scale_fill_green ------------------------------------------------------------
-# Function with slightly nicer colours than default; ensuring uniformity; +
-# ensuring positive values are always green through a test for the presence of
-# FALSE values
-scale_fill_green <- function(variable) {
-  scale_fill_met_d("Tsimshian",
-    direction = ifelse(
-      mean(variable,
-        na.rm = TRUE
-      ) == 1, -1, 1
-    )
-  )
-}
-
-## no_information plot ---------------------------------------------------------
-# Function to print a 'no information' plot, if necessary
-no_information <- function() {
-  empty_plot("No information available")
-}
-
-## bad_data plot ---------------------------------------------------------------
-# Function to print a 'bad data' plot, if necessary
-bad_data <- function() {
-  empty_plot("Something is wrong\nwith the Excel data.")
-}
-
-## empty_plot-------------------------------------------------------------------
-# Function to print an empty plot with a supplied message
-empty_plot <- function(warning) {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    annotate(
-      "text",
-      label = warning,
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
-}
-
-# Testing arguments ============================================================
-# One or more of the elements of this mutate() operation may be inserted at the
-# end of the read_proff() function to create various data problems for some
-# quick and dirty testing
-
-# mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
-#       leder_annen_godtgjorelse = NA,
-#       lederlonn = NA,
-#       arsresultat = NA,
-#       sum_egenkapital_og_gjeld = NA
-#     ) |>
-```
-
-```{r new_data_tests}
-html_test_create <- function(file, start, end) {
+```{r}
+html_testing <- function(file, start, end) {
   # Reading file and node
@@
 
-  # Checking if node exists
+  # Checking of node exists
   if (class(htmltest) == "xml_node") {
@@
   } else {
-    validate("Please upload the specified type of html document.")
+    validate("table error")
   }
@@
   } else {
-    validate("Please upload the specified type of html document.")
+    validate("regnskap error")
   }

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002.Rmd
Stylering.

@@
               revenue_investments, assets_cash_in_bank_and_cash_equivalents,
-                na.rm = TRUE) /
+              na.rm = TRUE
+            ) /
               sum(liability_short_term_grants_payable,
                 liability_revocable_endowments,
-                  liability_long_term_grants_payable, na.rm = TRUE) * 100
+                liability_long_term_grants_payable,
+                na.rm = TRUE
+              ) * 100
         ) |>
@@
 ```{r reactive_plots, include=FALSE}
-
 # Beginning of the second reactive chunk =======================================
@@
     # Testing and creating data frame
-    html_test_create(input$html$datapath,
-                 as.character(input$start), as.character(input$end))
+    html_test_create(
+      input$html$datapath,
+      as.character(input$start), as.character(input$end)
+    )
   })
@@
 }) |> bindEvent(input$click)
-
 ```

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002 currency testing.Rmd
Found a problem with the html code. fml.

--- 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002.Rmd
@@
   - A common measure of transparency, but should normally be evaluated in context.
+
+Row 
+-----------------------------------------------------------------------
+
+### Data
+```{r}
+tableOutput("htmltable")
+```
 
@@
 
+  output$htmltable <- renderTable({
+    htmldata()
+  })
+
   # Recreating, to make available to this chunk as well
@@
       # Checking for uniform currency code
-      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
-        # Using the currency code if uniform
-        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
-        # alternative prefix if multiple currencies in the data
+      labels = if (pull(.data = htmldata(), valutakode) |> length() |> unique() == 1) {
         label_dollar(prefix = "? ")
-      )
+      } else {
+        label_dollar(prefix = str_c(unique(na.omit(pull(.data = htmldata(), valutakode))), " "))
+      }
     )
@@
     ) |>
+    fill(valutakode) |> 
     # removing dates and redundant currency

### 2022-07-28 Old/Shiny/Test/HTML tests.Rmd
Found a problem with the html code. fml.

@@
 ```
+
+```{r}
+read_proff <- function(file, start, end) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    # removing empty column, using all_of(argument)
+    select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column
+    fill(
+      valutakode,
+      .direction = "updown"
+    ) |>
+    # removing dates and redundant currency
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+}
+```
+
+```{r}
+read_proff("Data/Kreftforeningen2002_2021 copy.html", "2002", "2021")
+
+valutatest <- read_html("Data/Kreftforeningen2002_2021 copy.html", encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |> 
+   select(
+      REGNSKAPSPERIODE:`2002`
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |> 
+  pivot_longer(
+      `2021`:`2002`,
+      names_to = "year"
+    ) |> 
+   mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |> 
+  fill(
+      valutakode,
+      .direction = "updown"
+    ) |>
+   distinct() |> 
+   pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |> 
+  clean_names() |> 
+  select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |> 
+  rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+
+valutatest
+
+unique(valutatest$valutakode)
+
+length(unique(valutatest$valutakode))
+
+na.omit(valutatest$valutakode) |> unique()
+
+ifelse(length(unique(valutatest$valutakode)) == 1, TRUE, FALSE)
+
+pull(.data = valutatest, valutakode) |> unique() |> length()
+
+unique(na.omit(pull(.data = valutatest, valutakode)))
+```
+
+```{r}
+read_proff <- function(file, start, end) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    # removing empty column, using all_of(argument)
+    select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column
+
+    ### NEED TO SPLIT HERE
+
+    fill(
+      valutakode,
+      .direction = "downup"
+    ) # |>
+    # removing dates and redundant currency
+    # filter(
+    #   REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+    #     REGNSKAPSPERIODE != "Valutakode"
+    # ) |>
+    # mutate(
+    #   # removing white space in numbers
+    #   value = str_replace_all(value, "\\s", ""),
+    #   # turning numbers into true numbers
+    #   value = as.numeric(value),
+    #   # removing values that are years
+    #   value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    # ) |>
+    # # removing any remaining duplicate rows
+    # distinct() |>
+    # # pivoting to create 'tidy' variables
+    # pivot_wider(
+    #   names_from = REGNSKAPSPERIODE, values_from = value
+    # ) |>
+    # # tidying names
+    # clean_names() |>
+    # # removing headings from the table
+    # select(
+    #   -lederlonn_i_hele_1000,
+    #   -resultatregnskap_i_hele_1000,
+    #   -balanseregnskap_i_hele_1000
+    # ) |>
+    # # making it clear what lonn refers to
+    # rename(
+    #   lederlonn = lonn
+    # ) |>
+    # # Creating new variables for plots
+    # mutate(
+    #   annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+    #   balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    # ) |>
+    # # arranging by year for practical display
+    # arrange(desc(year))
+}
+
+test <- read_proff("Data/Kreftforeningen2002_2021 copy.html", "2002", "2021")
+
+unique(test$valutakode)
+
+test
+```

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002.Rmd
Fixing the currency test.

@@
   ## Creating data frame from html file-----------------------------------------
-  htmldata <- reactive({
+  html_currency <- reactive({
     # validating file
@@
     )
+  })
+
+  htmldata <- reactive({
+    read_proff_two(html_currency())
   })
@@
       # Checking for uniform currency code
-      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+      labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
         # Using the currency code if uniform
-        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+        label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
         # alternative prefix if multiple currencies in the data
@@
     list(
-      x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
+      x = 1, y = -0.1, text = ifelse(length(unique(html_currency()$valutakode)) == 1,
         "", "Note: not all values are denominated in the same currency"
@@
     # Creating the data frame
-    read_proff(file, start, end)
+    read_proff_one(htmltable, start, end)
   } else {
@@
 
-read_proff <- function(file, start, end) {
-  read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame") |>
-    html_table() |>
+read_proff_one <- function(object, start, end) {
+    object |> 
   # removing empty column, using all_of(argument)
@@
     ) |> 
-    # filling the currency column
     fill(
       valutakode,
-      .direction = "updown"
-    ) |>
+      .direction = "downup"
+    )
+}
+
+read_proff_two <- function(object) {
+     object |>
     # removing dates and redundant currency
+    select(-valutakode) |> 
     filter(

### 2022-07-28 Old/Shiny/Test/HTML tests.Rmd
Fixing the currency test.

@@
 ```{r}
-read_proff <- function(file, start, end) {
+read_proff_one <- function(file, start, end) {
   read_html(file, encoding = "UTF-8") |>
@@
     ) |> 
-    # filling the currency column
+    fill(
+      valutakode,
+      .direction = "downup"
+    )
+}
 
@@
 
-    fill(
-      valutakode,
-      .direction = "downup"
-    ) # |>
+read_proff_two <- function(object) {
+     object |>
     # removing dates and redundant currency
-    # filter(
-    #   REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-    #     REGNSKAPSPERIODE != "Valutakode"
-    # ) |>
-    # mutate(
-    #   # removing white space in numbers
-    #   value = str_replace_all(value, "\\s", ""),
-    #   # turning numbers into true numbers
-    #   value = as.numeric(value),
-    #   # removing values that are years
-    #   value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-    # ) |>
-    # # removing any remaining duplicate rows
-    # distinct() |>
-    # # pivoting to create 'tidy' variables
-    # pivot_wider(
-    #   names_from = REGNSKAPSPERIODE, values_from = value
-    # ) |>
-    # # tidying names
-    # clean_names() |>
-    # # removing headings from the table
-    # select(
-    #   -lederlonn_i_hele_1000,
-    #   -resultatregnskap_i_hele_1000,
-    #   -balanseregnskap_i_hele_1000
-    # ) |>
-    # # making it clear what lonn refers to
-    # rename(
-    #   lederlonn = lonn
-    # ) |>
-    # # Creating new variables for plots
-    # mutate(
-    #   annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-    #   balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-    # ) |>
-    # # arranging by year for practical display
-    # arrange(desc(year))
+    select(-valutakode) |> 
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
 }
@@
 
+read_proff_two(test)
+
 test
 ```
+
+```{r}
+?validate
+```

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002.Rmd
Removing testing arguments.

@@
 }
-
-# Testing arguments ============================================================
-# One or more of the elements of this mutate() operation may be inserted at the
-# end of the read_proff() function to create various data problems for some
-# quick and dirty testing
-
-# mutate(
-#       valutakode = ifelse(year == 2019, "USD", valutakode),
-#       leder_annen_godtgjorelse = NA,
-#       lederlonn = NA,
-#       arsresultat = NA,
-#       sum_egenkapital_og_gjeld = NA
-#     ) |>
-```
+```

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha v003.Rmd
Comments and stylering.

--- 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002.Rmd
@@
       clean_names()
-    # Notfification if all columns are not present
+    # Notification if all columns are not present
   } else {
@@
     theme_void() +
+    # annotating the plot with the specified warning message
     annotate(

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha v003 value box test.Rmd
So much testing.

--- 2022-07-28 Dashboard/NGO Analysis dashboard alpha v003.Rmd
@@
     # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
+    if (exceldata() == FALSE) {
+      "?"
+    }else if (datatest() == TRUE) {
       NA
@@
   } else {
-    validate("Please upload data using the specified Excel template.")
+    FALSE
   }

### 2022-07-28 Old/Shiny/Test/NGO Analysis dashboard more html testing.Rmd
So much testing.

--- 2022-07-28 Dashboard/NGO Analysis dashboard alpha v002.Rmd
@@
     # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
+    if (datatest() == TRUE || datatest_box() == TRUE) {
       NA
@@
   })
+
+  ## Excel data test (box only) ------------------------------------------------
+  datatest_box <- reactive({
+    if (setequal(
+    names(exceldata()),
+    c("year", "firm", "category", "sub_category", "name", "value"))) {
+      FALSE
+    } else {
+      TRUE
+    }
+  })
+
   # the button click

### 2022-07-29 Presentasjon/presentasjon.Rmd
Utkast til presentasjon

---
title: "Presentasjon OPX Fiffig tittel"
author: "Gruppe 5"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation:
     css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

# Introduction

## <br> Introduction | OPX and their goal
- A startup developing solutions for socially responsible investment strategies
- The goal is to increase cash flow to charitable purposes 
- Donor's trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients

## <br> Introduction | How to assess financial performance?
- Data collection <br> OPX requested an automated solution to replace today's manual data collection
- Key Performance Indicators <br> OPX needed meaningful KPIs allowing to assess the relationship between assets and obligations

# Brainstorming

## <br> Brainstorming | Phase 1 – how do we harvest the data?
- Webcrawler? <br> this would be too difficult and it would only work with some organizations
-

## <br> Brainstorming | Phase 2 – how do we analyze the harvested data?

## <br> Brainstorming | Phase 3 - exploring three possible solutions

## <br> Brainstorming | Phase 4 - do we go with proff.no or Brønnøysund anyway?

## <br> Brainstorming | Phase 5 - throwing everything away

# Data collection

# Data description

# Findings

# Limitations and steps forward

### 2022-07-29 Report/Group 5 report.Rmd
La inn begynnelsen av rapport-arbeidet

---
title: "Group 5 Report"
author: "ISSSV1337 Group 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```

# Brainstorming

Our brainstorming process for this case was characterized by several
surprises and turns in the direction of the solution. As we dug deeper
into the solution opportunities, we were also met by new obstacles that
forced us to change course. In particular, there were four turning
points. Therefore, we have decided to divide the brainstorming into
three phases.

### Phase 1: How to harvest and analyze the data? {.smaller}

OPX wanted us to harvest the organizations' financial data from the web.
During the first stage of brainstorming, we wanted to find a simple and
automated way to do so.

We discussed creating a webcrawler that could collect data across
different websites, but quickly decided that this would be too
difficult, and that it would only work with selected organizations. The
suggested solution was to create a code that could read PDF documents
that OPX downloaded themselves. Our code would analyze and organize the
uploaded data from the PDF documents.

However, different organizations publish financial data in different
ways in their yearly reports -- some financial statements were included
as tables, some as pictures, etc. This led us to phase 2.

```{r}
knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene

# inkluder en med årsregnskapet som pdf her
```

### Phase 2: Exploring solutions {.smaller}

At this stage, we wanted to explore alternatives to analyzing PDF
documents. We discussed how to do so, and concluded that we had three
options: 
- Financial statements in yearly report PDF documents. 
- Webscrape Proff.no for financial statements. 
- Webscrape the Brønnøysund registry.

We divided the task among ourselves to explore whether these solutions
were feasible or not. The first option of using PDF documents still had
the issue of financial statements not always being published as text.
Also, we quickly found that Proff.no disallowed webscraping on their
website. The Brønnøysund registry did store downloadable XML documents
of all the financial statements on their website, but we discovered that
access to these files cost about half a million NOK.

We started to look at workarounds for this problem. It would still be
possible to download Proff.no HTML websites manually for each
organization. And, we attempted to request a temporary access to the
Brønnøysund registry's XML documents.

```{r}

```

The Brønnøysund registry and Proff.no had different divisions of
accounting items than the ones OPX wanted. We found that we needed to
create an item path from the webpage items to the OPX items, and we
attempted to do so in Excel at first, so that we later could code this
into R. While doing so, we realized that it would be practically
impossible to extract the correct sums for the OPX accounting items from
the Brønnøysund items. This was e.g. because OPX had split the operating
income into public and private contributions, while the Brønnøysund
registry and Proff.no had not.

```{r}
knitr::include_graphics("Images/info_kreft.png")
knitr::include_graphics("Images/regnskapssti.png")
```

When also discovering that not all organizations publish their financial
statements, and certainly not for every year, we decided to give up on
collecting financial data for OPX. This is when we moved on to the third
and final stage.

### Phase 3: Scrapping the scraping {.smaller}

The OPX case requested two tasks -- webscrape financial data, and create
KPIs. During phase three, we scrapped all ideas and progress from the
first two stages, and focused on creating KPIs.

We divided amongst ourselves that some would find useful KPIs from best
practices, while someone else found the accounting items in the OPX
structure that were needed in the calculations. Then, we coded the KPI
calculations into R.

In order to create a somewhat more challenging task for ourselves, we
also decided to create a shiny app that OPX could use, which would
display the KPIs as well as financial trends for the selected
organization. To use this, OPX needs to gather the financial data
themselves, then upload this into the shiny app as a x document

# Data collection

We wanted to use webscraping to collect the data, so that we could have
automated the data input process. As described, we explored different
potential solutions. We experimented on coding such that data could have
been imported from XML files. However, once we realized that the
financial data was of little ¨use, this experimentation stopped.

As we entered the third phase of brainstorming, we still needed some
example data so that we could test our KPI codes. For this, we used the
financial data that OPX had already gathered from different
organizations and organized into useful accounting items.

We copied these financial statements manually into an Excel document.
Perhaps a "simple" solution, however our code requires OPX to import
both an Excel document themselves into shiny, as well as a HTML document
from Proff.no.

If OPX wishes to develop our solution, we recommend that they purchase
API access from the Brønnøysund registry and/or Proff.no. They also need
to create e.g. a simple SQL database with its own data that can be used
by R.

### 2022-07-29 Dashboard/report testing.Rmd
In the process of testing subfig.

--- 2022-07-26 Dashboard/report.Rmd
@@
 date: "`r Sys.Date()`"
-output: pdf_document
+output: 
+  pdf_document:
+    extra_dependencies: "subfig"
 params:
@@
 # Creating general warning
-general_warning <- if (datatest == TRUE || sum(!is.na(htmldata$arsresultat)) == 0 || sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 || (sum(!is.na(htmldata$lederlonn)) == 0 &&
+general_warning <- if (datatest == TRUE ||
+  sum(!is.na(htmldata$arsresultat)) == 0 || 
+  sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 ||
+  (sum(!is.na(htmldata$lederlonn)) == 0 &&
     sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0)) {
@@
 
-<!--# Need to decide whether to underline the KPIs, as in the first one -->
-
 - **Fundraising ROI:** `r fundraising`
@@
 
+```{r revenue_expenses_plot_test, fig.asp=1, fig.show="hold", out.width="45%", fig.cap="Based on the categories \`Revenue' and \`Expenses' in the Excel sheet."}
+## Revenue--------------------------------------------------------------------
+revenue_plot <-
+  # testing for data quality
+  if (datatest == TRUE) {
+    # printing 'bad data' plot if data are bad
+    NA
+  } else {
+    exceldata |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        `Private contributions` = revenue_private_contributions,
+        `Membership fees` = revenue_membership_fees,
+        `Public grants` = revenue_public_grants,
+        Investments = revenue_investments,
+        `Operational activities` = revenue_operational_activities,
+        `Commercial activities` = revenue_commercial_activities,
+        Other = revenue_other
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(
+        c(`Private contributions`:Other),
+        names_to = "Revenue source"
+      ) |>
+      # Ordering for the plot
+      mutate(
+        `Revenue source` =
+          ordered(`Revenue source`,
+            levels =
+              c(
+                "Other", "Commercial activities",
+                "Operational activities", "Investments",
+                "Membership fees", "Private contributions",
+                "Public grants"
+              )
+          )
+      ) |>
+      ggplot(aes(Year, value, fill = `Revenue source`)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum)
+      ) +
+      # Specifying the number of values, as the base palette only includes 6
+      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
+      labs(x = "", y = "", title = "Revenue (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
+  }
+
+# Printing the plot if the data are OK
+if (datatest == FALSE) {
+  revenue_plot
+}
+
+## Expenses -------------------------------------------------------------------
+expenses_plot <-
+  # testing for data quality
+  if (datatest == TRUE) {
+    # printing 'bad data' plot if data are bad
+    NA
+  } else {
+    exceldata |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        Other = expenses_other,
+        Administrative = expenses_administrative,
+        Fundraising = expenses_fundraising,
+        `Program services` = expenses_program_services
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
+      # Ordering for the plot
+      mutate(
+        Expenses = ordered(Expenses,
+          levels =
+            c("Program services", "Other", "Fundraising", "Administrative")
+        )
+      ) |>
+      ggplot(aes(Year, value, fill = Expenses)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum)
+      ) +
+      scale_fill_met_d("Hokusai2") +
+      labs(x = "", y = "", title = "Expenses (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
+  }
+
+# Printing the plot if the data are OK
+if (datatest == FALSE) {
+  expenses_plot
+}
+```
+
 ```{r compensation_plot, out.width="100%", fig.cap="Based on the category \`Lederlønn' in the Proff.no data.

### 2022-07-29 Dashboard/report.Rmd
Deleting unneded comment.
Renaming to use the testing version with the dashboard.

@@
 
+```{r revenue_expenses_plot_test, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("(Based on the category \`Revenue' in the Excel sheet.)", "( Based on the category \`Expenses' in the Excel sheet.)")}
+## Revenue--------------------------------------------------------------------
+revenue_plot <-
+  # testing for data quality
+  if (datatest == TRUE) {
+    # printing 'bad data' plot if data are bad
+    NA
+  } else {
+    exceldata |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        `Private contributions` = revenue_private_contributions,
+        `Membership fees` = revenue_membership_fees,
+        `Public grants` = revenue_public_grants,
+        Investments = revenue_investments,
+        `Operational activities` = revenue_operational_activities,
+        `Commercial activities` = revenue_commercial_activities,
+        Other = revenue_other
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(
+        c(`Private contributions`:Other),
+        names_to = "Revenue source"
+      ) |>
+      # Ordering for the plot
+      mutate(
+        `Revenue source` =
+          ordered(`Revenue source`,
+            levels =
+              c(
+                "Other", "Commercial activities",
+                "Operational activities", "Investments",
+                "Membership fees", "Private contributions",
+                "Public grants"
+              )
+          )
+      ) |>
+      ggplot(aes(Year, value, fill = `Revenue source`)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum)
+      ) +
+      # Specifying the number of values, as the base palette only includes 6
+      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
+      labs(x = "", y = "", title = "Revenue (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
+  }
+
+# Printing the plot if the data are OK
+if (datatest == FALSE) {
+  revenue_plot
+}
+
+## Expenses -------------------------------------------------------------------
+expenses_plot <-
+  # testing for data quality
+  if (datatest == TRUE) {
+    # printing 'bad data' plot if data are bad
+    NA
+  } else {
+    exceldata |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        Other = expenses_other,
+        Administrative = expenses_administrative,
+        Fundraising = expenses_fundraising,
+        `Program services` = expenses_program_services
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
+      # Ordering for the plot
+      mutate(
+        Expenses = ordered(Expenses,
+          levels =
+            c("Program services", "Other", "Fundraising", "Administrative")
+        )
+      ) |>
+      ggplot(aes(Year, value, fill = Expenses)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum)
+      ) +
+      scale_fill_met_d("Hokusai2") +
+      labs(x = "", y = "", title = "Expenses (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
+  }
+
+# Printing the plot if the data are OK
+if (datatest == FALSE) {
+  expenses_plot
+}
+```
+
 ```{r compensation_plot, out.width="100%", fig.cap="Based on the category \`Lederlønn' in the Proff.no data.

### 2022-07-29 Report/KPIs and Visualisations.Rmd
Copying over latest KPI description from the dashboard.

@@
 
--   **Fundraising ROI**: $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
-    -   A basic measure of the ROI of direct fundraising expenses.
--   **Liquid assets/obligations**: $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
-    -   A measure of the relationship between the organisation's liquid assets and its obligations.
--   **Private donations growth**: $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
-    -   A calculation of the percentage change in private contributions based on the two most recent years.
--   **Public grants growth**: $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
-    -   A calculation of the percentage change in public grants received based on the two most recent years.
--   **Private/public proportion**: $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
-    -   Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants.
+- **Fundraising ROI**:
+  - A basic measure of the return on investment of direct fundraising expenses:
+
+  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
+
+- **Liquid assets/obligations**:
+  - A measure of the relationship between the organisation's liquid assets and its obligations:
+
+  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
+
+- **Private donations growth**:
+  - A calculation of the percentage change in private contributions based on the two most recent years:
+
+  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
+
+- **Public grants growth**:
+  - A calculation of the percentage change in public grants received based on the two most recent years:
+
+$$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+
+- **Private/public proportion**:
+  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
+
+$$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
+
+>NB: Koden er kanskje ikke oppdatert, men beskrivelsen over er den som er i dashboardet og som koden der er basert på. (Det kan være feil utregninger i koden i dette dokumentet akkurat nå.)
 

### 2022-07-29 Dashboard/report.Rmd
Think subfig is best. NB: Delete useless plot if else portions.

--- 2022-07-26 Dashboard/report.Rmd
@@
 
+```{r revenue_expenses_plot_test, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
+## Revenue--------------------------------------------------------------------
+revenue_plot <-
+  # testing for data quality
+  if (datatest == TRUE) {
+    # printing 'bad data' plot if data are bad
+    NA
+  } else {
+    exceldata |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        `Private contributions` = revenue_private_contributions,
+        `Membership fees` = revenue_membership_fees,
+        `Public grants` = revenue_public_grants,
+        Investments = revenue_investments,
+        `Operational activities` = revenue_operational_activities,
+        `Commercial activities` = revenue_commercial_activities,
+        Other = revenue_other
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(
+        c(`Private contributions`:Other),
+        names_to = "Revenue source"
+      ) |>
+      # Ordering for the plot
+      mutate(
+        `Revenue source` =
+          ordered(`Revenue source`,
+            levels =
+              c(
+                "Other", "Commercial activities",
+                "Operational activities", "Investments",
+                "Membership fees", "Private contributions",
+                "Public grants"
+              )
+          )
+      ) |>
+      ggplot(aes(Year, value, fill = `Revenue source`)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum)
+      ) +
+      # Specifying the number of values, as the base palette only includes 6
+      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
+      labs(x = "", y = "", title = "Revenue (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
+  }
+
+# Printing the plot if the data are OK
+if (datatest == FALSE) {
+  revenue_plot
+}
+
+## Expenses -------------------------------------------------------------------
+expenses_plot <-
+  # testing for data quality
+  if (datatest == TRUE) {
+    # printing 'bad data' plot if data are bad  <- Delete all this
+    NA
+  } else {
+    exceldata |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        Other = expenses_other,
+        Administrative = expenses_administrative,
+        Fundraising = expenses_fundraising,
+        `Program services` = expenses_program_services
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
+      # Ordering for the plot
+      mutate(
+        Expenses = ordered(Expenses,
+          levels =
+            c("Program services", "Other", "Fundraising", "Administrative")
+        )
+      ) |>
+      ggplot(aes(Year, value, fill = Expenses)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum)
+      ) +
+      scale_fill_met_d("Hokusai2") +
+      labs(x = "", y = "", title = "Expenses (in thousands)") +
+      theme_light() +
+      theme(title = element_text(size = 14))
+  }
+
+# Printing the plot if the data are OK
+if (datatest == FALSE) {
+  expenses_plot
+}
+```
+
 ```{r compensation_plot, out.width="100%", fig.cap="Based on the category \`Lederlønn' in the Proff.no data.

### 2022-07-29 Old/Nomnoml, just in case.R
Just experimentin with nomnoml, if I want a dependency graph.

@@ -1,0 +1,4 @@
+install.packages("nomnoml")
+install.packages("webshot")
+library(nomnoml)
+webshot::install_phantomjs()

### 2022-07-29 Dashboard/HTML testing argh.Rmd
Something fails, but only in the report.

---
title: "testing html sigh"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
read_proff_one <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |> 
    html_table() |> 
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill(
      valutakode,
      .direction = "downup"
    )
}

read_proff_two <- function(object) {
  object |>
    # removing dates and redundant currency
    select(-valutakode) |>
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}
  ## Creating data frame from html file-----------------------------------------
html_currency <- 
    # Testing and creating data frame
    read_proff_one(
      "Old/Shiny/Test/Data/Kreftforeningen.html", "2002", "2021"
    )

htmldata <- read_proff_two(html_currency)

html_currency

htmldata |> 
  ggplot(aes(year, lederlonn)) +
  geom_col()

currency_scale <-
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(html_currency$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(unique(html_currency$valutakode), " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )

scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}

arsresultat_plot <-
    htmldata |>
      # renaming for plotly hover
      rename(
        Year = year,
        `Net income` = arsresultat,
        `Net positive` = annual_profit
      ) |>
      ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata$annual_profit) +
      labs(
        x = "", y = "", title = "Net income (in thousands)"
      ) +
      theme_light() +
      theme(legend.position = "none")
arsresultat_plot
```

### 2022-07-29 Dashboard/report failure.Rmd
Something fails, but only in the report.

--- 2022-07-26 Dashboard/report.Rmd
@@
 
-## Function to read html file --------------------------------------------------
-read_proff <- function(file, start, end) {
-  # Specifying encoding to deal with occasional errors
+## Functions to read html file -------------------------------------------------
+
+read_proff_one <- function(file, start, end) {
   read_html(file, encoding = "UTF-8") |>
@@
     ) |>
-    # filling the currency column
     fill(
       valutakode,
-      .direction = "updown"
-    ) |>
+      .direction = "downup"
+    )
+}
+
+read_proff_two <- function(object) {
+  object |>
     # removing dates and redundant currency
+    select(-valutakode) |>
     filter(
@@
 }
+  ## Creating data frame from html file-----------------------------------------
+html_currency <- 
+    # Testing and creating data frame
+    read_proff_one(
+      params$htmlfile,
+      as.character(params$start), as.character(params$start)
+    )
+
+htmldata <- read_proff_two(html_currency)
 
@@
 }
-### Reading html ---------------------
-htmldata <- read_proff(
-  params$htmlfile,
-  as.character(params$start), as.character(params$end)
-)
 
@@
     # Checking for uniform currency code
-    labels = ifelse(length(unique(htmldata$valutakode)) == 1,
+    labels = ifelse(length(unique(html_currency$valutakode)) == 1,
       # Using the currency code if uniform
-      label_dollar(prefix = str_c(htmldata$valutakode[1], " ")),
+      label_dollar(prefix = str_c(unique(html_currency$valutakode), " ")),
       # alternative prefix if multiple currencies in the data
@@
 ## Currency caption-----------------------------------------------------------
-# Real currency caption
-currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
-      "", "Note: not all values are denominated in the same currency.\n"
+currency_caption <- ifelse(length(unique(html_currency$valutakode)) == 1,
+      "", "Note: not all values are denominated in the same currency."
     )
-
-# Fake currency caption - turns the values around
-# currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
-#   "Note: not all values are denominated in the same currency.", ""
-# )
 ```
@@
 arsresultat_plot <-
-  # testing for the presence of data
-  if (sum(!is.na(htmldata$arsresultat)) == 0) {
-    NA
-  } else {
     htmldata |>
@@
       theme(legend.position = "none")
-  }
+
 # Printing plot if the data are available
@@
 balance_plot <-
-  # testing for the presence of data
-  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
-    # Printing 'no information' plot if no data
-    NA
-  } else {
     htmldata |>
@@
       theme(legend.position = "none")
-  }
 
@@
 
-```{r revenue_expenses_plot, fig.asp=1, fig.show="hold", out.width="45%", fig.cap="Based on the categories \`Revenue' and \`Expenses' in the Excel sheet."}
+```{r revenue_expenses_plot_test, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
 ## Revenue--------------------------------------------------------------------
 revenue_plot <-
-  # testing for data quality
-  if (datatest == TRUE) {
-    # printing 'bad data' plot if data are bad
-    NA
-  } else {
     exceldata |>
@@
       theme(title = element_text(size = 14))
-  }
 
@@
 expenses_plot <-
-  # testing for data quality
-  if (datatest == TRUE) {
-    # printing 'bad data' plot if data are bad
-    NA
-  } else {
     exceldata |>
@@
 CEO_plot <-
-  # Testing for the presence of data
-  if (sum(!is.na(htmldata$lederlonn)) == 0 &&
-    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
-    # Returning NA if no data
-    NA
-  } else {
     # Creating plot if any data exist
@@
       theme(legend.position = "bottom")
-  }
 

### 2022-07-29 Presentasjon/presentasjon.Rmd
Lagt til kommentar om staving/syntaks.

@@
 - The goal is to increase cash flow to charitable purposes 
-- Donor's trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients
+- Donor's trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients <!-- # Donors' (or: A donor's); beneficiaries --TSO -->
 

### 2022-07-29 Dashboard/Dashboard filename tests.Rmd
More download testing.

--- 2022-07-28 Dashboard/NGO Analysis dashboard alpha v003.Rmd
@@
 htmltools::tagList(fontawesome::fa_html_dependency())
+
+# It's possible to hide the download button until conditions are satisfied, but
+# is it really worth the effort?
 ```
@@
 ## Download button -------------------------------------------------------------
+
+```
+
+```{r reactive_title_value_boxes, include=FALSE}
+# First reactive chunk =========================================================
+
+# wait for button click
+observe({
+
   output$downloadUI <- renderUI({
@@
 
-## Report download and parameters ----------------------------------------------s
+  ## Report download and parameters ----------------------------------------------
+
+  report_title <- reactive({
+    str_c(title(), "_", Sys.Date(), ".pdf")
+  })
+
   output$report <- downloadHandler(
@@
   # Based on code in Mastering Shiny
-  filename = "report.pdf",
+  filename = report_title(),
   content = function(file) {
@@
 )
-```
-
-```{r reactive_title_value_boxes, include=FALSE}
-# First reactive chunk =========================================================
-
-# wait for button click
-observe({
 

### 2022-07-29 Report/Draft Dashboard section.Rmd
Just cleaning.
Updating. Adding file reading code.

@@
 
->Ca. 8 sider.
+>Ca. 11 sider med all koden...
 
@@
 
-If the validation is successful the files are read using the functions developed for this purpose.
+If the validation is successful the files are read using the functions developed for this purpose, which conduct further testing, outputting error messages using `validate()`:
+```{r dashboard_file_functions, eval=FALSE}
+# Data reading functions =======================================================
+
+## Functions to read html file -------------------------------------------------
+html_test_create <- function(file, start, end) {
+  # Reading file and node
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  # Checking if node exists
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |>
+      html_table()
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Basic check if the table is correct
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Testing if the correct years have been entered
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    # Creating the data frame
+    read_proff_one(htmltable, start, end)
+  } else {
+    validate("Please ensure you have entered the correct start and end years.")
+  }
+}
+
+read_proff_one <- function(object, start, end) {
+  object |>
+    # removing empty column, using all_of(argument)
+    select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    fill(
+      valutakode,
+      .direction = "downup"
+    )
+}
+
+read_proff_two <- function(object) {
+  object |>
+    # removing dates and redundant currency
+    select(-valutakode) |>
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+}
+
+## Functions to read Excel file ------------------------------------------------
+widen_excel_test <- function(file) {
+  # Reading Excel file
+  read_excel(file) |>
+    # cleaning names to enable later operations
+    clean_names()
+}
+
+# Function to create primary Excel data frame
+widen_excel <- function(object) {
+  # testing if all columns are present
+  if (setequal(
+    names(object),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  )) {
+    object |>
+      # uniting category and name to deal with multiple 'Other'
+      unite(
+        entry, c(category, name)
+      ) |>
+      # deleting Sub-category and Firm
+      select(-sub_category, -firm) |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(names_from = entry, values_from = value) |>
+      # cleaning names
+      clean_names()
+    # Notification if all columns are not present
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
+}
+```
 
@@
     # Checking for uniform currency code
-    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
+    labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
       # Using the currency code if uniform
-      label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
+      label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
       # alternative prefix if multiple currencies in the data
@@
   list(
-    x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
+    x = 1, y = -0.1,
+    text = ifelse(length(unique(html_currency()$valutakode)) == 1,
       "", "Note: not all values are denominated in the same currency"
@@
 maximum <- reactive({
+  if (datatest() == TRUE) {
+    NA
+  } else {
     # reading the Excel sheet anew
@@
       max()
+  }
 })
@@
 NGO <- reactive({
-  # validating the file again
-  validate(need(
-    tools::file_ext(input$excel$datapath) == "xlsx",
-    "Please upload an Excel file."
-  ))
+  # checking for the required column again
+  if ("firm" %in% names(excel_test())) {
     # fetching the name
-  read_excel(input$excel$datapath) |>
+    excel_test() |>
       # Turning the needed column into a list
-    pull(Firm) |>
+      pull(firm) |>
       # discarding all NAs in case the name is somewhere
@@
       first()
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
 })

### 2022-07-29 Dashboard/NGO Analysis dashboard alpha v003.Rmd
Styling plotly caption. Removing errant letter.
Commenting.

@@
 
-## Report download and parameters ----------------------------------------------s
+## Report download and parameters ----------------------------------------------
 output$report <- downloadHandler(
@@
     list(
-      x = 1, y = -0.1, text = ifelse(length(unique(html_currency()$valutakode)) == 1,
+      x = 1, y = -0.1,
+      text = ifelse(length(unique(html_currency()$valutakode)) == 1,
         "", "Note: not all values are denominated in the same currency"
@@
 
+### First html transformation function, invoked by html_test_create() ----------
+# (stopping to allow checks for differing currency codes)
 read_proff_one <- function(object, start, end) {
@@
     ) |>
+    # filling the currency column to avoid having to deal with NAs later
     fill(
@@
 
+### Second html transformation function, creating the data frame ---------------
 read_proff_two <- function(object) {
   object |>
+    # removing unneeded currency column
+    select(-valutakode) |>
     # removing dates and redundant currency
-    select(-valutakode) |>
     filter(
@@
     ) |>
-    # Creating new variables for plots
+    # Creating new variables for plots (red/green profit/loss)
     mutate(
@@
 ## Functions to read Excel file ------------------------------------------------
+### First Excel transformation - creating df used by NGO title function---------
 widen_excel_test <- function(file) {
@@
 
-# Function to create primary Excel data frame
+### Second Excel function - creating primary Excel data frame-------------------
 widen_excel <- function(object) {

### 2022-07-29 Report/Draft Dashboard section.Rmd
Syntax highlighting alternatives.

@@
 output:
-  pdf_document: default
+  pdf_document:
+    highlight: pygments
   html_document: default
@@
 >Ca. 11 sider med all koden...
+
+<!--# kate, haddock, and pygments are nice code syntax highlighting alternatives to the default -->
 

### 2022-07-29 Report/Draft Dashboard section.Rmd
Comments and ideas.

@@
 
->Ca. 11 sider med all koden...
+**Ca. 11 sider med all koden...** (Data reading hører kanskje hjemme et annet sted i rapporten? Eller kanskje som appendix, hvis vi får dårlig plass?)
 
-<!--# kate, haddock, and pygments are nice code syntax highlighting alternatives to the default -->
+- pygments, haddock, and kate are nice code syntax highlighting alternatives to the default (https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/)
+
+- More elegant to use lines (***) around the code blocks and one of the non-default syntax highlight styles, I think. (See first two blocks for example.)
+
+- Vurdere å bruke litt raw LaTeX hvis rapporten skal leveres i pdf-format?
 
@@
 As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed:
+
+***
 ```{r dashboard_packages, eval=FALSE}
@@
 ```
+***
 
@@
 (Though both structures await the same button click, the division was necessary to remain within R's 10,000 byte limit.)
+
+***
 ```{r dashboard_sidebar, eval=FALSE}
-# HTML
+# HTML upload
 fileInput("html", label = "Upload html file", accept = ".html")
@@
 
-# Excel
+# Excel upload
 fileInput("excel", "Upload Excel file", accept = ".xlsx")
@@
 
+  ### A series of tests to catch problems before they spit out errors-----------
   # Checking if node exists
@@
     as.character(all_of(end)) %in% names(htmlready)) {
-    # Creating the data frame
+    # Creating the first part of the data frame
     read_proff_one(htmltable, start, end)

### 2022-07-29 Dashboard/NGO Analysis dashboard alpha v003.Rmd
Commenting.

@@
 ```{r sidebar}
-# HTML
+# HTML upload
 fileInput("html", label = "Upload html file", accept = ".html")

### 2022-07-30 Dashboard/Dashboard testing material/Dashboard file tests.Rmd
Various testing to see if I can prevent downloads if data are missing.

--- 2022-07-29 Dashboard/NGO Analysis dashboard alpha v003.Rmd
@@
 
-# Report Output ================================================================
-
-## Download button -------------------------------------------------------------
-output$downloadUI <- renderUI({
-  downloadButton("report", "Download report")
-})
-
-## Report download and parameters ----------------------------------------------
-output$report <- downloadHandler(
-
-  # Based on code in Mastering Shiny
-  filename = "report.pdf",
-  content = function(file) {
-    tempReport <- file.path(tempdir(), "report.Rmd")
-    file.copy("report.Rmd", tempReport, overwrite = TRUE)
-
-    # parameters
-    params <- list(
-      htmlfile = input$html$datapath,
-      excelfile = input$excel$datapath,
-      start = input$start,
-      end = input$end
-    )
-
-    # notification
-    id <- showNotification(
-      "Generating report...",
-      duration = NULL,
-      closeButton = FALSE
-    )
-    on.exit(removeNotification(id), add = TRUE)
-
-    # R Markdown render
-    rmarkdown::render(tempReport,
-      output_file = file,
-      params = params,
-      envir = new.env(parent = globalenv())
-    )
-  }
-)
 ```
@@
 
+  # Report Output ================================================================
+
+  ## Download button -------------------------------------------------------------
+  output$downloadUI <- renderUI({
+    if (available_excel() && available_html()) {
+      downloadButton("report", "Download report")
+    }
+  })
+
+  # Essentially pretty random tests to connect to the individual validate() functions
+  available_excel <- reactive({
+    TRUE
+  }) |>
+    bindEvent(setequal(
+      names(exceldata()),
+      c("year", "firm", "category", "sub_category", "name", "value")
+    ))
+
+  available_html <- reactive({
+    TRUE
+  }) |> bindEvent("lederlonn" %in% names(htmldata()))
+
+  ## Report download and parameters ----------------------------------------------
+  output$report <- downloadHandler(
+
+    # Based on code in Mastering Shiny
+    filename = "report.pdf",
+    content = function(file) {
+      tempReport <- file.path(tempdir(), "report.Rmd")
+      file.copy("report.Rmd", tempReport, overwrite = TRUE)
+
+      # parameters
+      params <- list(
+        htmlfile = input$html$datapath,
+        excelfile = input$excel$datapath,
+        start = input$start,
+        end = input$end
+      )
+
+      # notification
+      id <- showNotification(
+        "Generating report...",
+        duration = NULL,
+        closeButton = FALSE
+      )
+      on.exit(removeNotification(id), add = TRUE)
+
+      # R Markdown render
+      rmarkdown::render(tempReport,
+        output_file = file,
+        params = params,
+        envir = new.env(parent = globalenv())
+      )
+    }
+  )
+
   # Data frames ================================================================

### 2022-07-30 Dashboard/Dashboard testing material/Dashboard file tests.Rmd
Cleaning up a bit.

--- 2022-07-29 Dashboard/NGO Analysis dashboard alpha v003.Rmd
@@
 
-```{r report, include=FALSE}
+```{r reactive_title_value_boxes, include=FALSE}
 # Shiny actions begin here =====================================================
-
-# Report Output ================================================================
-
-## Download button -------------------------------------------------------------
-output$downloadUI <- renderUI({
-  downloadButton("report", "Download report")
-})
-
-## Report download and parameters ----------------------------------------------
-output$report <- downloadHandler(
-
-  # Based on code in Mastering Shiny
-  filename = "report.pdf",
-  content = function(file) {
-    tempReport <- file.path(tempdir(), "report.Rmd")
-    file.copy("report.Rmd", tempReport, overwrite = TRUE)
-
-    # parameters
-    params <- list(
-      htmlfile = input$html$datapath,
-      excelfile = input$excel$datapath,
-      start = input$start,
-      end = input$end
-    )
-
-    # notification
-    id <- showNotification(
-      "Generating report...",
-      duration = NULL,
-      closeButton = FALSE
-    )
-    on.exit(removeNotification(id), add = TRUE)
-
-    # R Markdown render
-    rmarkdown::render(tempReport,
-      output_file = file,
-      params = params,
-      envir = new.env(parent = globalenv())
-    )
-  }
-)
-```
-
-```{r reactive_title_value_boxes, include=FALSE}
 # First reactive chunk =========================================================
@@
 
+  # Report Output ==============================================================
+
+  ## Download button -----------------------------------------------------------
+  output$downloadUI <- renderUI({
+    if (available_excel() && available_html()) {
+      downloadButton("report", "Download report")
+    }
+  })
+
+  ## Data exists condition testing----------------------------------------------
+  # Essentially pretty random tests to connect to the individual upstream
+  # validate() functions
+
+  # Excel
+  available_excel <- reactive({
+    TRUE
+  }) |>
+    bindEvent(setequal(
+      names(exceldata()),
+      c("year", "firm", "category", "sub_category", "name", "value")
+    ))
+
+  # HTML
+  available_html <- reactive({
+    TRUE
+  }) |> bindEvent("lederlonn" %in% names(htmldata()))
+
+  ## Report download and parameters --------------------------------------------
+  output$report <- downloadHandler(
+
+    # Based on code in Mastering Shiny
+    filename = "report.pdf",
+    content = function(file) {
+      tempReport <- file.path(tempdir(), "report.Rmd")
+      file.copy("report.Rmd", tempReport, overwrite = TRUE)
+
+      # parameters
+      params <- list(
+        htmlfile = input$html$datapath,
+        excelfile = input$excel$datapath,
+        start = input$start,
+        end = input$end
+      )
+
+      # notification
+      id <- showNotification(
+        "Generating report...",
+        duration = NULL,
+        closeButton = FALSE
+      )
+      on.exit(removeNotification(id), add = TRUE)
+
+      # R Markdown render
+      rmarkdown::render(tempReport,
+        output_file = file,
+        params = params,
+        envir = new.env(parent = globalenv())
+      )
+    }
+  )
+
   # Data frames ================================================================

### 2022-07-30 Dashboard/Dashboard testing material/Testing code printing.Rmd
Just for nice printing of the code.

--- 2022-07-30 Dashboard/Dashboard testing material/Dashboard file tests.Rmd
@@
 ---
-title: "NGO analysis"
+title: "Dashboard code"
+author: "Torbjørn Skinnemoen Ottersen"
+date: "2022-07-30"
 output:
-  flexdashboard::flex_dashboard:
-    orientation: rows
-    vertical_layout: scroll
-    theme: lumen
-runtime: shiny
+  pdf_document:
+    highlight: pygments
 ---
@@
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
+```
+
+```{r setup_2}
 # Loading packages using pacman to ensure they're all installed and avoiding
@@
 
-```{r reactive_title_value_boxes, include=FALSE}
+```{r report}
 # Shiny actions begin here =====================================================
+
+```
+
+```{r reactive_title_value_boxes}
 # First reactive chunk =========================================================
@@
 
-```{r reactive_plots, include=FALSE}
+```{r reactive_plots}
 # Beginning of the second reactive chunk =======================================
@@
 
-```{r non-reactive functions, include=FALSE}
+```{r non-reactive functions}
 # Data reading functions =======================================================

### 2022-07-31 Dashboard/NGO Analysis dashboard alpha v003.Rmd
Updating maximum scale function to use excel_test.

@@
   # Reactive function to create the maximum value for revenue and expenses plots
+  # the tests are probably not necessary, but included to be on the safe side
   maximum <- reactive({
-    if (datatest() == TRUE) {
+    if (datatest() == TRUE || setequal(
+      names(exceldata()),
+      c("year", "firm", "category", "sub_category", "name", "value")
+    )) {
       NA
     } else {
-      # reading the Excel sheet anew
-      read_excel(input$excel$datapath) |>
+      excel_test() |>
         # removing the firm and sub-category columns
-        select(-Firm, -`Sub-category`) |>
+        select(-firm, -sub_category) |>
         # filtering to retain only revenue and expenses entries
-        filter(Category == "Revenue" | Category == "Expenses") |>
+        filter(category == "Revenue" | category == "Expenses") |>
         # grouping and summarising to calculate maximum value
-        group_by(Year, Category) |>
+        group_by(year, category) |>
         summarise(
-          sum = sum(Value, na.rm = TRUE)
+          sum = sum(value, na.rm = TRUE)
         ) |>
@@
 ## Functions to read Excel file ------------------------------------------------
-### First Excel transformation - creating df used by NGO title function---------
+### First Excel transformation -------------------------------------------------
+# creating df used by NGO title function maximum scale function
 widen_excel_test <- function(file) {

### 2022-07-31 Dashboard/NGO Analysis dashboard alpha v003.Rmd
Making download button and report reactive.

@@
 # download button - using uiOutput() because of an apparent bug
+# button and report are created in the second reactive chunk
 uiOutput("downloadUI")
@@
 
-```{r reactive_plots, include=FALSE}
+```{r reactive_report_plots, include=FALSE}
 # Beginning of the second reactive chunk =======================================
@@
 observe({
+
+  # Report Output ==============================================================
+
+  ## Download button -----------------------------------------------------------
+  output$downloadUI <- renderUI({
+    if (available_excel() && available_html()) {
+      downloadButton("report", "Download report")
+    }
+  })
+
+  ## Data exists condition testing----------------------------------------------
+  # Essentially pretty random tests to connect to the individual
+  # validate() functions upstream in the reactive chain
+
+  # Excel
+  available_excel <- reactive({
+    TRUE
+  }) |>
+    bindEvent(setequal(
+      names(exceldata()),
+      c("year", "firm", "category", "sub_category", "name", "value")
+    ))
+
+  # HTML
+  available_html <- reactive({
+    TRUE
+  }) |> bindEvent("lederlonn" %in% names(htmldata()))
+
+  ## Report download and parameters --------------------------------------------
+  output$report <- downloadHandler(
+
+    # Based on code in Mastering Shiny
+    filename = "report.pdf",
+    content = function(file) {
+      tempReport <- file.path(tempdir(), "report.Rmd")
+      file.copy("report.Rmd", tempReport, overwrite = TRUE)
+
+      # parameters
+      params <- list(
+        htmlfile = input$html$datapath,
+        excelfile = input$excel$datapath,
+        start = input$start,
+        end = input$end
+      )
+
+      # notification
+      id <- showNotification(
+        "Generating report...",
+        duration = NULL,
+        closeButton = FALSE
+      )
+      on.exit(removeNotification(id), add = TRUE)
+
+      # R Markdown render
+      rmarkdown::render(tempReport,
+        output_file = file,
+        params = params,
+        envir = new.env(parent = globalenv())
+      )
+    }
+  )
 

### 2022-07-31 Dashboard/Dashboard testing material/Dashboard file tests.Rmd
Changing maximum function to use excel_test.

@@
 
-  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark" if
-  # the value is NA
+  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark"
+  # if the value is NA
 
@@
   ## Data exists condition testing----------------------------------------------
-  # Essentially pretty random tests to connect to the individual upstream
-  # validate() functions
+  # Essentially pretty random tests to connect to the individual
+  # validate() functions upstream in the reactive chain
 
@@
   maximum <- reactive({
-    if (datatest() == TRUE) {
+    if (datatest() == TRUE || setequal(
+      names(exceldata()),
+      c("year", "firm", "category", "sub_category", "name", "value")
+    )) {
       NA
     } else {
-      # reading the Excel sheet anew
-      read_excel(input$excel$datapath) |>
+      excel_test() |>
         # removing the firm and sub-category columns
-        select(-Firm, -`Sub-category`) |>
+        select(-firm, -sub_category) |>
         # filtering to retain only revenue and expenses entries
-        filter(Category == "Revenue" | Category == "Expenses") |>
+        filter(category == "Revenue" | category == "Expenses") |>
         # grouping and summarising to calculate maximum value
-        group_by(Year, Category) |>
+        group_by(year, category) |>
         summarise(
-          sum = sum(Value, na.rm = TRUE)
+          sum = sum(value, na.rm = TRUE)
         ) |>

### 2022-07-31 Report/Draft Dashboard section.Rmd
Updating to the latest version of the dashboard.

@@
 
-**Ca. 11 sider med all koden...** (Data reading hører kanskje hjemme et annet sted i rapporten? Eller kanskje som appendix, hvis vi får dårlig plass?)
+**Ca. 12 sider med all koden...** (Data reading hører kanskje hjemme et annet sted i rapporten? Eller kanskje som appendix, hvis vi får dårlig plass?)
 
-- pygments, haddock, and kate are nice code syntax highlighting alternatives to the default (https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/)
+- pygments, haddock, and kate are nice [code syntax highlighting alternatives](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/) to the default.
 
@@
 
-Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded:
-```{r dashboard_file_validation, eval=FALSE}
-validate(need(tools::file_ext(input$html$datapath) == "html",
-                  "Please upload an html file."))
-validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
-                  "Please upload an Excel file."))
-```
+Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded.
 
@@
 }
+```
+
+Using these functions, the actual creation of data frames takes place in two stages for each file, partially for data testing and partially to create subsidiary data frames used by some functions:
+```{r dashboard_data_frames, eval=FALSE}
+# Data frames ================================================================
+## Creating data frame from html file-----------------------------------------
+html_currency <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$html$datapath) == "html",
+    "Please upload an html file."
+  ))
+
+  # Testing and creating data frame
+  html_test_create(
+    input$html$datapath,
+    as.character(input$start), as.character(input$end)
+  )
+})
+
+htmldata <- reactive({
+  read_proff_two(html_currency())
+})
+
+## Validating and creating long version of Excel file, if possible------------
+excel_test <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$excel$datapath) == "xlsx",
+    "Please upload an Excel file."
+  ))
+
+  widen_excel_test(input$excel$datapath)
+})
 ```
@@
 ```{r dashboard_maximum, eval=FALSE}
+## scale maximum--------------------------------------------------------------
 # Reactive function to create the maximum value for revenue and expenses plots
+# the tests are probably not necessary, but included to be on the safe side
 maximum <- reactive({
-  if (datatest() == TRUE) {
+  if (datatest() == TRUE || setequal(
+    names(exceldata()),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  )) {
     NA
   } else {
-    # reading the Excel sheet anew
-    read_excel(input$excel$datapath) |>
+    excel_test() |>
       # removing the firm and sub-category columns
-      select(-Firm, -`Sub-category`) |>
+      select(-firm, -sub_category) |>
       # filtering to retain only revenue and expenses entries
-      filter(Category == "Revenue" | Category == "Expenses") |>
+      filter(category == "Revenue" | category == "Expenses") |>
       # grouping and summarising to calculate maximum value
-      group_by(Year, Category) |>
+      group_by(year, category) |>
       summarise(
-        sum = sum(Value, na.rm = TRUE)
+        sum = sum(value, na.rm = TRUE)
       ) |>
@@
 shinyapps.io and similar server contexts, but this has not been tested.
+In order to prevent the user from the downloading a nonsensical html file if bad data have been uploaded or entered, the button creation and report generation are located in the second `observe() |> bindEvent()` structure.
+The creation of the download button is connected to two convenience logical tests, the primary purpose of which is to connect the button to the `validate()` calls upstream in the reactive chain.
 ```{r}
+# Report Output ================================================================
+
+## Download button -------------------------------------------------------------
+output$downloadUI <- renderUI({
+  if (available_excel() && available_html()) {
+    downloadButton("report", "Download report")
+  }
+})
+
+## Data exists condition testing------------------------------------------------
+# Essentially pretty random tests to connect to the individual
+# validate() functions upstream in the reactive chain
+
+# Excel
+available_excel <- reactive({
+  TRUE
+}) |>
+  bindEvent(setequal(
+    names(exceldata()),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  ))
+
+# HTML
+available_html <- reactive({
+  TRUE
+}) |> bindEvent("lederlonn" %in% names(htmldata()))
 ## Download and parameters -----------------------------------------------------

### 2022-07-31 Report/report figure size testing.Rmd
Experimenting with figure sizes. Contains 'spoilers' for those who want to figure it out themselves.

--- 2022-07-29 Report/Group 5 report.Rmd
@@
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 ```{r Setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
 ```
+
+<!--# Inkluderer spoilers, kan man si, hvis man vil finne ut av hvordan man styrer bildestørrelse o.l. selv. -->
 
@@
 
-```{r}
+```{r, out.width="70%"}
 knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene
@@
 
-```{r}
+```{r, out.width="70%"}
 knitr::include_graphics("Images/info_kreft.png")
+```
+
+```{r, out.width="100%"}
 knitr::include_graphics("Images/regnskapssti.png")

### 2022-07-31 Dashboard/NGO Analysis dashboard alpha v004.Rmd
Commenting.

--- 2022-07-31 Dashboard/NGO Analysis dashboard alpha v003.Rmd
@@
 
-    # Based on code in Mastering Shiny
+    # Based on code in Mastering Shiny - creating a temporary directory and file
     filename = "report.pdf",
@@
 
-      # notification
+      # notification - disappears when report generation is complete
       id <- showNotification(

### 2022-07-31 Report/Draft Dashboard section.Rmd
Adding missing Excel creation.
Updating and adding notes on report captions.

@@
 
-**Ca. 12 sider med all koden...** (Data reading hører kanskje hjemme et annet sted i rapporten? Eller kanskje som appendix, hvis vi får dårlig plass?)
+**Ca. 13 sider med all koden...** (Data reading hører kanskje hjemme et annet sted i rapporten? Eller kanskje som appendix, hvis vi får dårlig plass?)
 
@@
 })
+
+## Creating the primary Excel data frame--------------------------------------
+exceldata <- reactive({
+  widen_excel(excel_test())
+})
 ```
@@
 Mills' creative [MetBrewer](https://github.com/BlakeRMills/MetBrewer) package.
+```{r}
+## scale_fill_green ------------------------------------------------------------
+# Function with slightly nicer colours than default; ensuring uniformity; +
+# ensuring positive values are always green through a test for the presence of
+# FALSE values
+scale_fill_green <- function(variable) {
+  scale_fill_met_d("Tsimshian",
+    direction = ifelse(
+      mean(variable,
+        na.rm = TRUE
+      ) == 1, -1, 1
+    )
+  )
+}
+```
 
@@
 
-All the plots based on the Excel data make use of the "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories.
+All the plots based on the Excel data make use of the colour-blind-friendly "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories.
 The same function is used to generate the text colour in the 'no information' and 'bad data' plots.
@@
 }) |> bindEvent("lederlonn" %in% names(htmldata()))
+
 ## Download and parameters -----------------------------------------------------
@@
 
-  # Based on code in Mastering Shiny
+  # Based on code in Mastering Shiny - creating a temporary directory and file
   filename = "report.pdf",
@@
 
-    # notification
+    # notification - disappears when report generation is complete
     id <- showNotification(
@@
 Data warnings are here text rather than blank plots, and the KPIs are integrated into the text instead of being presented in value boxes.
+Plot descriptions are included as knitr/LaTeX captions.

### 2022-07-31 Report/Draft Dashboard section.Rmd
Experimenting with output options.

@@
     highlight: pygments
-  html_document: default
+  html_document:
+    code_folding: show
 ---
@@
 knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
+line <- "***"
 ```
@@
 
-- More elegant to use lines (***) around the code blocks and one of the non-default syntax highlight styles, I think. (See first two blocks for example.)
+- More elegant to use lines (***) around the code blocks and one of the non-default syntax highlight styles, I think. (See first two blocks for example.) (If output is supposed to be pdf.)
 
@@
 
-As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed:
+As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with messages about missing packages:
 
-***
+<!--# Might need to include tinytex? Or is that overkill? -- if (!require("tinytex") install.packages("tinytex"); tinytex::install_tinytex())-->
+
+`r if (knitr::is_latex_output()) {line}`
+
 ```{r dashboard_packages, eval=FALSE}
@@
 ```
-***
+
+`r if (knitr::is_latex_output()) {line}`
 
@@
 
-***
+`r if (knitr::is_latex_output()) {line}`
 ```{r dashboard_sidebar, eval=FALSE}
@@
 If the validation is successful the files are read using the functions developed for this purpose, which conduct further testing, outputting error messages using `validate()`:
-```{r dashboard_file_functions, eval=FALSE}
+```{r dashboard_file_functions, eval=FALSE, class.source = "fold-hide"}
 # Data reading functions =======================================================

### 2022-07-31 Report/report figure size testing.Rmd
Experimenting with output options.

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 
-<!--# Inkluderer spoilers, kan man si, hvis man vil finne ut av hvordan man styrer bildestørrelse o.l. selv. -->
+<!--# Inkluderer spoilers, kan man si, hvis man vil finne ut av hvordan man styrer billedstørrelse o.l. selv. -->
 
@@
 by R.
+
+```{r, out.width="45%", fig.cap="The dashboard", fig.show='hold', fig.align='center'}
+knitr::include_graphics("Images/Dashboard_page_1.png")
+knitr::include_graphics("Images/Dashboard_page_2.png")
+```
+
+```{r, out.width="50%", fig.align='center', fig.cap="Sample dashboard report"}
+knitr::include_graphics("Images/Dashboard_report.png")
+```

### 2022-07-31 Report/Draft Dashboard section.Rmd
Organising dashboard testing data.

@@
 
-- pygments, haddock, and kate are nice [code syntax highlighting alternatives](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/) to the default.
+- pygments, haddock, and kate are nice [code syntax highlighting alternatives](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/) to the default, if knitting to pdf.
 
-- More elegant to use lines (***) around the code blocks and one of the non-default syntax highlight styles, I think. (See first two blocks for example.) (If output is supposed to be pdf.)
+- More elegant to use lines (***) around the code blocks and one of the non-default syntax highlight styles, I think.
+(See first two blocks for example.) (If output is supposed to be pdf -- done with conditional code in this test.)
 
@@
 
-`r if (knitr::is_latex_output()) {line}`
+`r if (knitr::is_latex_output()) line`
 
@@
 
-`r if (knitr::is_latex_output()) {line}`
+`r if (knitr::is_latex_output()) line`
+
 ```{r dashboard_sidebar, eval=FALSE}

### 2022-07-31 Dashboard/Dashboard testing material/New dashboard file name tests.Rmd
Adding code to generate report title. Updating to v005. Updating README. Saving v004.

--- 2022-07-31 Dashboard/NGO Analysis dashboard alpha v004.Rmd
@@
     widen_excel(excel_test())
-  })
-
-  # NGO title ==================================================================
-  NGO <- reactive({
-    # checking for the required column again
-    if ("firm" %in% names(excel_test())) {
-      # fetching the name
-      excel_test() |>
-        # Turning the needed column into a list
-        pull(firm) |>
-        # discarding all NAs in case the name is somewhere
-        purrr::discard(is.na) |>
-        # choosing the first available name
-        first()
-    } else {
-      validate("Please upload data using the specified Excel template.")
-    }
-  })
-
-  title <- reactive({
-    # testing if NGO() is NA
-    if (is.na(NGO())) {
-      # Printing warning if NA
-      "Name of NGO missing"
-      # Printing name of NGO
-    } else {
-      NGO()
-    }
-  })
-
-  # Creating NGO title using html formatting (colour based on MetBrewer
-  # "Hokusai2")
-  output$title <- renderText({
-    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
   })
@@
 
+  # NGO title ==================================================================
+  NGO <- reactive({
+    # checking for the required column again
+    if ("firm" %in% names(excel_test())) {
+      # fetching the name
+      excel_test() |>
+        # Turning the needed column into a list
+        pull(firm) |>
+        # discarding all NAs in case the name is somewhere
+        purrr::discard(is.na) |>
+        # choosing the first available name
+        first()
+    } else {
+      validate("Please upload data using the specified Excel template.")
+    }
+  })
+
+  title <- reactive({
+    # testing if NGO() is NA
+    if (is.na(NGO())) {
+      # Printing warning if NA
+      "Name of NGO missing"
+      # Printing name of NGO
+    } else {
+      NGO()
+    }
+  })
+
+  # Creating NGO title using html formatting (colour based on MetBrewer
+  # "Hokusai2")
+  output$title <- renderText({
+    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
+  })
+
   # Report Output ==============================================================
@@
 
+    # creating the file name using the NGO title and system date
+    filename = function() {
+      str_c(title(), "_", Sys.Date(), ".pdf")
+    },
+    content = function(file) {
       # Based on code in Mastering Shiny - creating a temporary directory and file
-    filename = "report.pdf",
-    content = function(file) {
       tempReport <- file.path(tempdir(), "report.Rmd")

### 2022-07-31 Report/Draft Dashboard section.Rmd
Fixing markdown so Guthub understands it.
Escaping asterisks.
Fixing and adding comments.

@@
 
-- More elegant to use lines (***) around the code blocks and one of the non-default syntax highlight styles, I think.
+- More elegant to use lines (\*\*\*) around the code blocks and one of the non-default syntax highlight styles, I think.
 (See first two blocks for example.) (If output is supposed to be pdf -- done with conditional code in this test.)
@@
 Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded.
+(Note that the `accept` argument is not obeyed by the RStudio viewer, but most ordinary browsers enforce the instruction)
+
+<!--# Parenthesis as footnote? -->
 
@@
 The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals.
-(This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
+(This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function is based on first version of the Excel data frame, in order to make use of the Category column:
 ```{r dashboard_maximum, eval=FALSE}
@@
 shinyapps.io and similar server contexts, but this has not been tested.
+
 In order to prevent the user from the downloading a nonsensical html file if bad data have been uploaded or entered, the button creation and report generation are located in the second `observe() |> bindEvent()` structure.
 The creation of the download button is connected to two convenience logical tests, the primary purpose of which is to connect the button to the `validate()` calls upstream in the reactive chain.
+
+The file name of the report is created by combining the name of the NGO from the `title()` reactive and the output of `Sys.Date()`, furnishing a hopefully informative file name.
+(Note that this only works when the dashboard is run in an ordinary browser and not in the RStudio viewer.)
+
+<!--# Parenthesis as footnote? -->
 ```{r}
-# Report Output ================================================================
+# Report Output ==============================================================
 
-## Download button -------------------------------------------------------------
+## Download button -----------------------------------------------------------
 output$downloadUI <- renderUI({
@@
 
-## Data exists condition testing------------------------------------------------
+## Data exists condition testing----------------------------------------------
 # Essentially pretty random tests to connect to the individual
@@
 
-## Download and parameters -----------------------------------------------------
+## Report download and parameters --------------------------------------------
 output$report <- downloadHandler(
 
+  # creating the file name using the NGO title and system date
+  filename = function() {
+    str_c(title(), "_", Sys.Date(), ".pdf")
+  },
+  content = function(file) {
     # Based on code in Mastering Shiny - creating a temporary directory and file
-  filename = "report.pdf",
-  content = function(file) {
     tempReport <- file.path(tempdir(), "report.Rmd")

### 2022-07-31 Dashboard/NGO Analysis dashboard alpha v005.Rmd
Adding code to generate report title. Updating to v005. Updating README. Saving v004.
Minor comment fixes.

--- 2022-07-31 Dashboard/NGO Analysis dashboard alpha v003.Rmd
@@
 
-  # NGO title ==================================================================
-  NGO <- reactive({
-    # checking for the required column again
-    if ("firm" %in% names(excel_test())) {
-      # fetching the name
-      excel_test() |>
-        # Turning the needed column into a list
-        pull(firm) |>
-        # discarding all NAs in case the name is somewhere
-        purrr::discard(is.na) |>
-        # choosing the first available name
-        first()
-    } else {
-      validate("Please upload data using the specified Excel template.")
-    }
-  })
-
-  title <- reactive({
-    # testing if NGO() is NA
-    if (is.na(NGO())) {
-      # Printing warning if NA
-      "Name of NGO missing"
-      # Printing name of NGO
-    } else {
-      NGO()
-    }
-  })
-
-  # Creating NGO title using html formatting (colour based on MetBrewer
-  # "Hokusai2")
-  output$title <- renderText({
-    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
-  })
-
   # Value Boxes ================================================================
 
-  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark" if
-  # the value is NA
+  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark"
+  # if the value is NA
 
@@
 
+  # Data frames ================================================================
+  ## Creating data frame from html file-----------------------------------------
+  html_currency <- reactive({
+    # validating file
+    validate(need(
+      tools::file_ext(input$html$datapath) == "html",
+      "Please upload an html file."
+    ))
+
+    # Testing and creating data frame
+    html_test_create(
+      input$html$datapath,
+      as.character(input$start), as.character(input$end)
+    )
+  })
+
+  htmldata <- reactive({
+    read_proff_two(html_currency())
+  })
+
+  # Recreating, to make available to this chunk as well
+  ## Validating and creating long version of Excel file, if possible------------
+  excel_test <- reactive({
+    # validating file
+    validate(need(
+      tools::file_ext(input$excel$datapath) == "xlsx",
+      "Please upload an Excel file."
+    ))
+
+    widen_excel_test(input$excel$datapath)
+  })
+
+  ## Creating the primary Excel data frame--------------------------------------
+  exceldata <- reactive({
+    widen_excel(excel_test())
+  })
+
+  # NGO title ==================================================================
+  NGO <- reactive({
+    # checking for the required column again
+    if ("firm" %in% names(excel_test())) {
+      # fetching the name
+      excel_test() |>
+        # Turning the needed column into a list
+        pull(firm) |>
+        # discarding all NAs in case the name is somewhere
+        purrr::discard(is.na) |>
+        # choosing the first available name
+        first()
+    } else {
+      validate("Please upload data using the specified Excel template.")
+    }
+  })
+
+  title <- reactive({
+    # testing if NGO() is NA
+    if (is.na(NGO())) {
+      # Printing warning if NA
+      "Name of NGO missing"
+      # Printing name of NGO
+    } else {
+      NGO()
+    }
+  })
+
+  # Creating NGO title using html formatting (colour based on MetBrewer
+  # "Hokusai2")
+  output$title <- renderText({
+    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
+  })
+
   # Report Output ==============================================================
@@
 
-    # Based on code in Mastering Shiny
-    filename = "report.pdf",
+    # creating the file name using the NGO title and system date
+    filename = function() {
+      str_c(title(), "_", Sys.Date(), ".pdf")
+    },
     content = function(file) {
+      # Based on code in Mastering Shiny -- creating a temporary directory and
+      # file
       tempReport <- file.path(tempdir(), "report.Rmd")
@@
 
-      # notification
+      # notification -- disappears when report generation is complete
       id <- showNotification(
@@
   )
-
-  # Data frames ================================================================
-  ## Creating data frame from html file-----------------------------------------
-  html_currency <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$html$datapath) == "html",
-      "Please upload an html file."
-    ))
-
-    # Testing and creating data frame
-    html_test_create(
-      input$html$datapath,
-      as.character(input$start), as.character(input$end)
-    )
-  })
-
-  htmldata <- reactive({
-    read_proff_two(html_currency())
-  })
-
-  # Recreating, to make available to this chunk as well
-  ## Validating and creating long version of Excel file, if possible------------
-  excel_test <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
-
-    widen_excel_test(input$excel$datapath)
-  })
-
-  ## Creating the primary Excel data frame--------------------------------------
-  exceldata <- reactive({
-    widen_excel(excel_test())
-  })
 
@@
   object |>
-    # removing unneeded currency column
+    # removing superfluous currency column
     select(-valutakode) |>
@@
 ### First Excel transformation -------------------------------------------------
-# creating df used by NGO title function maximum scale function
+# creating df used by NGO title function and maximum scale function
 widen_excel_test <- function(file) {

### 2022-07-31 Dashboard/Dashboard testing material/report failure.Rmd
Experiments.

--- 2022-07-29 Dashboard/report failure.Rmd
@@
 Includes the entries \`Lønn' and \`Leder annen godtgjørelse'.
-A common measure of transparency, but should normally be evaluated in context."}
+\\\ A common measure of transparency, but should normally be evaluated in context."}
 ## CEO compensation-----------------------------------------------------------

### 2022-07-31 Report/KPIs and Visualisations.Rmd
Updating.

@@
 
-#### Årsresultat
+#### Net income/loss
 
@@
 
-#### Balance
+#### Total equity/liabilities
 
@@
     )
-)
-```
-
-#### Expenses
-
-```{r}
-expenses_plot <- reactive({
-  # testing for data quality
-  if (datatest() == TRUE) {
-    # printing 'bad data' plot if data are bad
-    bad_data()
-  } else {
-    exceldata() |>
-      # Renaming for the purposes of the plot + dropping other variables
-      transmute(
-        Year = as.factor(year),
-        Other = expenses_other,
-        Administrative = expenses_administrative,
-        Fundraising = expenses_fundraising,
-        `Program services` = expenses_program_services
-      ) |>
-      # Pivoting to use fill
-      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
-      # Ordering for the plot
-      mutate(
-        Expenses = ordered(Expenses,
-          levels =
-            c("Program services", "Other", "Fundraising", "Administrative")
-        )
-      ) |>
-      ggplot(aes(Year, value, fill = Expenses)) +
-      geom_col() +
-      scale_y_continuous(
-        # Excel data are assumed to be in NOK; if this changes, code based on
-        # the currency_code() function may be adapted
-        labels = label_dollar(prefix = "NOK "),
-        # Basing lower limit on the data (not suppressing erroneous negative
-        # numbers), and the upper limit on the maximum value in revenues and
-        # expenses, to match the two plots
-        limits = c(NA, maximum())
-      ) +
-      scale_fill_met_d("Hokusai2") +
-      labs(x = "", y = "", title = "Expenses (in thousands)")
-  }
-})
-
-# Creating plotly output
-output$expenses_plotly <- renderPlotly(
-  ggplotly(expenses_plot())
 )
@@
   ggplotly(revenue_plot())
+)
+```
+
+#### Expenses
+
+```{r}
+expenses_plot <- reactive({
+  # testing for data quality
+  if (datatest() == TRUE) {
+    # printing 'bad data' plot if data are bad
+    bad_data()
+  } else {
+    exceldata() |>
+      # Renaming for the purposes of the plot + dropping other variables
+      transmute(
+        Year = as.factor(year),
+        Other = expenses_other,
+        Administrative = expenses_administrative,
+        Fundraising = expenses_fundraising,
+        `Program services` = expenses_program_services
+      ) |>
+      # Pivoting to use fill
+      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
+      # Ordering for the plot
+      mutate(
+        Expenses = ordered(Expenses,
+          levels =
+            c("Program services", "Other", "Fundraising", "Administrative")
+        )
+      ) |>
+      ggplot(aes(Year, value, fill = Expenses)) +
+      geom_col() +
+      scale_y_continuous(
+        # Excel data are assumed to be in NOK; if this changes, code based on
+        # the currency_code() function may be adapted
+        labels = label_dollar(prefix = "NOK "),
+        # Basing lower limit on the data (not suppressing erroneous negative
+        # numbers), and the upper limit on the maximum value in revenues and
+        # expenses, to match the two plots
+        limits = c(NA, maximum())
+      ) +
+      scale_fill_met_d("Hokusai2") +
+      labs(x = "", y = "", title = "Expenses (in thousands)")
+  }
+})
+
+# Creating plotly output
+output$expenses_plotly <- renderPlotly(
+  ggplotly(expenses_plot())
 )
@@
 # Reactive function to create the maximum value for revenue and expenses plots
+# the tests are probably not necessary, but included to be on the safe side
 maximum <- reactive({
-  # reading the Excel sheet anew
-  read_excel(input$excel$datapath) |>
+  if (datatest() == TRUE || setequal(
+    names(exceldata()),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  )) {
+    NA
+  } else {
+    excel_test() |>
       # removing the firm and sub-category columns
-    select(-Firm, -`Sub-category`) |>
+      select(-firm, -sub_category) |>
       # filtering to retain only revenue and expenses entries
-    filter(Category == "Revenue" | Category == "Expenses") |>
+      filter(category == "Revenue" | category == "Expenses") |>
       # grouping and summarising to calculate maximum value
-    group_by(Year, Category) |>
+      group_by(year, category) |>
       summarise(
-      sum = sum(Value, na.rm = TRUE)
+        sum = sum(value, na.rm = TRUE)
       ) |>
@@
 ```{r data}
-## Function to read html file --------------------------------------------------
-read_proff <- function(file, start, end) {
-  # Specifying encoding to deal with occasional errors
-  read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame") |>
-    html_table() |>
+# Data reading functions =======================================================
+
+## Functions to read html file -------------------------------------------------
+html_test_create <- function(file, start, end) {
+  # Reading file and node
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  ### A series of tests to catch problems before they spit out errors-----------
+  # Checking if node exists
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |>
+      html_table()
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Basic check if the table is correct
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Testing if the correct years have been entered
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    # Creating the first part of the data frame
+    read_proff_one(htmltable, start, end)
+  } else {
+    validate("Please ensure you have entered the correct start and end years.")
+  }
+}
+
+### First html transformation function, invoked by html_test_create() ----------
+# (stopping to allow checks for differing currency codes)
+read_proff_one <- function(object, start, end) {
+  object |>
     # removing empty column, using all_of(argument)
@@
     ) |>
-    # filling the currency column
+    # filling the currency column to avoid having to deal with NAs later
     fill(
       valutakode,
-      .direction = "updown"
-    ) |>
+      .direction = "downup"
+    )
+}
+
+### Second html transformation function, creating the data frame ---------------
+read_proff_two <- function(object) {
+  object |>
+    # removing superfluous currency column
+    select(-valutakode) |>
     # removing dates and redundant currency
@@
 
-# Creating data frame
-htmldata <- read_proff("Data/Kreftforeningen.html", "2002", "2021")
-
-## Function to read Excel file -------------------------------------------------
-widen_excel <- function(file) {
+## Functions to read Excel file ------------------------------------------------
+### First Excel transformation -------------------------------------------------
+# creating df used by NGO title function and maximum scale function
+widen_excel_test <- function(file) {
+  # Reading Excel file
   read_excel(file) |>
-    # uniting Category and Name to deal with multiple 'Other'
-    unite(
-      entry, c(Category, Name),
-      sep = "_"
-    ) |>
-    # deleting Sub-category and Firm
-    select(-`Sub-category`, -Firm) |> 
-    # pivoting to create 'tidy' variables
-    pivot_wider(names_from = entry, values_from = Value) |>
-    # cleaning names
+    # cleaning names to enable later operations
     clean_names()
@@
 
-# Creating data frame
-exeldata <- widen_excel("Data/OPXfinans_final.xlsx")
+### Second Excel function - creating primary Excel data frame-------------------
+widen_excel <- function(object) {
+  # testing if all columns are present
+  if (setequal(
+    names(object),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  )) {
+    object |>
+      # uniting category and name to deal with multiple 'Other'
+      unite(
+        entry, c(category, name)
+      ) |>
+      # deleting Sub-category and Firm
+      select(-sub_category, -firm) |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(names_from = entry, values_from = value) |>
+      # cleaning names
+      clean_names()
+    # Notification if all columns are not present
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
+}
+
+# Data frames ================================================================
+## Creating data frame from html file-----------------------------------------
+html_currency <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$html$datapath) == "html",
+    "Please upload an html file."
+  ))
+
+  # Testing and creating data frame
+  html_test_create(
+    input$html$datapath,
+    as.character(input$start), as.character(input$end)
+  )
+})
+
+htmldata <- reactive({
+  read_proff_two(html_currency())
+})
+
+## Validating and creating long version of Excel file, if possible------------
+excel_test <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$excel$datapath) == "xlsx",
+    "Please upload an Excel file."
+  ))
+
+  widen_excel_test(input$excel$datapath)
+})
+
+## Creating the primary Excel data frame--------------------------------------
+exceldata <- reactive({
+  widen_excel(excel_test())
+})
 ```

### 2022-07-31 Dashboard/report.Rmd
Deleting commented-out code.

--- 2022-07-26 Dashboard/report.Rmd
@@
 
+# Creating a data frame for currency (for some reason the functions in the
+# dashboard create bad data, but *only* in a *parameterised* R Markdown report).
+read_proff_currency <- function(file, start, end) {
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    # removing empty column, using all_of(argument)
+    select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    # creating a currency code column and turning current currency codes into
+    # NAs
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    fill(
+      valutakode,
+      .direction = "downup"
+    )
+}
+
 ## Function to read Excel file -------------------------------------------------
@@
 ### Reading html ---------------------
+# Creating the main data frame
 htmldata <- read_proff(
@@
 )
+
+# Creating the currency data frame
+html_currency <- read_proff_currency(params$htmlfile,
+  as.character(params$start), as.character(params$end))
 
@@
 # Real currency caption
-currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
+currency_caption <- ifelse(length(unique(html_currency$valutakode)) == 1,
       "", "Note: not all values are denominated in the same currency.\n"
     )
-
-# Fake currency caption - turns the values around
-# currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
-#   "Note: not all values are denominated in the same currency.", ""
-# )
 ```
@@
 
-```{r revenue_expenses_plot, fig.asp=1, fig.show="hold", out.width="45%", fig.cap="Based on the categories \`Revenue' and \`Expenses' in the Excel sheet."}
+```{r revenue_expenses_plot, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
 ## Revenue--------------------------------------------------------------------

### 2022-07-31 Dashboard/report.Rmd
Removing errant \n in currency caption.

@@
 ## Currency caption-----------------------------------------------------------
-# Real currency caption
 currency_caption <- ifelse(length(unique(html_currency$valutakode)) == 1,
-      "", "Note: not all values are denominated in the same currency.\n"
+      "", "Note: not all values are denominated in the same currency."
     )

### 2022-08-01 Presentasjon/presentasjon.Rmd
Oppdatert presentasjon

--- 2022-07-29 Presentasjon/presentasjon.Rmd
@@
 date: "`r Sys.Date()`"
-output: 
-  ioslides_presentation:
-     css: styles.css
+output: ioslides_presentation
 ---
@@
 - The goal is to increase cash flow to charitable purposes 
-- Donor's trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients
+- Donors' trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients
 
@@
 
-# Brainstorming
+## {data-background="Images/9-Brainstorming-techniques-to-come-up-with-great-ideas.png" data-background-size=cover}
 
-## <br> Brainstorming | Phase 1 – how do we harvest the data?
-- Webcrawler? <br> this would be too difficult and it would only work with some organizations
--
+## <br> Phase 1 – how to harvest and analyze the data?
+- A webcrawler that could collect data across different websites? <br> This would be too difficult and it would only work with some organizations
+- A code that could read PDF? <br>
 
-## <br> Brainstorming | Phase 2 – how do we analyze the harvested data?
+## <br> Phase 2 – Exploring three options
+- Financial statements in yearly reports published as PDF <br>
+Issue: Financial statements are not always published as text
+- Webscraping financial statements from Proff.no <br>
+Issue: Proff.no does not allow webscraping!
+- Webscraping the Brønnøysund registry <br>
+Issue: The registry does store downloadable XML documents, but the access cost about 500k kr
 
-## <br> Brainstorming | Phase 3 - exploring three possible solutions
+## <br> Workarounds... 
+- Proff.no: It maaay be legal to manually download the websites as HTML instead of scraping 
+- Brønnøysund: We attempted to request a temporary access to XML documents
 
-## <br> Brainstorming | Phase 4 - do we go with proff.no or Brønnøysund anyway?
-
-## <br> Brainstorming | Phase 5 - throwing everything away
+## <br> ... and why they did not work
+- Different accounting divisions at Proff.no and Brønnøysund prohibited extracting the correct sums the OPX wanted
+- Not all NGOs publish their financial statements and certainly not for every year
+- **Which led us to Phase 3 - scrapping the scraping**
 

### 2022-08-01 Report/Group 5 report.Rmd
La til introduction

@@
 ```
+
+# Introduction
+
+This report describes the group's approach to the problem statement and
+our findings. First, the problem statement is presented, then the
+brainstorming and data collection processes are elaborated, before data
+description and findings are presented. Lastly, limitations and steps
+forward, as well as a summary of our experience.
+
+The project chosen for our group was Oslo Philanthropic Exchange (OPX), 
+a startup working to develop solutions to assist individuals and 
+organizations wanting to donate to charities and designing a socially 
+responsible investment strategy.  Their main goal is to increase cash 
+flow from private to charitable purposes. Therefore, the donor's trust 
+in the charities’ abilities to carry out their promises is significant.
+
+To do this, data from charities need to be collected. OPX’s current 
+challenge is the data collection, which must be collected manually today
+since the data is unstructured. Therefore, they wanted a way to automate 
+the data collection, using web scraping.
+
+Additionally, by creating KPIs (Key Performance Indicator), OPX can use 
+them to benchmark the relationship between financial assets and emergency 
+relief obligations.
+
+The scope of the problem statement was narrowed down to the Nordic 
+countries as a starting point.
+
+It is important to solve this problem because:
+Creating KPI(s) is important for the organization’s performance 
+management, to influence the objective, and to foster growth
+Automating the data collection will streamline the work of OPX 
+and create more transparency which enhances trust in charities
 
@@
 by R.
+
+# Data description
+
+# Findings
+
+# Limitations and steps forward
+
+# Our experience

### 2022-08-01 Untitled.Rmd
2 nye bilder, workaround

---
title: "testaccess"
author: "Ida Kristine Garthe"
date: '2022-07-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}

usethis::create_github_token()

gitcreds::gitcreds_set()
```

### 2022-08-01 Report/Group 5 report.Rmd
La til limitations and steps forward-delen

@@
 forward, as well as a summary of our experience.
+
+## Problem statement
 
@@
 
+During the case we realized that we had to limit our solution. The case 
+presented by OPX wanted an automated code which would retrieve the data 
+and calculate the different KPIs, which turned out not to be possible. 
+The limitations can be divided into problems regarding: 
+(i) the accessible data, and 
+(ii) the non-accessible data
+
+The limitation with the accessible data is first and foremost the lack of 
+relevant data in the annual reports from the NGOs.  Since there is no 
+obligation to provide annual accounts, they often have more information 
+about the organization in general rather than numbers. If they do provide 
+the annual accounts, they are not standardized. This means that different 
+NGOs do not have an established template for what the different posts in 
+the annual accounts should contain.
+
+Also, these annual reports themselves appear in different formats. If they 
+do appear in the same format, the accounts we want to retrieve are often 
+in an image format rather than text. The lack of data, standardization and 
+diverse formats made it impossible to create an automated solution for 
+collecting data based on the NGOs’ annual reports.
+
+The limitation with the non-accessible data is that the data OPX needs to 
+calculate the different KPIs do not exist in the annual reports.  We 
+experimented with an annual report from Brønnøysundregisteret. Although 
+the reports were standardized, the different posts in the report did not 
+match the template from OPX. Additionally the data that OPX is interested 
+in and which is necessary for calculating the relevant KPIs are not 
+available.
+
+Also, with limited experience with KPIs it was hard to sort out which ones 
+that would be of relevance and not. After examining the different 
+alternatives, we concluded that the NGOs could report in a standardized 
+excel document from which our code retrieves the data to calculate the KPIs.
+
+In addition, we looked at the draft of a new law on associations from 2021 
+by the Ministry of Trade, Industries, and Fisheries, to see if it contained 
+changes in the laws regarding annual reports, which would have made the data 
+needed available in the future. This turned out to be a dead end, as the 
+draft focused largely on smaller and local associations and contained no 
+changes in reporting obligations. Therefore, if this law were to be 
+implemented, it would not be of any relevance to OPX. As a result, due to 
+OPX’s need for this data to be available, we recommend discussions with 
+relevant politicians and ministry officials as a starting point.
+
+Based on the findings of working with this case, we would recommend looking 
+at best practices for further development. One good example could be charity 
+navigator.
+
 # Our experience

### 2022-08-01 Report/Group 5 report.Rmd
edit intro

@@
 It is important to solve this problem because:
-Creating KPI(s) is important for the organization’s performance 
+1. Creating KPI(s) is important for the organization’s performance 
 management, to influence the objective, and to foster growth
-Automating the data collection will streamline the work of OPX 
+2. Automating the data collection will streamline the work of OPX 
 and create more transparency which enhances trust in charities

### 2022-08-01 Presentasjon/presentasjon.Rmd
Endret navn på bildefil kpi

@@
 
-## <br> Introduction | OPX and their goal
+## <br> OPX and their goal
 - A startup developing solutions for socially responsible investment strategies
@@
 
-## <br> Introduction | How to assess financial performance?
-- Data collection <br> OPX requested an automated solution to replace today's manual data collection
-- Key Performance Indicators <br> OPX needed meaningful KPIs allowing to assess the relationship between assets and obligations
+## <br> How to assess financial performance?
+- Data collection 
+- Key Performance Indicators
+
+```{r, fig.align='center'}
+knitr::include_graphics("Images/kpi.png")
+
+```
 

### 2022-08-01 Presentasjon/presentasjon.Rmd
Flere enkel navn på bildefiler
Flere bildefiler

@@
 - The goal is to increase cash flow to charitable purposes 
-- Donors' trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients
+- Donors' trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients <br>
+
+```{r, out.height="128px", out.width="568px", fig.align='center'}
+knitr::include_graphics("Images/logo_opx.png")
+```
 
@@
 - Data collection 
-- Key Performance Indicators
+- Key Performance Indicators <br>
 
-```{r, fig.align='center'}
+```{r, out.height="268px", out.width="350px", fig.align='center'}
 knitr::include_graphics("Images/kpi.png")
@@
 
-## {data-background="Images/9-Brainstorming-techniques-to-come-up-with-great-ideas.png" data-background-size=cover}
+## {data-background="Images/brainstorming2.png" data-background-size=cover}
 

### 2022-08-01 Report/KPIs and Visualisations for report.Rmd

--- 2022-07-31 Dashboard/NGO Analysis dashboard alpha v004.Rmd
@@
 ---
-title: "NGO analysis"
+title: "KPIs and Visualisations for report"
+author: "Torbjørn Skinnemoen Ottersen"
+date: "2022-08-01"
 output:
-  flexdashboard::flex_dashboard:
-    orientation: rows
-    vertical_layout: scroll
-    theme: lumen
-runtime: shiny
+  pdf_document: default
+  html_document: default
 ---
@@
 ```{r setup, include=FALSE}
-# Loading packages using pacman to ensure they're all installed and avoiding
-# catastrophic dashboard errors if they're not available
-if (!require("pacman")) install.packages("pacman")
-library(pacman)
-p_load(
-  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-  flexdashboard, MetBrewer, fontawesome, htmltools
-)
-# Using the fontawesome package for more up-to-date (and currently supported)
-# Fontawesome icons
-htmltools::tagList(fontawesome::fa_html_dependency())
+knitr::opts_chunk$set(echo = TRUE)
 ```
-
-Sidebar {.sidebar}
-=======================================================================
-
-### File upload
-
-Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.
-
-Then upload an Excel file in the set format.
-
-Click 'Analyse' to analyse the data.
-
-```{r sidebar}
-# HTML upload
-fileInput("html", label = "Upload html file", accept = ".html")
-
-# Year inputs
-numericInput("start", label = "Start", value = 2002)
-numericInput("end", label = "End", value = 2021)
-
-# Excel upload
-fileInput("excel", "Upload Excel file", accept = ".xlsx")
-
-# Analyse button
-actionButton("click", "Analyse")
-```
-
-If desired, a copy of the analysis in PDF format may then be downloaded.
-
-```{r sidebar_button}
-# download button - using uiOutput() because of an apparent bug
-# button and report are created in the second reactive chunk
-uiOutput("downloadUI")
-```
-
-Dashboard
-=======================================================================
-
-Row 
------------------------------------------------------------------------
-
-### KPI Fundraising
-```{r}
-valueBoxOutput("KPI_box_fundraising")
-```
-
-### KPI OPX
-```{r}
-valueBoxOutput("KPI_box_OPX")
-```
-
-### KPI Private growth
-```{r}
-valueBoxOutput("KPI_box_private")
-```
-
-### KPI Public growth
-```{r}
-valueBoxOutput("KPI_box_public")
-```
-
-### KPI Private/public proportion
-```{r}
-valueBoxOutput("KPI_box_proportion")
-```
-
-Row
------------------------------------------------------------------------
-
-### Net income
-```{r}
-plotlyOutput("arsresultat_plotly")
-```
-
-### Total equity/liabilities
-```{r}
-plotlyOutput("balance_plotly")
-```
-
-Row 
------------------------------------------------------------------------
-
-### Revenues
-```{r}
-plotlyOutput("revenue_plotly")
-```
-
-### Expenses
-```{r}
-plotlyOutput("expenses_plotly")
-```
-
-Row 
------------------------------------------------------------------------
-
-### Transparency: CEO compensation
-```{r}
-plotlyOutput("CEO_plotly")
-```
-
-Details
-=======================================================================
-
-Row {data-height=60}
------------------------------------------------------------------------
-
-### NGO analysed
-
-```{r}
-# Using htmlOutput() to use html styling
-htmlOutput("title")
-```
-
-Row 
------------------------------------------------------------------------
 
@@
 
-The KPIs are based on the following calculations, using data derived from the Excel sheet:
+On the basis of such data as we could gather we developed five potential KPIs:
 
@@
 
+This KPI is not a true measure of fundraising ROI, as it only includes (as we understand OPX's data) the direct fundraising expenses, and does not include major indirect expenses such as salary costs.
+
+***
+
 - **Liquid assets/obligations**:
@@
   $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
+
+This KPI was suggested by OPX.
+
+***
 
@@
 
+***
+
 - **Public grants growth**:
@@
 $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+
+***
 
@@
 
-Row 
------------------------------------------------------------------------
+This KPI may say something about the NGO's sources of funding. There are prima facie no specific desirable funding mix, but this measure may be of interest to some donors.
 
-### Graphs
+***
 
-The dashboard includes five graphs, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.
+### Visualisations
+
+The dashboard includes five plots, three based on data from Proff.no and two based on data from the Excel sheet.
 
@@
 -   **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
-  - A common measure of transparency, but should normally be evaluated in context.
+    -   This appears to increasingly be treated as a necessary measure of transparency and performance, but the values should normally be evaluated in the context of the size of the NGO and the sector(s) within which it operates.
 
-```{r reactive_title_value_boxes, include=FALSE}
-# Shiny actions begin here =====================================================
-# First reactive chunk =========================================================
-
-# wait for button click
-observe({
-
-  # Excel data frame ===========================================================
-
-  ## Validating and creating long version of Excel file, if possible------------
-  excel_test <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
-
-    widen_excel_test(input$excel$datapath)
-  })
-
-  ## Creating the primary Excel data frame--------------------------------------
-  exceldata <- reactive({
-    widen_excel(excel_test())
-  })
-
-  # NGO title ==================================================================
-  NGO <- reactive({
-    # checking for the required column again
-    if ("firm" %in% names(excel_test())) {
-      # fetching the name
-      excel_test() |>
-        # Turning the needed column into a list
-        pull(firm) |>
-        # discarding all NAs in case the name is somewhere
-        purrr::discard(is.na) |>
-        # choosing the first available name
-        first()
-    } else {
-      validate("Please upload data using the specified Excel template.")
-    }
-  })
-
-  title <- reactive({
-    # testing if NGO() is NA
-    if (is.na(NGO())) {
-      # Printing warning if NA
-      "Name of NGO missing"
-      # Printing name of NGO
-    } else {
-      NGO()
-    }
-  })
-
-  # Creating NGO title using html formatting (colour based on MetBrewer
-  # "Hokusai2")
-  output$title <- renderText({
-    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
-  })
-
-  # Value Boxes ================================================================
-
-  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark" if
-  # the value is NA
-
-  ## Fundraising ROI------------------------------------------------------------
-  # Creating Fundraising ROI value
-  KPI_fundraising_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # most recent year
-        filter(year == max(year)) |>
-        transmute(
-          # Calculating the value as a percentage; removing all other
-          KPI_fundraising_num =
-            (sum(revenue_private_contributions,
-              revenue_membership_fees,
-              na.rm = TRUE
-            ) - expenses_fundraising) /
-              expenses_fundraising * 100
-        ) |>
-        # just want the single number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for fundraising ROI
-  # Unsure about colour levels
-  output$KPI_box_fundraising <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_fundraising_num()),
-      caption = "Fundraising ROI",
-      # Colour depends on value
-      color = if (is.na(KPI_fundraising_num())) {
-        "primary"
-      } else if (KPI_fundraising_num() < 100) {
-        "danger"
-      } else if (KPI_fundraising_num() < 200) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_fundraising_num())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fas fa-donate"
-      }
-    )
-  })
-
-  ## OPX KPI--------------------------------------------------------------------
-  # Creating OPX KPI
-  KPI_OPX_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # most recent year
-        filter(year == max(year)) |>
-        # Calculating the value as a %
-        transmute(
-          KPI_OPX =
-            sum(revenue_private_contributions,
-              revenue_investments, assets_cash_in_bank_and_cash_equivalents,
-              na.rm = TRUE
-            ) /
-              sum(liability_short_term_grants_payable,
-                liability_revocable_endowments,
-                liability_long_term_grants_payable,
-                na.rm = TRUE
-              ) * 100
-        ) |>
-        # Just the single number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for OPX KPI
-  output$KPI_box_OPX <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_OPX_num()),
-      caption = "Liquid assets / obligations",
-      # colour depends on value
-      color = if (is.na(KPI_OPX_num())) {
-        "primary"
-      } else if (KPI_OPX_num() < 100) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on value
-      icon = if (is.na(KPI_OPX_num())) {
-        "fa-solid fa-circle-xmark"
-      } else if (KPI_OPX_num() < 100) {
-        "fa-solid fa-scale-unbalanced"
-      } else {
-        "fa-solid fa-scale-balanced"
-      }
-    )
-  })
-
-  ## Private donations growth---------------------------------------------------
-  # Creating private donations growth
-  KPI_private_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # Calculating the two most recent years as a %
-        transmute(
-          growth =
-            (revenue_private_contributions[1] -
-              revenue_private_contributions[2]) /
-              revenue_private_contributions[2] * 100
-        ) |>
-        # Pulling the first instance of the number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for private donations growth
-  output$KPI_box_private <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_private_num()),
-      caption = "Private donation growth",
-      # colour depends on value
-      color = if (is.na(KPI_private_num())) {
-        "primary"
-      } else if (KPI_private_num() < 0) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_private_num())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fas fa-hands-helping"
-      }
-    )
-  })
-
-  ## Public grants growth-------------------------------------------------------
-  # Creating public grants growth
-  KPI_public_num <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # Calculating the two most recent years as a %
-        transmute(
-          growth =
-            (revenue_public_grants[1] - revenue_public_grants[2]) /
-              revenue_public_grants[2] * 100
-        ) |>
-        # Pulling the first instance of the number
-        pull() |>
-        first()
-    }
-  })
-
-  # Creating value box for public grants growth
-  output$KPI_box_public <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_public_num()),
-      caption = "Public grants growth",
-      # colour depends on value
-      color = if (is.na(KPI_public_num())) {
-        "primary"
-      } else if (KPI_public_num() < 0) {
-        "warning"
-      } else {
-        "success"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_public_num())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fa-solid fa-building-columns"
-      }
-    )
-  })
-
-  ## Public/private proportion--------------------------------------------------
-  # Creating public/private proportion
-  KPI_proportion <- reactive({
-    # testing for data quality, returning NA if bad
-    if (datatest() == TRUE) {
-      NA
-    } else {
-      exceldata() |>
-        # calculating the value as a %
-        transmute(
-          proportion =
-            sum(revenue_private_contributions,
-              revenue_membership_fees,
-              na.rm = TRUE
-            ) /
-              sum(revenue_public_grants, revenue_private_contributions,
-                revenue_membership_fees,
-                na.rm = TRUE
-              ) * 100
-        ) |>
-        pull() |>
-        # getting the most recent year
-        first()
-    }
-  })
-
-  # Creating value box for private/public proportion
-  output$KPI_box_proportion <- renderValueBox({
-    valueBox(
-      value = KPI_to_string(KPI_proportion()),
-      caption = "Private donations",
-      # Adding in the same check as elsewhere - colours match if all are NA
-      color = if (is.na(KPI_proportion())) {
-        "primary"
-      } else {
-        "info"
-      },
-      # icon depends on whether or not the value is NA
-      icon = if (is.na(KPI_proportion())) {
-        "fa-solid fa-circle-xmark"
-      } else {
-        "fa-solid fa-chart-pie"
-      }
-    )
-  })
-
-  ## Excel data test -----------------------------------------------------------
-  # (also below, in the second chunk)
-  # Reactive function to test for Excel data quality (basic)
-  datatest <- reactive({
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
-    }
-  })
-  # the button click
-}) |> bindEvent(input$click)
-```
-
-```{r reactive_report_plots, include=FALSE}
-# Beginning of the second reactive chunk =======================================
-# It's necessary to split the observers into separate parts, with all necessary
-# code duplicated, because of R's limitations.
-
-observe({
-
-  # Report Output ==============================================================
-
-  ## Download button -----------------------------------------------------------
-  output$downloadUI <- renderUI({
-    if (available_excel() && available_html()) {
-      downloadButton("report", "Download report")
-    }
-  })
-
-  ## Data exists condition testing----------------------------------------------
-  # Essentially pretty random tests to connect to the individual
-  # validate() functions upstream in the reactive chain
-
-  # Excel
-  available_excel <- reactive({
-    TRUE
-  }) |>
-    bindEvent(setequal(
-      names(exceldata()),
-      c("year", "firm", "category", "sub_category", "name", "value")
-    ))
-
-  # HTML
-  available_html <- reactive({
-    TRUE
-  }) |> bindEvent("lederlonn" %in% names(htmldata()))
-
-  ## Report download and parameters --------------------------------------------
-  output$report <- downloadHandler(
-
-    # Based on code in Mastering Shiny - creating a temporary directory and file
-    filename = "report.pdf",
-    content = function(file) {
-      tempReport <- file.path(tempdir(), "report.Rmd")
-      file.copy("report.Rmd", tempReport, overwrite = TRUE)
-
-      # parameters
-      params <- list(
-        htmlfile = input$html$datapath,
-        excelfile = input$excel$datapath,
-        start = input$start,
-        end = input$end
-      )
-
-      # notification - disappears when report generation is complete
-      id <- showNotification(
-        "Generating report...",
-        duration = NULL,
-        closeButton = FALSE
-      )
-      on.exit(removeNotification(id), add = TRUE)
-
-      # R Markdown render
-      rmarkdown::render(tempReport,
-        output_file = file,
-        params = params,
-        envir = new.env(parent = globalenv())
-      )
-    }
-  )
-
-  # Data frames ================================================================
-  ## Creating data frame from html file-----------------------------------------
-  html_currency <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$html$datapath) == "html",
-      "Please upload an html file."
-    ))
-
-    # Testing and creating data frame
-    html_test_create(
-      input$html$datapath,
-      as.character(input$start), as.character(input$end)
-    )
-  })
-
-  htmldata <- reactive({
-    read_proff_two(html_currency())
-  })
-
-  # Recreating, to make available to this chunk as well
-  ## Validating and creating long version of Excel file, if possible------------
-  excel_test <- reactive({
-    # validating file
-    validate(need(
-      tools::file_ext(input$excel$datapath) == "xlsx",
-      "Please upload an Excel file."
-    ))
-
-    widen_excel_test(input$excel$datapath)
-  })
-
-  ## Creating the primary Excel data frame--------------------------------------
-  exceldata <- reactive({
-    widen_excel(excel_test())
-  })
-
-  # Plots ======================================================================
-
-  ## Annual result--------------------------------------------------------------
-  arsresultat_plot <- reactive({
-    # testing for the presence of data
-    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      htmldata() |>
-        # renaming for plotly hover
-        rename(
-          Year = year,
-          `Net income` = arsresultat,
-          `Net positive` = annual_profit
-        ) |>
-        ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # using a function to ensure uniform colours irrespective of data
-        scale_fill_green(htmldata()$annual_profit) +
-        theme(legend.position = "none") +
-        labs(
-          x = "", y = "", title = "Net income (in thousands)"
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$arsresultat_plotly <- renderPlotly(
-    ggplotly(arsresultat_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  ## Balance--------------------------------------------------------------------
-  balance_plot <- reactive({
-    # testing for the presence of data
-    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      htmldata() |>
-        # rename for plotly hover
-        rename(
-          Year = year,
-          `Total equity/liabilities` = sum_egenkapital_og_gjeld,
-          `Positive equity` = balance_positive
-        ) |>
-        ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # using a function to ensure uniform colours irrespective of data
-        scale_fill_green(htmldata()$balance_positive) +
-        theme(legend.position = "none") +
-        labs(
-          x = "", y = "", title = "Total equity/liabilities (in thousands)",
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$balance_plotly <- renderPlotly(
-    ggplotly(balance_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  ## Expenses-------------------------------------------------------------------
-  expenses_plot <- reactive({
-    # testing for data quality
-    if (datatest() == TRUE) {
-      # printing 'bad data' plot if data are bad
-      bad_data()
-    } else {
-      exceldata() |>
-        # Renaming for the purposes of the plot + dropping other variables
-        transmute(
-          Year = as.factor(year),
-          Other = expenses_other,
-          Administrative = expenses_administrative,
-          Fundraising = expenses_fundraising,
-          `Program services` = expenses_program_services
-        ) |>
-        # Pivoting to use fill
-        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
-        # Ordering for the plot
-        mutate(
-          Expenses = ordered(Expenses,
-            levels =
-              c("Program services", "Other", "Fundraising", "Administrative")
-          )
-        ) |>
-        ggplot(aes(Year, value, fill = Expenses)) +
-        geom_col() +
-        scale_y_continuous(
-          # Excel data are assumed to be in NOK; if this changes, code based on
-          # the currency_code() function may be adapted
-          labels = label_dollar(prefix = "NOK "),
-          # Basing lower limit on the data (not suppressing erroneous negative
-          # numbers), and the upper limit on the maximum value in revenues and
-          # expenses, to match the two plots
-          limits = c(NA, maximum())
-        ) +
-        scale_fill_met_d("Hokusai2") +
-        labs(x = "", y = "", title = "Expenses (in thousands)")
-    }
-  })
-
-  # Creating plotly output
-  output$expenses_plotly <- renderPlotly(
-    ggplotly(expenses_plot())
-  )
-
-  ## Revenue--------------------------------------------------------------------
-  revenue_plot <- reactive({
-    # testing for data quality
-    if (datatest() == TRUE) {
-      # printing 'bad data' plot if data are bad
-      bad_data()
-    } else {
-      exceldata() |>
-        # Renaming for the purposes of the plot + dropping other variables
-        transmute(
-          Year = as.factor(year),
-          `Private contributions` = revenue_private_contributions,
-          `Membership fees` = revenue_membership_fees,
-          `Public grants` = revenue_public_grants,
-          Investments = revenue_investments,
-          `Operational activities` = revenue_operational_activities,
-          `Commercial activities` = revenue_commercial_activities,
-          Other = revenue_other
-        ) |>
-        # Pivoting to use fill
-        pivot_longer(
-          c(`Private contributions`:Other),
-          names_to = "Revenue source"
-        ) |>
-        # Ordering for the plot
-        mutate(
-          `Revenue source` =
-            ordered(`Revenue source`,
-              levels =
-                c(
-                  "Other", "Commercial activities",
-                  "Operational activities", "Investments",
-                  "Membership fees", "Private contributions",
-                  "Public grants"
-                )
-            )
-        ) |>
-        ggplot(aes(Year, value, fill = `Revenue source`)) +
-        geom_col() +
-        scale_y_continuous(
-          # Excel data are assumed to be in NOK; if this changes, code based on
-          # the currency_code() function may be adapted
-          labels = label_dollar(prefix = "NOK "),
-          # Basing lower limit on the data (not suppressing erroneous negative
-          # numbers), and the upper limit on the maximum value in revenues and
-          # expenses, to match the two plots
-          limits = c(NA, maximum())
-        ) +
-        # Specifying the number of values, as the base palette only includes 6
-        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
-        labs(x = "", y = "", title = "Revenue (in thousands)")
-    }
-  })
-
-  # Creating plotly output
-  output$revenue_plotly <- renderPlotly(
-    ggplotly(revenue_plot())
-  )
-
-  ## CEO compensation-----------------------------------------------------------
-  CEO_plot <- reactive({
-    # Testing for the presence of data
-    if (sum(!is.na(htmldata()$lederlonn)) == 0 &&
-      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
-      # Printing 'no information' plot if no data
-      no_information()
-    } else {
-      # Creating plot if any data exist
-      # Note that the plot will still look empty, but have a legend and scales,
-      # if all values are explicit zeroes
-      htmldata() |>
-        # preparing for pivoting
-        select(
-          year, lederlonn, leder_annen_godtgjorelse
-        ) |>
-        # Compensation in ordinary numbers
-        mutate(
-          lederlonn = lederlonn * 1000,
-          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-        ) |>
-        # renaming for the purposes of the plot
-        rename(
-          Year = year,
-          Salary = lederlonn,
-          `Other compensation` = leder_annen_godtgjorelse
-        ) |>
-        # pivoting to join the two values to use fill
-        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-        ggplot(aes(Year, value, fill = Compensation)) +
-        geom_col() +
-        # using currency scale function to add correct currency code or ?
-        currency_scale() +
-        # pretty colours
-        scale_fill_met_d("Hokusai2", direction = -1) +
-        labs(
-          x = "", y = "Total compensation", title = "CEO compensation"
-        )
-    }
-  })
-
-  # Creating plotly output
-  output$CEO_plotly <- renderPlotly(
-    ggplotly(CEO_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
-      layout(
-        annotations =
-          plotly_caption()
-      )
-  )
-
-  # Reactive functions =========================================================
-
-  ## Currency scale-------------------------------------------------------------
-  # Reactive function to check for currency and use currency code or ?
-  currency_scale <- reactive({
-    scale_y_continuous(
-      # Checking for uniform currency code
-      labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
-        # Using the currency code if uniform
-        label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
-        # alternative prefix if multiple currencies in the data
-        label_dollar(prefix = "? ")
-      )
-    )
-  })
-
-  ## Plotly caption-------------------------------------------------------------
-  # Reactive function to create plotly caption (ggplot captions and subtitles
-  # are lost in the conversion) (shamelessly stolen from the Internet, with
-  # minor modifications).
-  # Checks for uniform currency code and prints a warning
-  # if there are multiple currency codes.
-  plotly_caption <- reactive({
-    list(
-      x = 1, y = -0.1,
-      text = ifelse(length(unique(html_currency()$valutakode)) == 1,
-        "", "Note: not all values are denominated in the same currency"
-      ),
-      showarrow = F, xref = "paper", yref = "paper",
-      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
-      font = list(size = 11)
-    )
-  })
-
-  ## scale maximum--------------------------------------------------------------
-  # Reactive function to create the maximum value for revenue and expenses plots
-  # the tests are probably not necessary, but included to be on the safe side
-  maximum <- reactive({
-    if (datatest() == TRUE || setequal(
-      names(exceldata()),
-      c("year", "firm", "category", "sub_category", "name", "value")
-    )) {
-      NA
-    } else {
-      excel_test() |>
-        # removing the firm and sub-category columns
-        select(-firm, -sub_category) |>
-        # filtering to retain only revenue and expenses entries
-        filter(category == "Revenue" | category == "Expenses") |>
-        # grouping and summarising to calculate maximum value
-        group_by(year, category) |>
-        summarise(
-          sum = sum(value, na.rm = TRUE)
-        ) |>
-        # Creating list of sums
-        pull() |>
-        # selecting the highest value
-        max()
-    }
-  })
-
-  ## Excel data test (also above)-----------------------------------------------
-  # Reactive function to test for Excel data quality (basic)
-  datatest <- reactive({
-    # let me know if anyone has a better idea for testing
-    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
-      TRUE
-    } else {
-      FALSE
-    }
-  })
-  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
-}) |> bindEvent(input$click)
-```
-
-```{r non-reactive functions, include=FALSE}
-# Data reading functions =======================================================
-
-## Functions to read html file -------------------------------------------------
-html_test_create <- function(file, start, end) {
-  # Reading file and node
-  htmltest <- read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame")
-
-  ### A series of tests to catch problems before they spit out errors-----------
-  # Checking if node exists
-  if (class(htmltest) == "xml_node") {
-    htmltable <- htmltest |>
-      html_table()
-  } else {
-    validate("Please upload the specified type of html document.")
-  }
-
-  # Basic check if the table is correct
-  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
-    htmlready <- htmltable
-  } else {
-    validate("Please upload the specified type of html document.")
-  }
-
-  # Testing if the correct years have been entered
-  if (as.character(all_of(start)) %in% names(htmlready) &&
-    as.character(all_of(end)) %in% names(htmlready)) {
-    # Creating the first part of the data frame
-    read_proff_one(htmltable, start, end)
-  } else {
-    validate("Please ensure you have entered the correct start and end years.")
-  }
-}
-
-### First html transformation function, invoked by html_test_create() ----------
-# (stopping to allow checks for differing currency codes)
-read_proff_one <- function(object, start, end) {
-  object |>
-    # removing empty column, using all_of(argument)
-    select(
-      REGNSKAPSPERIODE:all_of(start)
-    ) |>
-    # removing white space
-    mutate(
-      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-    ) |>
-    # removing duplicate table sometimes created by read_html()
-    filter(
-      !row_number() > 176
-    ) |>
-    # removing duplicate rows
-    filter(
-      !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |>
-    # tidying data using pivot_longer() and all_of(argument)
-    pivot_longer(
-      all_of(end):all_of(start),
-      names_to = "year"
-    ) |>
-    # creating a currency code column and turning current currency codes into
-    # NAs
-    mutate(
-      # changing to real NAs
-      value = na_if(value, "-"),
-      # turning years into numbers
-      year = as.integer(year),
-      # creating a currency code column and turning current currency codes into
-      # NAs
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |>
-    # filling the currency column to avoid having to deal with NAs later
-    fill(
-      valutakode,
-      .direction = "downup"
-    )
-}
-
-### Second html transformation function, creating the data frame ---------------
-read_proff_two <- function(object) {
-  object |>
-    # removing unneeded currency column
-    select(-valutakode) |>
-    # removing dates and redundant currency
-    filter(
-      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-        REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-    mutate(
-      # removing white space in numbers
-      value = str_replace_all(value, "\\s", ""),
-      # turning numbers into true numbers
-      value = as.numeric(value),
-      # removing values that are years
-      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-    ) |>
-    # removing any remaining duplicate rows
-    distinct() |>
-    # pivoting to create 'tidy' variables
-    pivot_wider(
-      names_from = REGNSKAPSPERIODE, values_from = value
-    ) |>
-    # tidying names
-    clean_names() |>
-    # removing headings from the table
-    select(
-      -lederlonn_i_hele_1000,
-      -resultatregnskap_i_hele_1000,
-      -balanseregnskap_i_hele_1000
-    ) |>
-    # making it clear what lonn refers to
-    rename(
-      lederlonn = lonn
-    ) |>
-    # Creating new variables for plots (red/green profit/loss)
-    mutate(
-      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-    ) |>
-    # arranging by year for practical display
-    arrange(desc(year))
-}
-
-## Functions to read Excel file ------------------------------------------------
-### First Excel transformation -------------------------------------------------
-# creating df used by NGO title function maximum scale function
-widen_excel_test <- function(file) {
-  # Reading Excel file
-  read_excel(file) |>
-    # cleaning names to enable later operations
-    clean_names()
-}
-
-### Second Excel function - creating primary Excel data frame-------------------
-widen_excel <- function(object) {
-  # testing if all columns are present
-  if (setequal(
-    names(object),
-    c("year", "firm", "category", "sub_category", "name", "value")
-  )) {
-    object |>
-      # uniting category and name to deal with multiple 'Other'
-      unite(
-        entry, c(category, name)
-      ) |>
-      # deleting Sub-category and Firm
-      select(-sub_category, -firm) |>
-      # pivoting to create 'tidy' variables
-      pivot_wider(names_from = entry, values_from = value) |>
-      # cleaning names
-      clean_names()
-    # Notification if all columns are not present
-  } else {
-    validate("Please upload data using the specified Excel template.")
-  }
-}
-
-# Support functions ============================================================
-
-## KPI_to_string ---------------------------------------------------------------
-# Function to turn KPI values into NA or percentage string
-KPI_to_string <- function(variable) {
-  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
-  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
-  # avoid passing an NA to subsequent operations.)
-  if (is.na(variable)) {
-    NA
-  } else if (variable == Inf) {
-    NA
-  } else {
-    # Shortening (lengthening) the value, adding %
-    str_c(format(variable, digits = 3), "%")
-  }
-}
-
-## scale_fill_green ------------------------------------------------------------
-# Function with slightly nicer colours than default; ensuring uniformity; +
-# ensuring positive values are always green through a test for the presence of
-# FALSE values
-scale_fill_green <- function(variable) {
-  scale_fill_met_d("Tsimshian",
-    direction = ifelse(
-      mean(variable,
-        na.rm = TRUE
-      ) == 1, -1, 1
-    )
-  )
-}
-
-## no_information plot ---------------------------------------------------------
-# Function to print a 'no information' plot, if necessary
-no_information <- function() {
-  empty_plot("No information available")
-}
-
-## bad_data plot ---------------------------------------------------------------
-# Function to print a 'bad data' plot, if necessary
-bad_data <- function() {
-  empty_plot("Something is wrong\nwith the Excel data.")
-}
-
-## empty_plot-------------------------------------------------------------------
-# Function to print an empty plot with a supplied message
-empty_plot <- function(warning) {
-  ggplot() +
-    # text geom with a space to create the plot, but otherwise show nothing
-    geom_text(aes(0, 0, label = " ")) +
-    # as blank a slate as possible
-    theme_void() +
-    # annotating the plot with the specified warning message
-    annotate(
-      "text",
-      label = warning,
-      # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
-}
-```
+The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.

### 2022-08-01 Dashboard/Dashboard testing material/Dashboard testing zero plot.Rmd
Zero values testing.
Excel zero plot working.

--- 2022-07-31 Dashboard/NGO Analysis dashboard alpha v005.Rmd
@@
 
-Then upload an Excel file in the set format.
+Then upload an Excel file in the set format. <!--# Need to insert link when repository is public -->
 
@@
       bad_data()
+    } else if (excel_zero_check("Expenses") == 0) {
+      zero_data()
     } else {
@@
       bad_data()
+    } else if (excel_zero_check("Revenue") == 0) {
+      zero_data()
     } else {
@@
   })
+
+  ## Excel zero test------------------------------------------------------------
+  excel_zero_check <- function(type) {
+    excel_test() |>
+      filter(
+        category == as.character(all_of(type))
+      ) |>
+      summarise(
+        zero = sum(value != 0, na.rm = TRUE)
+      ) |>
+      pull()
+  }
+
   # Let the magic/disaster happen (i.e. execute when the user clicks the button)
@@
   empty_plot("Something is wrong\nwith the Excel data.")
+}
+
+## zero_data_plot --------------------------------------------------------------
+
+zero_data <- function() {
+  empty_plot("All values are zero.\nThis is unusual and may\nindicate a data problem.")
 }

### 2022-08-01 Dashboard/Dashboard testing material/Dashboard testing zero plot.Rmd
Working on zero testing.

@@
 ### First Excel transformation -------------------------------------------------
-# creating df used by NGO title function and maximum scale function
+# creating df used by NGO title function, maximum scale function, and excel zero check
 widen_excel_test <- function(file) {

### 2022-08-01 Dashboard/report.Rmd
Working on zero testing.

@@
 ```{r data_warnings}
-# Excel data test =============================================================
-# Function to test for Excel data quality (basic)
+# Excel data tests =============================================================
+# Functions to test for Excel data quality (basic)
 datatest <-
@@
   if (datatest == TRUE) {
-    # printing 'bad data' plot if data are bad
     NA

### 2022-08-01 Dashboard/report.Rmd
Updating with two Excel functions; adding zero tests and warnings.

@@
 
-## Function to read Excel file -------------------------------------------------
-widen_excel <- function(file) {
-  read_excel(file) |>
-    # uniting Category and Name to deal with multiple 'Other'
-    unite(
-      entry, c(Category, Name),
-      sep = "_"
-    ) |>
-    # deleting Sub-category and Firm
-    select(-`Sub-category`, -Firm) |>
-    # pivoting to create 'tidy' variables
-    pivot_wider(names_from = entry, values_from = Value) |>
-    # cleaning names
-    clean_names()
-}
 ### Reading html ---------------------
@@
 
+## Functions to read Excel file ------------------------------------------------
+### First Excel transformation -------------------------------------------------
+# creating df used by NGO title function and maximum scale function
+widen_excel_test <- function(file) {
+  # Reading Excel file
+  read_excel(file) |>
+    # cleaning names to enable later operations
+    clean_names()
+}
+
+### Second Excel function - creating primary Excel data frame-------------------
+widen_excel <- function(object) {
+  object |>
+    # uniting category and name to deal with multiple 'Other'
+    unite(
+      entry, c(category, name),
+      sep = "_"
+    ) |>
+    # deleting sub_category and firm
+    select(-sub_category, -firm) |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(names_from = entry, values_from = value) |>
+    # cleaning names
+    clean_names()
+}
+
 ### Reading excel ---------------------
-exceldata <- widen_excel(params$excelfile)
+excel_test <- widen_excel_test(params$excelfile)
+
+exceldata <- widen_excel(excel_test)
 
@@
 
-NGO <- read_excel(params$excelfile) |>
+NGO <- excel_test |>
   # Turning the needed column into a list
-  pull(Firm) |>
+  pull(firm) |>
   # discarding all NAs in case the name is somewhere
@@
 ```{r data_warnings}
-# Excel data tests =============================================================
+# Excel data tests and warnings ================================================
 # Functions to test for Excel data quality (basic)
+## Excel datatest---------------------------------------------------------------
 datatest <-
@@
 
-## Creating the Excel warning text --------------------------------------------
+## Excel zero test------------------------------------------------------------
+excel_zero_check <- function(type) {
+  excel_test |>
+    filter(
+      category == as.character(all_of(type))
+    ) |>
+    summarise(
+      zero = sum(value != 0, na.rm = TRUE)
+    ) |>
+    pull()
+}
+
+## Creating the Excel warning texts --------------------------------------------
+# Excel data test warning
 excel_warning <- if (datatest == TRUE) {
@@
 
-# HTML warnings ===============================================================
+# Excel zero revenues warning
+excel_zero_revenue_warning <- if (excel_zero_check("Revenue") == 0) {
+  "*    All revenue values are zero or missing. This is unusual and may indicate a data problem.*"
+} else {
+  ""
+}
+
+# Excel zero expenses warning
+excel_zero_expenses_warning <- if (excel_zero_check("Expenses") == 0) {
+  "*    All expenses values are zero or missing. This is unusual and may indicate a data problem.*"
+} else {
+  ""
+}
+
+# HTML warnings ================================================================
 # Creating the arsresultat warning text
@@
 } else {
-  # reading the Excel sheet anew
-  read_excel(params$excelfile) |>
+  excel_test |>
     # removing the firm and sub-category columns
-    select(-Firm, -`Sub-category`) |>
+    select(-firm, -sub_category) |>
     # filtering to retain only revenue and expenses entries
-    filter(Category == "Revenue" | Category == "Expenses") |>
+    filter(category == "Revenue" | category == "Expenses") |>
     # grouping and summarising to calculate maximum value
-    group_by(Year, Category) |>
+    group_by(year, category) |>
     summarise(
-      sum = sum(Value, na.rm = TRUE)
+      sum = sum(value, na.rm = TRUE)
     ) |>
@@
 } else {
-  {
   exceldata |>
@@
     pull() |>
-      first()
-  } |> KPI_to_string()
+    first() |>
+    KPI_to_string()
 }
@@
 
+`r excel_zero_revenue_warning`
+
+`r excel_zero_expenses_warning`
+
 `r arsresultat_plot_warning`
@@
 # Printing the plot if the data are OK
-if (datatest == FALSE) {
+if (datatest == FALSE && excel_zero_check("Revenue") != 0) {
   revenue_plot
@@
 # Printing the plot if the data are OK
-if (datatest == FALSE) {
+if (datatest == FALSE && excel_zero_check("Expenses") != 0) {
   expenses_plot

### 2022-08-01 Dashboard/report.Rmd
Stylering.

@@
 # Creating the currency data frame
-html_currency <- read_proff_currency(params$htmlfile,
-  as.character(params$start), as.character(params$end))
+html_currency <- read_proff_currency(
+  params$htmlfile,
+  as.character(params$start), as.character(params$end)
+)
 
@@
 # Testing for the presence of NAs; printing warning if there are any
-KPI_warning <- if (is.na(fundraising) || is.na(KPI_OPX) || is.na( KPI_private) ||
+KPI_warning <- if (is.na(fundraising) || is.na(KPI_OPX) || is.na(KPI_private) ||
   is.na(KPI_public) || is.na(KPI_proportion)) {

### 2022-08-01 Dashboard/Dashboard testing material/report safe version 005.Rmd
Calling it v006-

--- 2022-07-26 Dashboard/report.Rmd
@@
 # Real currency caption
-currency_caption <- ifelse(length(unique(htmldata$valutakode)) == 1,
+currency_caption <- ifelse(length(unique(html_currency$valutakode)) == 1,
       "", "Note: not all values are denominated in the same currency.\n"
@@
 The KPIs are based on the following calculations, using data derived from the Excel sheet:
-
-<!--# Need to decide whether to underline the KPIs, as in the first one -->
 

### 2022-08-01 Dashboard/NGO Analysis dashboard v006.Rmd
Calling it v006-

--- 2022-08-01 Dashboard/Dashboard testing material/Dashboard testing zero plot.Rmd
@@
 zero_data <- function() {
-  empty_plot("All values are zero.\nThis is unusual and may\nindicate a data problem.")
+  empty_plot("All values are zero or missing.\nThis is unusual and may\nindicate a data problem.")
 }
@@
       # Using the same palette as the other non-profit/loss plots
-      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
-    )
-}
-```
+      x = 0, y = 0, size = 6, colour = met.brewer("Hokusai2", 1)
+    )
+}
+```

### 2022-08-01 Presentasjon/presentasjon.Rmd
Flere bilder
Spelling and comment.

@@
 
-# Introduction
+## {data-background="Images/start1.png" data-background-size=cover}
 
@@
 - The goal is to increase cash flow to charitable purposes 
-- Donors' trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients <br>
+- Donors' trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficiaries <br>
 
@@
 - Webscraping the Brønnøysund registry <br>
-Issue: The registry does store downloadable XML documents, but the access cost about 500k kr
+Issue: The registry does offer API access, but access costs about 500k kr
 
@@
 - Brønnøysund: We attempted to request a temporary access to XML documents
+<!--# We received some sample XML data from Statistics Norway -- we did not talk to Brønnøysund at all-->
 
 ## <br> ... and why they did not work
-- Different accounting divisions at Proff.no and Brønnøysund prohibited extracting the correct sums the OPX wanted
+- Different accounting divisions at Proff.no and Brønnøysund prohibited extracting the correct sums  OPX wanted
 - Not all NGOs publish their financial statements and certainly not for every year

### 2022-08-01 Report/Group 5 report.Rmd
edit limitations
Setting out.width = "100%" and pdf as default.

@@
 ```{r Setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE, out.width = "100%")
 ```
@@
 
-```{r}
+```{r, echo=FALSE}
 knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene
@@
 
-```{r}
+```{r, echo=FALSE}
 knitr::include_graphics("Images/info_kreft.png")
@@
 
-The limitation with the accessible data is first and foremost the lack of 
-relevant data in the annual reports from the NGOs.  Since there is no 
-obligation to provide annual accounts, they often have more information 
+The limitation with the accessible data is first and foremost the lack 
+of relevant data in the annual reports from the NGOs.  Since there is 
+no obligation to provide annual accounts, they often have more information 
 about the organization in general rather than numbers. If they do provide 
@@
 
-In addition, we looked at the draft of a new law on associations from 2021 
-by the Ministry of Trade, Industries, and Fisheries, to see if it contained 
-changes in the laws regarding annual reports, which would have made the data 
-needed available in the future. This turned out to be a dead end, as the 
-draft focused largely on smaller and local associations and contained no 
-changes in reporting obligations. Therefore, if this law were to be 
-implemented, it would not be of any relevance to OPX. As a result, due to 
-OPX’s need for this data to be available, we recommend discussions with 
-relevant politicians and ministry officials as a starting point.
+In addition, we looked at the draft of a new law on associations from 
+2021 by the Ministry of Trade, Industries, and Fisheries, to see if it 
+contained changes in the laws regarding annual reports, which would have 
+made the data needed available in the future. This turned out to be a dead 
+end, as the draft focused largely on smaller and local associations and 
+contained no changes in reporting obligations. Therefore, if this law were 
+to be implemented, it would not be of any relevance to OPX. As a result, 
+due to OPX’s need for this data to be available, we recommend discussions 
+with relevant politicians and ministry officials as a starting point.
 

### 2022-08-01 Report/Group 5 report.Rmd
Copying over KPIs and Visualisations text.

@@
 It is important to solve this problem because:
+
 1. Creating KPI(s) is important for the organization’s performance 
-management, to influence the objective, and to foster growth
+management, to influence the objective, and to foster growth.
+
 2. Automating the data collection will streamline the work of OPX 
-and create more transparency which enhances trust in charities
+and create more transparency which enhances trust in charities.
 
@@
 
+## KPIs and visualisations
+
+### Key Performance Indicators
+
+On the basis of such data as we could gather we developed five potential KPIs:
+
+- **Fundraising ROI**:
+  - A basic measure of the return on investment of direct fundraising expenses:
+
+  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
+
+This KPI is not a true measure of fundraising ROI, as it only includes (as we understand OPX's data) the direct fundraising expenses, and does not include major indirect expenses such as salary costs.
+
+***
+
+- **Liquid assets/obligations**:
+  - A measure of the relationship between the organisation's liquid assets and its obligations:
+
+  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
+
+This KPI was suggested by OPX.
+
+***
+
+- **Private donations growth**:
+  - A calculation of the percentage change in private contributions based on the two most recent years:
+
+  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
+
+***
+
+- **Public grants growth**:
+  - A calculation of the percentage change in public grants received based on the two most recent years:
+
+$$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
+
+***
+
+- **Private/public proportion**:
+  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
+
+$$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
+
+This KPI may say something about the NGO's sources of funding. There are prima facie no specific desirable funding mix, but this measure may be of interest to some donors.
+
+***
+
+### Visualisations
+
+The dashboard includes five plots, three based on data from Proff.no and two based on data from the Excel sheet.
+
+-   **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
+-   **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
+-   **Revenues**: Based on the category *Revenue* in the Excel sheet.
+-   **Expenses**: Based on the category *Expenses* in the Excel sheet.
+-   **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
+    -   This appears to increasingly be treated as a necessary measure of transparency and performance, but the values should normally be evaluated in the context of the size of the NGO and the sector(s) within which it operates.
+
+The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.
+
 # Limitations and steps forward

### 2022-08-01 Dashboard/NGO Analysis dashboard v006.Rmd
Deleting unnecessary tests. Formatting.

@@
   # Reactive function to create the maximum value for revenue and expenses plots
-  # the tests are probably not necessary, but included to be on the safe side
   maximum <- reactive({
-    if (datatest() == TRUE || setequal(
-      names(exceldata()),
-      c("year", "firm", "category", "sub_category", "name", "value")
-    )) {
-      NA
-    } else {
       excel_test() |>
@@
         max()
-    }
   })
@@
   datatest <- reactive({
-    # let me know if anyone has a better idea for testing
+    # Testing if there are any characters instead of numbers (data probably
+    # hopelessly corrupted)
     if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
@@
     excel_test() |>
+      # Choosing the specified category (i.e. Revenue or Expenses)
       filter(
@@
       ) |>
+      # Calculating if there are any zeroes
       summarise(
@@
       ) |>
+      # just the single number
       pull()
@@
 ## zero_data_plot --------------------------------------------------------------
-
 zero_data <- function() {
-  empty_plot("All values are zero or missing.\nThis is unusual and may\nindicate a data problem.")
+  empty_plot(
+    "All values are zero or missing.
+    This is unusual and may
+    indicate a data problem.")
 }

### 2022-08-01 Report/Draft Dashboard section.Rmd
Updating with zero check and removing unnecessary check.

@@
   pdf_document:
-    highlight: pygments
+    highlight: default
   html_document:
@@
 
+They also make use of a function that checks that any data are available in the respective category:
+```{r dashboard_zero_check, eval=FALSE}
+## Excel zero test------------------------------------------------------------
+excel_zero_check <- function(type) {
+  excel_test() |>
+    # Choosing the specified category (i.e. Revenue or Expenses)
+    filter(
+      category == as.character(all_of(type))
+    ) |>
+    # Calculating if there are any zeroes
+    summarise(
+      zero = sum(value != 0, na.rm = TRUE)
+    ) |>
+    # just the single number
+    pull()
+}
+```
+
+Through a function that passes information to the `empty_plot()` function the user receives a warning if an entire category consists of only zeroes or NA:
+```{r}
+## zero_data_plot --------------------------------------------------------------
+zero_data <- function() {
+  empty_plot(
+    "All values are zero or missing.
+    This is unusual and may
+    indicate a data problem.")
+}
+```
+
 Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting - in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order.
@@
 # Reactive function to create the maximum value for revenue and expenses plots
-# the tests are probably not necessary, but included to be on the safe side
 maximum <- reactive({
-  if (datatest() == TRUE || setequal(
-    names(exceldata()),
-    c("year", "firm", "category", "sub_category", "name", "value")
-  )) {
-    NA
-  } else {
     excel_test() |>
@@
       max()
-  }
 })

### 2022-08-01 Dashboard/NGO Analysis dashboard v006.Rmd
Removing redundant comments.

@@
     ) |>
-    # creating a currency code column and turning current currency codes into
-    # NAs
     mutate(
@@
 ### First Excel transformation -------------------------------------------------
-# creating df used by NGO title function, maximum scale function, and excel zero check
+# creating df used by NGO title function, maximum scale function, and excel zero
+# check
 widen_excel_test <- function(file) {

### 2022-08-01 Dashboard/report.Rmd
Tidying. Changing to is.ggplot() test.

@@
 ### First Excel transformation -------------------------------------------------
-# creating df used by NGO title function and maximum scale function
+# creating df used by NGO title function, maximum scale function, and excel zero
+# check
 widen_excel_test <- function(file) {
@@
 # Creating general warning
-general_warning <- if (datatest == TRUE ||
+general_warning <- if (
+  datatest == TRUE ||
     sum(!is.na(htmldata$arsresultat)) == 0 ||
@@
     (sum(!is.na(htmldata$lederlonn)) == 0 &&
-    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0)) {
+      sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) ||
+    excel_zero_check("Revenue") == 0 ||
+    excel_zero_check("Expenses") == 0
+) {
   "## Note:"
@@
   if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
-    # Printing 'no information' plot if no data
     NA
@@
 # Printing the plot if the data are available
-if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) != 0) {
+if (is.ggplot(balance_plot)) {
   balance_plot
@@
   # testing for data quality
-  if (datatest == TRUE) {
-    # printing 'bad data' plot if data are bad
+  if (datatest == TRUE || excel_zero_check("Revenue") == 0) {
     NA
@@
 # Printing the plot if the data are OK
-if (datatest == FALSE && excel_zero_check("Revenue") != 0) {
+if (is.ggplot(revenue_plot)) {
   revenue_plot
@@
   # testing for data quality
-  if (datatest == TRUE) {
+  if (datatest == TRUE || excel_zero_check("Expenses") == 0) {
     NA
@@
 # Printing the plot if the data are OK
-if (datatest == FALSE && excel_zero_check("Expenses") != 0) {
+if (is.ggplot(expenses_plot)) {
   expenses_plot
@@
 # Printing the plot if the data are available
-if (sum(!is.na(htmldata$lederlonn)) != 0 ||
-  sum(!is.na(htmldata$leder_annen_godtgjorelse)) != 0) {
+if (is.ggplot(CEO_plot)) {
   CEO_plot

### 2022-08-01 Report/Draft Dashboard section.Rmd
Deleting comments, minor change in revenue and expenses section.

@@
   pdf_document:
-    highlight: default
+    highlight: haddock
   html_document:
@@
 As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with messages about missing packages:
-
-<!--# Might need to include tinytex? Or is that overkill? -- if (!require("tinytex") install.packages("tinytex"); tinytex::install_tinytex())-->
 
@@
 The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals.
-(This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function is based on first version of the Excel data frame, in order to make use of the Category column:
+(This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function is based on first version of the Excel data frame, in order to make use of the 'category' column:
 ```{r dashboard_maximum, eval=FALSE}

### 2022-08-01 Dashboard/report.Rmd
Manner of printing of revenues and expenses plots conditional on the presence of the other.

@@
 
-```{r revenue_expenses_plot, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
+```{r revenue_expenses_plots_creation}
 ## Revenue--------------------------------------------------------------------
@@
 
-# Printing the plot if the data are OK
-if (is.ggplot(revenue_plot)) {
-  revenue_plot
-}
-
 ## Expenses -------------------------------------------------------------------
@@
   }
+```
 
-# Printing the plot if the data are OK
-if (is.ggplot(expenses_plot)) {
+```{r revenue_only, fig.asp=1, out.width="50%", fig.cap="Based on the category \`Revenue' in the Excel sheet."}
+# Printing the plot if ONLY the revenue data are OK
+if (is.ggplot(revenue_plot) && !is.ggplot(expenses_plot)) {
+  revenue_plot
+}
+```
+
+```{r expenses_only, fig.asp=1, out.width="50%", fig.cap="Based on the category \`Expenses' in the Excel sheet."}
+# Printing the plot if ONLY the expenses data are ok
+if (!is.ggplot(revenue_plot) && is.ggplot(expenses_plot)) {
+  expenses_plot
+}
+```
+
+```{r revenue_expenses_plot_both, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
+
+# Printing the plots if the data are OK
+if (is.ggplot(revenue_plot) && is.ggplot(expenses_plot)) {
+  revenue_plot
+}
+
+if (is.ggplot(revenue_plot) && is.ggplot(expenses_plot)) {
   expenses_plot

### 2022-08-02 Report/Group 5 report.Rmd
Legger til toc.
Figure sizing.

@@
 output:
-  pdf_document: default
+  pdf_document: 
+    toc: true
   html_document: default
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, out.width="70%"}
 knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, out.width="70%"}
 knitr::include_graphics("Images/info_kreft.png")
+```
+
+```{r, echo=FALSE, out.width="100%"}
 knitr::include_graphics("Images/regnskapssti.png")

### 2022-08-02 Report/Group 5 report.Rmd
Comment onn Brønnøysund.
Brønnøysund update.

@@
 Also, we quickly found that Proff.no disallowed webscraping on their
-website. The Brønnøysund registry did store downloadable XML documents
-of all the financial statements on their website, but we discovered that
-access to these files cost about half a million NOK.
+website. The Brønnøysund registry does offer API access to all the financial statements, but we discovered that access to costs NOK 480,000.
 
@@
 organization. And, we attempted to request a temporary access to the
-Brønnøysund registry's XML documents.
+Brønnøysund registry's XML documents. <!-- # Ikke Brønnøysund XML -- de har API. Vi fikk en XML-dump fra SSB, men det er ikke noe OPX kan bruke.-->
 

### 2022-08-02 Report/Group 5 report.Rmd
Copying over dashboard section.

@@
     toc: true
+    highlight: pygments
   html_document: default
@@
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE, out.width = "100%")
+line <- "***"
 ```
@@
 
+## The Dashboard
+
+One of the outputs we have created is an example dashboard that calculates some KPIs and creates a few visualisations based on user-uploaded data from Proff.no and the Excel template we developed.
+The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'under the hood' was created using Shiny.
+Additionally, the dashboard provides the user with the option of downloading a pdf version of the analysis, using a parameterised R Markdown document.
+
+As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with messages about missing packages:
+
+`r if (knitr::is_latex_output()) line`
+
+```{r dashboard_packages, eval=FALSE}
+if (!require("pacman")) install.packages("pacman")
+library(pacman)
+p_load(
+  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
+  flexdashboard, MetBrewer, fontawesome, htmltools
+)
+```
+
+`r if (knitr::is_latex_output()) line`
+
+The dashboard consists of three main parts: a sidebar, the dashboard proper, and a 'Details' tab that provides basic information about the KPIs and plots.
+
+### The Sidebar
+
+The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data, and an 'Analyse' button which is connected to two Shiny `observe() |> bindEvent()` structures, ensuring that most of the code is not executed until the user has finished entering data.
+(Though both structures await the same button click, the division was necessary to remain within R's 10,000 byte limit.)
+
+`r if (knitr::is_latex_output()) line`
+
+```{r dashboard_sidebar, eval=FALSE}
+# HTML upload
+fileInput("html", label = "Upload html file", accept = ".html")
+
+# Year inputs
+numericInput("start", label = "Start", value = 2002)
+numericInput("end", label = "End", value = 2021)
+
+# Excel upload
+fileInput("excel", "Upload Excel file", accept = ".xlsx")
+
+# Button
+actionButton("click", "Analyse")
+```
+
+`r if (knitr::is_latex_output()) line`
+
+Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded.^[Note that the `accept` argument is not obeyed by the RStudio viewer, but most ordinary browsers enforce the instruction.]
+
+<!--# Parenthesis as footnote? -->
+
+If the validation is successful the files are read using the functions developed for this purpose, which conduct further testing, outputting error messages using `validate()`.
+Using these functions, the actual creation of data frames takes place in two stages for each file, partially for data testing and partially to create subsidiary data frames used by some functions.^[The data reading code is included as an appendix because of its length.]
+
+### The dashboard proper
+
+#### Key Performance Indicators
+
+The main section of the dashboard has a row-based layout, with a top row containing five 'value boxes' displaying KPIs derived from data in the Excel sheet, calculated and formatted in separate reactive functions.
+As Flexdashboard fails particularly gracelessly when errors occur with the value boxes, each function relies on some basic checks carried out by separate functions created for this purpose.
+In the case of both this is a simple function that checks if any of the numeric columns contain characters (in which case the data frame is probably hopelessly corrupted):
+```{r dashboard_datatest, eval=FALSE}
+datatest <- reactive({
+    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
+      TRUE
+    } else {
+      FALSE
+    }
+  })
+```
+
+If there are any characters, the calculation outputs `NA`, otherwise it carries out the calculation. One example:
+```{r dashboard_KPI_calculation, eval=FALSE}
+# Creating Fundraising ROI value
+KPI_fundraising_num <- reactive({
+  # testing for data quality, returning NA if bad
+  if (datatest() == TRUE) {
+    NA
+  } else {
+    exceldata() |>
+      # most recent year
+      filter(year == max(year)) |>
+      transmute(
+        # Calculating the value as a percentage; removing all other
+        KPI_fundraising_num =
+          (sum(revenue_private_contributions,
+            revenue_membership_fees,
+            na.rm = TRUE
+          ) - expenses_fundraising) /
+            expenses_fundraising * 100
+      ) |>
+      # just want the single number
+      pull() |>
+      first()
+  }
+})
+```
+
+The value box then also relies on a function that converts the number into a percentage string, checking in the process if the number is `NA` or `Inf` (indicating a division by zero, and thus either a data problem or data not suited to this calculation).
+```{r dashboard_KPI_string, eval=FALSE}
+# Function to turn KPI values into NA or percentage string
+KPI_to_string <- function(variable) {
+  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
+  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
+  # avoid passing an NA to subsequent operations.)
+  if (is.na(variable)) {
+    NA
+  } else if (variable == Inf) {
+    NA
+  } else {
+    # Shortening (lengthening) the value, adding %
+    str_c(format(variable, digits = 3), "%")
+  }
+}
+```
+
+In the case of some value boxes, such as the Fundraising ROI one, the function includes logic that changes the colour and icon of the box based on the calculated number.
+The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
+```{r dashboard_valuebox, eval=FALSE}
+# Creating value box for fundraising ROI
+output$KPI_box_fundraising <- renderValueBox({
+  valueBox(
+    value = KPI_to_string(KPI_fundraising_num()),
+    caption = "Fundraising ROI",
+    # Colour depends on value
+    color = if (is.na(KPI_fundraising_num())) {
+      "primary"
+    } else if (KPI_fundraising_num() < 100) {
+      "danger"
+    } else if (KPI_fundraising_num() < 200) {
+      "warning"
+    } else {
+      "success"
+    },
+    # icon depends on whether or not the value is NA
+    icon = if (is.na(KPI_fundraising_num())) {
+      "fa-solid fa-circle-xmark"
+    } else {
+      "fas fa-donate"
+    }
+  )
+})
+```
+
+All the boxes also include, as mentioned, a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1.
+All boxes use the theme colour 'primary' and the icon 'fa-solid fa-circle-xmark' if the value is NA.
+```{r dashboard_fontawesome, eval=FALSE}
+# Using fontawesome for more up-to-date (and currently supported) fontawesome
+# icons
+htmltools::tagList(fontawesome::fa_html_dependency())
+```
+
+#### Plots
+
+Below the row of KPI boxes, the dashboard contains three more rows containing five plots, three based on Proff.no data and two based on Excel data.
+The html-based plots include a basic test of whether the necessary data are available, as some or all may sometimes be missing, printing a 'no information' plot using a separate function if the data is missing.
+```{r html_datatest, eval=FALSE}
+# testing for the presence of data
+if (sum(!is.na(htmldata()$arsresultat)) == 0) {
+  # Printing 'no information' plot if no data
+  no_information()
+} else {
+  ggplot() # create plot
+}
+```
+
+The 'no information' plot is an essentially empty plot displaying a notice to the user, and is created through a function that passes label information into a more general blank plot function:
+```{r dashboard_no_information, eval=FALSE}
+## no_information plot ---------------------------------------------------------
+# Function to print a 'no information' plot, if necessary
+no_information <- function() {
+  empty_plot("No information available")
+}
+
+## empty_plot-------------------------------------------------------------------
+# Function to print an empty plot with a supplied message
+empty_plot <- function(warning) {
+  ggplot() +
+    # text geom with a space to create the plot, but otherwise show nothing
+    geom_text(aes(0, 0, label = " ")) +
+    # as blank a slate as possible
+    theme_void() +
+    annotate(
+      "text",
+      label = warning,
+      # Using the same palette as the other non-profit/loss plots
+      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
+    )
+}
+```
+
+The top row contains two plots showing the annual result and balance data for the years covered by Proff.no.
+These are fairly simple bar plots, including a `fill` parameter colouring the bars depending on whether the numbers are positive or negative.
+They do, however, rely on two simple additional functions, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, or, otherwise, uses a question mark instead of a currency code, both using `scales::label_dollar()`.
+```{r dashboard_currency, eval=FALSE}
+# Reactive function to check for currency and use currency code or ?
+currency_scale <- reactive({
+  scale_y_continuous(
+    # Checking for uniform currency code
+    labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
+      # Using the currency code if uniform
+      label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
+      # alternative prefix if multiple currencies in the data
+      label_dollar(prefix = "? ")
+    )
+  )
+})
+```
+
+A second function is used for the `fill` parameter in order to (1) ensure that positive numbers are always green, irrespective of the presence of negative numbers, and vice versa, and (2) to use slightly nicer colours than the default ggplot2 ones, making use of the 'Tsimshian' palette from Blake R.
+Mills' creative [MetBrewer](https://github.com/BlakeRMills/MetBrewer) package.
+```{r}
+## scale_fill_green ------------------------------------------------------------
+# Function with slightly nicer colours than default; ensuring uniformity; +
+# ensuring positive values are always green through a test for the presence of
+# FALSE values
+scale_fill_green <- function(variable) {
+  scale_fill_met_d("Tsimshian",
+    direction = ifelse(
+      mean(variable,
+        na.rm = TRUE
+      ) == 1, -1, 1
+    )
+  )
+}
+```
+
+The plots are converted to plotly using `ggplotly()`.
+As subtitles and captions do not survive the conversion, a separate function outputs a plotly annotation if the data contains multiple currency codes:
+```{r dashboard_plotly, eval=FALSE}
+# Creating plotly output
+output$arsresultat_plotly <- renderPlotly(
+  ggplotly(arsresultat_plot()) |>
+    ## ggplot subtitles and captions are lost when using ggplotly(), a
+    ## currency warning caption is therefore added in plotly, if needed, using
+    ## a separate function
+    layout(
+      annotations =
+        plotly_caption()
+    )
+)
+
+# Reactive function to create plotly caption (ggplot captions and subtitles
+# are lost in the conversion) (shamelessly stolen from the Internet, with
+# minor modifications).
+# Checks for uniform currency code and prints a warning
+# if there are multiple currency codes.
+plotly_caption <- reactive({
+  list(
+    x = 1, y = -0.1,
+    text = ifelse(length(unique(html_currency()$valutakode)) == 1,
+      "", "Note: not all values are denominated in the same currency"
+    ),
+    showarrow = F, xref = "paper", yref = "paper",
+    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
+    font = list(size = 11)
+  )
+})
+```
+
+Below these plots are two plots based on Excel data.
+These both use the same check for data quality as the value box, outputting a 'bad data' plot if the data are bad using a function that passes label information to the `empty_plot()` function:
+```{r dashboard_excel_datatest, eval=FALSE}
+## bad_data plot ---------------------------------------------------------------
+# Function to print a 'bad data' plot, if necessary
+bad_data <- function() {
+  empty_plot("Something is wrong\nwith the Excel data.")
+}
+```
+
+They also make use of a function that checks that any data are available in the respective category:
+```{r dashboard_zero_check, eval=FALSE}
+## Excel zero test------------------------------------------------------------
+excel_zero_check <- function(type) {
+  excel_test() |>
+    # Choosing the specified category (i.e. Revenue or Expenses)
+    filter(
+      category == as.character(all_of(type))
+    ) |>
+    # Calculating if there are any zeroes
+    summarise(
+      zero = sum(value != 0, na.rm = TRUE)
+    ) |>
+    # just the single number
+    pull()
+}
+```
+
+Through a function that passes information to the `empty_plot()` function the user receives a warning if an entire category consists of only zeroes or NA:
+```{r}
+## zero_data_plot --------------------------------------------------------------
+zero_data <- function() {
+  empty_plot(
+    "All values are zero or missing.
+    This is unusual and may
+    indicate a data problem.")
+}
+```
+
+Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting - in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order.
+Excel data are presumed to be supplied in NOK.
+```{r dashboard_excelplot, eval=FALSE}
+exceldata() |>
+  # Renaming for the purposes of the plot + dropping other variables
+  transmute(
+    Year = as.factor(year),
+    `Private contributions` = revenue_private_contributions,
+    `Membership fees` = revenue_membership_fees,
+    `Public grants` = revenue_public_grants,
+    Investments = revenue_investments,
+    `Operational activities` = revenue_operational_activities,
+    `Commercial activities` = revenue_commercial_activities,
+    Other = revenue_other
+  ) |>
+  # Pivoting to use fill
+  pivot_longer(
+    c(`Private contributions`:Other),
+    names_to = "Revenue source"
+  ) |>
+  # Ordering for the plot
+  mutate(
+    `Revenue source` =
+      ordered(`Revenue source`,
+        levels =
+          c(
+            "Other", "Commercial activities",
+            "Operational activities", "Investments",
+            "Membership fees", "Private contributions",
+            "Public grants"
+          )
+      )
+  ) |>
+  ggplot(aes(Year, value, fill = `Revenue source`)) +
+  geom_col() +
+  scale_y_continuous(
+    # Excel data are assumed to be in NOK; if this changes, code based on
+    # the currency_code() function may be adapted
+    labels = label_dollar(prefix = "NOK "),
+    # Basing lower limit on the data (not suppressing erroneous negative
+    # numbers), and the upper limit on the maximum value in revenues and
+    # expenses, to match the two plots
+    limits = c(NA, maximum())
+  ) +
+  # Specifying the number of values, as the base palette only includes 6
+  scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
+  labs(x = "", y = "", title = "Revenue (in thousands)")
+```
+
+The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data.
+The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals.
+(This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function is based on first version of the Excel data frame, in order to make use of the 'category' column:
+```{r dashboard_maximum, eval=FALSE}
+## scale maximum--------------------------------------------------------------
+# Reactive function to create the maximum value for revenue and expenses plots
+maximum <- reactive({
+    excel_test() |>
+      # removing the firm and sub-category columns
+      select(-firm, -sub_category) |>
+      # filtering to retain only revenue and expenses entries
+      filter(category == "Revenue" | category == "Expenses") |>
+      # grouping and summarising to calculate maximum value
+      group_by(year, category) |>
+      summarise(
+        sum = sum(value, na.rm = TRUE)
+      ) |>
+      # Creating list of sums
+      pull() |>
+      # selecting the highest value
+      max()
+})
+```
+
+All the plots based on the Excel data make use of the colour-blind-friendly "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories.
+The same function is used to generate the text colour in the 'no information' and 'bad data' plots.
+
+Below these two plots is a plot showing the total reported CEO compensation in the years covered by the html data, where this information is available.
+
+### Details tab
+
+Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots. At the top it includes a heading containing the name of the NGO as written in the Excel file:
+```{r dashboard_title_creation, eval=FALSE}
+# NGO title ==================================================================
+NGO <- reactive({
+  # checking for the required column again
+  if ("firm" %in% names(excel_test())) {
+    # fetching the name
+    excel_test() |>
+      # Turning the needed column into a list
+      pull(firm) |>
+      # discarding all NAs in case the name is somewhere
+      purrr::discard(is.na) |>
+      # choosing the first available name
+      first()
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
+})
+
+title <- reactive({
+  # testing if NGO() is NA
+  if (is.na(NGO())) {
+    # Printing warning if NA
+    "Name of NGO missing"
+    # Printing name of NGO
+  } else {
+    NGO()
+  }
+})
+```
+
+The title is output using `htmlOutput()` to use some basic html styling:
+```{r dashboard_title, eval=FALSE}
+# Creating NGO title using html formatting (colour based on MetBrewer
+# "Hokusai2")
+output$title <- renderText({
+  paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
+})
+```
+
+### The parameterised R Markdown report
+
+The dashboard also allows the user to download a pdf version of the analysis through clicking on a button.
+A `downloadHandler()` function passes the necessary parameters to the R Markdown document and displays a temporary notification while the file is being created.
+The function includes temporary file creation arguments that should enable it to function on e.g.
+shinyapps.io and similar server contexts, but this has not been tested.
+
+In order to prevent the user from the downloading a nonsensical html file if bad data have been uploaded or entered, the button creation and report generation are located in the second `observe() |> bindEvent()` structure.
+The creation of the download button is connected to two convenience logical tests, the primary purpose of which is to connect the button to the `validate()` calls upstream in the reactive chain.
+
+The file name of the report is created by combining the name of the NGO from the `title()` reactive and the output of `Sys.Date()`, furnishing a hopefully informative file name.^[Note that this only works when the dashboard is run in an ordinary browser and not in the RStudio viewer.]
+
 # Limitations and steps forward

### 2022-08-02 Report/Group 5 report.Rmd
Fixing XML.

@@
 options:
-- Financial statements in yearly report PDF documents. 
+
+- Financial statements in annual report PDF documents. 
 - Webscrape Proff.no for financial statements. 
@@
 possible to download Proff.no HTML websites manually for each
-organization. And, we attempted to request a temporary access to the
-Brønnøysund registry's XML documents. <!-- # Ikke Brønnøysund XML -- de har API. Vi fikk en XML-dump fra SSB, men det er ikke noe OPX kan bruke.-->
+organization. We did receive some XML files from Statistics Norway, but using these is ultimately not a viable solution for OPX.
 

### 2022-08-02 Report/Group 5 report.Rmd
Uploading appendix.

@@
 # Our experience
+
+# Appendix: Data reading and cleaning code
+```{r dashboard_file_functions, eval=FALSE}
+# Data reading functions =======================================================
+
+## Functions to read html file -------------------------------------------------
+html_test_create <- function(file, start, end) {
+  # Reading file and node
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  ### A series of tests to catch problems before they spit out errors-----------
+  # Checking if node exists
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |>
+      html_table()
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Basic check if the table is correct
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Testing if the correct years have been entered
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    # Creating the first part of the data frame
+    read_proff_one(htmltable, start, end)
+  } else {
+    validate("Please ensure you have entered the correct start and end years.")
+  }
+}
+
+### First html transformation function, invoked by html_test_create() ----------
+# (stopping to allow checks for differing currency codes)
+read_proff_one <- function(object, start, end) {
+  object |>
+    # removing empty column, using all_of(argument)
+    select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column to avoid having to deal with NAs later
+    fill(
+      valutakode,
+      .direction = "downup"
+    )
+}
+
+### Second html transformation function, creating the data frame ---------------
+read_proff_two <- function(object) {
+  object |>
+    # removing unneeded currency column
+    select(-valutakode) |>
+    # removing dates and redundant currency
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots (red/green profit/loss)
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+}
+
+## Functions to read Excel file ------------------------------------------------
+### First Excel transformation - creating df used by NGO title function---------
+widen_excel_test <- function(file) {
+  # Reading Excel file
+  read_excel(file) |>
+    # cleaning names to enable later operations
+    clean_names()
+}
+
+### Second Excel function - creating primary Excel data frame-------------------
+widen_excel <- function(object) {
+  # testing if all columns are present
+  if (setequal(
+    names(object),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  )) {
+    object |>
+      # uniting category and name to deal with multiple 'Other'
+      unite(
+        entry, c(category, name)
+      ) |>
+      # deleting Sub-category and Firm
+      select(-sub_category, -firm) |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(names_from = entry, values_from = value) |>
+      # cleaning names
+      clean_names()
+    # Notification if all columns are not present
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
+}
+
+# Data frames ================================================================
+## Creating data frame from html file-----------------------------------------
+html_currency <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$html$datapath) == "html",
+    "Please upload an html file."
+  ))
+
+  # Testing and creating data frame
+  html_test_create(
+    input$html$datapath,
+    as.character(input$start), as.character(input$end)
+  )
+})
+
+htmldata <- reactive({
+  read_proff_two(html_currency())
+})
+
+## Validating and creating long version of Excel file, if possible------------
+excel_test <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$excel$datapath) == "xlsx",
+    "Please upload an Excel file."
+  ))
+
+  widen_excel_test(input$excel$datapath)
+})
+
+## Creating the primary Excel data frame--------------------------------------
+exceldata <- reactive({
+  widen_excel(excel_test())
+})
+```

### 2022-08-02 Report/Group 5 report.Rmd
endret på intro
Adding dashboard image.

@@
 
+It is important to solve this problem because:
+
+1. Automating the data collection will streamline the work of OPX 
+and create more transparency which enhances trust in charities.
+
+2. Creating KPI(s) is important for the organization’s performance 
+management, to influence the objective, and to foster growth.
+
 The scope of the problem statement was narrowed down to the Nordic 
 countries as a starting point.
-
-It is important to solve this problem because:
-
-1. Creating KPI(s) is important for the organization’s performance 
-management, to influence the objective, and to foster growth.
-
-2. Automating the data collection will streamline the work of OPX 
-and create more transparency which enhances trust in charities.
 
@@
 ### The dashboard proper
+
+```{r, out.width="65%", fig.cap="The dashboard", fig.align='center', echo=FALSE}
+knitr::include_graphics("Images/Dashboard_page_1.png")
+```
 
@@
 The function includes temporary file creation arguments that should enable it to function on e.g.
-shinyapps.io and similar server contexts, but this has not been tested.
+[shinyapps.io](https://www.shinyapps.io/) and similar server contexts.
+It has been tested on shinyapps.io, but not on any private server.
 

### 2022-08-02 Presentasjon/presentasjon.Rmd
Flere bilder

--- 2022-08-01 Presentasjon/presentasjon.Rmd
@@
 
-# Introduction
+## {data-background="Images/start1.png" data-background-size=cover}
 
 ## <br> OPX and their goal
-- A startup developing solutions for socially responsible investment strategies
-- The goal is to increase cash flow to charitable purposes 
-- Donors' trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficients <br>
+- Develop solutions for socially responsible investment strategies
+- Increase cash flow to charitable purposes <br>
 
@@
 
+## {data-background="Images/fcuk.png" data-background-size=cover}
+
 # Data collection
@@
 # Limitations and steps forward
+
+## <iframe width="1280" height="720" src="Images/henrik.mp4" align="middle" frameborder="0" allowfullscreen></iframe>

### 2022-08-02 Report/Group 5 report.Rmd
endring under data description
Images.
flyttet bilde
Adding images.

--- 2022-08-02 Report/Group 5 report.Rmd
@@
 
+<!-- Insert images here -->
+
 # Findings
@@
 
-<!--# Parenthesis as footnote? -->
-
 If the validation is successful the files are read using the functions developed for this purpose, which conduct further testing, outputting error messages using `validate()`.
@@
 
+```{r, out.width="50%", fig.cap="The main dashboard page", fig.align='center', echo=FALSE}
+knitr::include_graphics("Images/Dashboard_page_1.png")
+```
 #### Key Performance Indicators
@@
 ### Details tab
+```{r dashboard_page_2, out.width="50%", fig.cap="The details tab of the dashboard.", echo=FALSE}
+knitr::include_graphics("Images/Dashboard_page_2.png")
+```
 
@@
 ### The parameterised R Markdown report
+
+```{r dashboard_report_image, echo=FALSE, out.width="50%"}
+knitr::include_graphics("Images/Dashboard_report.png")
+```
 

### 2022-08-02 Dashboard/Dashboard testing material/zero testing.Rmd

---
title: "Zero testing"
author: "Torbjørn Skinnemoen Ottersen"
date: "2022-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/ISSSV1337/ISSSV1337-Case/Dashboard")
```

```{r}
x <- c(5, NA, 0, 0, 10, -15)

sum(x == 0, na.rm = TRUE)

warning_message <- "ALl values are zero. This is unusual and may indicate a data problem."

?pull

setwd("~/ISSSV1337/ISSSV1337-Case/Dashboard")

exceltest <- read_excel("Data/Kreftforeningen.xlsx") |> clean_names()

exceltest

revenue_zero <- exceltest |>
  filter(category == "Revenue") |>
  summarise(
    zero = sum(value != 0, na.rm = TRUE)
  ) |> 
  pull()

revenue_zero

excel_zero_check <- function(type) {
  exceltest |>
    filter(
      category == as.character(all_of(type))
    ) |>
    summarise(
      zero = sum(value != 0, na.rm = TRUE)
    ) |>
    pull()
}

excel_zero_check("Revenue")
```

### 2022-08-02 Dashboard/report.Rmd

@@
 exceldata <- widen_excel(excel_test)
+```
 
+```{r ngo_title}
 # NGO title ==================================================================
-
 NGO <- excel_test |>
@@
 
-title <- # testing if NGO() is NA
+title <-
+  # testing if NGO() is NA
   if (is.na(NGO)) {
@@
 
-# Creating general warning
+## Creating general warning-----------------------------------------------------
 general_warning <- if (
@@
 # Function to create the maximum value for revenue and expenses plots
+# Testing, because this function cannot rely on the reactive chain to stop
 maximum <- if (datatest == TRUE) {
@@
 ```{r revenue_expenses_plot_both, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
-
 # Printing the plots if the data are OK

### 2022-08-02 Report/Group 5 report.Rmd
Image centred.
fig.align.

--- 2022-08-02 Report/Group 5 report.Rmd
@@
 organization. We did receive some XML files from Statistics Norway, but using these is ultimately not a viable solution for OPX.
-
-```{r}
-
-```
 
@@
 # Data description
+
+<!--Legg inn bilder her-->
 
@@
 ### Details tab
+```{r dashboard_page_2, out.width="50%", fig.cap="The details tab of the dashboard.", fig.align='center', echo=FALSE}
+knitr::include_graphics("Images/Dashboard_page_2.png")
+```
 
@@
 ### The parameterised R Markdown report
+
+```{r dashboard_report_image, echo=FALSE, out.width="50%", fig.align='center'}
+knitr::include_graphics("Images/Dashboard_report.png")
+```
 

### 2022-08-02 Report/Group 5 report.Rmd
Adding lines between code blocks.

@@
 In the case of both this is a simple function that checks if any of the numeric columns contain characters (in which case the data frame is probably hopelessly corrupted):
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_datatest, eval=FALSE}
@@
 ```
+
+`r if (knitr::is_latex_output()) line`
 
@@
 
+`r if (knitr::is_latex_output()) line`
+
 The value box then also relies on a function that converts the number into a percentage string, checking in the process if the number is `NA` or `Inf` (indicating a division by zero, and thus either a data problem or data not suited to this calculation).
@@
 
+`r if (knitr::is_latex_output()) line`
+
 All the boxes also include, as mentioned, a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1.
@@
 
+`r if (knitr::is_latex_output()) line`
+
 The top row contains two plots showing the annual result and balance data for the years covered by Proff.no.
@@
 They do, however, rely on two simple additional functions, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, or, otherwise, uses a question mark instead of a currency code, both using `scales::label_dollar()`.
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_currency, eval=FALSE}
@@
 
+`r if (knitr::is_latex_output()) line`
+
 A second function is used for the `fill` parameter in order to (1) ensure that positive numbers are always green, irrespective of the presence of negative numbers, and vice versa, and (2) to use slightly nicer colours than the default ggplot2 ones, making use of the 'Tsimshian' palette from Blake R.
 Mills' creative [MetBrewer](https://github.com/BlakeRMills/MetBrewer) package.
+
+`r if (knitr::is_latex_output()) line`
+
 ```{r}
@@
 
+`r if (knitr::is_latex_output()) line`
+
 The plots are converted to plotly using `ggplotly()`.
 As subtitles and captions do not survive the conversion, a separate function outputs a plotly annotation if the data contains multiple currency codes:
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_plotly, eval=FALSE}
@@
 
+`r if (knitr::is_latex_output()) line`
+
 Below these plots are two plots based on Excel data.
@@
 
+`r if (knitr::is_latex_output()) line`
+
 They also make use of a function that checks that any data are available in the respective category:
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_zero_check, eval=FALSE}
@@
 
+`r if (knitr::is_latex_output()) line`
+
 Through a function that passes information to the `empty_plot()` function the user receives a warning if an entire category consists of only zeroes or NA:
+
+`r if (knitr::is_latex_output()) line`
 ```{r}
@@
 
+`r if (knitr::is_latex_output()) line`
+
 The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data.
@@
 (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function is based on first version of the Excel data frame, in order to make use of the 'category' column:
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_maximum, eval=FALSE}
@@
 
+`r if (knitr::is_latex_output()) line`
+
 All the plots based on the Excel data make use of the colour-blind-friendly "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories.
@@
 Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots. At the top it includes a heading containing the name of the NGO as written in the Excel file:
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_title_creation, eval=FALSE}
@@
 
+`r if (knitr::is_latex_output()) line`
+
 The title is output using `htmlOutput()` to use some basic html styling:

### 2022-08-02 Report/Group 5 report.Rmd
Adding missing line.

@@
 If there are any characters, the calculation outputs `NA`, otherwise it carries out the calculation. One example:
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_KPI_calculation, eval=FALSE}

### 2022-08-02 Report/Group 5 report.Rmd
Logical headings.
Removing empty lines at end of document.
Starting report on p. 2.
Captions comment.
Numbered lines.

@@
 
+<!--# Starting report on second page to avoid a tiny text portion below the toc-->
+\newpage
 # Introduction
@@
 
-### Phase 1: How to harvest and analyze the data? {.smaller}
+## Phase 1: How to harvest and analyze the data? {.smaller}
 
@@
 
+<!-- # All images need captions-->
+
 ```{r, echo=FALSE, out.width="70%"}
@@
 
-### Phase 2: Exploring solutions {.smaller}
+## Phase 2: Exploring solutions {.smaller}
 
@@
 
-### Phase 3: Scrapping the scraping {.smaller}
+## Phase 3: Scrapping the scraping {.smaller}
 
@@
 The limitations can be divided into problems regarding:
-(i) the accessible data, and 
-(ii) the non-accessible data
+
+1. the accessible data, and 
+2. the non-accessible data
 

### 2022-08-02 Report/Group 5 report.Rmd
Full stop.
Copying over Our experience.

@@
 1. the accessible data, and 
-2. the non-accessible data
+2. the non-accessible data.
 
@@
 # Our experience
+
+In this last section, we would like to describe how we experienced this process and what we have learned.
+As explained, the main challenge with the task was highlighted by an unsolvable due to non-existent data.
+In cases where data were available, they were inadequate.
+
+In addition, parts of the case demanded knowledge from the field of economics, in which we were lucky to have Amalie on the team; this was, however, still an area that was new to all of us.
+Without detailed experience in the NGO sector, it was challenging to find credible KPIs and interpret these correctly.
+
+The lectures allowed us to learn about many interesting topics.
+However, as the connection between what we learned there and what we got to utilize in the case was distant, and parts were too advanced for some of us, especially those without any particular experience within statistics.
+
+Despite this, we have done our best to work with the resources at hand.
+As shown in this presentation, trying out solutions and later scrapping them has been a big part of the problem-solving process.
+We also got an insight into real-life work situations, such as the challenges that come when what is desired clashes with what is possible.
+We are certain that these experiences will be useful for all of us in the future, and we are thankful for being given this opportunity.
 

### 2022-08-02 Report/Group 5 report.Rmd
Adding libraries.

@@
 knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE, out.width = "100%")
+```
+
+```{r libraries, include=FALSE}
+library(tidyverse)
+library(janitor)
+library(rvest)
+library(MetBrewer)
+```
+
+```{r line, include=FALSE}
 line <- "***"
+```
+
+```{r}
+
 ```

### 2022-08-02 Report/Graphs for final report.Rmd
Document to work on graphs for report.

--- 2022-08-01 Dashboard/report.Rmd
@@
 ---
-date: "`r Sys.Date()`"
-output: 
-  pdf_document:
-    extra_dependencies: "subfig"
-params:
-  htmlfile: NA
-  excelfile: NA
-  start: NA
-  end: NA
+title: "Graphs for final report"
+author: "Torbjørn Skinnemoen Ottersen"
+date: "2022-08-02"
+output: html_document
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 0.618, out.width = "75%", fig.align = "center")
-
-# Repeating loading the libraries, in case the information isn't passed on from
-# the dashboard
-if (!require("pacman")) install.packages("pacman")
-library(pacman)
-p_load(
-  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
-  flexdashboard, MetBrewer, fontawesome, htmltools, purrr
-)
+knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(janitor)
+library(rvest)
+library(MetBrewer)
 ```
 
-<!--# The title element of the YAML block is further down, as it relies on data from the Excel sheet. -->
-
-```{r data}
+```{r}
 # Data reading =================================================================
@@
 
-# Creating a data frame for currency (for some reason the functions in the
-# dashboard create bad data, but *only* in a *parameterised* R Markdown report).
-read_proff_currency <- function(file, start, end) {
-  read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame") |>
-    html_table() |>
-    # removing empty column, using all_of(argument)
-    select(
-      REGNSKAPSPERIODE:all_of(start)
-    ) |>
-    # removing white space
-    mutate(
-      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-    ) |>
-    # removing duplicate table sometimes created by read_html()
-    filter(
-      !row_number() > 176
-    ) |>
-    # removing duplicate rows
-    filter(
-      !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |>
-    # tidying data using pivot_longer() and all_of(argument)
-    pivot_longer(
-      all_of(end):all_of(start),
-      names_to = "year"
-    ) |>
-    # creating a currency code column and turning current currency codes into
-    # NAs
-    mutate(
-      # changing to real NAs
-      value = na_if(value, "-"),
-      # turning years into numbers
-      year = as.integer(year),
-      # creating a currency code column and turning current currency codes into
-      # NAs
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |>
-    fill(
-      valutakode,
-      .direction = "downup"
-    )
-}
-
-### Reading html ---------------------
-# Creating the main data frame
-htmldata <- read_proff(
-  params$htmlfile,
-  as.character(params$start), as.character(params$end)
-)
-
-# Creating the currency data frame
-html_currency <- read_proff_currency(
-  params$htmlfile,
-  as.character(params$start), as.character(params$end)
-)
-
-## Functions to read Excel file ------------------------------------------------
-### First Excel transformation -------------------------------------------------
-# creating df used by NGO title function, maximum scale function, and excel zero
-# check
-widen_excel_test <- function(file) {
-  # Reading Excel file
-  read_excel(file) |>
-    # cleaning names to enable later operations
-    clean_names()
-}
-
-### Second Excel function - creating primary Excel data frame-------------------
-widen_excel <- function(object) {
-  object |>
-    # uniting category and name to deal with multiple 'Other'
-    unite(
-      entry, c(category, name),
-      sep = "_"
-    ) |>
-    # deleting sub_category and firm
-    select(-sub_category, -firm) |>
-    # pivoting to create 'tidy' variables
-    pivot_wider(names_from = entry, values_from = value) |>
-    # cleaning names
-    clean_names()
-}
-
-### Reading excel ---------------------
-excel_test <- widen_excel_test(params$excelfile)
-
-exceldata <- widen_excel(excel_test)
-
-# NGO title ==================================================================
-
-NGO <- excel_test |>
-  # Turning the needed column into a list
-  pull(firm) |>
-  # discarding all NAs in case the name is somewhere
-  purrr::discard(is.na) |>
-  # choosing the first available name
-  first()
-
-title <- # testing if NGO() is NA
-  if (is.na(NGO)) {
-    # Printing warning if NA
-    "Name of NGO missing"
-    # Printing name of NGO
-  } else {
-    NGO
-  }
+kreftforeningen <- read_proff("Data/Kreftforeningen.html", "2002", "2021")
 ```
 
-<!--# Report title-->
-
----
-title: "NGO Analysis: `r title`"
----
-
-```{r data_warnings}
-# Excel data tests and warnings ================================================
-# Functions to test for Excel data quality (basic)
-## Excel datatest---------------------------------------------------------------
-datatest <-
-  # let me know if anyone has a better idea for testing
-  if (mean(unlist(lapply(exceldata, is.character))) != 0) {
-    TRUE
-  } else {
-    FALSE
-  }
-
-## Excel zero test------------------------------------------------------------
-excel_zero_check <- function(type) {
-  excel_test |>
-    filter(
-      category == as.character(all_of(type))
-    ) |>
-    summarise(
-      zero = sum(value != 0, na.rm = TRUE)
-    ) |>
-    pull()
-}
-
-## Creating the Excel warning texts --------------------------------------------
-# Excel data test warning
-excel_warning <- if (datatest == TRUE) {
-  "***    Warning: something is wrong with the Excel data.***"
-} else {
-  ""
-}
-
-# Excel zero revenues warning
-excel_zero_revenue_warning <- if (excel_zero_check("Revenue") == 0) {
-  "*    All revenue values are zero or missing. This is unusual and may indicate a data problem.*"
-} else {
-  ""
-}
-
-# Excel zero expenses warning
-excel_zero_expenses_warning <- if (excel_zero_check("Expenses") == 0) {
-  "*    All expenses values are zero or missing. This is unusual and may indicate a data problem.*"
-} else {
-  ""
-}
-
-# HTML warnings ================================================================
-# Creating the arsresultat warning text
-arsresultat_plot_warning <- if (sum(!is.na(htmldata$arsresultat)) == 0) {
-  "*    No information about net income/loss is available.*"
-} else {
-  ""
-}
-
-# Creating the balance warning text
-balance_plot_warning <- if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
-  "*    No information about total equity/liabilities is available.*"
-} else {
-  ""
-}
-
-# Creating the CEO_plot warning text
-CEO_plot_warning <- if (sum(!is.na(htmldata$lederlonn)) == 0 &&
-  sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
-  "*    No information about CEO compensation is available.*"
-} else {
-  ""
-}
-
-# Creating general warning
-general_warning <- if (
-  datatest == TRUE ||
-    sum(!is.na(htmldata$arsresultat)) == 0 ||
-    sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 ||
-    (sum(!is.na(htmldata$lederlonn)) == 0 &&
-      sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) ||
-    excel_zero_check("Revenue") == 0 ||
-    excel_zero_check("Expenses") == 0
-) {
-  "## Note:"
-} else {
-  ""
-}
-```
-
-<!--# Functions that don't rely on any of the parameters passed from the dashboard -->
-
-```{r static_functions}
-## KPI_to_string ---------------------------------------------------------------
-# Function to turn KPI values into NA or percentage string
-KPI_to_string <- function(variable) {
-  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
-  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
-  # avoid passing an NA to subsequent operations.)
-  if (is.na(variable)) {
-    NA
-  } else if (variable == Inf) {
-    NA
-  } else {
-    # Shortening (lengthening) the value, adding %; making the text bold
-    str_c("**", format(variable, digits = 3), "%**")
-  }
-}
-
+```{r}
 ## scale_fill_green ------------------------------------------------------------
@@
 }
+
 ```
 
-<!--# Functions that rely on parameters from the dashboard -->
-
-```{r functions_reliant}
-## scale maximum--------------------------------------------------------------
-# Function to create the maximum value for revenue and expenses plots
-maximum <- if (datatest == TRUE) {
-  NA
-} else {
-  excel_test |>
-    # removing the firm and sub-category columns
-    select(-firm, -sub_category) |>
-    # filtering to retain only revenue and expenses entries
-    filter(category == "Revenue" | category == "Expenses") |>
-    # grouping and summarising to calculate maximum value
-    group_by(year, category) |>
-    summarise(
-      sum = sum(value, na.rm = TRUE)
-    ) |>
-    # Creating list of sums
-    pull() |>
-    # selecting the highest value
-    max()
-}
-
-## Currency scale-------------------------------------------------------------
-# Function to check for currency and use currency code or ?
-currency_scale <-
-  scale_y_continuous(
-    # Checking for uniform currency code
-    labels = ifelse(length(unique(html_currency$valutakode)) == 1,
-      # Using the currency code if uniform
-      label_dollar(prefix = str_c(unique(html_currency$valutakode), " ")),
-      # alternative prefix if multiple currencies in the data
-      label_dollar(prefix = "? ")
-    )
-  )
-
-## Currency caption-----------------------------------------------------------
-currency_caption <- ifelse(length(unique(html_currency$valutakode)) == 1,
-  "", "Note: not all values are denominated in the same currency."
-)
-```
-
-<!--# Creating KPIs -->
-
-```{r KPIs}
-## Fundraising ROI------------------------------------------------------------
-# Creating Fundraising ROI value
-# testing for data quality, returning NA if bad
-fundraising <- if (datatest == TRUE) {
-  NA
-} else {
-  exceldata |>
-    # most recent year
-    filter(year == max(year)) |>
-    transmute(
-      # Calculating the value as a percentage; removing all other
-      KPI_fundraising_num =
-        (sum(revenue_private_contributions,
-          revenue_membership_fees,
-          na.rm = TRUE
-        ) - expenses_fundraising) /
-          expenses_fundraising * 100
-    ) |>
-    # just want the single number
-    pull() |>
-    first() |>
-    KPI_to_string()
-}
-
-## OPX KPI--------------------------------------------------------------------
-# Creating OPX KPI
-KPI_OPX <-
-  # testing for data quality, returning NA if bad
-  if (datatest == TRUE) {
-    NA
-  } else {
-    exceldata |>
-      # most recent year
-      filter(year == max(year)) |>
-      # Calculating the value as a %
-      transmute(
-        KPI_OPX =
-          sum(revenue_private_contributions,
-            revenue_investments, assets_cash_in_bank_and_cash_equivalents,
-            na.rm = TRUE
-          ) /
-            sum(liability_short_term_grants_payable,
-              liability_revocable_endowments,
-              liability_long_term_grants_payable,
-              na.rm = TRUE
-            ) * 100
-      ) |>
-      # Just the single number
-      pull() |>
-      first() |>
-      KPI_to_string()
-  }
-
-## Private donations ---------------------------------------------------------
-# Creating private donations growth
-KPI_private <-
-  # testing for data quality, returning NA if bad
-  if (datatest == TRUE) {
-    NA
-  } else {
-    exceldata |>
-      # Calculating the two most recent years as a %
-      transmute(
-        growth =
-          (revenue_private_contributions[1] -
-            revenue_private_contributions[2]) /
-            revenue_private_contributions[2] * 100
-      ) |>
-      # Pulling the first instance of the number
-      pull() |>
-      first() |>
-      KPI_to_string()
-  }
-
-## Public grants growth-------------------------------------------------------
-# Creating public grants growth
-KPI_public <-
-  # testing for data quality, returning NA if bad
-  if (datatest == TRUE) {
-    NA
-  } else {
-    exceldata |>
-      # Calculating the two most recent years as a %
-      transmute(
-        growth =
-          (revenue_public_grants[1] - revenue_public_grants[2]) /
-            revenue_public_grants[2] * 100
-      ) |>
-      # Pulling the first instance of the number
-      pull() |>
-      first() |>
-      KPI_to_string()
-  }
-
-## Public/private proportion--------------------------------------------------
-# Creating public/private proportion
-KPI_proportion <-
-  # testing for data quality, returning NA if bad
-  if (datatest == TRUE) {
-    NA
-  } else {
-    exceldata |>
-      transmute(
-        proportion =
-          sum(revenue_private_contributions,
-            revenue_membership_fees,
-            na.rm = TRUE
-          ) /
-            sum(revenue_public_grants, revenue_private_contributions,
-              revenue_membership_fees,
-              na.rm = TRUE
-            ) * 100
-      ) |>
-      pull() |>
-      first() |>
-      KPI_to_string()
-  }
-
-### NA warning-----------------------------------------------------------------
-# Testing for the presence of NAs; printing warning if there are any
-KPI_warning <- if (is.na(fundraising) || is.na(KPI_OPX) || is.na(KPI_private) ||
-  is.na(KPI_public) || is.na(KPI_proportion)) {
-  "*A value of 'NA' indicates that there is something wrong with the Excel data, that the necessary data are not available, or that the data are unsuited to an automatic calculation of the KPI.*"
-} else {
-  ""
-}
-```
-
-# Key Performance Indicators
-
-The KPIs are based on the following calculations, using data derived from the Excel sheet:
-
-- **Fundraising ROI:** `r fundraising`
-  - A basic measure of the return on investment of direct fundraising expenses:
-  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$
-
-- **Liquid assets/obligations:** `r KPI_OPX`
-  - A measure of the relationship between the organisation's liquid assets and its obligations:
-  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$
-
-- **Private donations growth:** `r KPI_private`
-  - A calculation of the percentage change in private contributions based on the two most recent years:
-  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$
-
-- **Public grants growth:** `r KPI_public`
-  - A calculation of the percentage change in public grants received based on the two most recent years:
-  $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$
-
-- **Private/public proportion:** `r KPI_proportion`
-  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
-  $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
-
-`r KPI_warning`
-
-<!--# Latex command to force new page - ensures the plot heading doesn't move up to the first page-->
-
-\newpage
-
-# Graphs
-
-<!--# Warning header, if necessary -->
-
-`r general_warning`
-
-<!--# specific alerts -->
-
-`r excel_warning`
-
-`r excel_zero_revenue_warning`
-
-`r excel_zero_expenses_warning`
-
-`r arsresultat_plot_warning`
-
-`r balance_plot_warning`
-
-`r CEO_plot_warning`
-
-<!--# Creating plots -->
-
-<!-- But of course a bit of raw LaTeX is necessary to print a left quotation mark -->
-
-```{r arsresultat_plot, fig.cap="Based on the entry \`Årsresultat' in the Proff.no data."}
-## Annual result--------------------------------------------------------------
-arsresultat_plot <-
-  # testing for the presence of data
-  if (sum(!is.na(htmldata$arsresultat)) == 0) {
-    NA
-  } else {
-    htmldata |>
-      # renaming for plotly hover
+```{r}
+kreftforeningen |>
       rename(
@@
       geom_col() +
-      # using currency scale function to add correct currency code or ?
-      currency_scale +
-      # using a function to ensure uniform colours irrespective of data
-      scale_fill_green(htmldata$annual_profit) +
+      scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+      scale_fill_green(kreftforeningen$annual_profit) +
       labs(
-        x = "", y = "", caption = currency_caption, title = "Net income (in thousands)"
+        x = "", y = "", title = "Net income (in thousands)"
       ) +
@@
       theme(legend.position = "none")
-  }
-# Printing plot if the data are available
-if (sum(!is.na(htmldata$arsresultat)) != 0) {
-  arsresultat_plot
-}
 ```
-
-```{r balance_plot, fig.cap="Based on the entry \`Sum egenkapital og gjeld' in the Proff.no data."}
-## Balance--------------------------------------------------------------------
-balance_plot <-
-  # testing for the presence of data
-  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
-    NA
-  } else {
-    htmldata |>
-      # rename for plotly hover
-      rename(
-        Year = year,
-        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
-        `Positive equity` = balance_positive
-      ) |>
-      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
-      geom_col() +
-      # using currency scale function to add correct currency code or ?
-      currency_scale +
-      # using a function to ensure uniform colours irrespective of data
-      scale_fill_green(htmldata$balance_positive) +
-      labs(
-        x = "", y = "",
-        caption = currency_caption, title = "Total equity/liabilities (in thousands)"
-      ) +
-      theme_light() +
-      theme(legend.position = "none")
-  }
-
-# Printing the plot if the data are available
-if (is.ggplot(balance_plot)) {
-  balance_plot
-}
-```
-
-```{r revenue_expenses_plot, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
-## Revenue--------------------------------------------------------------------
-revenue_plot <-
-  # testing for data quality
-  if (datatest == TRUE || excel_zero_check("Revenue") == 0) {
-    NA
-  } else {
-    exceldata |>
-      # Renaming for the purposes of the plot + dropping other variables
-      transmute(
-        Year = as.factor(year),
-        `Private contributions` = revenue_private_contributions,
-        `Membership fees` = revenue_membership_fees,
-        `Public grants` = revenue_public_grants,
-        Investments = revenue_investments,
-        `Operational activities` = revenue_operational_activities,
-        `Commercial activities` = revenue_commercial_activities,
-        Other = revenue_other
-      ) |>
-      # Pivoting to use fill
-      pivot_longer(
-        c(`Private contributions`:Other),
-        names_to = "Revenue source"
-      ) |>
-      # Ordering for the plot
-      mutate(
-        `Revenue source` =
-          ordered(`Revenue source`,
-            levels =
-              c(
-                "Other", "Commercial activities",
-                "Operational activities", "Investments",
-                "Membership fees", "Private contributions",
-                "Public grants"
-              )
-          )
-      ) |>
-      ggplot(aes(Year, value, fill = `Revenue source`)) +
-      geom_col() +
-      scale_y_continuous(
-        # Excel data are assumed to be in NOK; if this changes, code based on
-        # the currency_code() function may be adapted
-        labels = label_dollar(prefix = "NOK "),
-        # Basing lower limit on the data (not suppressing erroneous negative
-        # numbers), and the upper limit on the maximum value in revenues and
-        # expenses, to match the two plots
-        limits = c(NA, maximum)
-      ) +
-      # Specifying the number of values, as the base palette only includes 6
-      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
-      labs(x = "", y = "", title = "Revenue (in thousands)") +
-      theme_light() +
-      theme(title = element_text(size = 14))
-  }
-
-# Printing the plot if the data are OK
-if (is.ggplot(revenue_plot)) {
-  revenue_plot
-}
-
-## Expenses -------------------------------------------------------------------
-expenses_plot <-
-  # testing for data quality
-  if (datatest == TRUE || excel_zero_check("Expenses") == 0) {
-    NA
-  } else {
-    exceldata |>
-      # Renaming for the purposes of the plot + dropping other variables
-      transmute(
-        Year = as.factor(year),
-        Other = expenses_other,
-        Administrative = expenses_administrative,
-        Fundraising = expenses_fundraising,
-        `Program services` = expenses_program_services
-      ) |>
-      # Pivoting to use fill
-      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
-      # Ordering for the plot
-      mutate(
-        Expenses = ordered(Expenses,
-          levels =
-            c("Program services", "Other", "Fundraising", "Administrative")
-        )
-      ) |>
-      ggplot(aes(Year, value, fill = Expenses)) +
-      geom_col() +
-      scale_y_continuous(
-        # Excel data are assumed to be in NOK; if this changes, code based on
-        # the currency_code() function may be adapted
-        labels = label_dollar(prefix = "NOK "),
-        # Basing lower limit on the data (not suppressing erroneous negative
-        # numbers), and the upper limit on the maximum value in revenues and
-        # expenses, to match the two plots
-        limits = c(NA, maximum)
-      ) +
-      scale_fill_met_d("Hokusai2") +
-      labs(x = "", y = "", title = "Expenses (in thousands)") +
-      theme_light() +
-      theme(title = element_text(size = 14))
-  }
-
-# Printing the plot if the data are OK
-if (is.ggplot(expenses_plot)) {
-  expenses_plot
-}
-```
-
-```{r compensation_plot, out.width="100%", fig.cap="Based on the category \`Lederlønn' in the Proff.no data.
-Not in thousands.
-Includes the entries \`Lønn' and \`Leder annen godtgjørelse'.
-A common measure of transparency, but should normally be evaluated in context."}
-## CEO compensation-----------------------------------------------------------
-CEO_plot <-
-  # Testing for the presence of data
-  if (sum(!is.na(htmldata$lederlonn)) == 0 &&
-    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
-    # Returning NA if no data
-    NA
-  } else {
-    # Creating plot if any data exist
-    # Note that the plot will still look empty, but have a legend and scales,
-    # if all values are explicit zeroes
-    htmldata |>
-      # preparing for pivoting
-      select(
-        year, lederlonn, leder_annen_godtgjorelse
-      ) |>
-      # Compensation in ordinary numbers
-      mutate(
-        lederlonn = lederlonn * 1000,
-        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-      ) |>
-      # renaming for the purposes of the plot
-      rename(
-        Year = year,
-        Salary = lederlonn,
-        `Other compensation` = leder_annen_godtgjorelse
-      ) |>
-      # pivoting to join the two values to use fill
-      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-      ggplot(aes(Year, value, fill = Compensation)) +
-      geom_col() +
-      # using currency scale function to add correct currency code or ?
-      currency_scale +
-      # pretty colours
-      scale_fill_met_d("Hokusai2", direction = -1) +
-      labs(
-        x = "", y = "", title = "CEO compensation"
-      ) +
-      theme_light() +
-      theme(legend.position = "bottom")
-  }
-
-# Printing the plot if the data are available
-if (is.ggplot(CEO_plot)) {
-  CEO_plot
-}
-```

### 2022-08-02 Report/Graphs for final report.Rmd
Adding a plot.

@@
 ```
+
+```{r}
+kreftforeningen |>
+      rename(
+        Year = year,
+        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
+        `Positive equity` = balance_positive
+      ) |>
+      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
+      geom_col() +
+      # using currency scale function to add correct currency code or ?
+      scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+      # using a function to ensure uniform colours irrespective of data
+      scale_fill_green(kreftforeningen$balance_positive) +
+      labs(
+        x = "", y = "", title = "Total equity/liabilities (in thousands)"
+      ) +
+      theme_light() +
+      theme(legend.position = "none")
+```

### 2022-08-02 Report/Group 5 report.Rmd
Adding a plot.

@@
 
-```{r}
-
+```{r data_and_functions, include=FALSE}
+# Data reading =================================================================
+
+## Function to read html file --------------------------------------------------
+read_proff <- function(file, start, end) {
+  # Specifying encoding to deal with occasional errors
+  read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame") |>
+    html_table() |>
+    # removing empty column, using all_of(argument)
+    select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column
+    fill(
+      valutakode,
+      .direction = "updown"
+    ) |>
+    # removing dates and redundant currency
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+}
+
+## Creating df------------------------------------------------------------------
+kreftforeningen <- read_proff("Data/Kreftforeningen.html", "2002", "2021")
+
+## scale_fill_green ------------------------------------------------------------
+# Function with slightly nicer colours than default; ensuring uniformity; +
+# ensuring positive values are always green through a test for the presence of
+# FALSE values
+scale_fill_green <- function(variable) {
+  scale_fill_met_d("Tsimshian",
+    direction = ifelse(
+      mean(variable,
+        na.rm = TRUE
+      ) == 1, -1, 1
+    )
+  )
+}
 ```
@@
 <!--Legg inn bilder her-->
+
+<!-- # Legger inn to grafer som viser manglende data i Proffs årsregnskap-->
+
+```{r arsresultat, echo=FALSE, out.width="70%", fig.align='center', fig.cap="Net income/loss data for the Norwegian Cancer Society from Proff.no"}
+kreftforeningen |>
+  rename(
+    Year = year,
+    `Net income` = arsresultat,
+    `Net positive` = annual_profit
+  ) |>
+  ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+  scale_fill_green(kreftforeningen$annual_profit) +
+  labs(
+    x = "", y = "", title = "Net income (in thousands)"
+  ) +
+  theme_light() +
+  theme(legend.position = "none")
+```
 

### 2022-08-02 Presentasjon/presentasjon.Rmd
Sletter og renamer bildefiler, adder css fil

--- 2022-08-01 Presentasjon/presentasjon.Rmd
@@
 author: "Gruppe 5"
-date: "`r Sys.Date()`"
-output: ioslides_presentation
+date: "4.8.2022"
+output: 
+  ioslides_presentation: 
+    css: stylesss.css
 ---
@@
 
-## {data-background="Images/start1.png" data-background-size=cover}
-
 ## <br> OPX and their goal
-- A startup developing solutions for socially responsible investment strategies
-- The goal is to increase cash flow to charitable purposes 
-- Donors' trust in NGOs' financial performance and their ability to deliver results is crucial for their choice of beneficiaries <br>
+- Develop solutions for socially responsible investment strategies
+- Increase cash flow to charitable purposes <br>
 
@@
 
+----
+```{r, out.height="360", out.width="612", fig.align='center'}
+knitr::include_graphics("Images/fcuk.png")
+
+```
+
 # Data collection
@@
 # Limitations and steps forward
+
+## <iframe width="1280" height="720" src="Images/henrik.mp4" align="middle" frameborder="0" allowfullscreen></iframe>
+
+## {data-background="Images/tenke.png" data-background-size=cover}
+
+## <br> Our experience
+
+- Unsolvable case due to non-existent data
+- Hard to create credible KPIs due to limited knowledge of economics and the NGO sector
+- **Insight into real-life work challenges - the clash between is desired and what is possible**

### 2022-08-02 Report/Graphs for final report.Rmd
Adding graphs.

@@
 kreftforeningen <- read_proff("Data/Kreftforeningen.html", "2002", "2021")
+
+care <- read_proff("Data/CARENorge2010_2021.html", "2010", "2021")
 ```
@@
 ```
+
+```{r}
+care |>
+  rename(
+    Year = year,
+    `Total equity/liabilities` = sum_egenkapital_og_gjeld,
+    `Positive equity` = balance_positive
+  ) |>
+  ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
+  geom_col() +
+  # using currency scale function to add correct currency code or ?
+  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+  # using a function to ensure uniform colours irrespective of data
+  scale_fill_green(kreftforeningen$balance_positive) +
+  labs(
+    x = "", y = "", title = "Total equity/liabilities (in thousands)"
+  ) +
+  theme_light() +
+  theme(legend.position = "none")
+```
+
+```{r}
+kreftforeningen |>
+  # preparing for pivoting
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |>
+  # Compensation in ordinary numbers
+  mutate(
+    lederlonn = lederlonn * 1000,
+    leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
+  ) |>
+  # renaming for the purposes of the plot
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |>
+  # pivoting to join the two values to use fill
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  # using currency scale function to add correct currency code or ?
+  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+  # pretty colours
+  scale_fill_met_d("Hokusai2", direction = -1) +
+  labs(
+    x = "", y = "", title = "CEO compensation"
+  ) +
+  theme_light() +
+  theme(legend.position = "bottom")
+```

### 2022-08-02 Report/Group 5 report.Rmd
Adding graphs.

@@
 
-<!-- # Legger inn to grafer som viser manglende data i Proffs årsregnskap-->
-
-```{r arsresultat, echo=FALSE, out.width="70%", fig.align='center', fig.cap="Net income/loss data for the Norwegian Cancer Society from Proff.no"}
+<!-- # Legger inn tre grafer som viser manglende data i Proffs årsregnskap-->
+
+```{r arsresultat, echo=FALSE, out.width="70%", fig.align='center', fig.asp=0.5, fig.cap="Net income/loss data for the Norwegian Cancer Society from Proff.no"}
 kreftforeningen |>
@@
   labs(
-    x = "", y = "", title = "Net income (in thousands)"
+    x = "", y = ""
   ) +
@@
   theme(legend.position = "none")
+```
+
+```{r egenkapital, echo=FALSE, out.width="70%", fig.align='center', fig.asp=0.5, fig.cap="Total equity/liabilities data for the Norwegian Cancer Society from Proff.no"}
+kreftforeningen |>
+  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+  scale_fill_green(kreftforeningen$balance_positive) +
+  labs(
+    x = "", y = "", title = "Total equity/liabilities (in thousands)"
+  ) +
+  theme_light() +
+  theme(legend.position = "none")
+```
+
+```{r CEO, echo=FALSE, out.width="70%", fig.align='center', fig.asp=0.618, fig.cap="CEO compensation data for the Norwegian Cancer Society from Proff.no"}
+kreftforeningen |>
+  # preparing for pivoting
+  select(
+    year, lederlonn, leder_annen_godtgjorelse
+  ) |>
+  # Compensation in ordinary numbers
+  mutate(
+    lederlonn = lederlonn * 1000,
+    leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
+  ) |>
+  # renaming for the purposes of the plot
+  rename(
+    Year = year,
+    Salary = lederlonn,
+    `Other compensation` = leder_annen_godtgjorelse
+  ) |>
+  # pivoting to join the two values to use fill
+  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
+  ggplot(aes(Year, value, fill = Compensation)) +
+  geom_col() +
+  # using currency scale function to add correct currency code or ?
+  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+  # pretty colours
+  scale_fill_met_d("Hokusai2", direction = -1) +
+  labs(
+    x = "", y = "", title = "CEO compensation"
+  ) +
+  theme_light() +
+  theme(legend.position = "bottom")
 ```

### 2022-08-02 Report/Group 5 report.Rmd
Adding newpage before Appendix

@@
 
+\newpage
 # Appendix: Data reading and cleaning code

### 2022-08-02 Report/Group 5 report.Rmd
Changing date.

@@
 author: "ISSSV1337 Group 5"
-date: "`r Sys.Date()`"
+date: "2022-08-03"
 output:

### 2022-08-02 Report/Group 5 report.Rmd
Note to include captions.

@@
 
-```{r, echo=FALSE, out.width="70%"}
+```{r, echo=FALSE, out.width="70%", fig.cap="Need caption"}
 knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene
@@
 
-```{r, echo=FALSE, out.width="70%"}
+```{r, echo=FALSE, out.width="70%", fig.cap="Need caption"}
 knitr::include_graphics("Images/info_kreft.png")
@@
 
-```{r, echo=FALSE, out.width="100%"}
+```{r, echo=FALSE, out.width="100%", fig.cap="Need caption"}
 knitr::include_graphics("Images/regnskapssti.png")

### 2022-08-02 Report/Group 5 report.Rmd
Setting out.width="70%" as default.

@@
 ```{r Setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE, out.width = "100%")
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE, out.width = "70%")
 ```
@@
 
-```{r, echo=FALSE, out.width="70%", fig.cap="Need caption"}
+```{r, echo=FALSE, fig.cap="Need caption"}
 knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene
@@
 
-```{r, echo=FALSE, out.width="70%", fig.cap="Need caption"}
+```{r, echo=FALSE, fig.cap="Need caption"}
 knitr::include_graphics("Images/info_kreft.png")
@@
 
-```{r arsresultat, echo=FALSE, out.width="70%", fig.align='center', fig.asp=0.5, fig.cap="Net income/loss data for the Norwegian Cancer Society from Proff.no"}
+```{r arsresultat, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Net income/loss data for the Norwegian Cancer Society from Proff.no"}
 kreftforeningen |>
@@
 
-```{r egenkapital, echo=FALSE, out.width="70%", fig.align='center', fig.asp=0.5, fig.cap="Total equity/liabilities data for the Norwegian Cancer Society from Proff.no"}
+```{r egenkapital, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Total equity/liabilities data for the Norwegian Cancer Society from Proff.no"}
 kreftforeningen |>
@@
 
-```{r CEO, echo=FALSE, out.width="70%", fig.align='center', fig.asp=0.618, fig.cap="CEO compensation data for the Norwegian Cancer Society from Proff.no"}
+```{r CEO, echo=FALSE, fig.align='center', fig.asp=0.618, fig.cap="CEO compensation data for the Norwegian Cancer Society from Proff.no"}
 kreftforeningen |>

### 2022-08-02 Report/Group 5 report.Rmd
Proofreading and editing.

@@
 brainstorming and data collection processes are elaborated, before data
-description and findings are presented. Lastly, limitations and steps
-forward, as well as a summary of our experience.
+description and findings are presented. Lastly, we discuss limitations and steps
+forward, and provide a brief summary of our experience.
 
@@
 a startup working to develop solutions to assist individuals and 
-organizations wanting to donate to charities and designing a socially 
+organisations wanting to donate to charities and designing a socially 
 responsible investment strategy. Their main goal is to increase cash 
-flow from private to charitable purposes. Therefore, the donor's trust 
-in the charities’ abilities to carry out their promises is significant.
+flow from private groups and individuals to charitable purposes. Therefore, the donor's trust 
+in the charities’ abilities to carry out their promises is of paramount significance.
 
 To do this, data from charities need to be collected. OPX’s current 
-challenge is the data collection, which must be collected manually today
-since the data is unstructured. Therefore, they wanted a way to automate 
-the data collection, using web scraping.
-
-Additionally, by creating KPIs (Key Performance Indicator), OPX can use 
+challenge is data acquisition, as data currently must be collected manually. They therefore wanted a way to automate data collection, for example using web scraping.
+
+ <!-- Needs work -->
+Additionally, by creating Key Performance Indicators based on these data, OPX can use 
 them to benchmark the relationship between financial assets and emergency 
@@
 1. Automating the data collection will streamline the work of OPX 
-and create more transparency which enhances trust in charities.
+and create more transparency, which enhances trust in charities.
 
 2. Creating KPI(s) is important for the organization’s performance 
-management, to influence the objective, and to foster growth.
+management, to influence its objectives, and to foster growth.
 
@@
 
-Our brainstorming process for this case was characterized by several
+Our brainstorming process for this case was characterised by several
 surprises and turns in the direction of the solution. As we dug deeper
-into the solution opportunities, we were also met by new obstacles that
+into the solution opportunities we were also met by new obstacles that
 forced us to change course. In particular, there were four turning
-points. Therefore, we have decided to divide the brainstorming into
+points. Therefore, we have decided to divide the brainstorming section into
 three phases.
 
-## Phase 1: How to harvest and analyze the data? {.smaller}
-
-OPX wanted us to harvest the organizations' financial data from the web.
+## Phase 1: How to harvest and analyse the data? {.smaller}
+
+OPX wanted us to harvest the organisations' financial data from the web.
 During the first stage of brainstorming, we wanted to find a simple and
@@
 different websites, but quickly decided that this would be too
-difficult, and that it would only work with selected organizations. The
-suggested solution was to create a code that could read PDF documents
-that OPX downloaded themselves. Our code would analyze and organize the
+difficult, and that it would only work with selected organizations. The next
+suggested solution was to create code that could read PDF documents
+that OPX downloaded themselves. Our code would analyse and organise the
 uploaded data from the PDF documents.
 
-However, different organizations publish financial data in different
-ways in their yearly reports -- some financial statements were included
+However, different organisations publish financial data in different
+ways in their annual reports -- some financial statements were included
 as tables, some as pictures, etc. This led us to phase 2.
@@
 
-At this stage, we wanted to explore alternatives to analyzing PDF
+At this stage, we wanted to explore alternatives to analysing PDF
 documents. We discussed how to do so, and concluded that we had three
@@
 were feasible or not. The first option of using PDF documents still had
-the issue of financial statements not always being published as text.
-Also, we quickly found that Proff.no disallowed webscraping on their
-website. The Brønnøysund registry does offer API access to all the financial statements, but we discovered that access to costs NOK 480,000.
-
-We started to look at workarounds for this problem. It would still be
+the issue of financial statements not always being published as text, as well as their often wildly differing formats.
+Unfortunately, we quickly found that Proff.no disallows any automated collection of data from their website.
+The Brønnøysund registry does offer API access to all the financial statements, but we discovered that access currently costs NOK 480,000.
+
+We started to look at workarounds for this problem. It could still be
 possible to download Proff.no HTML websites manually for each
-organization. We did receive some XML files from Statistics Norway, but using these is ultimately not a viable solution for OPX.
+organisation. We did receive some XML files from Statistics Norway, but using these is ultimately not a viable solution for OPX.
 
@@
 attempted to do so in Excel at first, so that we later could code this
-into R. While doing so, we realized that it would be practically
+into R. While doing so, we realised that it would be practically
 impossible to extract the correct sums for the OPX accounting items from
@@
 
-When also discovering that not all organizations publish their financial
+When also discovering that not all organisations publish their financial
 statements, and certainly not for every year, we decided to give up on
-collecting financial data for OPX. This is when we moved on to the third
+collecting financial data for OPX, as the task turned out to impossible given current reporting requirements and practices. This is when we moved on to the third
 and final stage.
@@
 In order to create a somewhat more challenging task for ourselves, we
-also decided to create a shiny app that OPX could use, which would
+also decided to create an interactive dashboard based on Shiny that OPX could use, which would
 display the KPIs as well as financial trends for the selected
-organization. To use this, OPX needs to gather the financial data
-themselves, then upload this into the shiny app as a x document
+organisation. To use this, OPX needs to gather the financial data
+themselves, as well as download data manually from Proff.no, and then upload this into the dashboard in the form of Excel and HTML documents.
 
@@
 potential solutions. We experimented on coding such that data could have
-been imported from XML files. However, once we realized that the
-financial data was of little ¨use, this experimentation stopped.
+been imported from XML files. However, once we realised that the
+financial data was of little use, this experimentation stopped.
 

### 2022-08-02 Dashboard/NGO Analysis dashboard v006.Rmd
Cleaning up a comment.

@@
     ggplotly(arsresultat_plot()) |>
-      ## ggplot subtitles and captions are lost when using ggplotly(), a
-      ## currency warning caption is therefore added in plotly, if needed, using
-      ## a separate function
+      # ggplot subtitles and captions are lost when using ggplotly(), a
+      # currency warning caption is therefore added in plotly, if needed, using
+      # a separate function
       layout(

### 2022-08-02 Presentasjon/presentasjon.Rmd
La til kommentar om slide.

@@
 
+<!-- Hvis du kan legge inn en slide med webadressen til appen *etter* presentasjonen min hadde det vært fint: https://torbjornso.shinyapps.io/NGOanalysis/ (veldig begrenset med prosessortid på gratisversjonen, men ta en titt hvis du vil.
+Tenkte "førstemann til mølla" med den gjenstående prosessortiden etter presentasjonen-->
+
 # Limitations and steps forward

### 2022-08-02 Report/Group 5 report.Rmd
subfig, tidying.

@@
     highlight: pygments
+    extra_dependencies: "subfig"
   html_document: default
@@
 
-```{r, echo=FALSE, fig.cap="Need caption"}
+```{r, echo=FALSE, fig.cap="Need caption", fig.align='center'}
 knitr::include_graphics("Images/info_kreft.png")
@@
 
-```{r arsresultat, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Net income/loss data for the Norwegian Cancer Society from Proff.no"}
+```{r arsresultat, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Net income/loss (in NOK 1,000) data for the Norwegian Cancer Society from Proff.no"}
 kreftforeningen |>
@@
 
-```{r egenkapital, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Total equity/liabilities data for the Norwegian Cancer Society from Proff.no"}
+```{r egenkapital, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Total equity/liabilities (in NOK 1,000) data for the Norwegian Cancer Society from Proff.no"}
 kreftforeningen |>
@@
   labs(
-    x = "", y = "", title = "Total equity/liabilities (in thousands)"
+    x = "", y = ""
   ) +
@@
   labs(
-    x = "", y = "", title = "CEO compensation"
+    x = "", y = ""
   ) +
@@
 
-This KPI may say something about the NGO's sources of funding. There are prima facie no specific desirable funding mix, but this measure may be of interest to some donors.
+This KPI may say something about the NGO's sources of funding. There is *prima facie* no specific desirable funding mix, but this measure may be of interest to some donors.
 
@@
 
-```{r, out.width="50%", fig.cap="The main dashboard page", fig.align='center', echo=FALSE}
+```{r dashboard_images, out.width="50%", fig.cap="The dashboard.", fig.subcap=c("The main page.", "The details tab."), echo=FALSE}
 knitr::include_graphics("Images/Dashboard_page_1.png")
+knitr::include_graphics("Images/Dashboard_page_2.png")
 ```
@@
 The value box then also relies on a function that converts the number into a percentage string, checking in the process if the number is `NA` or `Inf` (indicating a division by zero, and thus either a data problem or data not suited to this calculation).
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_KPI_string, eval=FALSE}
@@
 The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_valuebox, eval=FALSE}
@@
 All boxes use the theme colour 'primary' and the icon 'fa-solid fa-circle-xmark' if the value is NA.
-```{r dashboard_fontawesome, eval=FALSE}
-# Using fontawesome for more up-to-date (and currently supported) fontawesome
-# icons
-htmltools::tagList(fontawesome::fa_html_dependency())
-```
 
@@
 The html-based plots include a basic test of whether the necessary data are available, as some or all may sometimes be missing, printing a 'no information' plot using a separate function if the data is missing.
+
+`r if (knitr::is_latex_output()) line`
 ```{r html_datatest, eval=FALSE}
@@
 The 'no information' plot is an essentially empty plot displaying a notice to the user, and is created through a function that passes label information into a more general blank plot function:
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_no_information, eval=FALSE}
@@
 ```{r}
-## scale_fill_green ------------------------------------------------------------
 # Function with slightly nicer colours than default; ensuring uniformity; +
@@
   ggplotly(arsresultat_plot()) |>
-    ## ggplot subtitles and captions are lost when using ggplotly(), a
-    ## currency warning caption is therefore added in plotly, if needed, using
-    ## a separate function
+    # ggplot subtitles and captions are lost when using ggplotly(), a
+    # currency warning caption is therefore added in plotly, if needed, using
+    # a separate function
     layout(
@@
 These both use the same check for data quality as the value box, outputting a 'bad data' plot if the data are bad using a function that passes label information to the `empty_plot()` function:
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_excel_datatest, eval=FALSE}
-## bad_data plot ---------------------------------------------------------------
 # Function to print a 'bad data' plot, if necessary
@@
 ```{r dashboard_zero_check, eval=FALSE}
-## Excel zero test------------------------------------------------------------
 excel_zero_check <- function(type) {
@@
 ```{r}
-## zero_data_plot --------------------------------------------------------------
 zero_data <- function() {
@@
 Excel data are presumed to be supplied in NOK.
+
+`r if (knitr::is_latex_output()) line`
 ```{r dashboard_excelplot, eval=FALSE}
@@
 ```{r dashboard_maximum, eval=FALSE}
-## scale maximum--------------------------------------------------------------
 # Reactive function to create the maximum value for revenue and expenses plots
@@
 ### Details tab
-```{r dashboard_page_2, out.width="50%", fig.cap="The details tab of the dashboard.", fig.align='center', echo=FALSE}
-knitr::include_graphics("Images/Dashboard_page_2.png")
-```
 
@@
 ```{r dashboard_title_creation, eval=FALSE}
-# NGO title ==================================================================
 NGO <- reactive({
@@
 
-```{r dashboard_report_image, echo=FALSE, out.width="50%", fig.align='center'}
-knitr::include_graphics("Images/Dashboard_report.png")
+```{r dashboard_report_image, echo=FALSE, out.width="50%", fig.align='center', fig.cap="Sample dashboard report."}
+knitr::include_graphics("Images/Dashboard_report_cropped.png")
 ```

### 2022-08-02 Report/Group 5 report.Rmd
Adding dashboard report code. Editing throughout.

@@
     extra_dependencies: "subfig"
-  html_document: default
 ---
@@
 
-We discussed creating a webcrawler that could collect data across
+<!--# En webcrawler er en bot -->
+
+We discussed creating web scraping code that could collect data across
 different websites, but quickly decided that this would be too
-difficult, and that it would only work with selected organizations. The next
+difficult, and that it would only work with selected organisations. The next
 suggested solution was to create code that could read PDF documents
@@
 - Financial statements in annual report PDF documents. 
-- Webscrape Proff.no for financial statements. 
-- Webscrape the Brønnøysund registry.
+- Web scrape Proff.no for financial statements. 
+- Web scrape the Brønnøysund registry.
 
@@
 
-The OPX case requested two tasks -- webscrape financial data, and create
+The OPX case requested two tasks -- web scrape financial data, and create
 KPIs. During phase three, we scrapped all ideas and progress from the
@@
 
-We wanted to use webscraping to collect the data, so that we could have
+We wanted to use web scraping to collect the data, so that we could have
 automated the data input process. As described, we explored different
 potential solutions. We experimented on coding such that data could have
-been imported from XML files. However, once we realised that the
+been imported from XML files. However, once we realised that the available
 financial data was of little use, this experimentation stopped.
@@
 We copied these financial statements manually into an Excel document.
-Perhaps a "simple" solution, however our code requires OPX to import
+Perhaps a 'simple' solution, however our code requires OPX to import
 both an Excel document themselves into shiny, as well as a HTML document
@@
 # Data description
+
+<!--# Fra Visualisations (nedenfor): The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.-->
 
@@
 
-As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with messages about missing packages:
+As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with (cryptic) messages about missing functions:
 
@@
 
-The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data, and an 'Analyse' button which is connected to two Shiny `observe() |> bindEvent()` structures, ensuring that most of the code is not executed until the user has finished entering data.
+The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data (needed because of the highly variable availability of historical data), and an 'Analyse' button which is connected to two Shiny `observe() |> bindEvent()` structures, ensuring that most of the code is not executed until the user has finished entering data.
 (Though both structures await the same button click, the division was necessary to remain within R's 10,000 byte limit.)
@@
 ```{r dashboard_no_information, eval=FALSE}
-## no_information plot ---------------------------------------------------------
 # Function to print a 'no information' plot, if necessary
@@
 
-## empty_plot-------------------------------------------------------------------
 # Function to print an empty plot with a supplied message
@@
     This is unusual and may
-    indicate a data problem.")
-}
-```
-
-`r if (knitr::is_latex_output()) line`
-
-Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting - in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order.
+    indicate a data problem."
+  )
+}
+```
+
+`r if (knitr::is_latex_output()) line`
+
+Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting -- in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order.
 Excel data are presumed to be supplied in NOK.
@@
 
-All the plots based on the Excel data make use of the colour-blind-friendly "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories.
-The same function is used to generate the text colour in the 'no information' and 'bad data' plots.
+All the plots based on the Excel data make use of the colour-blind-friendly 'Hokusai2' palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories.
+The same function is used to generate the text colour in the 'no information' and 'bad data' plots in order to present the user with a somewhat unified plot aesthetic.
 
@@
 
+`r if (knitr::is_latex_output()) line`
+```{r dashboard_download, eval=FALSE}
+output$report <- downloadHandler(
+
+  # creating the file name using the NGO title and system date
+  filename = function() {
+    str_c(title(), "_", Sys.Date(), ".pdf")
+  },
+  content = function(file) {
+    # Based on code in Mastering Shiny -- creating a temporary directory and
+    # file
+    tempReport <- file.path(tempdir(), "report.Rmd")
+    file.copy("report.Rmd", tempReport, overwrite = TRUE)
+
+    # parameters passed to the R Markdown document
+    params <- list(
+      htmlfile = input$html$datapath,
+      excelfile = input$excel$datapath,
+      start = input$start,
+      end = input$end
+    )
+
+    # notification -- disappears when report generation is complete
+    id <- showNotification(
+      "Generating report...",
+      duration = NULL,
+      closeButton = FALSE
+    )
+    on.exit(removeNotification(id), add = TRUE)
+
+    # R Markdown render
+    rmarkdown::render(tempReport,
+      output_file = file,
+      params = params,
+      envir = new.env(parent = globalenv())
+    )
+  }
+)
+```
+
+`r if (knitr::is_latex_output()) line`
+
 In order to prevent the user from the downloading a nonsensical html file if bad data have been uploaded or entered, the button creation and report generation are located in the second `observe() |> bindEvent()` structure.
 The creation of the download button is connected to two convenience logical tests, the primary purpose of which is to connect the button to the `validate()` calls upstream in the reactive chain.
+
+`r if (knitr::is_latex_output()) line`
+```{r dashboard_download_button, eval=FALSE}
+## Download button -----------------------------------------------------------
+output$downloadUI <- renderUI({
+  if (available_excel() && available_html()) {
+    downloadButton("report", "Download report")
+  }
+})
+
+## Data exists condition testing----------------------------------------------
+# Essentially pretty random tests to connect to the individual
+# validate() functions upstream in the reactive chain
+
+# Excel
+available_excel <- reactive({
+  TRUE
+}) |>
+  bindEvent(setequal(
+    names(exceldata()),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  ))
+
+# HTML
+available_html <- reactive({
+  TRUE
+}) |> bindEvent("lederlonn" %in% names(htmldata()))
+```
+
+`r if (knitr::is_latex_output()) line`
 
@@
 The limitation with the accessible data is first and foremost the lack 
-of relevant data in the annual reports from the NGOs.  Since there is 
+of relevant data in the annual reports from the NGOs. Since there is often 
 no obligation to provide annual accounts, they often have more information 
-about the organization in general rather than numbers. If they do provide 
+about the organisation in general rather than numbers. If they do provide 
 the annual accounts, they are not standardized. This means that different 
@@
 do appear in the same format, the accounts we want to retrieve are often 
-in an image format rather than text. The lack of data, standardization and 
-diverse formats made it impossible to create an automated solution for 
+in an image format rather than text. The lack of data, standardisation and 
+diverse formats made it impossible to create an automated solution for (safely)
 collecting data based on the NGOs’ annual reports.
@@
 calculate the different KPIs do not exist in the annual reports.  We 
-experimented with an annual report from Brønnøysundregisteret. Although 
-the reports were standardized, the different posts in the report did not 
+experimented with an annual report from Brønnøysund. Although 
+the reports were standardised, the different posts in the report did not 
 match the template from OPX. Additionally the data that OPX is interested 
@@
 that would be of relevance and not. After examining the different 
-alternatives, we concluded that the NGOs could report in a standardized 
+alternatives, we concluded that the NGOs could report in a standardised 
 excel document from which our code retrieves the data to calculate the KPIs.
@@
 to be implemented, it would not be of any relevance to OPX. As a result, 
-due to OPX’s need for this data to be available, we recommend discussions 
+due to OPX’s need for this data to be available, we weould suggest discussions 
 with relevant politicians and ministry officials as a starting point.
 
-Based on the findings of working with this case, we would recommend looking 
-at best practices for further development. One good example could be charity 
-navigator.
+Based on the findings of working with this case, we would recommend looking at best practices for further development. One well-known established actor in the field is Charity Navigator.
 
@@
 In this last section, we would like to describe how we experienced this process and what we have learned.
-As explained, the main challenge with the task was highlighted by an unsolvable due to non-existent data.
+As explained, the main challenge with the task was thatb it was in many ways unsolvable due to non-existent data.
 In cases where data were available, they were inadequate.
@@
 The lectures allowed us to learn about many interesting topics.
-However, as the connection between what we learned there and what we got to utilize in the case was distant, and parts were too advanced for some of us, especially those without any particular experience within statistics.
+However, as the connection between what we learned there and what we got to utilise in the case was distant, and parts were too advanced for some of us, especially those without any particular experience within statistics.
 
@@
 We also got an insight into real-life work situations, such as the challenges that come when what is desired clashes with what is possible.
-We are certain that these experiences will be useful for all of us in the future, and we are thankful for being given this opportunity.
+We are certain that these experiences will be useful for all of us in the future, and we are grateful for being given this opportunity.
 

### 2022-08-02 Dashboard/NGO Analysis dashboard v006.Rmd
Added source code note. Commenting.

@@
 htmltools::tagList(fontawesome::fa_html_dependency())
+
+# source_code: embed OR Github link
 ```
@@
 
-      # parameters
+      # parameters passed to the R Markdown document
       params <- list(

### 2022-08-02 Report/Group 5 report.Rmd
Adding info on library calls in shinyapps.io version.

@@
 
-As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with (cryptic) messages about missing functions:
+As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with (cryptic) messages about missing functions when running the dashboard locally:^[A test version uploaded to [shinyapps.io](https://www.shinyapps.io) uses normal serial library calls, as the server then automatically installs the necessary packages during deployment.]
 

### 2022-08-02 Report/Old/Draft Dashboard section.Rmd
Cleaning.

--- 2022-08-01 Report/Draft Dashboard section.Rmd
@@
 
-# The Dashboard
+## The Dashboard
 
@@
 
-## The Sidebar
+### The Sidebar
 
@@
 
-Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded.
-(Note that the `accept` argument is not obeyed by the RStudio viewer, but most ordinary browsers enforce the instruction)
+Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded.^[Note that the `accept` argument is not obeyed by the RStudio viewer, but most ordinary browsers enforce the instruction.]
 
@@
 
-If the validation is successful the files are read using the functions developed for this purpose, which conduct further testing, outputting error messages using `validate()`:
-```{r dashboard_file_functions, eval=FALSE, class.source = "fold-hide"}
-# Data reading functions =======================================================
+If the validation is successful the files are read using the functions developed for this purpose, which conduct further testing, outputting error messages using `validate()`.
+Using these functions, the actual creation of data frames takes place in two stages for each file, partially for data testing and partially to create subsidiary data frames used by some functions.^[The data reading code is included as an appendix because of its length.]
 
-## Functions to read html file -------------------------------------------------
-html_test_create <- function(file, start, end) {
-  # Reading file and node
-  htmltest <- read_html(file, encoding = "UTF-8") |>
-    html_node("#inner-frame")
+### The dashboard proper
 
-  ### A series of tests to catch problems before they spit out errors-----------
-  # Checking if node exists
-  if (class(htmltest) == "xml_node") {
-    htmltable <- htmltest |>
-      html_table()
-  } else {
-    validate("Please upload the specified type of html document.")
-  }
-
-  # Basic check if the table is correct
-  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
-    htmlready <- htmltable
-  } else {
-    validate("Please upload the specified type of html document.")
-  }
-
-  # Testing if the correct years have been entered
-  if (as.character(all_of(start)) %in% names(htmlready) &&
-    as.character(all_of(end)) %in% names(htmlready)) {
-    # Creating the first part of the data frame
-    read_proff_one(htmltable, start, end)
-  } else {
-    validate("Please ensure you have entered the correct start and end years.")
-  }
-}
-
-### First html transformation function, invoked by html_test_create() ----------
-# (stopping to allow checks for differing currency codes)
-read_proff_one <- function(object, start, end) {
-  object |>
-    # removing empty column, using all_of(argument)
-    select(
-      REGNSKAPSPERIODE:all_of(start)
-    ) |>
-    # removing white space
-    mutate(
-      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
-    ) |>
-    # removing duplicate table sometimes created by read_html()
-    filter(
-      !row_number() > 176
-    ) |>
-    # removing duplicate rows
-    filter(
-      !grepl("Lukk", REGNSKAPSPERIODE)
-    ) |>
-    # tidying data using pivot_longer() and all_of(argument)
-    pivot_longer(
-      all_of(end):all_of(start),
-      names_to = "year"
-    ) |>
-    mutate(
-      # changing to real NAs
-      value = na_if(value, "-"),
-      # turning years into numbers
-      year = as.integer(year),
-      # creating a currency code column and turning current currency codes into
-      # NAs
-      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
-    ) |>
-    # filling the currency column to avoid having to deal with NAs later
-    fill(
-      valutakode,
-      .direction = "downup"
-    )
-}
-
-### Second html transformation function, creating the data frame ---------------
-read_proff_two <- function(object) {
-  object |>
-    # removing unneeded currency column
-    select(-valutakode) |>
-    # removing dates and redundant currency
-    filter(
-      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
-        REGNSKAPSPERIODE != "Valutakode"
-    ) |>
-    mutate(
-      # removing white space in numbers
-      value = str_replace_all(value, "\\s", ""),
-      # turning numbers into true numbers
-      value = as.numeric(value),
-      # removing values that are years
-      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
-    ) |>
-    # removing any remaining duplicate rows
-    distinct() |>
-    # pivoting to create 'tidy' variables
-    pivot_wider(
-      names_from = REGNSKAPSPERIODE, values_from = value
-    ) |>
-    # tidying names
-    clean_names() |>
-    # removing headings from the table
-    select(
-      -lederlonn_i_hele_1000,
-      -resultatregnskap_i_hele_1000,
-      -balanseregnskap_i_hele_1000
-    ) |>
-    # making it clear what lonn refers to
-    rename(
-      lederlonn = lonn
-    ) |>
-    # Creating new variables for plots (red/green profit/loss)
-    mutate(
-      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
-      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
-    ) |>
-    # arranging by year for practical display
-    arrange(desc(year))
-}
-
-## Functions to read Excel file ------------------------------------------------
-### First Excel transformation - creating df used by NGO title function---------
-widen_excel_test <- function(file) {
-  # Reading Excel file
-  read_excel(file) |>
-    # cleaning names to enable later operations
-    clean_names()
-}
-
-### Second Excel function - creating primary Excel data frame-------------------
-widen_excel <- function(object) {
-  # testing if all columns are present
-  if (setequal(
-    names(object),
-    c("year", "firm", "category", "sub_category", "name", "value")
-  )) {
-    object |>
-      # uniting category and name to deal with multiple 'Other'
-      unite(
-        entry, c(category, name)
-      ) |>
-      # deleting Sub-category and Firm
-      select(-sub_category, -firm) |>
-      # pivoting to create 'tidy' variables
-      pivot_wider(names_from = entry, values_from = value) |>
-      # cleaning names
-      clean_names()
-    # Notification if all columns are not present
-  } else {
-    validate("Please upload data using the specified Excel template.")
-  }
-}
-```
-
-Using these functions, the actual creation of data frames takes place in two stages for each file, partially for data testing and partially to create subsidiary data frames used by some functions:
-```{r dashboard_data_frames, eval=FALSE}
-# Data frames ================================================================
-## Creating data frame from html file-----------------------------------------
-html_currency <- reactive({
-  # validating file
-  validate(need(
-    tools::file_ext(input$html$datapath) == "html",
-    "Please upload an html file."
-  ))
-
-  # Testing and creating data frame
-  html_test_create(
-    input$html$datapath,
-    as.character(input$start), as.character(input$end)
-  )
-})
-
-htmldata <- reactive({
-  read_proff_two(html_currency())
-})
-
-## Validating and creating long version of Excel file, if possible------------
-excel_test <- reactive({
-  # validating file
-  validate(need(
-    tools::file_ext(input$excel$datapath) == "xlsx",
-    "Please upload an Excel file."
-  ))
-
-  widen_excel_test(input$excel$datapath)
-})
-
-## Creating the primary Excel data frame--------------------------------------
-exceldata <- reactive({
-  widen_excel(excel_test())
-})
-```
-
-## The dashboard proper
-
-### Key Performance Indicators
+#### Key Performance Indicators
 
@@
 
-### Plots
+#### Plots
 
@@
 
-## Details tab
+### Details tab
 
@@
 
-## The parameterised R Markdown report
+### The parameterised R Markdown report
 
@@
 
-The file name of the report is created by combining the name of the NGO from the `title()` reactive and the output of `Sys.Date()`, furnishing a hopefully informative file name.
-(Note that this only works when the dashboard is run in an ordinary browser and not in the RStudio viewer.)
+The file name of the report is created by combining the name of the NGO from the `title()` reactive and the output of `Sys.Date()`, furnishing a hopefully informative file name.^[Note that this only works when the dashboard is run in an ordinary browser and not in the RStudio viewer.]
 
-<!--# Parenthesis as footnote? -->
 ```{r}
@@
 Plot descriptions are included as knitr/LaTeX captions.
+
+# Appendix: Data reading and cleaning code
+```{r dashboard_file_functions, eval=FALSE}
+# Data reading functions =======================================================
+
+## Functions to read html file -------------------------------------------------
+html_test_create <- function(file, start, end) {
+  # Reading file and node
+  htmltest <- read_html(file, encoding = "UTF-8") |>
+    html_node("#inner-frame")
+
+  ### A series of tests to catch problems before they spit out errors-----------
+  # Checking if node exists
+  if (class(htmltest) == "xml_node") {
+    htmltable <- htmltest |>
+      html_table()
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Basic check if the table is correct
+  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
+    htmlready <- htmltable
+  } else {
+    validate("Please upload the specified type of html document.")
+  }
+
+  # Testing if the correct years have been entered
+  if (as.character(all_of(start)) %in% names(htmlready) &&
+    as.character(all_of(end)) %in% names(htmlready)) {
+    # Creating the first part of the data frame
+    read_proff_one(htmltable, start, end)
+  } else {
+    validate("Please ensure you have entered the correct start and end years.")
+  }
+}
+
+### First html transformation function, invoked by html_test_create() ----------
+# (stopping to allow checks for differing currency codes)
+read_proff_one <- function(object, start, end) {
+  object |>
+    # removing empty column, using all_of(argument)
+    select(
+      REGNSKAPSPERIODE:all_of(start)
+    ) |>
+    # removing white space
+    mutate(
+      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
+    ) |>
+    # removing duplicate table sometimes created by read_html()
+    filter(
+      !row_number() > 176
+    ) |>
+    # removing duplicate rows
+    filter(
+      !grepl("Lukk", REGNSKAPSPERIODE)
+    ) |>
+    # tidying data using pivot_longer() and all_of(argument)
+    pivot_longer(
+      all_of(end):all_of(start),
+      names_to = "year"
+    ) |>
+    mutate(
+      # changing to real NAs
+      value = na_if(value, "-"),
+      # turning years into numbers
+      year = as.integer(year),
+      # creating a currency code column and turning current currency codes into
+      # NAs
+      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
+    ) |>
+    # filling the currency column to avoid having to deal with NAs later
+    fill(
+      valutakode,
+      .direction = "downup"
+    )
+}
+
+### Second html transformation function, creating the data frame ---------------
+read_proff_two <- function(object) {
+  object |>
+    # removing unneeded currency column
+    select(-valutakode) |>
+    # removing dates and redundant currency
+    filter(
+      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
+        REGNSKAPSPERIODE != "Valutakode"
+    ) |>
+    mutate(
+      # removing white space in numbers
+      value = str_replace_all(value, "\\s", ""),
+      # turning numbers into true numbers
+      value = as.numeric(value),
+      # removing values that are years
+      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
+    ) |>
+    # removing any remaining duplicate rows
+    distinct() |>
+    # pivoting to create 'tidy' variables
+    pivot_wider(
+      names_from = REGNSKAPSPERIODE, values_from = value
+    ) |>
+    # tidying names
+    clean_names() |>
+    # removing headings from the table
+    select(
+      -lederlonn_i_hele_1000,
+      -resultatregnskap_i_hele_1000,
+      -balanseregnskap_i_hele_1000
+    ) |>
+    # making it clear what lonn refers to
+    rename(
+      lederlonn = lonn
+    ) |>
+    # Creating new variables for plots (red/green profit/loss)
+    mutate(
+      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
+      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
+    ) |>
+    # arranging by year for practical display
+    arrange(desc(year))
+}
+
+## Functions to read Excel file ------------------------------------------------
+### First Excel transformation - creating df used by NGO title function---------
+widen_excel_test <- function(file) {
+  # Reading Excel file
+  read_excel(file) |>
+    # cleaning names to enable later operations
+    clean_names()
+}
+
+### Second Excel function - creating primary Excel data frame-------------------
+widen_excel <- function(object) {
+  # testing if all columns are present
+  if (setequal(
+    names(object),
+    c("year", "firm", "category", "sub_category", "name", "value")
+  )) {
+    object |>
+      # uniting category and name to deal with multiple 'Other'
+      unite(
+        entry, c(category, name)
+      ) |>
+      # deleting Sub-category and Firm
+      select(-sub_category, -firm) |>
+      # pivoting to create 'tidy' variables
+      pivot_wider(names_from = entry, values_from = value) |>
+      # cleaning names
+      clean_names()
+    # Notification if all columns are not present
+  } else {
+    validate("Please upload data using the specified Excel template.")
+  }
+}
+
+# Data frames ================================================================
+## Creating data frame from html file-----------------------------------------
+html_currency <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$html$datapath) == "html",
+    "Please upload an html file."
+  ))
+
+  # Testing and creating data frame
+  html_test_create(
+    input$html$datapath,
+    as.character(input$start), as.character(input$end)
+  )
+})
+
+htmldata <- reactive({
+  read_proff_two(html_currency())
+})
+
+## Validating and creating long version of Excel file, if possible------------
+excel_test <- reactive({
+  # validating file
+  validate(need(
+    tools::file_ext(input$excel$datapath) == "xlsx",
+    "Please upload an Excel file."
+  ))
+
+  widen_excel_test(input$excel$datapath)
+})
+
+## Creating the primary Excel data frame--------------------------------------
+exceldata <- reactive({
+  widen_excel(excel_test())
+})
+```

### 2022-08-02 Report/Old/Summary.Rmd
Cleaning.

---
title: "Summary"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-12'
output: 
  html_document: 
    code_folding: hide
    number_sections: yes
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r nb, include=FALSE}
# Best read as a knitted HTML document!
```

# Data acquisition

## Official (Brønnøysund) data are not easily accessible

### Brønnøysund API access is expensive

[API access](https://www.brreg.no/en/products-and-services-2/order-products/subscription/subscription-to-annual-accounts/?nocache=1657649463638) to annual accounts data costs NOK 480,000.

The [experimental free API](https://data.norge.no/dataservices/4dabc5ea-b57b-3b68-9e48-2ff993be3f6e) no longer works.

### Brønnøysund pdfs must be ordered manually, and data cannot safely be extracted automatically

Pdf copies of annual accounts may be ordered manually. Data cannot safely be extracted automatically.

### Proff.no HTML is an option, but data must be downloaded manually

Brønnøysund data can be accessed manually at [Proff.no](https://www.proff.no). They do, however, prohibit automated access/web scraping. Data can, with some provisos, safely be extracted from downloaded HTML copies of individual web pages. Sample code is included below. The legality of this approach may be questionable, and, as noted below, Brønnøysund data does not contain all the necessary information.

```{r proff, eval=FALSE}
library(tidyverse)
read_proff <- function(file, start, end){
  # the user must supply file name and start and end years
  read_html(file, encoding = "UTF-8") |> # reading HTML
  html_node("#inner-frame") |> 
  html_table() |> 
  select( # removing empty column
    REGNSKAPSPERIODE:start) |> 
  mutate( # removing whitespace
    REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
  ) |> 
  filter( # removing duplicate table
      !row_number() > 176
    ) |>
  filter( # removing duplicate rows
    !grepl("Lukk", REGNSKAPSPERIODE)
    ) |> 
  pivot_longer( # tidying data
    end:start, names_to = "year") |> 
  mutate( # changing to real NAs and turning years into numbers
    value = na_if(value, "-"),
    year = as.integer(year)) |> 
  mutate( # adding currency column
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |> 
  fill( # filling the currency column
    valutakode, .direction = "updown"
    ) |> 
  filter( # removing dates and redundant currency
    REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
      REGNSKAPSPERIODE != "Valutakode") |> 
  mutate( # removing whitespace in numbers
    value = str_replace_all(value, "\\s", "")
  ) |> 
  mutate( # turning into numbers
    value = as.numeric(value)
  ) |> 
  mutate( # removing years as values
    value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
  )
}
```

>**Conclusion**: Manual downloads from Proff.no is the only functional option in the absence of API access. But this is only a partial solution.

## Brønnøysund data does not contain all the necessary information

A further problem is that the accounting data that must be supplied to Brønnøysund does not contain all the data necessary to calculate most relevant KPIs.

### Annual reports are not always available, and cannot be read automatically

Even in the case of NGOs that make pdf copies of their annual reports available, these may not contain the necessary data, and even if they do, these cannot safely be extracted automatically. (In some instances they may be extracted with the help of [Tabula](https://tabula.technology), but this process cannot be automated.)

>**Conclusion**: The only possible data source is manual reporting by organisations. We propose using a standardised Excel form, which would enable automatic data extraction and calculation of KPIs, etc. (Example below.) This solution may be combined with data taken from Proff.no.

```{r excel, message=FALSE}
library(readxl)
excel <- read_excel("Data/OPXfinans2.xlsx")
excel
```

# KPIs and visualisations

We can calculate the suggested KPI on the basis of the Excel. We are investigating other KPIs, and have some suggestions.

>**To do**: decide on KPIs. Explore if any visualisations would be useful.

# Final products

## Shiny app?
A Shiny app that accepts uploaded Excel and/or html files and automatically calculates the desired KPIs and displays them (+ one or more plots, if desired)?

Problems:

- hosting?
- does this actually add any value beyond what they already have?
- Need to discuss deliverables with Bosse and Solveig.
- Do we want to learn Shiny? (Does anyone know it already?)

## R Markdown report (*obligatory!*)
Is this an outline of the process, a report to OPX, or?

### 2022-08-02 Report/Graphs for final report.Rmd
CARE net income/loss graph.

@@
 ```
+
+```{r}
+care |>
+  rename(
+    Year = year,
+    `Net income` = arsresultat,
+    `Net positive` = annual_profit
+  ) |>
+  ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+  scale_fill_green(kreftforeningen$annual_profit) +
+  labs(
+    x = "", y = "", title = "Net income (in thousands)"
+  ) +
+  theme_light() +
+  theme(legend.position = "none")
+```

### 2022-08-02 Report/Group 5 report.Rmd
Commen on names.
Added Innsamlingskontrollen. Added CARE net income/loss graph. Changing text on Fundraising ROI.

@@
 
+<!--# Vi burde vel ha navnene våre, siden rapporten skal lastes opp på Inspera?-->
+<!--# Legg til et sted: The Norwegian Control Committee for Fundraising (Innsamlingskontrollen) also has some very basic financial data and information available about selected charitable organisations.-->
+
 ```{r libraries, include=FALSE}
@@
 
-## Creating df------------------------------------------------------------------
+## Creating dfs-----------------------------------------------------------------
 kreftforeningen <- read_proff("Data/Kreftforeningen.html", "2002", "2021")
+care <- read_proff("Data/CARENorge2010_2021.html", "2010", "2021")
 
@@
 
-<!--# En webcrawler er en bot -->
-
 We discussed creating web scraping code that could collect data across
@@
 
-<!-- # Legger inn tre grafer som viser manglende data i Proffs årsregnskap-->
+<!-- # Legger inn fire grafer som viser manglende data i Proffs årsregnskap-->
 
@@
   theme_light() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "top")
+```
+
+```{r care_arsresultat, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Net income/loss (in NOK 1,000) data for CARE Norway from Proff.no"}
+care |>
+  rename(
+    Year = year,
+    `Net income` = arsresultat,
+    `Net positive` = annual_profit
+  ) |>
+  ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
+  scale_fill_green(kreftforeningen$annual_profit) +
+  labs(
+    x = "", y = ""
+  ) +
+  theme_light() +
+  theme(legend.position = "none")
 ```
@@
 - **Fundraising ROI**:
-  - A basic measure of the return on investment of direct fundraising expenses:
+  - A basic measure of the return on investment of fundraising expenses:
 
@@
 
-This KPI is not a true measure of fundraising ROI, as it only includes (as we understand OPX's data) the direct fundraising expenses, and does not include major indirect expenses such as salary costs.
+Whether this is a true measure of fundraising ROI depends on how the underlying data have been calculated -- e.g. if they include salary costs, etc.
 

### 2022-08-03 Presentasjon/presentasjon.Rmd
Pres test

--- 2022-08-02 Presentasjon/presentasjon.Rmd
@@
 
-# Findings
-
-# Limitations and steps forward
-
 ## <iframe width="1280" height="720" src="Images/henrik.mp4" align="middle" frameborder="0" allowfullscreen></iframe>
 
-## {data-background="Images/tenke.png" data-background-size=cover}
+----
+```{r, out.height="500", out.width="500", fig.align='center'}
+knitr::include_graphics("Images/tenke.png")
+
+```
 
 ## <br> Our experience
-
 - Unsolvable case due to non-existent data
 - Hard to create credible KPIs due to limited knowledge of economics and the NGO sector
-- **Insight into real-life work challenges - the clash between is desired and what is possible**
+- **Insight into real-life work challenges - the clash between what is desired and what is possible**

### 2022-08-03 Presentasjon/presentasjon.Rmd
Changing slides

@@
 - Brønnøysund: We attempted to request a temporary access to XML documents
-<!--# We received some sample XML data from Statistics Norway -- we did not talk to Brønnøysund at all-->
 
@@
 
+# Findings
+
+# Limitations and steps forward
+
 ## <iframe width="1280" height="720" src="Images/henrik.mp4" align="middle" frameborder="0" allowfullscreen></iframe>
 
-----
-```{r, out.height="500", out.width="500", fig.align='center'}
-knitr::include_graphics("Images/tenke.png")
-
-```
+## {data-background="Images/tenke.png" data-background-size=cover}
 

### 2022-08-03 Report/Group 5 report v2.Rmd
Kopi

--- 2022-08-02 Report/Group 5 report.Rmd
@@
----
+----
 title: "Group 5 Report"
@@
 <!--# Vi burde vel ha navnene våre, siden rapporten skal lastes opp på Inspera?-->
+
 <!--# Legg til et sted: The Norwegian Control Committee for Fundraising (Innsamlingskontrollen) also has some very basic financial data and information available about selected charitable organisations.-->
@@
 
-The project chosen for our group was Oslo Philanthropic Exchange (OPX), 
+The project assigned to our group was Oslo Philanthropic Exchange (OPX), 
 a startup working to develop solutions to assist individuals and 
@@
 
-To do this, data from charities need to be collected. OPX’s current 
-challenge is data acquisition, as data currently must be collected manually. They therefore wanted a way to automate data collection, for example using web scraping.
+To do this, data from charities need to be collected. One of OPX’s current 
+challenges is data acquisition, as data currently must be collected manually. They therefore wanted a way to automate data collection, for example using web scraping.
 

### 2022-08-03 Report/Group 5 report v2.Rmd
Update brainstorming and data

@@
 
-OPX wanted us to harvest the organisations' financial data from the web.
+OPX wanted us to harvest the organizations' financial data from the web.
 During the first stage of brainstorming, we wanted to find a simple and
@@
 
-We discussed creating web scraping code that could collect data across
+We discussed creating a webcrawler that could collect data across
 different websites, but quickly decided that this would be too
-difficult, and that it would only work with selected organisations. The next
-suggested solution was to create code that could read PDF documents
-that OPX downloaded themselves. Our code would analyse and organise the
+difficult, and that it would only work with selected organizations. The
+suggested solution was to create a code that could read PDF documents
+that OPX downloaded themselves. Our code would analyze and organize the
 uploaded data from the PDF documents.
 
-However, different organisations publish financial data in different
-ways in their annual reports -- some financial statements were included
+However, different organizations publish financial data in different
+ways in their yearly reports -- some financial statements were included
 as tables, some as pictures, etc. This led us to phase 2.
@@
 ```{r, echo=FALSE, fig.cap="Need caption"}
-knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene
-
-# inkluder en med årsregnskapet som pdf her
+# Illustration some text some pictures in financial statements
+knitr::include_graphics("Images/lagret_tekst.png") # Kreftforeningen 201?
+knitr::include_graphics("Images/lagret_bilde.png") # Regnskogfondet 2018
 ```
@@
 
-At this stage, we wanted to explore alternatives to analysing PDF
+At this stage, we wanted to explore alternatives to analyzing PDF
 documents. We discussed how to do so, and concluded that we had three
 options: 
-
-- Financial statements in annual report PDF documents. 
-- Web scrape Proff.no for financial statements. 
-- Web scrape the Brønnøysund registry.
+- Financial statements in yearly report PDF documents. 
+- Webscrape Proff.no for financial statements. 
+- Webscrape the Brønnøysund registry.
 
@@
 were feasible or not. The first option of using PDF documents still had
-the issue of financial statements not always being published as text, as well as their often wildly differing formats.
-Unfortunately, we quickly found that Proff.no disallows any automated collection of data from their website.
-The Brønnøysund registry does offer API access to all the financial statements, but we discovered that access currently costs NOK 480,000.
-
-We started to look at workarounds for this problem. It could still be
+the issue of financial statements not always being published as text.
+Also, we quickly found that Proff.no disallowed webscraping on their
+website. The Brønnøysund registry did store downloadable XML documents
+of all the financial statements on their website, but we discovered that
+access to these files cost about half a million NOK.
+
+We started to look at workarounds for this problem. It would still be
 possible to download Proff.no HTML websites manually for each
-organisation. We did receive some XML files from Statistics Norway, but using these is ultimately not a viable solution for OPX.
+organization. And, we attempted to request a temporary access to the
+Brønnøysund registry's XML documents.
 
@@
 attempted to do so in Excel at first, so that we later could code this
-into R. While doing so, we realised that it would be practically
+into R. While doing so, we realized that it would be practically
 impossible to extract the correct sums for the OPX accounting items from
@@
 ```{r, echo=FALSE, fig.cap="Need caption", fig.align='center'}
-knitr::include_graphics("Images/info_kreft.png")
-```
+# Illustration brns vs OPX accounting items
+knitr::include_graphics("Images/Brns_items.png")
+knitr::include_graphics("Images/OPX_items.png")
+knitr::include_graphics("Images/Marriage.png")
+```
+
+When also discovering that not all organizations publish their financial
+statements, and certainly not for every year, we decided to give up on
+collecting financial data for OPX. This is when we moved on to the third
+and final stage.
 
 ```{r, echo=FALSE, out.width="100%", fig.cap="Need caption"}
-knitr::include_graphics("Images/regnskapssti.png")
-```
-
-When also discovering that not all organisations publish their financial
-statements, and certainly not for every year, we decided to give up on
-collecting financial data for OPX, as the task turned out to impossible given current reporting requirements and practices. This is when we moved on to the third
-and final stage.
+# Illustration only published for some years example
+knitr::include_graphics("Images/Regnskogfondet_stops2018.png")
+knitr::include_graphics("Images/Kreftforeningen_starts2019.png")
+```
 
@@
 
-The OPX case requested two tasks -- web scrape financial data, and create
+The OPX case requested two tasks -- webscrape financial data, and create
 KPIs. During phase three, we scrapped all ideas and progress from the
@@
 In order to create a somewhat more challenging task for ourselves, we
-also decided to create an interactive dashboard based on Shiny that OPX could use, which would
-display the KPIs as well as financial trends for the selected
-organisation. To use this, OPX needs to gather the financial data
-themselves, as well as download data manually from Proff.no, and then upload this into the dashboard in the form of Excel and HTML documents.
+also decided to create an interactive dashboard that OPX could use, which 
+would display the KPIs as well as financial trends for the selected
+organization. To use this, OPX needs to gather the financial data
+themselves, then upload this into the shiny app as an excel and an 
+html document.
 
@@
 
-We wanted to use web scraping to collect the data, so that we could have
+We wanted to use webscraping to collect the data, so that we could have
 automated the data input process. As described, we explored different
 potential solutions. We experimented on coding such that data could have
-been imported from XML files. However, once we realised that the available
+been imported from XML files. However, once we realized that the
 financial data was of little use, this experimentation stopped.
@@
 We copied these financial statements manually into an Excel document.
-Perhaps a 'simple' solution, however our code requires OPX to import
+Perhaps a "simple" solution, however our code requires OPX to import
 both an Excel document themselves into shiny, as well as a HTML document
 from Proff.no.
+
+```{r, echo=FALSE, out.width="100%", fig.cap="Need caption"}
+# Bilde av den "endelige" excel-versjonen som brukes til å laste inn i shiny
+knitr::include_graphics("Images/Excel_final.png")
+```
 
@@
 # Data description
+
+ur data was gathered from OPXs own web page, and from Proff.no. As described 
+in the previous parts, we wanted to gather data from the yearly reports or by 
+scraping the Brønnøyund registry or Proff.no, but found this to be close to 
+impossible. The data used is the one that OPX has gathered themselves directly 
+from the organizations. This data was also used to test that our codes in R 
+were correct.
+
+We have also used Proff.no to visualize the annual incomes and some KPIs. 
+This is data that the organizations have reported themselves to Proff.no. 
+Since there are no regulations regarding reporting data for NGOs in Norway, 
+the organizations have not been consistent with the data reporting. This 
+explains the holes in the presented data.
 

### 2022-08-03 Presentasjon/presentasjon.Rmd
Nye stikkord

@@
 ## <br> Phase 1 – how to harvest and analyze the data?
-- A webcrawler that could collect data across different websites? <br> This would be too difficult and it would only work with some organizations
+- OPX wanted us to harvest data --> how? <br> 
+- A webcrawler? <br> 
+- No: Would not work for every organization <br> 
 - A code that could read PDF? <br> 
+- Perhaps: Financial statements as text tables and as images
 
 ## <br> Phase 2 – Exploring three options
-- Financial statements in yearly reports published as PDF <br>
-Issue: Financial statements are not always published as text
-- Webscraping financial statements from Proff.no <br>
-Issue: Proff.no does not allow webscraping!
-- Webscraping the Brønnøysund registry <br>
-Issue: The registry does offer API access, but access costs about 500k kr
+- 1. Reading PDFs <br>
+Issue: Not always published as text
+- 2. Webscraping from Proff.no <br>
+Issue: Proff.no does not allow webscraping
+- 3. Webscraping the Brønnøysund registry <br>
+Issue: No API access, and access costs ~ kr 500K
 
 ## <br> Workarounds... 
-- Proff.no: It maaay be legal to manually download the websites as HTML instead of scraping 
-- Brønnøysund: We attempted to request a temporary access to XML documents
+- Proff.no: Download HTML?
+- Brønnøysund: Request temporary access
 
 ## <br> ... and why they did not work
-- Different accounting divisions at Proff.no and Brønnøysund prohibited extracting the correct sums  OPX wanted
-- Not all NGOs publish their financial statements and certainly not for every year
+- Different accounting items used
+- Incomplete financial reporting from NGOs
+
 - **Which led us to Phase 3 - scrapping the scraping**
@@
 
-# Data collection
+## <br> Phase 3 – Scrapping the scraping
+- OPX wanted KPIs <br>
+- Found useful KPIs <br>
+- Created an interactive dashboard
 
-# Data description
+## <br> Data collection & data description
+- Used data from OPX and Proff.no <br>
+- OPX: From their webpage. <br>
+- Used to create KPIs and check our codes <br>
+- Proff.no: Downloaded HTML documents. <br>
+- Used to create visualizations.
 

### 2022-08-03 Report/Old/KPIs and Visualisations information.Rmd

--- 2022-07-31 Report/KPIs and Visualisations.Rmd
@@
 
-En idé om hvordan pakken [MetBrewer](https://github.com/BlakeRMills/MetBrewer), som jeg har brukt for farger, virker i praksis kan man se hvis man kjører "Knit" på dokumentet "Colourful Plots" i Shiny/Test.
+En idé om hvordan pakken [MetBrewer](https://github.com/BlakeRMills/MetBrewer), som jeg har brukt for farger, virker i praksis kan man se hvis man kjører "Knit" på dokumentet "Colourful Plots".
 
@@
 
-- **Liquid assets/obligations**:
+- **Liquid assets/obligations**: [Eneste KPI OPX hadde]
   - A measure of the relationship between the organisation's liquid assets and its obligations:
@@
 $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$
-
->NB: Koden er kanskje ikke oppdatert, men beskrivelsen over er den som er i dashboardet og som koden der er basert på. (Det kan være feil utregninger i koden i dette dokumentet akkurat nå.)
 

### 2022-08-03 Presentasjon/Presentasjon v3.Rmd
Pres v3
Presentasjon oppdatering

--- 2022-08-03 Presentasjon/presentasjon.Rmd
@@
 - A webcrawler? <br> 
-- No: Would not work for every organization <br> 
+No: Would not work for every organization 
 - A code that could read PDF? <br> 
-- Perhaps: Financial statements as text tables and as images
+Perhaps: Financial statements as text tables and as images
 
@@
 
-- **Which led us to Phase 3 - scrapping the scraping**
+#brns_items
+#opx_items
 
@@
 
-## <br> Phase 3 – Scrapping the scraping
-- OPX wanted KPIs <br>
-- Found useful KPIs <br>
+## <br> Which led us to Phase 3 – Scrapping the scraping
+- OPX wanted KPIs
+- Found useful KPIs
 - Created an interactive dashboard
@@
 ## <br> Data collection & data description
-- Used data from OPX and Proff.no <br>
-- OPX: From their webpage. <br>
-- Used to create KPIs and check our codes <br>
-- Proff.no: Downloaded HTML documents. <br>
-- Used to create visualizations.
-
-# Findings
-
-# Limitations and steps forward
+- Used data from OPX and Proff.no 
+- OPX: From their webpage
+- Used to create KPIs and check our codes
+- Proff.no: Downloaded HTML documents
+- Used to create visualizations
 

### 2022-08-03 Report/Group 5 report v2.Rmd
Bilder + La inn en "O"
Fixing YAML block

@@
-----
+---
 title: "Group 5 Report"
@@
 
-ur data was gathered from OPXs own web page, and from Proff.no. As described 
+Our data was gathered from OPXs own web page, and from Proff.no. As described 
 in the previous parts, we wanted to gather data from the yearly reports or by 

### 2022-08-03 Report/Group 5 report v2.Rmd
Image captions

@@
 # Illustration brns vs OPX accounting items
-knitr::include_graphics("Images/Brns_items.png")
-knitr::include_graphics("Images/OPX_items.png")
+knitr::include_graphics("Images/Brns_items1.png")
+knitr::include_graphics("Images/OPX_items1.png")
 knitr::include_graphics("Images/Marriage.png")

### 2022-08-03 Report/Group 5 report v2.Rmd
saved this time

--- 2022-08-03 Report/Group 5 report v2.Rmd
@@
 
-```{r, echo=FALSE, fig.cap="Need caption"}
+```{r, echo=FALSE, fig.cap="Kreftforeningen 2021: The financial report saved as text in the PDF"}
 # Illustration some text some pictures in financial statements
-knitr::include_graphics("Images/lagret_tekst.png") # Kreftforeningen 201?
+knitr::include_graphics("Images/lagret_tekst.png") # Kreftforeningen 2021
+```
+
+```{r, echo=FALSE, fig.cap="Regnskogfondet 2018: The financial report saved as an image in the PDF"}
+# Illustration some text some pictures in financial statements
 knitr::include_graphics("Images/lagret_bilde.png") # Regnskogfondet 2018
@@
 
-```{r, echo=FALSE, fig.cap="Need caption", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Accounting items for Kreftforeningen 2020 and Diabetesforbunet 2020, gathered from the Brønnøysund registry", fig.align='center'}
 # Illustration brns vs OPX accounting items
-knitr::include_graphics("Images/Brns_items.png")
-knitr::include_graphics("Images/OPX_items.png")
+knitr::include_graphics("Images/Brns_items1.png")
+```
+
+```{r, echo=FALSE, fig.cap="Accounting items for Kreftforeningen 2020 and Diabetesforbunet 2020, gathered from OPX", fig.align='center'}
+# Illustration brns vs OPX accounting items
+knitr::include_graphics("Images/OPX_items1.png")
+```
+
+```{r, echo=FALSE, fig.cap="An attempt at marrying the two sets of accounting items", fig.align='center'}
+# Illustration brns vs OPX accounting items
 knitr::include_graphics("Images/Marriage.png")
@@
 
-```{r, echo=FALSE, out.width="100%", fig.cap="Need caption"}
+```{r, echo=FALSE, out.width="100%", fig.cap="Regnskogfondet stopped posting financial statement PDFs on their website in 2018"}
 # Illustration only published for some years example
 knitr::include_graphics("Images/Regnskogfondet_stops2018.png")
+```
+
+```{r, echo=FALSE, out.width="100%", fig.cap="Kreftforeningen started posting financial statement PDFs on their website in 2019"}
+# Illustration only published for some years example
 knitr::include_graphics("Images/Kreftforeningen_starts2019.png")
@@
 
-```{r, echo=FALSE, out.width="100%", fig.cap="Need caption"}
+```{r, echo=FALSE, out.width="100%", fig.cap="The data format used to upload into shiny"}
 # Bilde av den "endelige" excel-versjonen som brukes til å laste inn i shiny
@@
   theme(legend.position = "none")
-```
-
-```{r egenkapital, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Total equity/liabilities (in NOK 1,000) data for the Norwegian Cancer Society from Proff.no"}
-kreftforeningen |>
-  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
-  scale_fill_green(kreftforeningen$balance_positive) +
-  labs(
-    x = "", y = ""
-  ) +
-  theme_light() +
-  theme(legend.position = "none")
-```
-
-```{r CEO, echo=FALSE, fig.align='center', fig.asp=0.618, fig.cap="CEO compensation data for the Norwegian Cancer Society from Proff.no"}
-kreftforeningen |>
-  # preparing for pivoting
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |>
-  # Compensation in ordinary numbers
-  mutate(
-    lederlonn = lederlonn * 1000,
-    leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-  ) |>
-  # renaming for the purposes of the plot
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |>
-  # pivoting to join the two values to use fill
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  # using currency scale function to add correct currency code or ?
-  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
-  # pretty colours
-  scale_fill_met_d("Hokusai2", direction = -1) +
-  labs(
-    x = "", y = ""
-  ) +
-  theme_light() +
-  theme(legend.position = "top")
 ```

### 2022-08-03 Presentasjon/Presentasjon v3.Rmd
Presentasjon mock
Oppdatert mock

@@
 title: "Presentasjon OPX Fiffig tittel"
-author: "Gruppe 5"
-date: "4.8.2022"
+author: "Group 5: Aga, Amalie, Henrik, Ida, Torbjørn"
+date: "2022-08-04"
 output: 
   ioslides_presentation:
+    widescreen: TRUE 
     css: stylesss.css
@@
 
-#brns_items
-#opx_items
+## Items Brønnøysund
+```{r, out.height="470", out.width="800", fig.align='center'}
+knitr::include_graphics("Images/Brns_items.png")
+
+```
+
+## Items OPX
+```{r, out.height="514", out.width="710", fig.align='center'}
+knitr::include_graphics("Images/OPX_items.png")
+
+```
 
@@
 
+# Dashboard
+
+# Limitations and steps forward
+
 ## <iframe width="1280" height="720" src="Images/henrik.mp4" align="middle" frameborder="0" allowfullscreen></iframe>

### 2022-08-03 Presentasjon/Presentasjon v3.Rmd
Presentasjon foreløpig

@@
 ---
-title: "Presentasjon OPX Fiffig tittel"
+title: "Presentation: OPX"
 author: "Group 5: Aga, Amalie, Henrik, Ida, Torbjørn"
@@
 
-## <br> Phase 1 – how to harvest and analyze the data?
-- OPX wanted us to harvest data --> how? <br> 
-- A webcrawler? <br> 
-No: Would not work for every organization 
-- A code that could read PDF? <br> 
+## <br> How to harvest and analyze the data?
+- OPX wanted data collection to be automated <br> 
+- Webscraping? <br> 
+No: Different HTML and PDF structures
+- A code that could read PDF documents? <br> 
 Perhaps: Financial statements as text tables and as images
 
-## <br> Phase 2 – Exploring three options
+## <br> Exploring three options
 - 1. Reading PDFs <br>
@@
 - 2. Webscraping from Proff.no <br>
-Issue: Proff.no does not allow webscraping
+Issue: Proff.no does not allow (automated) webscraping
 - 3. Webscraping the Brønnøysund registry <br>
-Issue: No API access, and access costs ~ kr 500K
+Issue: No API access, and access costs ~ NOK 500 000
+
+#BYTTE UT MED 1 2 3
 
 ## <br> Workarounds... 
-- Proff.no: Download HTML?
-- Brønnøysund: Request temporary access
+- Proff.no: Download the webpage as HTML document?
+- Brønnøysund: Received XML files from Statistics Norway
 
 ## <br> ... and why they did not work
-- Different accounting items used
-- Incomplete financial reporting from NGOs
+- XML not a viable long term solution for OPX 
+- Incomplete financial reporting from NGOs -- no legal obligations
+- Proff.no vs. OPX: Different accounting items used <br> <br> <br>
+
+```{r, out.height="188", out.width="500", fig.align='center'}
+knitr::include_graphics("Images/FRI.png")
+
+```
 
@@
 
-## <br> Which led us to Phase 3 – Scrapping the scraping
-- OPX wanted KPIs
+## <br> Which led us to scrapping the scraping <br>
 - Found useful KPIs
@@
 ## <br> Data collection & data description
-- Used data from OPX and Proff.no 
-- OPX: From their webpage
-- Used to create KPIs and check our codes
-- Proff.no: Downloaded HTML documents
-- Used to create visualizations
+- Proff.no: Downloaded HTML documents <br>
+Used to create visualizations
+- OPX: From their webpage <br>
+Used to create KPIs and test our codes
 
 # Dashboard
+
+## <br> <br> <br> <br> torbjornso.shinyapps.io/NGOanalysis/
 
@@
 ## <br> Our experience
-- Unsolvable case due to non-existent data
+- Unsolvable case due to lack of accessible and relevant data
 - Hard to create credible KPIs due to limited knowledge of economics and the NGO sector

### 2022-08-03 Presentasjon/Presentasjon v3.Rmd
Presetnasjon endelig

@@
 ## <br> Exploring three options
-- 1. Reading PDFs <br>
+1. Reading PDFs <br>
 Issue: Not always published as text
-- 2. Webscraping from Proff.no <br>
+2. Webscraping from Proff.no <br>
 Issue: Proff.no does not allow (automated) webscraping
-- 3. Webscraping the Brønnøysund registry <br>
+3. Webscraping the Brønnøysund registry <br>
 Issue: No API access, and access costs ~ NOK 500 000
-
-#BYTTE UT MED 1 2 3
 
@@
 
-## {data-background="Images/tenke.png" data-background-size=cover}
+## {data-background="Images/upbeat_and_happy.png" data-background-size=cover}
 
@@
 - **Insight into real-life work challenges - the clash between what is desired and what is possible**
+
+## <br> <br> <br> The End
+**Thank you for this opportunity!** <br> <br> <br>
+- Questions?

### 2022-08-03 Report/Group 5 report v2.Rmd
Authors, editing.

@@
 ---
-title: "Group 5 Report"
-author: "ISSSV1337 Group 5"
+title: "ISSSV1337 Group 5 Report"
+author:    | 
+    | Amalie Halle Christensen, Ida Kristine Garthe
+    | Torbjørn Skinnemoen Ottersen, Henrik Waage Rui
+    | Agnieszka Sadlowska
 date: "2022-08-03"
@@
 ```
-
-<!--# Vi burde vel ha navnene våre, siden rapporten skal lastes opp på Inspera?-->
-
-<!--# Legg til et sted: The Norwegian Control Committee for Fundraising (Innsamlingskontrollen) also has some very basic financial data and information available about selected charitable organisations.-->
 
@@
 responsible investment strategy. Their main goal is to increase cash 
-flow from private groups and individuals to charitable purposes. Therefore, the donor's trust 
-in the charities’ abilities to carry out their promises is of paramount significance.
+flow from private groups and individuals to charitable purposes. Therefore, the donors' trust in the charities’ abilities to carry out their promises is of paramount significance.
 
@@
 
- <!-- Needs work -->
-Additionally, by creating Key Performance Indicators based on these data, OPX can use 
-them to benchmark the relationship between financial assets and emergency 
-relief obligations.
+Additionally, by creating Key Performance Indicators based on these data, OPX can use them to benchmark the relationship between financial assets and emergency relief obligations.
 
@@
 
-2. Creating KPI(s) is important for the organization’s performance 
+2. Creating KPI(s) is important for the organisation’s performance 
 management, to influence its objectives, and to foster growth.
@@
 
-## Phase 1: How to harvest and analyse the data? {.smaller}
+## Phase 1: How to harvest and analyse the data?
 
@@
 
-We discussed creating a webcrawler that could collect data across
+We discussed creating web scraping code that could collect data across
 different websites, but quickly decided that this would be too
 difficult, and that it would only work with selected organizations. The
-suggested solution was to create a code that could read PDF documents
-that OPX downloaded themselves. Our code would analyze and organize the
+suggested solution was to create code that could read PDF documents
+that OPX downloaded themselves. Our code would analyse and organise the
 uploaded data from the PDF documents.
 
-However, different organizations publish financial data in different
-ways in their yearly reports -- some financial statements were included
+However, different organisations publish financial data in different
+ways in their annual reports -- some financial statements were included
 as tables, some as pictures, etc. This led us to phase 2.
@@
 
-```{r, echo=FALSE, fig.cap="Kreftforeningen 2021: The financial report saved as text in the PDF"}
+```{r, echo=FALSE, fig.cap="Norwegian Cancer Society 2021: The financial report saved as text in the PDF."}
 # Illustration some text some pictures in financial statements
@@
 
-```{r, echo=FALSE, fig.cap="Regnskogfondet 2018: The financial report saved as an image in the PDF"}
+```{r, echo=FALSE, fig.cap="Rainforest Foundation 2018: The financial report saved as an image in the PDF."}
 # Illustration some text some pictures in financial statements
@@
 
-## Phase 2: Exploring solutions {.smaller}
-
-At this stage, we wanted to explore alternatives to analyzing PDF
+## Phase 2: Exploring solutions
+
+At this stage, we wanted to explore alternatives to analysing PDF
 documents. We discussed how to do so, and concluded that we had three
 options:
-- Financial statements in yearly report PDF documents. 
-- Webscrape Proff.no for financial statements. 
-- Webscrape the Brønnøysund registry.
+
+- Financial statements in anual report PDF documents. 
+- Web scrape Proff.no for financial statements. 
+- Web scrape the Brønnøysund registry.
 
@@
 the issue of financial statements not always being published as text.
-Also, we quickly found that Proff.no disallowed webscraping on their
-website. The Brønnøysund registry did store downloadable XML documents
-of all the financial statements on their website, but we discovered that
-access to these files cost about half a million NOK.
+Also, we quickly found that Proff.no disallowed web scraping on their
+website. The Brønnøysund registry did offer API access to all the financial statements on their website, but we discovered that
+access to these data costs about half a million NOK.
 
 We started to look at workarounds for this problem. It would still be
-possible to download Proff.no HTML websites manually for each
-organization. And, we attempted to request a temporary access to the
-Brønnøysund registry's XML documents.
+possible to download Proff.no HTML web pages manually for each
+organisation. We also received access to some XML files from Statistics Norway, but this would not be a viable solution for OPX in the long term.
 
@@
 
-```{r, echo=FALSE, fig.cap="Accounting items for Kreftforeningen 2020 and Diabetesforbunet 2020, gathered from the Brønnøysund registry", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Accounting items for the Norwegian Cancer Society 2020 and the Norwegian Diabetes Association 2020, gathered from the Brønnøysund registry.", fig.align='center'}
 # Illustration brns vs OPX accounting items
@@
 
-```{r, echo=FALSE, fig.cap="Accounting items for Kreftforeningen 2020 and Diabetesforbunet 2020, gathered from OPX", fig.align='center'}
+```{r, echo=FALSE, fig.cap="Accounting items for the Norwegian Cancer Society 2020 and the Norwegian Diabetes Association 2020, gathered from OPX.", fig.align='center'}
 # Illustration brns vs OPX accounting items
@@
 
-```{r, echo=FALSE, fig.cap="An attempt at marrying the two sets of accounting items", fig.align='center'}
+```{r, echo=FALSE, fig.cap="An attempt at marrying the two sets of accounting items.", fig.align='center'}
 # Illustration brns vs OPX accounting items
@@
 
-When also discovering that not all organizations publish their financial
+When we also discovered that not all organisations publish their financial
 statements, and certainly not for every year, we decided to give up on
@@
 
-```{r, echo=FALSE, out.width="100%", fig.cap="Regnskogfondet stopped posting financial statement PDFs on their website in 2018"}
+```{r, echo=FALSE, out.width="100%", fig.cap="The Rainforest Foundation stopped posting financial statement PDFs on their website in 2018."}
 # Illustration only published for some years example
@@
 
-```{r, echo=FALSE, out.width="100%", fig.cap="Kreftforeningen started posting financial statement PDFs on their website in 2019"}
+```{r, echo=FALSE, out.width="50%", fig.cap="The Norwegian Cancer Society started posting financial statement PDFs on their website in 2019."}
 # Illustration only published for some years example
@@
 
-## Phase 3: Scrapping the scraping {.smaller}
-
-The OPX case requested two tasks -- webscrape financial data, and create
+## Phase 3: Scrapping the scraping
+
+The OPX case requested two tasks -- web scrape financial data, and create
 KPIs. During phase three, we scrapped all ideas and progress from the
@@
 
-We wanted to use webscraping to collect the data, so that we could have
+We wanted to use web scraping to collect the data, so that we could have
 automated the data input process. As described, we explored different
 potential solutions. We experimented on coding such that data could have
-been imported from XML files. However, once we realized that the
+been imported from XML files. However, once we realised that the
 financial data was of little use, this experimentation stopped.
@@
 financial data that OPX had already gathered from different
-organizations and organized into useful accounting items.
+organisations and organised into useful accounting items.
 
 We copied these financial statements manually into an Excel document.
-Perhaps a "simple" solution, however our code requires OPX to import
-both an Excel document themselves into shiny, as well as a HTML document
-from Proff.no.
-
-```{r, echo=FALSE, out.width="100%", fig.cap="The data format used to upload into shiny"}
+Perhaps a 'simple' solution, however our code requires OPX to upload
+both an Excel document, as well as a HTML document
+from Proff.no, to the dashboard.
+
+```{r, echo=FALSE, out.width="100%", fig.cap="The data format used to upload to the dashboard."}
 # Bilde av den "endelige" excel-versjonen som brukes til å laste inn i shiny
@@
 
-Our data was gathered from OPXs own web page, and from Proff.no. As described 
-in the previous parts, we wanted to gather data from the yearly reports or by 
-scraping the Brønnøyund registry or Proff.no, but found this to be close to 
-impossible. The data used is the one that OPX has gathered themselves directly 
-from the organizations. This data was also used to test that our codes in R 
+Our data was gathered from OPX's own web page, and from Proff.no. As described 
+in the previous parts, we wanted to gather data from the annual reports or by 
+scraping the Brønnøyund registry or Proff.no, but found this to be impossible. The data used is the one that OPX has gathered themselves directly 
+from the organisations. This data was also used to test that our codes in R 
 were correct.
 
-We have also used Proff.no to visualize the annual incomes and some KPIs. 
-This is data that the organizations have reported themselves to Proff.no. 
+We have also used Proff.no to visualise the annual incomes and some KPIs. 
+This is data that the organisations have reported themselves to Proff.no. 
 Since there are no regulations regarding reporting data for NGOs in Norway, 
-the organizations have not been consistent with the data reporting. This 
-explains the holes in the presented data.
-
-<!--# Fra Visualisations (nedenfor): The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.-->
-
-<!--Legg inn bilder her-->
-
-<!-- # Legger inn fire grafer som viser manglende data i Proffs årsregnskap-->
+the organisations have not been consistent with the data reporting. This may 
+explain the lacunae in the presented data (see e.g. figures 9 and 10).
 
@@
 
+\newpage
 # Findings

### 2022-08-03 Report/Group 5 report v2.Rmd
Centred image.

@@
 
-```{r, echo=FALSE, out.width="50%", fig.cap="The Norwegian Cancer Society started posting financial statement PDFs on their website in 2019."}
+```{r, echo=FALSE, out.width="50%", fig.align='center', fig.cap="The Norwegian Cancer Society started posting financial statement PDFs on their website in 2019."}
 # Illustration only published for some years example
@@
 
+\newpage
 - **Private/public proportion**:

### 2022-08-03 Old/Nomnoml, just in case.R
Cleaning.

@@ -1,4 +1,0 @@
-install.packages("nomnoml")
-install.packages("webshot")
-library(nomnoml)
-webshot::install_phantomjs()

### 2022-08-03 Report/Group 5 report.Rmd
Adding Github link.

@@
 
+**Github**: [https://github.com/torbjornso/ISSSV1337group5](https://github.com/torbjornso/ISSSV1337group5)
+
 <!--# Starting report on second page to avoid a tiny text portion below the toc-->
@@
 
-## Phase 1: How to harvest and analyse the data? {.smaller}
-
-OPX wanted us to harvest the organisations' financial data from the web.
+## Phase 1: How to harvest and analyse the data?
+
+OPX wanted us to harvest the organizations' financial data from the web.
 During the first stage of brainstorming, we wanted to find a simple and
@@
 different websites, but quickly decided that this would be too
-difficult, and that it would only work with selected organisations. The next
+difficult, and that it would only work with selected organizations. The
 suggested solution was to create code that could read PDF documents
@@
 
-```{r, echo=FALSE, fig.cap="Need caption"}
-knitr::include_graphics("Images/lagret_tekst.png") # må finne ut av hvordan vi deler disse bildene
-
-# inkluder en med årsregnskapet som pdf her
-```
-
-## Phase 2: Exploring solutions {.smaller}
+```{r, echo=FALSE, fig.cap="Norwegian Cancer Society 2021: The financial report saved as text in the PDF."}
+# Illustration some text some pictures in financial statements
+knitr::include_graphics("Images/lagret_tekst.png") # Kreftforeningen 2021
+```
+
+```{r, echo=FALSE, fig.cap="Rainforest Foundation 2018: The financial report saved as an image in the PDF."}
+# Illustration some text some pictures in financial statements
+knitr::include_graphics("Images/lagret_bilde.png") # Regnskogfondet 2018
+```
+
+## Phase 2: Exploring solutions
 
@@
 
-- Financial statements in annual report PDF documents. 
+- Financial statements in anual report PDF documents. 
 - Web scrape Proff.no for financial statements. 
@@
 were feasible or not. The first option of using PDF documents still had
-the issue of financial statements not always being published as text, as well as their often wildly differing formats.
-Unfortunately, we quickly found that Proff.no disallows any automated collection of data from their website.
-The Brønnøysund registry does offer API access to all the financial statements, but we discovered that access currently costs NOK 480,000.
-
-We started to look at workarounds for this problem. It could still be
-possible to download Proff.no HTML websites manually for each
-organisation. We did receive some XML files from Statistics Norway, but using these is ultimately not a viable solution for OPX.
+the issue of financial statements not always being published as text.
+Also, we quickly found that Proff.no disallowed web scraping on their
+website. The Brønnøysund registry did offer API access to all the financial statements on their website, but we discovered that
+access to these data costs about half a million NOK.
+
+We started to look at workarounds for this problem. It would still be
+possible to download Proff.no HTML web pages manually for each
+organisation. We also received access to some XML files from Statistics Norway, but this would not be a viable solution for OPX in the long term.
 
@@
 
-```{r, echo=FALSE, fig.cap="Need caption", fig.align='center'}
-knitr::include_graphics("Images/info_kreft.png")
-```
-
-```{r, echo=FALSE, out.width="100%", fig.cap="Need caption"}
-knitr::include_graphics("Images/regnskapssti.png")
-```
-
-When also discovering that not all organisations publish their financial
+```{r, echo=FALSE, fig.cap="Accounting items for the Norwegian Cancer Society 2020 and the Norwegian Diabetes Association 2020, gathered from the Brønnøysund registry.", fig.align='center'}
+# Illustration brns vs OPX accounting items
+knitr::include_graphics("Images/Brns_items1.png")
+```
+
+```{r, echo=FALSE, fig.cap="Accounting items for the Norwegian Cancer Society 2020 and the Norwegian Diabetes Association 2020, gathered from OPX.", fig.align='center'}
+# Illustration brns vs OPX accounting items
+knitr::include_graphics("Images/OPX_items1.png")
+```
+
+```{r, echo=FALSE, fig.cap="An attempt at marrying the two sets of accounting items.", fig.align='center'}
+# Illustration brns vs OPX accounting items
+knitr::include_graphics("Images/Marriage.png")
+```
+
+When we also discovered that not all organisations publish their financial
 statements, and certainly not for every year, we decided to give up on
-collecting financial data for OPX, as the task turned out to impossible given current reporting requirements and practices. This is when we moved on to the third
+collecting financial data for OPX. This is when we moved on to the third
 and final stage.
 
-## Phase 3: Scrapping the scraping {.smaller}
+```{r, echo=FALSE, out.width="100%", fig.cap="The Rainforest Foundation stopped posting financial statement PDFs on their website in 2018."}
+# Illustration only published for some years example
+knitr::include_graphics("Images/Regnskogfondet_stops2018.png")
+```
+
+```{r, echo=FALSE, out.width="50%", fig.align='center', fig.cap="The Norwegian Cancer Society started posting financial statement PDFs on their website in 2019."}
+# Illustration only published for some years example
+knitr::include_graphics("Images/Kreftforeningen_starts2019.png")
+```
+
+## Phase 3: Scrapping the scraping
 
@@
 potential solutions. We experimented on coding such that data could have
-been imported from XML files. However, once we realised that the available
+been imported from XML files. However, once we realised that the
 financial data was of little use, this experimentation stopped.
@@
 financial data that OPX had already gathered from different
-organizations and organized into useful accounting items.
+organisations and organised into useful accounting items.
 
 We copied these financial statements manually into an Excel document.
-Perhaps a 'simple' solution, however our code requires OPX to import
-both an Excel document themselves into shiny, as well as a HTML document
-from Proff.no.
+Perhaps a 'simple' solution, however our code requires OPX to upload
+both an Excel document, as well as a HTML document
+from Proff.no, to the dashboard.
+
+```{r, echo=FALSE, out.width="100%", fig.cap="The data format used to upload to the dashboard."}
+# Bilde av den "endelige" excel-versjonen som brukes til å laste inn i shiny
+knitr::include_graphics("Images/Excel_final.png")
+```
 
@@
 
-<!--# Fra Visualisations (nedenfor): The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.-->
-
-<!--Legg inn bilder her-->
-
-<!-- # Legger inn fire grafer som viser manglende data i Proffs årsregnskap-->
+Our data was gathered from OPX's own web page, and from Proff.no. As described 
+in the previous parts, we wanted to gather data from the annual reports or by 
+scraping the Brønnøyund registry or Proff.no, but found this to be impossible. The data used is the one that OPX has gathered themselves directly 
+from the organisations. This data was also used to test that our codes in R 
+were correct.
+
+We have also used Proff.no to visualise the annual incomes and some KPIs. 
+This is data that the organisations have reported themselves to Proff.no. 
+Since there are no regulations regarding reporting data for NGOs in Norway, 
+the organisations have not been consistent with the data reporting. This may 
+explain the lacunae in the presented data (see e.g. figures 9 and 10).
 
@@
 
-```{r egenkapital, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Total equity/liabilities (in NOK 1,000) data for the Norwegian Cancer Society from Proff.no"}
-kreftforeningen |>
-  ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
-  scale_fill_green(kreftforeningen$balance_positive) +
-  labs(
-    x = "", y = ""
-  ) +
-  theme_light() +
-  theme(legend.position = "none")
-```
-
-```{r CEO, echo=FALSE, fig.align='center', fig.asp=0.618, fig.cap="CEO compensation data for the Norwegian Cancer Society from Proff.no"}
-kreftforeningen |>
-  # preparing for pivoting
-  select(
-    year, lederlonn, leder_annen_godtgjorelse
-  ) |>
-  # Compensation in ordinary numbers
-  mutate(
-    lederlonn = lederlonn * 1000,
-    leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
-  ) |>
-  # renaming for the purposes of the plot
-  rename(
-    Year = year,
-    Salary = lederlonn,
-    `Other compensation` = leder_annen_godtgjorelse
-  ) |>
-  # pivoting to join the two values to use fill
-  pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
-  ggplot(aes(Year, value, fill = Compensation)) +
-  geom_col() +
-  # using currency scale function to add correct currency code or ?
-  scale_y_continuous(labels = scales::label_dollar(prefix = "NOK ")) +
-  # pretty colours
-  scale_fill_met_d("Hokusai2", direction = -1) +
-  labs(
-    x = "", y = ""
-  ) +
-  theme_light() +
-  theme(legend.position = "top")
-```
-
 ```{r care_arsresultat, echo=FALSE, fig.align='center', fig.asp=0.5, fig.cap="Net income/loss (in NOK 1,000) data for CARE Norway from Proff.no"}

### 2022-08-03 Dashboard/NGO Analysis dashboard v006.Rmd
Adding links.

@@
     theme: lumen
+    source_code: https://github.com/torbjornso/ISSSV1337group5
 runtime: shiny
@@
 
-Then upload an Excel file in the set format. <!--# Need to insert link when repository is public -->
+Then upload an Excel file in the [set format](https://github.com/torbjornso/ISSSV1337group5/blob/27d1d4c25882091d6aef51956000f7ef3fefff75/Dashboard/Data/NGO_template.xltx).
 

### 2022-08-03 Report/Group 5 report.Rmd
A space, for safety.

@@
 ```
+
 #### Key Performance Indicators
